### 2020-04-28 README.Rmd
Add intro code

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# tempCycleways

<!-- badges: start -->
<!-- badges: end -->

The goal of this project is to flag roads on which there is

- high cycling potential
- road space that could be re-allocated

in the context of increased for demand for cycling to keyworker workplaces.

It is based on an analysis of data generated for the Department for Transport funded projects the Propensity to Cycle Tool (PCT) and the Cycling Infrastructure Prioritisation Toolkit (CyIPT).

```{r cars}
# Aim: identify wide roads that have high cycling potential

# setup -------------------------------------------------------------------

library(tidyverse)
library(sf)
library(tmap)
tmap_mode("view")

# input data --------------------------------------------------------------

up = "https://github.com/cyipt/cyipt-bigdata/raw/master/osm-prep/Leeds/osm-lines.Rds"
roads = readRDS(url(up))
names(roads)
summary(roads$width)
wide_potential1 = roads %>% filter(width > 10, pct.dutch > 100)
mapview::mapview(wide_potential1)
wide_potential2 = roads %>% filter(lanes.forward > 1, pct.dutch > 100)
mapview::mapview(wide_potential2)
wide_potential3 = roads %>% filter(lanes.forward + lanes.backward > 2, pct.dutch > 100)
mapview::mapview(wide_potential3)

# relationship between cycling and 2 way ----------------------------------
cor(roads$lanes.forward, roads$pct.census) # there is a small correlation
cor(roads$lanes.forward, roads$pct.dutch) # larger for go dutch scenario
cor(roads$width, roads$pct.census, use = "complete.obs") # also a positive correlation
plot(roads$width, roads$lanes.forward + roads$lanes.backward)

# out-takes ---------------------------------------------------------------

# u = "https://github.com/cyipt/cyipt-bigdata/raw/master/osm-clean/Leeds/osm-lines.Rds"
# osm_lines = readRDS(url(u))
# names(osm_lines)
# plot(osm_lines) # long time to plot...
# 
# uz = "https://github.com/cyipt/cyipt-bigdata/raw/master/forDB/roads.zip"
# d = file.path(tempdir(), "roads.zip")
# download.file(uz, d)
# r = readr::read_csv(d)
# head(r$elevation)

```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!

### 2020-04-28 README.Rmd
Add get-db script

@@
 up = "https://github.com/cyipt/cyipt-bigdata/raw/master/osm-prep/Leeds/osm-lines.Rds"
+up = "https://github.com/cyipt/cyipt-bigdata/raw/master/osm-prep/Cambridge/osm-lines.Rds"
+up = "https://github.com/cyipt/cyipt-bigdata/raw/master/osm-prep/Cambridge/osm-lines.Rds"
+up = "https://github.com/cyipt/cyipt-bigdata/raw/master/osm-recc/Hereford/osm-lines.Rds"
+
 roads = readRDS(url(up))
@@
 mapview::mapview(wide_potential1)
-wide_potential2 = roads %>% filter(lanes.forward > 1, pct.dutch > 100)
+wide_potential2 = roads %>% filter(
+  lanes.forward > 1 |
+  lanes.backward > 1,
+  pct.dutch > 50)
 mapview::mapview(wide_potential2)

### 2020-04-28 get-data-db.R

@@ -1,0 +1,46 @@
+# Aim: get, filter and re-upload data for regions
+
+library(sf)
+library(dplyr)
+
+# get large datasets ------------------------------------------------------
+
+old = setwd("~/cyipt/cyipt-server-data/malcolm/")
+zip("roads.zip", "roads.csv")
+piggyback::pb_upload("roads.zip", "cyipt/cyipt-phase-1-data")
+f = list.files(pattern = "csv")
+f = f[-5]
+z = gsub(pattern = ".csv", replacement = ".zip", f)
+for(i in seq_along(f)) {
+  zip(z[i], f[i])
+  piggyback::pb_upload(z[i], "cyipt/cyipt-phase-1-data")
+}
+setwd(old)
+
+# test read-in as geographic object ---------------------------------------
+
+r = data.table::fread("roads.csv")
+names(r)
+r1 = r[1, ]
+r1$geotext
+r1_sf = sf::st_as_sfc(r1$geotext, crs = 4326)
+# time-consuming
+# r_sfc_all = sf::st_as_sfc(r$geotext, crs = 4326)
+
+# subset for regions of interest ------------------------------------------
+
+regions_of_interest = 
+u = "https://github.com/npct/pct-shiny/raw/master/regions_www/www/la-map-resources/la.js"
+ul = readLines(url(u))
+ul[1]
+ul[1] = "{"
+writeLines(ul, "/tmp/ul.json")
+pct_la_summaries = jsonlite::read_json("/tmp/ul.json")
+str(pct_la_summaries)
+pct_la_summaries = sf::read_sf("/tmp/ul.json")
+head(pct_la_summaries)
+
+summary()
+
+regions = unique(f$region)
+dput(regions)

### 2020-04-28 get-data-db.R
Update get db code

@@ -29,7 +29,13 @@ r1_sf = sf::st_as_sfc(r1$geotext, crs = 4326)
 
 # subset for regions of interest ------------------------------------------
 
-regions_of_interest = 
+regions = unique(r$region)
+regions_cyipt_df = data.frame(regions)
+View(regions_cyipt_df)
+regions_of_interest = c(
+  "Leeds", "Liverpool", "London", "Bristol",
+  "Birmingham", "Manchester", "Newcastle"
+  )
 u = "https://github.com/npct/pct-shiny/raw/master/regions_www/www/la-map-resources/la.js"
 ul = readLines(url(u))
 ul[1]
@@ -40,7 +46,57 @@ str(pct_la_summaries)
 pct_la_summaries = sf::read_sf("/tmp/ul.json")
 head(pct_la_summaries)
 
-summary()
+summary((s = regions_of_interest %in% pct_la_summaries$name))
+regions_of_interest[!s]
+summary((s2 = regions_of_interest %in% pct_la_summaries$name))
+pct_la_summaries$name[grepl(pattern = "Bristol", pct_la_summaries$name)] = "Bristol"
+pct_la_summaries$name[grepl(pattern = "Newcastle u", pct_la_summaries$name)] = "Newcastle"
+pct_la_summaries$name[grepl(pattern = "Hull", pct_la_summaries$name)] = "Hull"
+summary((s2 = regions_of_interest %in% pct_la_summaries$name))
+summary(s2) # 6 out of 8
 
-regions = unique(f$region)
-dput(regions)
+lnd = spData::lnd
+
+summary(pct_la_summaries$name %in% lnd$NAME)
+summary(lnd$NAME %in% pct_la_summaries$name)
+lnd$NAME[!lnd$NAME %in% pct_la_summaries$name]
+pct_centroids = sf::st_centroid(pct_la_summaries)
+pct_london_points = pct_centroids[lnd, ]
+pct_london = pct_la_summaries %>% 
+  filter(name %in% pct_london_points$name)
+plot(pct_london)
+pct_london_aggregated = pct_london %>% 
+  summarise_if(is.numeric, sum) %>% 
+  add_column(name = "London", .before = 1) %>% 
+  add_column(CODE = "1", .before = 1)
+setdiff(names(pct_london_aggregated), names(las_top))
+mapview::mapview(pct_london_aggregated)  
+
+las_top = pct_la_summaries %>% 
[…]

### 2020-04-28 README.Rmd
Add updated readme

--- 2020-04-28 README.Rmd
@@
 knitr::opts_chunk$set(
+  warning = FALSE,
+  echo = FALSE,
   collapse = TRUE,
@@
 )
+library(tidyverse)
+library(sf)
+library(tmap)
+tmap_mode("view")
+piggyback::pb_download("las_top.Rds")
+las_top = readRDS("las_top.Rds")
+# piggyback::pb_download("rsf.Rds", "cyipt/cyipt-phase-1-data")
 ```
@@
 
-```{r cars}
+The first stage was to identify cities with high cycling potential.
+We did this by analysing data from the PCT project and selecting the top 10 cities, presented in the table below and figure below:
+
+```{r cities}
+las_top %>% 
+  select(name, all, bicycle, dutch_slc) %>% 
+  sf::st_drop_geometry() %>%
+  knitr::kable(caption = "Top 10 cities by cycling potential in England, with 'all' representing all commuters in the 2011 Census, 'bicycle' representing the number who cycled to work and 'dutch_slc' the number who could cycle to work under a 'Go Dutch' scenario of cycling uptake.")
+```
+
+```{r}
+tm_shape(las_top) +
+  tm_polygons("blue")
+```
+
+# London
+
+# Leeds
+
+```{r cycling}
 # Aim: identify wide roads that have high cycling potential
-
-# setup -------------------------------------------------------------------
-
-library(tidyverse)
-library(sf)
-library(tmap)
-tmap_mode("view")
 
@@
 roads = readRDS(url(up))
-names(roads)
-summary(roads$width)
+# names(roads)
+# summary(roads$width)
 wide_potential1 = roads %>% filter(width > 10, pct.dutch > 100)
-mapview::mapview(wide_potential1)
-wide_potential2 = roads %>% filter(lanes.forward > 1, pct.dutch > 100)
+wide_potential2 = roads %>% filter(
+  lanes.forward > 1 |
+  lanes.backward > 1,
+  pct.dutch > 50)
 mapview::mapview(wide_potential2)
-wide_potential3 = roads %>% filter(lanes.forward + lanes.backward > 2, pct.dutch > 100)
-mapview::mapview(wide_potential3)
 
-# relationship between cycling and 2 way ----------------------------------
-cor(roads$lanes.forward, roads$pct.census) # there is a small correlation
-cor(roads$lanes.forward, roads$pct.dutch) # larger for go dutch scenario
-cor(roads$width, roads$pct.census, use = "complete.obs") # also a positive correlation
-plot(roads$width, roads$lanes.forward + roads$lanes.backward)
+# # relationship between cycling and 2 way ----------------------------------
+# cor(roads$lanes.forward, roads$pct.census) # there is a small correlation
+# cor(roads$lanes.forward, roads$pct.dutch) # larger for go dutch scenario
+# cor(roads$width, roads$pct.census, use = "complete.obs") # also a positive correlation
+# plot(roads$width, roads$lanes.forward + roads$lanes.backward)
+#
 

### 2020-04-28 get-data-db.R

@@ -72,31 +72,112 @@ pct_london_aggregated = pct_london %>%
 setdiff(names(pct_london_aggregated), names(las_top))
 mapview::mapview(pct_london_aggregated)  
 
+pct_la_summaries %>% 
+  filter(!grepl(pattern = "Wand|West|Southw|Lamb|Hack|Towe|Eal|Houns|Wilt|Cheshire E|Sef|Isl|Riding", name)) %>% 
+  select(name, all, bicycle, dutch_slc) %>% 
+  arrange(desc(dutch_slc)) %>%
+  head(15) 
+
 las_top = pct_la_summaries %>% 
   filter(!grepl(pattern = "Wand|West|Southw|Lamb|Hack|Towe|Eal|Houns|Wilt|Cheshire E|Sef|Isl|Riding", name)) %>% 
   arrange(desc(dutch_slc)) %>%
   head(8) 
 
 las_other = pct_la_summaries %>% 
-  filter(name == "Newcastle")
+  filter(name == "Newcastle" | name == "Gateshead") %>% 
+  summarise_if(is.numeric, sum) %>% 
+  add_column(name = "Newcastle", .before = 1) %>% 
+  add_column(CODE = "10", .before = 1)
+
+las_other$dutch_slc
 
 names(pct_london_aggregated)
 names(las_top)
-las_top = rbind(pct_london_aggregated, las_top, las_other)
+las_top = rbind(pct_london_aggregated, las_top, las_other) %>% 
+  arrange(desc(dutch_slc))
 
 las_top %>% 
   select(name, all, bicycle, dutch_slc) %>% 
   sf::st_drop_geometry() %>%
   knitr::kable()
 
+pct_la_summaries %>% 
+  select(name, all, bicycle, dutch_slc) %>% 
+  sf::st_drop_geometry() %>%
+  filter(name == "Cambridge") %>% 
+  knitr::kable()
 
 sum(las_top$all) / sum(pct_la_summaries$all) # 9%
 mapview::mapview(las_top)
 
 regions_of_interest[!regions_of_interest %in% regions] # just Cardiff not in there
-regions_of_interest2 = 
 regions2 = c(regions_of_interest, las_top$name[!las_top$name %in% regions_of_interest])
 length(regions2)
 regions[!regions %in% regions2]
 regions2
 regions2[!regions2 %in% regions]
+regions2[!regions2 %in% las_top$name]
+
+r_regions = r %>% 
+  filter(region %in% regions2)
+
+nrow(r_regions) / nrow(r)
[…]

### 2020-04-28 README.Rmd
Update results

@@
 las_top = readRDS("las_top.Rds")
+rsf = readRDS("rsf_grouped_long.Rds")
 # piggyback::pb_download("rsf.Rds", "cyipt/cyipt-phase-1-data")
@@
 
-# tempCycleways
+```{r, eval=FALSE}
+# general analysis of data
+rsf = readRDS("rsf.Rds")
+names(rsf)
+rtid = readr::read_csv("~/cyipt/cyipt-server-data/malcolm/roadtypes.csv")
+rsf = left_join(rsf, rtid)
+rsf_spare_lane = rsf %>% filter(
+  (lanespsvforward + lanesforward) > 1 |
+  (lanespsvbackward + lanesbackward) > 1,
+  ! grepl(pattern = "motorway", highway)
+  ) 
+saveRDS(rsf_spare_lane, "rsf_spare_line.Rds")
+
+rsf = readRDS("rsf_spare_line.Rds")
+summary(rsf$length)
+unique(rsf$region)
+head(rsf$name)
+rsf %>% 
+  select(calcwidthnow) %>% 
+  sf::st_drop_geometry() %>%
+  group_by(calcwidthnow) %>% 
+  summarise(n = n()) %>% 
+  filter(n > 500)
+wide_potential1 = rsf %>% filter(width > 10, pct.dutch > 100)
+wide_potential2 = rsf %>% filter(
+  (lanespsvforward + lanesforward) > 1 |
+  (lanespsvbackward + lanesbackward) > 1,
+  pctdutch > 50)
+# mapview::mapview(wide_potential2) # very patchy
+rsf_grouped = rsf %>% 
+  group_by(name, ref, highway, region) %>% 
+  summarise(pctgov = mean(pctgov)) %>% 
+  ungroup()
+
+rsf_grouped$length = sf::st_length(rsf_grouped) %>% as.numeric()
+summary(rsf_grouped$length)
+rsf_grouped_long = rsf_grouped %>% filter(length > 100)
+# mapview::mapview(rsf_grouped_long)
+rsf_grouped_long %>%
+  st_drop_geometry() %>% 
+  arrange(desc(pctgov)) %>%
+  group_by(region) %>% 
+  slice(1:10) 
+tm_shape(rsf_grouped_long) + tm_lines("pctgov", lwd = 3, palette = "viridis")  
+saveRDS(rsf_grouped_long, "rsf_grouped_long.Rds")
+```
+
+# Data analysis to support temporary cycleways
 
@@
 
+```{r}
+roads = rsf %>% filter(region == "London", pctgov > 500)
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_view(set.zoom.limits = c(12, 13)) + tm_scale_bar()
+```
+
+The top 10 roads:
+
+```{r}
+roads_top = roads %>%
+  filter(length > 500) %>% 
+  arrange(desc(pctgov)) %>%
+  slice(1:10) 
+tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_text("name") +
+  tm_view(set.zoom.limits = c(13, 14)) + tm_scale_bar()
+
+roads_top %>%
+  st_drop_geometry() %>% 
+  knitr::kable(digits = 0)
+```
+
 # Leeds
 
-```{r cycling}
-# Aim: identify wide roads that have high cycling potential
-
-# input data --------------------------------------------------------------
-
-up = "https://github.com/cyipt/cyipt-bigdata/raw/master/osm-prep/Leeds/osm-lines.Rds"
-roads = readRDS(url(up))
-# names(roads)
-# summary(roads$width)
-wide_potential1 = roads %>% filter(width > 10, pct.dutch > 100)
-wide_potential2 = roads %>% filter(
-  lanes.forward > 1 |
-  lanes.backward > 1,
-  pct.dutch > 50)
-mapview::mapview(wide_potential2)
-
-# # relationship between cycling and 2 way ----------------------------------
-# cor(roads$lanes.forward, roads$pct.census) # there is a small correlation
-# cor(roads$lanes.forward, roads$pct.dutch) # larger for go dutch scenario
-# cor(roads$width, roads$pct.census, use = "complete.obs") # also a positive correlation
-# plot(roads$width, roads$lanes.forward + roads$lanes.backward)
-#
-
-# out-takes ---------------------------------------------------------------
-
-# u = "https://github.com/cyipt/cyipt-bigdata/raw/master/osm-clean/Leeds/osm-lines.Rds"
-# osm_lines = readRDS(url(u))
-# names(osm_lines)
-# plot(osm_lines) # long time to plot...
-# 
-# uz = "https://github.com/cyipt/cyipt-bigdata/raw/master/forDB/roads.zip"
-# d = file.path(tempdir(), "roads.zip")
-# download.file(uz, d)
-# r = readr::read_csv(d)
-# head(r$elevation)
-
+```{r}
+roads = rsf %>% filter(region == "Leeds", pctgov > 100)
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_view(set.zoom.limits = c(13, 14)) + tm_scale_bar()
 ```
 
-In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!
+The top 10 roads:
+
+```{r}
+roads_top = roads %>%
+  filter(length > 500) %>% 
+  arrange(desc(pctgov)) %>%
+  slice(1:10) 
+tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_text("name") +
+  tm_view(set.zoom.limits = c(13, 14)) + tm_scale_bar()
+
+roads_top %>%
+  st_drop_geometry() %>% 
+  knitr::kable(digits = 0)
+```
+
+# Birmingham
+
+```{r}
+roads = rsf %>% filter(region == "Birmingham", pctgov > 100)
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_view(set.zoom.limits = c(13, 14)) + tm_scale_bar()
+```
+
+The top 10 roads:
+
+```{r}
+roads_top = roads %>%
+  filter(length > 500) %>% 
+  arrange(desc(pctgov)) %>%
+  slice(1:10) 
+tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_text("name") +
+  tm_view(set.zoom.limits = c(13, 14)) + tm_scale_bar()
+
+roads_top %>%
+  st_drop_geometry() %>% 
+  knitr::kable(digits = 0)
+```

### 2020-04-28 README.Rmd
Update results with Cambridge and Sheffield

@@
 The first stage was to identify cities with high cycling potential.
-We did this by analysing data from the PCT project and selecting the top 10 cities, presented in the table below and figure below:
+We did this by analysing data from the PCT project and selecting the top 10 cities in terms of long term cycling potential, plus Sheffield and Cambridge, presented in the table and figure below:
 
@@
 ```
+
+These 12 study cities coincidentally represent almost exactly 1/4 of the population in England.
+Welsh and Scottish cities with high cycling potential such as Cardiff and Edinbugh were not included in the analysis because the CyIPT does not currently have data outside of England, although we could extend the methods to cover all UK cities at some point.
+
+Below we show the results for a selection of cities, with cycling potential on the road network visualised under the 'Government Target' scenario, which represents a doubling in cycling compared with 2011 levels.
+London is close to meeting this target already.
+
+We filtered-out roads with low levels of cycling potential and focus only on roads that have at least one 'spare lane', defined as having more than 1 lane in either direction.
+Roads that could be made oneway, or that could be converted into 'liveable streets' by preventing through traffic were not considered.
 

### 2020-04-28 get-data-db.R

@@ -1,7 +1,7 @@
 # Aim: get, filter and re-upload data for regions
 
 library(sf)
-library(dplyr)
+library(tidyverse)
 
 # get large datasets ------------------------------------------------------
 
@@ -41,8 +41,6 @@ ul = readLines(url(u))
 ul[1]
 ul[1] = "{"
 writeLines(ul, "/tmp/ul.json")
-pct_la_summaries = jsonlite::read_json("/tmp/ul.json")
-str(pct_la_summaries)
 pct_la_summaries = sf::read_sf("/tmp/ul.json")
 head(pct_la_summaries)
 
@@ -88,12 +86,15 @@ las_other = pct_la_summaries %>%
   summarise_if(is.numeric, sum) %>% 
   add_column(name = "Newcastle", .before = 1) %>% 
   add_column(CODE = "10", .before = 1)
-
 las_other$dutch_slc
+
+las_other2 = pct_la_summaries %>% 
+  filter(name %in% c("Cambridge", "Sheffield"))
+
 
 names(pct_london_aggregated)
 names(las_top)
-las_top = rbind(pct_london_aggregated, las_top, las_other) %>% 
+las_top = rbind(pct_london_aggregated, las_top, las_other, las_other2) %>% 
   arrange(desc(dutch_slc))
 
 las_top %>% 

### 2020-04-28 README.Rmd
Update with results for Cambridge
Heads-up @mvl22

@@
 saveRDS(rsf_spare_lane, "rsf_spare_line.Rds")
+# piggyback::pb_upload("rsf_spare_line.Rds", "cyipt/cyipt-phase-1-data")
 
@@
 
-# Birmingham
+# Cambridge
 
 ```{r}
-roads = rsf %>% filter(region == "Birmingham", pctgov > 100)
-tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_view(set.zoom.limits = c(13, 14)) + tm_scale_bar()
+roads = rsf %>% filter(region == "Cambridge", pctgov > 400)
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis")
 ```
@@
 roads_top = roads %>%
-  filter(length > 500) %>% 
+  filter(length > 100) %>% 
   arrange(desc(pctgov)) %>%

### 2020-04-28 get-data-db.R

@@ -15,11 +15,11 @@ for(i in seq_along(f)) {
   zip(z[i], f[i])
   piggyback::pb_upload(z[i], "cyipt/cyipt-phase-1-data")
 }
+r = data.table::fread("roads.csv")
 setwd(old)
 
 # test read-in as geographic object ---------------------------------------
 
-r = data.table::fread("roads.csv")
 names(r)
 r1 = r[1, ]
 r1$geotext
@@ -31,7 +31,7 @@ r1_sf = sf::st_as_sfc(r1$geotext, crs = 4326)
 
 regions = unique(r$region)
 regions_cyipt_df = data.frame(regions)
-View(regions_cyipt_df)
+# View(regions_cyipt_df)
 regions_of_interest = c(
   "Leeds", "Liverpool", "London", "Bristol",
   "Birmingham", "Manchester", "Newcastle"
@@ -90,7 +90,6 @@ las_other$dutch_slc
 
 las_other2 = pct_la_summaries %>% 
   filter(name %in% c("Cambridge", "Sheffield"))
-
 
 names(pct_london_aggregated)
 names(las_top)

### 2020-04-28 README.Rmd
Update results for Leeds

@@
   tm_text("name") +
-  tm_view(set.zoom.limits = c(13, 14)) + tm_scale_bar()
+  tm_view(set.zoom.limits = c(12, 13)) + tm_scale_bar()
 
@@
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_view(set.zoom.limits = c(13, 14)) + tm_scale_bar()
+  tm_view(set.zoom.limits = c(12, 13)) + tm_scale_bar()
 ```
@@
 roads = rsf %>% filter(region == "Cambridge", pctgov > 400)
-tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis")
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_view(set.zoom.limits = c(13, 14)) + tm_scale_bar()
 ```
@@
   tm_text("name") +
-  tm_view(set.zoom.limits = c(13, 14)) + tm_scale_bar()
+  tm_view(set.zoom.limits = c(14, 15)) + tm_scale_bar()
 

### 2020-04-29 README.Rmd
Add results for Bristol, heads-up @joeytalbot

@@
 knitr::opts_chunk$set(
+  message = FALSE,
   warning = FALSE,
@@
 tmap_mode("view")
-piggyback::pb_download("las_top.Rds")
+# piggyback::pb_download("las_top.Rds")
+pct_las = sf::read_sf("https://github.com/cyipt/tempCycleways/releases/download/0.1/pct_la_summaries.geojson")
 las_top = readRDS("las_top.Rds")
 rsf = readRDS("rsf_grouped_long.Rds")
-# piggyback::pb_download("rsf.Rds", "cyipt/cyipt-phase-1-data")
 ```
@@
 # general analysis of data
+# piggyback::pb_download("rsf.Rds", "cyipt/cyipt-phase-1-data")
 rsf = readRDS("rsf.Rds")
@@
 saveRDS(rsf_grouped_long, "rsf_grouped_long.Rds")
+piggyback::pb_upload("rsf_grouped_long.Rds")
 ```
@@
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_view(set.zoom.limits = c(13, 14)) + tm_scale_bar()
+  tm_scale_bar()
 ```
@@
   tm_text("name") +
-  tm_view(set.zoom.limits = c(14, 15)) + tm_scale_bar()
+  tm_view(set.zoom.limits = c(14, 150)) + tm_scale_bar()
 
@@
 ```
+
+# Bristol
+
+```{r}
+region = pct_las %>% filter(name == "Bristol")
+region = sf::st_make_valid(region)
+roads = rsf %>% filter(region == "Bristol", pctgov > 100)
+roads = roads[region, ]
+roads = st_intersection(roads, region)
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_shape(region) + tm_borders() +
+  tm_scale_bar() 
+```
+
+The top 10 roads:
+
+```{r}
+bristol_centre = tmaptools::geocode_OSM("Bristol", as.sf = TRUE)
+bristol_buff = stplanr::geo_buffer(bristol_centre, dist = 1500)
+roads_top = roads[bristol_buff, ] %>%
+  filter(length > 200) %>% 
+  arrange(desc(pctgov)) %>%
+  slice(2:10) %>% 
+  distinct(name, .keep_all = TRUE)
+tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_text("name") +
+  tm_scale_bar()
+
+roads_top %>%
+  st_drop_geometry() %>% 
+  knitr::kable(digits = 0)
+```

### 2020-04-29 README.Rmd
Update list of cities, now 5 + 5

@@
 
-The first stage was to identify cities with high cycling potential.
-We did this by analysing data from the PCT project and selecting the top 10 cities in terms of long term cycling potential, plus Sheffield and Cambridge, presented in the table and figure below:
+To overcome issues associated with doing this at a national level and elicit feedback on the methods and preliminary results, we ran the analysis on a sample of cities.
+We chose the top 5 in terms of absolute long-term cycling potential (London, Birmingham, Manchester, Leeds, Liverpool) plus an additional 5 cities that have active advocacy groups (Newcastle, Sheffield, Cambridge, Bristol, Leicester).
+Estimates of current and potential numbers who could cycle to work in these cities is presented in the table below.
+<!-- We did this by analysing data from the PCT project and selecting the top 10 cities in terms of long term cycling potential, plus Sheffield and Cambridge, : -->
 
@@
   sf::st_drop_geometry() %>%
-  knitr::kable(caption = "Top 10 cities by cycling potential in England, with 'all' representing all commuters in the 2011 Census, 'bicycle' representing the number who cycled to work and 'dutch_slc' the number who could cycle to work under a 'Go Dutch' scenario of cycling uptake.")
+  knitr::kable(caption = "Selection of 10 cities in England with high cycling potential or active adovcacy groups.
+'All' represents all commuters in the 2011 Census, 'bicycle' represents the number who cycled to work and 'dutch_slc' the number who could cycle to work under a 'Go Dutch' scenario of cycling uptake.")
 ```
+
+The geographic distribution of these cities is shown in the map below:
 

### 2020-04-29 get-data-db.R

@@ -76,24 +76,31 @@ pct_la_summaries %>%
   arrange(desc(dutch_slc)) %>%
   head(15) 
 
+pct_la_summaries %>% 
+  filter(!grepl(pattern = "Wand|West|Southw|Lamb|Hack|Towe|Eal|Houns|Wilt|Cheshire E|Sef|Isl|Riding", name)) %>% 
+  arrange(desc(govtarget_slc)) %>%
+  select(name, all, bicycle, dutch_slc) %>% 
+  st_drop_geometry() %>% 
+  head(15) 
+
 las_top = pct_la_summaries %>% 
   filter(!grepl(pattern = "Wand|West|Southw|Lamb|Hack|Towe|Eal|Houns|Wilt|Cheshire E|Sef|Isl|Riding", name)) %>% 
   arrange(desc(dutch_slc)) %>%
-  head(8) 
+  head(4) 
 
-las_other = pct_la_summaries %>% 
-  filter(name == "Newcastle" | name == "Gateshead") %>% 
-  summarise_if(is.numeric, sum) %>% 
-  add_column(name = "Newcastle", .before = 1) %>% 
-  add_column(CODE = "10", .before = 1)
-las_other$dutch_slc
+# las_other = pct_la_summaries %>% 
+#   filter(name == "Newcastle" | name == "Gateshead") %>% 
+#   summarise_if(is.numeric, sum) %>% 
+#   add_column(name = "Newcastle", .before = 1) %>% 
+#   add_column(CODE = "10", .before = 1)
+# las_other$dutch_slc
 
 las_other2 = pct_la_summaries %>% 
-  filter(name %in% c("Cambridge", "Sheffield"))
+  filter(name %in% c("Cambridge", "Sheffield", "Newcastle", "Bristol", "Leicester"))
 
 names(pct_london_aggregated)
 names(las_top)
-las_top = rbind(pct_london_aggregated, las_top, las_other, las_other2) %>% 
+las_top = rbind(pct_london_aggregated, las_top, las_other2) %>% 
   arrange(desc(dutch_slc))
 
 las_top %>% 

### 2020-04-29 README.Rmd
Update results for Leeds
Heads-up @joeytalbot

@@
   st_drop_geometry() %>% 
-  knitr::kable(digits = 0)
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
 ```
@@
 roads = rsf %>% filter(region == "Leeds", pctgov > 100)
+region_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
+region_buff = stplanr::geo_buffer(region_centre, dist = 3000)
+roads = st_intersection(roads, region_buff)
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_view(set.zoom.limits = c(12, 13)) + tm_scale_bar()
+  tm_scale_bar()
 ```
@@
 ```{r}
+region_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
+region_buff = stplanr::geo_buffer(region_centre, dist = 3000)
 roads_top = roads %>%
-  filter(length > 500) %>% 
+  filter(length > 300) %>% 
   arrange(desc(pctgov)) %>%
@@
   tm_text("name") +
-  tm_view(set.zoom.limits = c(12, 13)) + tm_scale_bar()
+  tm_view(set.view = 13) + tm_scale_bar()
 

### 2020-04-29 README.Rmd
Update results for Bristol

@@
 roads_top = roads[bristol_buff, ] %>%
-  filter(length > 200) %>% 
+  filter(length > 300) %>% 
+  filter(!name == "Bond Street", !name == "") %>% 
   arrange(desc(pctgov)) %>%
-  slice(2:10) %>% 
+  slice(2:11) %>% 
   distinct(name, .keep_all = TRUE)

### 2020-04-29 README.Rmd
Add first pass for all cities

@@
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_view(set.zoom.limits = c(12, 13)) + tm_scale_bar()
+  tm_view(set.view = 11) + tm_scale_bar()
 ```
@@
   filter(length > 500) %>% 
+  filter(!name == "Talgarth Road", length > 500) %>% 
   arrange(desc(pctgov)) %>%
@@
   tm_text("name") +
-  tm_view(set.zoom.limits = c(12, 13)) + tm_scale_bar()
+  tm_view(set.view = 12) + tm_scale_bar()
+
+roads_top %>%
+  st_drop_geometry() %>% 
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
+```
+
+# Birmingham
+
+```{r}
+roads = rsf %>% filter(region == "Birmingham", pctgov > 100)
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_view(set.view = 12) + tm_scale_bar()
+```
+
+The top 10 roads:
+
+```{r}
+roads_top = roads %>%
+  filter(length > 300) %>% 
+  arrange(desc(pctgov)) %>%
+  slice(1:10) 
+tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_text("name") +
+  tm_view(set.view = 13) + tm_scale_bar()
+
+roads_top %>%
+  st_drop_geometry() %>% 
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
+```
+
+# Manchester
+
+```{r}
+roads = rsf %>% filter(region == "Manchester", pctgov > 100)
+region_centre = tmaptools::geocode_OSM("Manchester", as.sf = TRUE)
+region_buff = stplanr::geo_buffer(region_centre, dist = 7000)
+roads = st_intersection(roads, region_buff)
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_view(set.view = 12) + tm_scale_bar()
+```
+
+The top 10 roads:
+
+```{r}
+roads_top = roads %>%
+  filter(length > 300) %>% 
+  arrange(desc(pctgov)) %>%
+  slice(1:10) 
+tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_text("name") +
+  tm_view(set.view = 13) + tm_scale_bar()
 
@@
 
-# Cambridge
+# Liverpool
 
 ```{r}
-roads = rsf %>% filter(region == "Cambridge", pctgov > 400)
+roads = rsf %>% filter(region == "Liverpool", pctgov > 100)
+region_centre = tmaptools::geocode_OSM("Liverpool", as.sf = TRUE)
+region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
+roads = st_intersection(roads, region_buff)
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_scale_bar()
+  tm_view(set.view = 12) + tm_scale_bar()
 ```
@@
 roads_top = roads %>%
-  filter(length > 100) %>% 
+  filter(length > 300) %>% 
   arrange(desc(pctgov)) %>%
@@
   tm_text("name") +
-  tm_view(set.zoom.limits = c(14, 150)) + tm_scale_bar()
+  tm_view(set.view = 12) + tm_scale_bar()
 
@@
 ```
+
+# Leicester
+
+```{r}
+roads = rsf %>% filter(region == "Leicester", pctgov > 100)
+region_centre = tmaptools::geocode_OSM("Leicester", as.sf = TRUE)
+region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
+roads = st_intersection(roads, region_buff)
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_view(set.view = 12) + tm_scale_bar()
+```
+
+The top 10 roads:
+
+```{r}
+roads_top = roads %>%
+  filter(length > 300) %>% 
+  arrange(desc(pctgov)) %>%
+  slice(1:10) 
+tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_text("name") +
+  tm_view(set.view = 12) + tm_scale_bar()
+
+roads_top %>%
+  st_drop_geometry() %>% 
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
+```
+
+# Sheffield
+
+```{r}
+roads = rsf %>% filter(region == "Sheffield", pctgov > 100)
+region_centre = tmaptools::geocode_OSM("Sheffield", as.sf = TRUE)
+region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
+roads = st_intersection(roads, region_buff)
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_view(set.view = 12) + tm_scale_bar()
+```
+
+The top 10 roads:
+
+```{r}
+roads_top = roads %>%
+  filter(length > 600) %>% 
+  arrange(desc(pctgov)) %>%
+  slice(1:10) 
+tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_text("name") +
+  tm_view(set.view = 12) + tm_scale_bar()
+
+roads_top %>%
+  st_drop_geometry() %>% 
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
+```
+
+# Newcastle
+
+```{r}
+roads = rsf %>% filter(region == "Newcastle", pctgov > 100)
+region_centre = tmaptools::geocode_OSM("Newcastle", as.sf = TRUE)
+region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
+roads = st_intersection(roads, region_buff)
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_view(set.view = 12) + tm_scale_bar()
+```
+
+The top 10 roads:
+
+```{r}
+roads_top = roads %>%
+  filter(length > 600) %>% 
+  arrange(desc(pctgov)) %>%
+  slice(1:10) 
+tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_text("name") +
+  tm_view(set.view = 12) + tm_scale_bar()
+
+roads_top %>%
+  st_drop_geometry() %>% 
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
+```
+
+# Cambridge
+
+```{r}
+roads = rsf %>% filter(region == "Cambridge", pctgov > 400)
+tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_view(set.view = 13) + tm_scale_bar()
+```
+
+The top 10 roads:
+
+```{r}
+roads_top = roads %>%
+  filter(length > 100) %>% 
+  arrange(desc(pctgov)) %>%
+  slice(1:10) 
+tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  tm_text("name") +
+  tm_view(set.view = 14) + tm_scale_bar()
+
+roads_top %>%
+  st_drop_geometry() %>% 
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
+```

### 2020-04-29 README.Rmd
Update results for all 10 cities
Close #3 for now, thanks for feedback @joeytalbot, @mem48 and @mvl22

@@
   filter(length > 600) %>% 
+  filter(!name == "") %>% 
   arrange(desc(pctgov)) %>%
+  group_by(name, ref, highway) %>% 
+  summarise(pctgov = mean(pctgov), length = mean(length)) %>% 
+  ungroup() %>% 
   slice(1:10) 

### 2020-04-29 README.Rmd
Builtup area buffers

@@
 tmap_mode("view")
-# piggyback::pb_download("las_top.Rds")
+piggyback::pb_download("las_top.Rds")
 pct_las = sf::read_sf("https://github.com/cyipt/tempCycleways/releases/download/0.1/pct_la_summaries.geojson")
 las_top = readRDS("las_top.Rds")
+st_make_valid(las_top)
+piggyback::pb_download("rsf_grouped_long.Rds")
 rsf = readRDS("rsf_grouped_long.Rds")
+piggyback::pb_download("Builtup_Areas_December_2011_Boundaries_V2.geojson")
+builtup = sf::read_sf("Builtup_Areas_December_2011_Boundaries_V2.geojson") %>% sf::st_transform(27700)
 ```
@@
 ```{r}
+##Mask not required for London, so no need to slow down the code unnecessarily
+# london_centroids = get_pct_centroids("london",geography = "lsoa") %>% sf::st_transform(27700)
+# london_lsoa_buff = st_buffer(x = london_centroids, dist = 1000)
+# london_res_buff = st_intersection(builtup,london_lsoa_buff) %>% st_union()
 roads = rsf %>% filter(region == "London", pctgov > 500)
+# roads = roads %>% st_transform(27700) %>% st_intersection(london_res_buff) %>% st_transform(4326)
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
@@
 ```{r}
+birmingham_centroids = get_pct_centroids("west-midlands",geography = "lsoa") %>% sf::st_transform(27700)
+birmingham_lsoa_buff = st_buffer(x = birmingham_centroids, dist = 1000)
+birmingham_res_buff = st_intersection(builtup,birmingham_lsoa_buff) %>% st_union()
 roads = rsf %>% filter(region == "Birmingham", pctgov > 100)
+roads = roads %>% st_transform(27700) %>% st_intersection(birmingham_res_buff) %>% st_transform(4326)
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
@@
 ```{r}
+manchester_centroids = get_pct_centroids("greater-manchester",geography = "lsoa") %>% sf::st_transform(27700)
+manchester_lsoa_buff = st_buffer(x = manchester_centroids, dist = 1000)
+manchester_res_buff = st_intersection(builtup,manchester_lsoa_buff) %>% st_union()
 roads = rsf %>% filter(region == "Manchester", pctgov > 100)
+roads = roads %>% st_transform(27700) %>% st_intersection(manchester_res_buff) %>% st_transform(4326)
 region_centre = tmaptools::geocode_OSM("Manchester", as.sf = TRUE)
@@
 ```{r}
+leeds_centroids = get_pct_centroids("west-yorkshire",geography = "lsoa") %>% sf::st_transform(27700)
+leeds_lsoa_buff = st_buffer(x = leeds_centroids$geometry, dist = 1000)
+leeds_res_buff = st_intersection(builtup,leeds_lsoa_buff) %>% st_union()
 roads = rsf %>% filter(region == "Leeds", pctgov > 100)
+roads = roads %>% st_transform(27700) %>% st_intersection(leeds_res_buff) %>% st_transform(4326)
 region_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
@@
 ```{r}
+liverpool_centroids = get_pct_centroids("liverpool-city-region",geography = "lsoa") %>% sf::st_transform(27700)
+liverpool_lsoa_buff = st_buffer(x = liverpool_centroids, dist = 1000)
+liverpool_res_buff = st_intersection(builtup,liverpool_lsoa_buff) %>% st_union()
 roads = rsf %>% filter(region == "Liverpool", pctgov > 100)
@@
 roads = st_intersection(roads, region_buff)
+roads = roads %>% st_transform(27700) %>% st_intersection(liverpool_res_buff) %>% st_transform(4326)
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
@@
 ```{r}
+bristol_centroids = get_pct_centroids("avon",geography = "lsoa") %>% sf::st_transform(27700)
+bristol_lsoa_buff = st_buffer(x = bristol_centroids, dist = 1000)
+bristol_res_buff = st_intersection(builtup,bristol_lsoa_buff) %>% st_union()
 region = pct_las %>% filter(name == "Bristol")
@@
 roads = st_intersection(roads, region)
+roads = roads %>% st_transform(27700) %>% st_intersection(bristol_res_buff) %>% st_transform(4326)
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
@@
 ```{r}
+leicester_centroids = get_pct_centroids("leicestershire",geography = "lsoa") %>% sf::st_transform(27700)
+leicester_lsoa_buff = st_buffer(x = leicester_centroids, dist = 1000)
+leicester_res_buff = st_intersection(builtup,leicester_lsoa_buff) %>% st_union()
 roads = rsf %>% filter(region == "Leicester", pctgov > 100)
@@
 roads = st_intersection(roads, region_buff)
+roads = roads %>% st_transform(27700) %>% st_intersection(leicester_res_buff) %>% st_transform(4326)
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
@@
 ```{r}
+sheffield_centroids = get_pct_centroids("south-yorkshire",geography = "lsoa") %>% sf::st_transform(27700)
+sheffield_lsoa_buff = st_buffer(x = sheffield_centroids, dist = 1000)
+sheffield_res_buff = st_intersection(builtup,sheffield_lsoa_buff) %>% st_union()
 roads = rsf %>% filter(region == "Sheffield", pctgov > 100)
@@
 roads = st_intersection(roads, region_buff)
+roads = roads %>% st_transform(27700) %>% st_intersection(sheffield_res_buff) %>% st_transform(4326)
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
@@
 ```{r}
+newcastle_centroids = get_pct_centroids("north-east",geography = "lsoa") %>% sf::st_transform(27700)
+newcastle_lsoa_buff = st_buffer(x = newcastle_centroids, dist = 1000)
+newcastle_res_buff = st_intersection(builtup,newcastle_lsoa_buff) %>% st_union()
 roads = rsf %>% filter(region == "Newcastle", pctgov > 100)
@@
 roads = st_intersection(roads, region_buff)
+roads = roads %>% st_transform(27700) %>% st_intersection(newcastle_res_buff) %>% st_transform(4326)
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
@@
 ```{r}
+cambridge_centroids = get_pct_centroids("cambridgeshire",geography = "lsoa") %>% sf::st_transform(27700)
+cambridge_lsoa_buff = st_buffer(x = cambridge_centroids, dist = 1000)
+cambridge_res_buff = st_intersection(builtup,cambridge_lsoa_buff) %>% st_union()
 roads = rsf %>% filter(region == "Cambridge", pctgov > 400)
+roads = roads %>% st_transform(27700) %>% st_intersection(cambridge_res_buff) %>% st_transform(4326)
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +

### 2020-04-29 README.Rmd
Update with changes form @joeytalbot

@@
 )
+library(pct)
+library(sf)
 library(tidyverse)
-library(sf)
 library(tmap)
@@
 pct_las = sf::read_sf("https://github.com/cyipt/tempCycleways/releases/download/0.1/pct_la_summaries.geojson")
-las_top = readRDS("las_top.Rds")
-st_make_valid(las_top)
+las_top = sf::read_sf("https://github.com/cyipt/tempCycleways/releases/download/0.1/las_top.geojson")
+# las_top = readRDS("las_top.Rds")
+# las_top = st_make_valid(las_top)
+# sf::write_sf(las_top, "las_top.geojson")
+# piggyback::pb_upload("las_top.geojson")
 piggyback::pb_download("rsf_grouped_long.Rds")

### 2020-04-29 README.Rmd
Update with new mapping approach
Heads-up @mem48 here is a new viz idea linked to your comment, any better?

--- 2020-04-29 README.Rmd
@@
 )
+library(pct)
+library(sf)
 library(tidyverse)
-library(sf)
 library(tmap)
 tmap_mode("view")
-# piggyback::pb_download("las_top.Rds")
+piggyback::pb_download("las_top.Rds")
 pct_las = sf::read_sf("https://github.com/cyipt/tempCycleways/releases/download/0.1/pct_la_summaries.geojson")
-las_top = readRDS("las_top.Rds")
+las_top = sf::read_sf("https://github.com/cyipt/tempCycleways/releases/download/0.1/las_top.geojson")
+# las_top = readRDS("las_top.Rds")
+# las_top = st_make_valid(las_top)
+# sf::write_sf(las_top, "las_top.geojson")
+# piggyback::pb_upload("las_top.geojson")
+piggyback::pb_download("rsf_grouped_long.Rds")
 rsf = readRDS("rsf_grouped_long.Rds")
+piggyback::pb_download("Builtup_Areas_December_2011_Boundaries_V2.geojson")
+# could use in subsequent analysis - removes part of Liverpool tunnel?
+# builtup = sf::read_sf("Builtup_Areas_December_2011_Boundaries_V2.geojson") %>% sf::st_transform(27700)
 ```
@@
     (lanespsvbackward + lanesbackward) > 1,
-  ! grepl(pattern = "motorway", highway)
+  !grepl(pattern = "motorway", highway)
 )
@@
 ```{r}
+## Mask not required for London, so no need to slow down the code unnecessarily
+# london_centroids = get_pct_centroids("london",geography = "lsoa") %>% sf::st_transform(27700)
+# london_lsoa_buff = st_buffer(x = london_centroids, dist = 1000)
+# london_res_buff = st_intersection(builtup,london_lsoa_buff) %>% st_union()
 roads = rsf %>% filter(region == "London", pctgov > 500)
+# roads = roads %>% st_transform(27700) %>% st_intersection(london_res_buff) %>% st_transform(4326)
 tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
@@
   st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
+  knitr::kable(digits = 0)
 ```
@@
 ```{r}
+# idea: use buffers of LSOA centroids to supplent built-up areas
+# birmingham_centroids = get_pct_centroids("west-midlands", geography = "lsoa")
+# birmingham_lsoa_buff = stplanr::geo_buffer(birmingham_centroids, dist = 1000)
+# mapview::mapview(birmingham_lsoa_buff)
+# birmingham_res_buff = st_intersection(builtup, birmingham_lsoa_buff) %>% st_union()
+# roads = roads %>%
+#   st_transform(27700) %>%
+#   st_intersection(birmingham_res_buff) %>%
+#   st_transform(4326)
+# idea: look for other basemaps
+s = c(
+  "Esri.WorldGrayCanvas",
+  "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
+  "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
+)
+region = las_top %>% filter(name == "Birmingham")
 roads = rsf %>% filter(region == "Birmingham", pctgov > 100)
-tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_view(set.view = 12) + tm_scale_bar()
+roads = st_intersection(roads, region) # remove roads outside region
 ```
@@
   slice(1:10)
-tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
+  tm_shape(region) + tm_borders() +
+  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
   tm_text("name") +
-  tm_view(set.view = 13) + tm_scale_bar()
+  tm_view(set.view = 11) + tm_scale_bar() +
+  tm_basemap(server = s)
 
@@
 roads = st_intersection(roads, region_buff)
-tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_view(set.view = 12) + tm_scale_bar()
 ```
@@
   slice(1:10)
-tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
+  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
   tm_text("name") +
-  tm_view(set.view = 13) + tm_scale_bar()
-
+  tm_view(set.view = 12) + tm_scale_bar() +
+  tm_basemap(server = s)
 roads_top %>%
   st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
+  knitr::kable(digits = 0)
 ```
@@
 region_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 3000)
-roads = st_intersection(roads, region_buff)
-tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_scale_bar()
-```
-
-The top 10 roads:
-
-```{r}
-region_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 3000)
-roads_top = roads %>%
-  filter(length > 300) %>% 
-  arrange(desc(pctgov)) %>%
-  slice(1:10) 
-tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_text("name") +
-  tm_view(set.view = 13) + tm_scale_bar()
-
-roads_top %>%
-  st_drop_geometry() %>% 
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
-```
-
-# Liverpool
-
-```{r}
-roads = rsf %>% filter(region == "Liverpool", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Liverpool", as.sf = TRUE)
 region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
 roads = st_intersection(roads, region_buff)
-tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_view(set.view = 12) + tm_scale_bar()
 ```
@@
   filter(length > 300) %>%
+  filter(name != "Broadway") %>%
   arrange(desc(pctgov)) %>%
   slice(1:10)
-tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
+  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
   tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar()
+  tm_view(set.view = 12) + tm_scale_bar() +
+  tm_basemap(server = s)
 
@@
   st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
+  knitr::kable(digits = 0)
+```
+
+# Liverpool
+
+```{r}
+roads = rsf %>% filter(region == "Liverpool", pctgov > 100)
+region = las_top %>% filter(name == "Liverpool")
+roads = st_intersection(roads, region)
+# region_centre = tmaptools::geocode_OSM("Liverpool", as.sf = TRUE)
+# region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
+# tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+#   tm_view(set.view = 10) + tm_scale_bar() +
+#   tm_shape(region) + tm_borders()
+```
+
+The top 10 roads:
+
+```{r}
+roads_top = roads %>%
+  filter(length > 300) %>%
+  filter(name != "Queensway (Mersey Tunnel)") %>%
+  filter(name != "Queensway") %>%
+  arrange(desc(pctgov)) %>%
+  slice(1:10)
+tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
+  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
+  tm_text("name") +
+  tm_view(set.view = 12) + tm_scale_bar() +
+  tm_basemap(server = s)
+
+roads_top %>%
+  st_drop_geometry() %>%
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
+  knitr::kable(digits = 0)
 ```
@@
 roads = rsf %>% filter(region == "Bristol", pctgov > 100)
-roads = roads[region, ]
 roads = st_intersection(roads, region)
-tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_shape(region) + tm_borders() +
-  tm_scale_bar() 
 ```
@@
   filter(length > 300) %>%
-  filter(!name == "Bond Street", !name == "") %>% 
+  filter(name != "Temple Way Underpass") %>%
+  filter(!name == "Newfoundland Street", !name == "") %>%
   arrange(desc(pctgov)) %>%
-  slice(2:11) %>% 
-  distinct(name, .keep_all = TRUE)
-tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+  slice(1:10)
+
+tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
+  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
   tm_text("name") +
-  tm_scale_bar()
+  tm_view(set.view = 12) + tm_scale_bar() +
+  tm_basemap(server = s)
 
@@
   slice(1:10)
-tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+
+tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
+  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
   tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar()
+  tm_view(set.view = 12) + tm_scale_bar() +
+  tm_basemap(server = s)
 
@@
 roads_top = roads %>%
-  filter(length > 600) %>% 
+  filter(length > 600) %>% # remove 500m roundabouts - is there a better way?
   arrange(desc(pctgov)) %>%
   slice(1:10)
-tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+
+tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
+  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
   tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar()
+  tm_view(set.view = 12) + tm_scale_bar() +
+  tm_basemap(server = s)
 
@@
   filter(!name == "") %>%
+  filter(!name == "Redheugh Bridge") %>%
   arrange(desc(pctgov)) %>%
-  group_by(name, ref, highway) %>% 
-  summarise(pctgov = mean(pctgov), length = mean(length)) %>% 
-  ungroup() %>% 
   slice(1:10)
-tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+
+tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
+  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
   tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar()
-
+  tm_view(set.view = 12) + tm_scale_bar() +
+  tm_basemap(server = s)
 roads_top %>%
   st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>% knitr::kable(digits = 0)
+  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
+  knitr::kable(digits = 0)
 ```
@@
 ```{r}
-roads = rsf %>% filter(region == "Cambridge", pctgov > 400)
-tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_view(set.view = 13) + tm_scale_bar()
+roads = rsf %>% filter(region == "Cambridge", pctgov > 100)
+region_centre = tmaptools::geocode_OSM("Cambridge", as.sf = TRUE)
+region_buff = stplanr::geo_buffer(region_centre, dist = 3000)
+roads = st_intersection(roads, region_buff)
 ```
@@
   filter(length > 100) %>%
+  # filter(name != "Chesterton Road") %>%
   arrange(desc(pctgov)) %>%
   slice(1:10)
-tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
+  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
   tm_text("name") +
-  tm_view(set.view = 14) + tm_scale_bar()
+  tm_view(set.view = 13) + tm_scale_bar() +
+  tm_basemap(server = s)
 

### 2020-04-29 README.Rmd
Changes to maps

@@
 ---
-output: github_document
+output: html_document
+always_allow_html: yes
 ---
@@
 ```{r, eval=FALSE}
+webshot::install_phantomjs()
 # general analysis of data
@@
 # london_res_buff = st_intersection(builtup,london_lsoa_buff) %>% st_union()
+s = c(
+  "Esri.WorldGrayCanvas",
+  "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
+  "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
+)
+region = las_top %>% filter(name == "London")
 roads = rsf %>% filter(region == "London", pctgov > 500)
 # roads = roads %>% st_transform(27700) %>% st_intersection(london_res_buff) %>% st_transform(4326)
-tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_view(set.view = 11) + tm_scale_bar()
-```
 
-The top 10 roads:
-
-```{r}
 roads_top = roads %>%
@@
   filter(!name == "Talgarth Road", length > 500) %>%
+  filter(!name == "Whitechapel Road", length > 500) %>%
   arrange(desc(pctgov)) %>%
   slice(1:10)
-tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+# tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
+#   tm_text("name") +
+#   tm_view(set.view = 12) + tm_scale_bar()
+tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
+  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
   tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar()
-
+  tm_view(set.view = 12) +
+  tm_scale_bar() +
+  tm_basemap(server = s)
 roads_top %>%
@@
 # idea: look for other basemaps
-s = c(
-  "Esri.WorldGrayCanvas",
-  "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
-  "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
-)
 region = las_top %>% filter(name == "Birmingham")
@@
 roads = st_intersection(roads, region) # remove roads outside region
-```
 
-The top 10 roads:
-
-```{r}
 roads_top = roads %>%
@@
 tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(region) + tm_borders() +
   tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
@@
 roads = st_intersection(roads, region_buff)
-```
 
-The top 10 roads:
-
-```{r}
 roads_top = roads %>%
@@
 #   tm_shape(region) + tm_borders()
-```
 
-The top 10 roads:
-
-```{r}
 roads_top = roads %>%
@@
 ```{r}
-region = pct_las %>% filter(name == "Bristol")
+region = las_top %>% filter(name == "Bristol")
 region = sf::st_make_valid(region)
@@
 roads = st_intersection(roads, region)
-```
 
-The top 10 roads:
-
-```{r}
 bristol_centre = tmaptools::geocode_OSM("Bristol", as.sf = TRUE)
@@
 roads = st_intersection(roads, region_buff)
-tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-  tm_view(set.view = 12) + tm_scale_bar()
-```
 
-The top 10 roads:
-
-```{r}
 roads_top = roads %>%
@@
 roads = st_intersection(roads, region_buff)
-```
 
-The top 10 roads:
-
-```{r}
 roads_top = roads %>%
-  filter(length > 600) %>% # remove 500m roundabouts - is there a better way?
+  filter(length > 200) %>% 
+  filter(!name == "Tinsley Roundabout")  %>%
+  filter(!name == "Meadowhead Roundabout")  %>%
   arrange(desc(pctgov)) %>%
@@
 region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-roads = st_intersection(roads, region_buff)
-```
+region = las_top %>% filter(name == "Newcastle")
+region = st_intersection(region_buff, region)
+roads = st_intersection(roads, region)
 
-The top 10 roads:
-
-```{r}
 roads_top = roads %>%
-  filter(length > 600) %>%
+  filter(length > 250) %>%
   filter(!name == "") %>%
   filter(!name == "Redheugh Bridge")  %>%
+  filter(!name == "Derwenthaugh Road", !name == "") %>%
+  filter(!name == "East Denton Interchange", !name == "") %>%
   arrange(desc(pctgov)) %>%
@@
 roads = st_intersection(roads, region_buff)
-```
 
-The top 10 roads:
-
-```{r}
 roads_top = roads %>%
   filter(length > 100) %>%
-  # filter(name != "Chesterton Road") %>%
+  filter(name != "") %>%
   arrange(desc(pctgov)) %>%

### 2020-04-30 README.Rmd
Update with new methods section and maps
Heads-up @joeytalbot, @mvl22 and @mem48, look better?

Any further suggestions on description or results very welcome

@@
 ---
-output: html_document
-always_allow_html: yes
+output: github_document
+# uncomment the next line if no phantomjs
+# always_allow_html: no
 ---
@@
 
-These 12 study cities coincidentally represent almost exactly 1/4 of the population in England.
+These cities represent around 1/4 of the population in England.
 Welsh and Scottish cities with high cycling potential such as Cardiff and Edinbugh were not included in the analysis because the CyIPT does not currently have data outside of England, although we could extend the methods to cover all UK cities at some point.
 
-Below we show the results for a selection of cities, with cycling potential on the road network visualised under the 'Government Target' scenario, which represents a doubling in cycling compared with 2011 levels.
+# Method
+
+To identify streets that may be strong candidates for the provision of temporary or 'pop-up' cycleways, building on data from the CyIPT and PCT projects, three filtering methods were used:
+
+- Number of lanes: roads with more than 1 lane in either direction were identified
+- Cycling potential: only roads with high cycling potential, of 100 or more in the long-term Go Dutch scenario, were selected
+- Length of road: only road sections at least 100m long were included in the analysis, in some cities this threshold was raised to identify the 'top 10' roads for new cycleways
+
+More criteria such as road width and proximity to key services such as hospitals could be added at a later date.
+A final stage involved manually removing road sections such as roads on which there is already good quality dedicated infrastructure and roundabouts.
+This final stage could be automated in future work.
+
+The cycling potential of the 'top 10' streets is presented in terms of the 'Government Target' scenario, which represents a doubling in cycling compared with 2011 levels.
 London is close to meeting this target already.
 
-We filtered-out roads with low levels of cycling potential and focus only on roads that have at least one 'spare lane', defined as having more than 1 lane in either direction.
-Roads that could be made oneway, or that could be converted into 'liveable streets' by preventing through traffic were not considered.
+<!-- We filtered-out roads with low levels of cycling potential and focus only on roads that have at least one 'spare lane', defined as having more than 1 lane in either direction. -->
+Roads that could be made oneway, or that could be converted into 'liveable streets' by preventing through traffic have not been considered at this stage.
+Providing evidence to identify such roads, e.g.
+where there is high demand but little road width, suggesting a strong argument for making them one-way during a period of vital physical distancing, in addition to the roads with 'spare lanes' identified below, could be an aim of future work.
 
@@
   tm_text("name") +
-  tm_view(set.view = 11) + tm_scale_bar() +
+  tm_view(set.view = 12) + tm_scale_bar() +
   tm_basemap(server = s)
@@
   tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
+  tm_view(set.view = 13) + tm_scale_bar() +
   tm_basemap(server = s)

### 2020-04-30 README.Rmd
Explain map colours

@@
 
+# Interpreting the results
+
+The results are not intended to be a definitive list of places where popup cycleways should be prioritised.
+There may be 'false positives' where there are strong reasons for not converting a 'spare lane' into a cycleway or temporary pavement.
+There will be many road sections that would benefit from interventions not shown in the maps below: roads with only one lane in each direction could be made oneway temporarily, creating a spare lane for cycleways or extra pavement width.
+
 <!-- We filtered-out roads with low levels of cycling potential and focus only on roads that have at least one 'spare lane', defined as having more than 1 lane in either direction. -->
-Roads that could be made oneway, or that could be converted into 'liveable streets' by preventing through traffic have not been considered at this stage.
+Such roads and roads that could be converted into 'liveable streets' by preventing through traffic have not been considered at this stage.
 Providing evidence to identify such roads, e.g.
 where there is high demand but little road width, suggesting a strong argument for making them one-way during a period of vital physical distancing, in addition to the roads with 'spare lanes' identified below, could be an aim of future work.
+
+## What the maps show
+
+The results below show all roads with a 'spare lane' in the low average traffic conditions created by the partial lockdown (light blue) plus the top 10 roads of a minimum length in terms of estimated cycling potential under the Government Target scenario in the PCT.
 

### 2020-04-30 README.Rmd
Add table of contents

@@
 ---
-output: github_document
+title: Data analysis to support temporary cycleways
+output:
+  github_document:
+    toc: true
+    toc_depth: 2
 # uncomment the next line if no phantomjs
@@
 
-# Data analysis to support temporary cycleways
+# Introduction
 

### 2020-04-30 README.Rmd
Update description of maps

@@
 
-The results below show all roads with a 'spare lane' in the low average traffic conditions created by the partial lockdown (light blue) plus the top 10 roads of a minimum length in terms of estimated cycling potential under the Government Target scenario in the PCT.
+The results below show all roads with a 'spare lane' in the low average traffic conditions created by the partial lockdown (light blue) based on the three criteria listed above (lanes, potential and length) plus the top 10 roads of a minimum length in terms of estimated cycling potential under the Government Target scenario in the PCT (dark blue).
 

### 2020-04-30 README.Rmd
Leicester adjustment

@@
 - Number of lanes: roads with more than 1 lane in either direction were identified
-- Cycling potential: only roads with high cycling potential, of 100 or more in the long-term Go Dutch scenario, were selected
+- Cycling potential: only roads with high cycling potential, of 100 or more in the long-term Government Target scenario, were selected
 - Length of road: only road sections at least 100m long were included in the analysis, in some cities this threshold was raised to identify the 'top 10' roads for new cycleways
@@
 roads_top = roads %>%
-  filter(length > 300) %>%
+  filter(length > 200) %>%
   filter(name != "Queensway (Mersey Tunnel)") %>%
@@
 roads_top = roads %>%
-  filter(length > 300) %>%
+  filter(length > 200) %>%
   arrange(desc(pctgov)) %>%

### 2020-04-30 README.Rmd
Recentre Liverpool map

@@
   tm_text("name") +
-  tm_view(set.view = 13) + tm_scale_bar() +
+  tm_view(set.view = c(-2.95, 53.4084, 13)) + 
+  tm_scale_bar() +
   tm_basemap(server = s)

### 2020-04-30 README.Rmd
Recentre Leeds and Bristol maps

@@
   tm_text("name") +
-  tm_view(set.view = 13) + tm_scale_bar() +
+  tm_view(set.view = c(-1.549, 53.805, 13)) + 
+  tm_scale_bar() +
   tm_basemap(server = s)
@@
   tm_text("name") +
-  tm_view(set.view = 13) + tm_scale_bar() +
+  tm_view(set.view = c(-2.5879, 51.454, 13)) + 
+  tm_scale_bar() +
   tm_basemap(server = s)

### 2020-05-04 README.Rmd
Add readme files

@@
 output:
-  github_document:
+  html_document:
+  # github_document:
     toc: true
@@
 ```
+
+```{r, engine='zsh', eval=FALSE}
+# deploy
+
+```

### 2020-05-04 README.Rmd
Changes for deployed version

--- 2020-04-30 README.Rmd
@@
 output:
+  # html_document:
   github_document:
@@
 # builtup = sf::read_sf("Builtup_Areas_December_2011_Boundaries_V2.geojson") %>% sf::st_transform(27700)
+## Mask not required for London, so no need to slow down the code unnecessarily
+s = c(
+  "Esri.WorldGrayCanvas",
+  "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
+  "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
+)
 ```
@@
 ```{r, eval=FALSE}
+
+rmarkdown::render("london.Rmd") # check
+rmarkdown_files = paste0(tolower(las_top$name), ".Rmd")
+lapply(rmarkdown_files, rmarkdown::render)
 webshot::install_phantomjs()
@@
 
-# London
+```{r, child="london.Rmd"}
 
-```{r}
-## Mask not required for London, so no need to slow down the code unnecessarily
-# london_centroids = get_pct_centroids("london",geography = "lsoa") %>% sf::st_transform(27700)
-# london_lsoa_buff = st_buffer(x = london_centroids, dist = 1000)
-# london_res_buff = st_intersection(builtup,london_lsoa_buff) %>% st_union()
-s = c(
-  "Esri.WorldGrayCanvas",
-  "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
-  "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
-)
-region = las_top %>% filter(name == "London")
-roads = rsf %>% filter(region == "London", pctgov > 500)
-# roads = roads %>% st_transform(27700) %>% st_intersection(london_res_buff) %>% st_transform(4326)
-
-roads_top = roads %>%
-  filter(length > 500) %>%
-  filter(!name == "Talgarth Road", length > 500) %>%
-  filter(!name == "Whitechapel Road", length > 500) %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-# tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-#   tm_text("name") +
-#   tm_view(set.view = 12) + tm_scale_bar()
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) +
-  tm_scale_bar() +
-  tm_basemap(server = s)
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Birmingham
+See [london.html](london.html) for the interactive version of the map above.
 
-```{r}
-# idea: use buffers of LSOA centroids to supplent built-up areas
-# birmingham_centroids = get_pct_centroids("west-midlands", geography = "lsoa")
-# birmingham_lsoa_buff = stplanr::geo_buffer(birmingham_centroids, dist = 1000)
-# mapview::mapview(birmingham_lsoa_buff)
-# birmingham_res_buff = st_intersection(builtup, birmingham_lsoa_buff) %>% st_union()
-# roads = roads %>%
-#   st_transform(27700) %>%
-#   st_intersection(birmingham_res_buff) %>%
-#   st_transform(4326)
-# idea: look for other basemaps
-region = las_top %>% filter(name == "Birmingham")
-roads = rsf %>% filter(region == "Birmingham", pctgov > 100)
-roads = st_intersection(roads, region) # remove roads outside region
+```{r, child="birmingham.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 300) %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Manchester
+See [birmingham.html](birmingham.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Manchester", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Manchester", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 7000)
-roads = st_intersection(roads, region_buff)
+```{r, child="manchester.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 300) %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Leeds
+See [manchester.html](manchester.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Leeds", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-roads = st_intersection(roads, region_buff)
+```{r, child="leeds.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 300) %>%
-  filter(name != "Broadway") %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 13) + tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Liverpool
+See [leeds.html](leeds.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Liverpool", pctgov > 100)
-region = las_top %>% filter(name == "Liverpool")
-roads = st_intersection(roads, region)
-# region_centre = tmaptools::geocode_OSM("Liverpool", as.sf = TRUE)
-# region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-# tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-#   tm_view(set.view = 10) + tm_scale_bar() +
-#   tm_shape(region) + tm_borders()
+```{r, child="liverpool.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 200) %>%
-  filter(name != "Queensway (Mersey Tunnel)") %>%
-  filter(name != "Queensway") %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = c(-2.95, 53.4084, 13)) + 
-  tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Bristol
+See [liverpool.html](liverpool.html) for the interactive version of the map above.
 
-```{r}
-region = las_top %>% filter(name == "Bristol")
-region = sf::st_make_valid(region)
-roads = rsf %>% filter(region == "Bristol", pctgov > 100)
-roads = st_intersection(roads, region)
+```{r, child="bristol.Rmd"}
 
-bristol_centre = tmaptools::geocode_OSM("Bristol", as.sf = TRUE)
-bristol_buff = stplanr::geo_buffer(bristol_centre, dist = 1500)
-roads_top = roads[bristol_buff, ] %>%
-  filter(length > 300) %>%
-  filter(name != "Temple Way Underpass") %>%
-  filter(!name == "Newfoundland Street", !name == "") %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 13) + tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Leicester
+See [bristol.html](bristol.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Leicester", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Leicester", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-roads = st_intersection(roads, region_buff)
+```{r, child="leicester.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 200) %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Sheffield
+See [leicester.html](leicester.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Sheffield", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Sheffield", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-roads = st_intersection(roads, region_buff)
+```{r, child="sheffield.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 200) %>% 
-  filter(!name == "Tinsley Roundabout")  %>%
-  filter(!name == "Meadowhead Roundabout")  %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Newcastle
+See [sheffield.html](sheffield.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Newcastle", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Newcastle", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-region = las_top %>% filter(name == "Newcastle")
-region = st_intersection(region_buff, region)
-roads = st_intersection(roads, region)
+```{r, child="newcastle.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 250) %>%
-  filter(!name == "") %>%
-  filter(!name == "Redheugh Bridge")  %>%
-  filter(!name == "Derwenthaugh Road", !name == "") %>%
-  filter(!name == "East Denton Interchange", !name == "") %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Cambridge
+See [newcastle.html](newcastle.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Cambridge", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Cambridge", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 3000)
-roads = st_intersection(roads, region_buff)
+```{r, child="cambridge.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 100) %>%
-  filter(name != "") %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 13) + tm_scale_bar() +
-  tm_basemap(server = s)
+```
 
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
+See [cambridge.html](cambridge.html) for the interactive version of the map above.
+
+```{r, engine='zsh', eval=FALSE}
+# deploy
+git clone git@github.com:cyipt/tempCycleways
+cd tempCycleways
+git checkout --orphan gh-pages
+git rm -rf .
+git commit --allow-empty -m 'Initial gh-pages commit'
+# git push origin gh-pages
+# git checkout master
+# cp -Rv ../README* . # it's 100 MB!
+cp -Rv ../*.html . # it's 100 MB!
+git status
+ls -hal
+mv README.html index.html
+firefox index.html
+git status
+git push origin gh-pages
+cd ..
+echo '*.html' >> .gitignore
+echo 'tempCycleways/' >> .gitignore
+pandoc -o README.html README.md
 ```

### 2020-05-04 birmingham.Rmd
Changes for deployed version

# Birmingham

```{r birmingham}
# idea: use buffers of LSOA centroids to supplent built-up areas
# birmingham_centroids = get_pct_centroids("west-midlands", geography = "lsoa")
# birmingham_lsoa_buff = stplanr::geo_buffer(birmingham_centroids, dist = 1000)
# mapview::mapview(birmingham_lsoa_buff)
# birmingham_res_buff = st_intersection(builtup, birmingham_lsoa_buff) %>% st_union()
# roads = roads %>%
#   st_transform(27700) %>%
#   st_intersection(birmingham_res_buff) %>%
#   st_transform(4326)
# idea: look for other basemaps
region = las_top %>% filter(name == "Birmingham")
roads = rsf %>% filter(region == "Birmingham", pctgov > 100)
roads = st_intersection(roads, region) # remove roads outside region

roads_top = roads %>%
  filter(length > 300) %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = 12) + tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```

### 2020-05-04 cambridge.Rmd
Changes for deployed version

# Cambridge

```{r cambridge}
roads = rsf %>% filter(region == "Cambridge", pctgov > 100)
region_centre = tmaptools::geocode_OSM("Cambridge", as.sf = TRUE)
region_buff = stplanr::geo_buffer(region_centre, dist = 3000)
roads = st_intersection(roads, region_buff)

roads_top = roads %>%
  filter(length > 100) %>%
  filter(name != "") %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = 13) + tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```

### 2020-05-04 leeds.Rmd
Changes for deployed version

# Leeds

```{r leeds}
roads = rsf %>% filter(region == "Leeds", pctgov > 100)
region_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
roads = st_intersection(roads, region_buff)

roads_top = roads %>%
  filter(length > 300) %>%
  filter(name != "Broadway") %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = c(-1.549, 53.805, 13)) + 
  tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```

### 2020-05-04 README.Rmd
leeds and cambridge streets

--- 2020-04-30 README.Rmd
@@
 ```{r}
+leeds_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
+leeds_buff = stplanr::geo_buffer(leeds_centre, dist = 4000)
 roads = rsf %>% filter(region == "Leeds", pctgov > 100)
@@
 
-roads_top = roads %>%
+roads_top = roads[leeds_buff, ] %>%
   filter(length > 300) %>%
   filter(name != "Broadway") %>%
+  filter(name != "Wellington Street") %>%
   arrange(desc(pctgov)) %>%
@@
   filter(name != "") %>%
+  filter(!name == "Hills Road")  %>%
   arrange(desc(pctgov)) %>%

### 2020-05-04 README.Rmd
Minor text changes

@@
 
-in the context of increased for demand for cycling to keyworker workplaces.
+in the context of increased demand for cycling to keyworker workplaces.
 
-It is based on an analysis of data generated for the Department for Transport funded projects the Propensity to Cycle Tool (PCT) and the Cycling Infrastructure Prioritisation Toolkit (CyIPT).
+It is based on an analysis of data generated for Department for Transport funded projects the Propensity to Cycle Tool (PCT) and the Cycling Infrastructure Prioritisation Toolkit (CyIPT).
 
-To overcome issues associated with doing this at a national level and elicit feedback on the methods and preliminary results, we ran the analysis on a sample of cities.
-We chose the top 5 in terms of absolute long-term cycling potential (London, Birmingham, Manchester, Leeds, Liverpool) plus an additional 5 cities that have active advocacy groups (Newcastle, Sheffield, Cambridge, Bristol, Leicester).
-Estimates of current and potential numbers who could cycle to work in these cities is presented in the table below.
+As an initial analysis, to elicit feedback on the methods and preliminary results, we have focused on a sample of major cities.
+We hope this can be further developed and expanded in due course to provide nationwide coverage.
+
+We chose the top 5 cities in terms of absolute long-term cycling potential (London, Birmingham, Manchester, Leeds, Liverpool) plus an additional 5 cities that have active advocacy groups (Newcastle, Sheffield, Cambridge, Bristol, Leicester).
+Estimates of current and potential numbers of commuters who could cycle to work in these cities are presented in the table below.
 <!-- We did this by analysing data from the PCT project and selecting the top 10 cities in terms of long term cycling potential, plus Sheffield and Cambridge, : -->
@@
 
-These cities represent around 1/4 of the population in England.
+These cities represent around 1/4 of the population of England.
 Welsh and Scottish cities with high cycling potential such as Cardiff and Edinbugh were not included in the analysis because the CyIPT does not currently have data outside of England, although we could extend the methods to cover all UK cities at some point.
@@
 - Cycling potential: only roads with high cycling potential, of 100 or more in the long-term Government Target scenario, were selected
-- Length of road: only road sections at least 100m long were included in the analysis, in some cities this threshold was raised to identify the 'top 10' roads for new cycleways
+- Length of road: only road sections at least 100m long were included in the analysis
+
+From the resulting selections we then identified the 'top 10' routes in each city based on cycling potential.
+In most cities, only road sections longer than a threshold of 200-300m were considered for entry to this 'top 10' list.
 
@@
 
-The results are not intended to be a definitive list of places where popup cycleways should be prioritised.
+The results are not intended to be a definitive list of places where pop-up cycleways should be prioritised.
 There may be 'false positives' where there are strong reasons for not converting a 'spare lane' into a cycleway or temporary pavement.
-There will be many road sections that would benefit from interventions not shown in the maps below: roads with only one lane in each direction could be made oneway temporarily, creating a spare lane for cycleways or extra pavement width.
+There will be many road sections that would benefit from interventions not shown in the maps below: for example roads with only one lane in each direction could be made oneway temporarily, creating a spare lane for cycleways or extra pavement width.
 

### 2020-05-06 README.Rmd
Update with definition of cycling potential

@@
 
-The results are not intended to be a definitive list of places where pop-up cycleways should be prioritised.
-There may be 'false positives' where there are strong reasons for not converting a 'spare lane' into a cycleway or temporary pavement.
+The results are not a definitive list of places where pop-up cycleways should be prioritised but a 'starter for 10' highlighting roads that good candidates for 'pop-up' active transport infrastructure through reallocation of road space, as has happened in [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e).
+The results highlight roads that have cycling potential and at least one spare lane, meaning a 2 lanes in one direction.
 There will be many road sections that would benefit from interventions not shown in the maps below: for example roads with only one lane in each direction could be made oneway temporarily, creating a spare lane for cycleways or extra pavement width.
+Prevent through-flow in residential areas, as is happening in [Lewisham](https://853.london/2020/05/05/lewisham-council-asks-which-streets-it-should-open-up-to-walkers-and-cyclists/) is another option that can complement road reallocation.
 
 <!-- We filtered-out roads with low levels of cycling potential and focus only on roads that have at least one 'spare lane', defined as having more than 1 lane in either direction. -->
-Such roads and roads that could be converted into 'liveable streets' by preventing through traffic have not been considered at this stage.
-Providing evidence to identify such roads, e.g.
-where there is high demand but little road width, suggesting a strong argument for making them one-way during a period of vital physical distancing, in addition to the roads with 'spare lanes' identified below, could be an aim of future work.
+<!-- Such roads that could be converted into 'liveable streets' by preventing through traffic have not been considered at this stage. -->
+Evidence supporting other types of interventions, for example where there is high demand for access to key workplaces but little space for walking and cycling, could be an aim of future work.
 
@@
 
-The results below show all roads with a 'spare lane' in the low average traffic conditions created by the partial lockdown (light blue) based on the three criteria listed above (lanes, potential and length) plus the top 10 roads of a minimum length in terms of estimated cycling potential under the Government Target scenario in the PCT (dark blue).
+The results below show all roads with a 'spare lane' in **light blue** based on the three criteria listed above (lanes, potential and length).
+The top 10 roads in terms of cycling potential are shown in **dark blue**.
+Cycling potential refers to the number of commuters who would cycle along the road (either to or from work) under the Government Target based on data from the Department for Transport funded Propensity to Cycle Tool (see [www.pct.bike](https://www.pct.bike/)).
 

### 2020-05-06 README.Rmd

@@
 
-# London
+```{r, child="london.Rmd"}
 
-```{r}
-## Mask not required for London, so no need to slow down the code unnecessarily
-# london_centroids = get_pct_centroids("london",geography = "lsoa") %>% sf::st_transform(27700)
-# london_lsoa_buff = st_buffer(x = london_centroids, dist = 1000)
-# london_res_buff = st_intersection(builtup,london_lsoa_buff) %>% st_union()
-s = c(
-  "Esri.WorldGrayCanvas",
-  "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
-  "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
-)
-region = las_top %>% filter(name == "London")
-roads = rsf %>% filter(region == "London", pctgov > 500)
-# roads = roads %>% st_transform(27700) %>% st_intersection(london_res_buff) %>% st_transform(4326)
-
-roads_top = roads %>%
-  filter(length > 500) %>%
-  filter(!name == "Talgarth Road", length > 500) %>%
-  filter(!name == "Whitechapel Road", length > 500) %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-# tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-#   tm_text("name") +
-#   tm_view(set.view = 12) + tm_scale_bar()
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) +
-  tm_scale_bar() +
-  tm_basemap(server = s)
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Birmingham
+See [london.html](https://cyipt.github.io/tempCycleways/london.html) for the interactive version of the map above.
 
-```{r}
-# idea: use buffers of LSOA centroids to supplent built-up areas
-# birmingham_centroids = get_pct_centroids("west-midlands", geography = "lsoa")
-# birmingham_lsoa_buff = stplanr::geo_buffer(birmingham_centroids, dist = 1000)
-# mapview::mapview(birmingham_lsoa_buff)
-# birmingham_res_buff = st_intersection(builtup, birmingham_lsoa_buff) %>% st_union()
-# roads = roads %>%
-#   st_transform(27700) %>%
-#   st_intersection(birmingham_res_buff) %>%
-#   st_transform(4326)
-# idea: look for other basemaps
-region = las_top %>% filter(name == "Birmingham")
-roads = rsf %>% filter(region == "Birmingham", pctgov > 100)
-roads = st_intersection(roads, region) # remove roads outside region
+```{r, child="birmingham.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 300) %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Manchester
+See [birmingham.html](https://cyipt.github.io/tempCycleways/birmingham.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Manchester", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Manchester", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 7000)
-roads = st_intersection(roads, region_buff)
+```{r, child="manchester.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 300) %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Leeds
+See [manchester.html](https://cyipt.github.io/tempCycleways/manchester.html) for the interactive version of the map above.
 
-```{r}
-leeds_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
-leeds_buff = stplanr::geo_buffer(leeds_centre, dist = 4000)
-roads = rsf %>% filter(region == "Leeds", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-roads = st_intersection(roads, region_buff)
+```{r, child="leeds.Rmd"}
 
-roads_top = roads[leeds_buff, ] %>%
-  filter(length > 300) %>%
-  filter(name != "Broadway") %>%
-  filter(name != "Wellington Street") %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = c(-1.549, 53.805, 13)) + 
-  tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Liverpool
+See [leeds.html](https://cyipt.github.io/tempCycleways/leeds.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Liverpool", pctgov > 100)
-region = las_top %>% filter(name == "Liverpool")
-roads = st_intersection(roads, region)
-# region_centre = tmaptools::geocode_OSM("Liverpool", as.sf = TRUE)
-# region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-# tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-#   tm_view(set.view = 10) + tm_scale_bar() +
-#   tm_shape(region) + tm_borders()
+```{r, child="liverpool.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 200) %>%
-  filter(name != "Queensway (Mersey Tunnel)") %>%
-  filter(name != "Queensway") %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = c(-2.95, 53.4084, 13)) + 
-  tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Bristol
+See [liverpool.html](https://cyipt.github.io/tempCycleways/liverpool.html) for the interactive version of the map above.
 
-```{r}
-region = las_top %>% filter(name == "Bristol")
-region = sf::st_make_valid(region)
-roads = rsf %>% filter(region == "Bristol", pctgov > 100)
-roads = st_intersection(roads, region)
+```{r, child="bristol.Rmd"}
 
-bristol_centre = tmaptools::geocode_OSM("Bristol", as.sf = TRUE)
-bristol_buff = stplanr::geo_buffer(bristol_centre, dist = 1500)
-roads_top = roads[bristol_buff, ] %>%
-  filter(length > 300) %>%
-  filter(name != "Temple Way Underpass") %>%
-  filter(!name == "Newfoundland Street", !name == "") %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = c(-2.5879, 51.454, 13)) + 
-  tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Leicester
+See [bristol.html](https://cyipt.github.io/tempCycleways/bristol.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Leicester", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Leicester", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-roads = st_intersection(roads, region_buff)
+```{r, child="leicester.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 200) %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Sheffield
+See [leicester.html](https://cyipt.github.io/tempCycleways/leicester.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Sheffield", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Sheffield", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-roads = st_intersection(roads, region_buff)
+```{r, child="sheffield.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 200) %>% 
-  filter(!name == "Tinsley Roundabout")  %>%
-  filter(!name == "Meadowhead Roundabout")  %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Newcastle
+See [sheffield.html](https://cyipt.github.io/tempCycleways/sheffield.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Newcastle", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Newcastle", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-region = las_top %>% filter(name == "Newcastle")
-region = st_intersection(region_buff, region)
-roads = st_intersection(roads, region)
+```{r, child="newcastle.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 250) %>%
-  filter(!name == "") %>%
-  filter(!name == "Redheugh Bridge")  %>%
-  filter(!name == "Derwenthaugh Road", !name == "") %>%
-  filter(!name == "East Denton Interchange", !name == "") %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
 ```
 
-# Cambridge
+See [newcastle.html](https://cyipt.github.io/tempCycleways/newcastle.html) for the interactive version of the map above.
 
-```{r}
-roads = rsf %>% filter(region == "Cambridge", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Cambridge", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 3000)
-roads = st_intersection(roads, region_buff)
+```{r, child="cambridge.Rmd"}
 
-roads_top = roads %>%
-  filter(length > 100) %>%
-  filter(name != "") %>%
-  filter(!name == "Hills Road")  %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 13) + tm_scale_bar() +
-  tm_basemap(server = s)
+```
 
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
+See [cambridge.html](https://cyipt.github.io/tempCycleways/cambridge.html) for the interactive version of the map above.
+
+```{r, engine='zsh', eval=FALSE}
+# deploy
+git clone git@github.com:cyipt/tempCycleways
+cd tempCycleways
+git checkout --orphan gh-pages
+git rm -rf .
+git commit --allow-empty -m 'Initial gh-pages commit'
+# git push origin gh-pages
+# git checkout master
+# cp -Rv ../README* . # it's 100 MB!
+cp -Rv ../*.html . # it's 100 MB!
+git status
+ls -hal
+mv README.html index.html
+firefox index.html
+git status
+git push origin gh-pages
+cd ..
+echo '*.html' >> .gitignore
+echo 'tempCycleways/' >> .gitignore
+pandoc -o README.html README.md
 ```

### 2020-05-06 cambridge.Rmd

@@
   filter(name != "") %>%
+  filter(!name == "Hills Road")  %>% 
   arrange(desc(pctgov)) %>%

### 2020-05-06 london.Rmd

# London

```{r london}
region = las_top %>% filter(name == "London")
roads = rsf %>% filter(region == "London", pctgov > 500)
# roads = roads %>% st_transform(27700) %>% st_intersection(london_res_buff) %>% st_transform(4326)

roads_top = roads %>%
  filter(length > 500) %>%
  filter(!name == "Talgarth Road", length > 500) %>%
  filter(!name == "Whitechapel Road", length > 500) %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)
# tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
#   tm_text("name") +
#   tm_view(set.view = 12) + tm_scale_bar()
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = 12) +
  tm_scale_bar() +
  tm_basemap(server = s)
roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```

### 2020-05-06 newcastle.Rmd

# Newcastle

```{r newcastle}
roads = rsf %>% filter(region == "Newcastle", pctgov > 100)
region_centre = tmaptools::geocode_OSM("Newcastle", as.sf = TRUE)
region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
region = las_top %>% filter(name == "Newcastle")
region = st_intersection(region_buff, region)
roads = st_intersection(roads, region)

roads_top = roads %>%
  filter(length > 250) %>%
  filter(!name == "") %>%
  filter(!name == "Redheugh Bridge")  %>%
  filter(!name == "Derwenthaugh Road", !name == "") %>%
  filter(!name == "East Denton Interchange", !name == "") %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)

tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = 12) + tm_scale_bar() +
  tm_basemap(server = s)
roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```

### 2020-05-06 sheffield.Rmd

# Sheffield

```{r sheffield}
roads = rsf %>% filter(region == "Sheffield", pctgov > 100)
region_centre = tmaptools::geocode_OSM("Sheffield", as.sf = TRUE)
region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
roads = st_intersection(roads, region_buff)

roads_top = roads %>%
  filter(length > 200) %>% 
  filter(!name == "Tinsley Roundabout")  %>%
  filter(!name == "Meadowhead Roundabout")  %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)

tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = 12) + tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```

### 2020-05-07 README.Rmd
conversation

@@
 cd tempCycleways
-git checkout --orphan gh-pages
+git checkout gh-pages
 git rm -rf .
@@
 git status
+git commit -am 'Update interactive maps'
 git push origin gh-pages
@@
 echo 'tempCycleways/' >> .gitignore
-pandoc -o README.html README.md
+pandoc -o README.html ../README.md
 ```

### 2020-05-07 conversation.Rmd
conversation

---
title: "Pop-up cycleways and pedestrian spaces: a peak into post Covid-19 transport futures"
---

The Covid-19 pandemic has transformed all sectors of the economy, not least transport. Long distance travel has been dramatically altered, as [demand plummets](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies go bust](https://arxiv.org/abs/2004.08460). Yet short distance travel patterns have also changed greatly. Forced to stay close to home, many people have started walking and cycling more. From a health and well-being perspective, this may be a more important shift, especially in the long term as coronavirus measures are gradually eased.

Local and national governments worldwide are still working out how best to respond to these changes to lif. Hundreds of kilometres of temporary infrastructure being put in place in dozens of cities worldwide, from [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e "") to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/ "") to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf ""), never has there been such a need for affordable, healthy and socially isolated transport that does not create the fumes that emerging evidence links to coronavirus susceptibility. With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised. As with the medical science, research is needed now. Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in now is effective, safe, and placed where it is most needed.

<!-- # Where is pop-up infrastructure most needed? -->

<!-- # Story  -->

<!-- The Covid-19 pandemic has transformed all sectors of the economy. -->
<!-- Transport patterns have been dramatically altered, with much discussion of plummeting demand for long distance travel. -->
<!-- Yet short-distance travel patterns have also changed greatly and, from a health and well-being perspective, this may be a more important shift. -->
<!-- With temporary infrastructure being put in place in dozens of cities worldwide, from [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e) to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf), never has there been such a need for affordable, healthy and socially isolated transport that does not create the fumes that emerging evidence links to coronavirus susceptibility. -->
<!-- With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised. -->
<!-- As with the medical science, research is needed now. -->
<!-- Methods developed to identify locations of high walking and cycling potential can help ensure that the 'pop-up' infrastructure that goes in now is effective, safe, and placed where it is most needed. -->

<!-- (this would be two to three sentences summarizing the article’s focus) -->

<!-- # Significance  -->

<!-- <!-- (why is this story interesting/relevant? Particularly to general audience, why should they care?)  --> -->
<!-- Politicians of all stripes recognise the importance of acting now to prevent a surge in car use and resulting air pollution when the lockdown is gradually released. -->
<!-- Sadiq Kahn has advocated re-allocating road space for walking and cycling, and Boris Johnson has [talked](https://twitter.com/SomethingELLIS/status/1258031387334053888) of a post Covid-19 "new golden age for cycling". -->
<!-- Without evidence there is a risk that these words are not supported by effective and long-term interventions. -->
<!-- Our research into prioritising the right kind of cycling infrastructure where it is most needed can ensure that decisions made now are of maximum benefit for the coming months when physical distancing will need to continue. -->
<!-- Polls showing people do not want to 'return to normal' also suggest that the Covid-19 crisis represents an historic opportunity to readjust transport cities in favour of health. -->
<!-- If pop-up cycleways go where there is *long term need* there is a greater chance that people will demand they become permanent, as happened with the experimental cycleway in Torrington Place in London, where a 2 way road for motor vehicles has been converted into a thriving 2 way cycleway, enabling social distancing and zero emissions transport for keyworkers in central London. -->

<!-- # Key points -->

<!-- - Our analysis of data generated for the Department for Transport funded Propensity to Cycle Tool (PCT) and Cycling Infrastructure Prioritisation Toolkit (CyIPT) has already flagged dozens of roads in 10 major cities across England where a 'spare lane' could be replaced by low-cost bidirectional temporary infrastructure.  -->
<!-- - A method highlights roads based on number of lanes, cycling potential and length. -->
<!-- - Feedback from cycle campaigners coordinated by CyclingUK suggests the results for 10 cities (London, Birmingham, Manchester, Leeds, Liverpool, Newcastle, Sheffield, Cambridge, Bristol, Leicester) match intuition and can support local knowledge and community engagement in support of temporary infrastructure. -->
<!-- - Further work, which we are discussing with the Department for Transport, will develop more sophisticated methods to provide evidence that is urgently needed to support pop-up spaces in the UK and beyond. -->

<!-- CyclingUK are putting out a press release as an exclusive to national media organisations. -->
<!-- This article will provide a more detailed view to go out the following day while interest in the topic is at a peak. -->

<!-- Results for Manchester shown below, which highlights Priness Street, one of the roads where [space for active travel will be increased](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175) as part of the post Covid-19 recovery package in Greater Manchester. -->

<!-- ![](https://raw.githubusercontent.com/cyipt/tempCycleways/master/README_files/figure-gfm/manchester-1.png) -->

<!-- Other key points(anything additional that you want to make sure they consider/know about) Highlighting the website further etc and we would also mention the BBC exclusive here. I’ve never had a problem with the Conversation agreeing to hold off on publishing an article until an embargo or exclusive deadline has been met. However it would be good to make clear how the Conversation article will differ from the BBC coverage.   -->

### 2020-05-07 article.Rmd
Article template

---
title: "Methods to support cycleways and pedestrianised spaces in response to the Covid-19 pandemic: overview and findings from 10 UK cities"
author: "Robin Lovelace"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Guidance from https://transportfindings.org/for-authors -->

<!-- Transport Findings welcomes research findings in the broad field of transport. Articles must either pose a New Question,  present a New Method, employ New Data (including New Contexts or Locations),  discover a New Finding (i.e. it can almost exactly replicate a previous study and find something different), or some combination of the above. -->
<!-- Scope -->

<!-- You may find yourself asking if your paper is within the scope of Transport Findings. -->

<!--     Is there a hypothesis somehow related to transport? -->
<!--     Is there a (scientifically valid, replicable) methodology? -->
<!--     Is there a finding? -->

<!-- If you can answer yes to these questions, it is within scope. -->
<!-- Article Types -->

<!--     Findings - where the object of study is nature -->
<!--     Syntheses - where the object of study is the literature -->
<!--     Cases - where the objects of study are particular sites or projects, and methods may be more qualitative -->

<!-- Sections -->

<!-- All articles shall have 3 sections, and only 3 sections, titled as follows: -->

<!--     RESEARCH QUESTION[S] AND HYPOTHESIS[ES] -->
<!--     METHODS AND DATA -->
<!--     FINDINGS -->

<!-- There shall be no introduction, “road-map paragraph,” literature review, conclusions, speculations, or  policy implications beyond what is included above. Focus on what you found, not why you found it. -->
<!-- Submissions -->

<!-- The manuscript submission must include the following: -->
<!-- TITLE -->
<!-- AUTHORS (NAME, AFFILIATION, CONTACT) -->
<!-- ABSTRACT -->
<!-- KEYWORDS -->
<!-- ARTICLE (Sections 1, 2, 3) -->
<!-- ACKNOWLEDGMENTS -->
<!-- REFERENCES -->

<!-- Manuscript submissions may include SUPPLEMENTAL INFORMATION in separate files that do not count against article length. This information should not be essential for the general understanding of the manuscript. -->
<!-- Style -->
<!-- Focus and Parsimony -->

<!-- Papers should be focused and to the point, and not begin with trite observations like “Congestion is a problem the world over.” Usually you can delete your opening paragraph if it begins like that, and the reader is no worse off. As Strunk and White say: “Omit Needless Words”. The Abstract should not say the same thing as the Introduction. -->
<!-- Transparency and Replicability -->

<!-- A minimum standard for a good paper is transparency and replicability: Can the reader understand what you did, and repeat it, and get the same answer? -->
<!-- Mathematical Conventions -->

<!-- Each variable shall have one, and only one, definition per document. -->

<!-- Each defined term in the document shall be represented by one and only one variable. -->

<!-- Lowercase and uppercase versions of the same letter should be logically related. For instance, use lowercase letters to define the PDF (probability distribution function) or individual instance, and uppercase letters the CDF (cumulative distribution function) or population, so when you sum:  i=1 to I, k=1 to K, etc. -->

<!-- All variables shall be a single letter or symbol. Double or triple letter variables can be confused with multiplication. If you have more than 52 symbols in your paper (26 letters for both lower and upper case), consider (a) there are too many, and (b) using Greek or Hebrew characters. -->

<!-- Use subscripts liberally to differentiate things that, for instance, are of a class but measured differently, or computed with different assumptions. -->

<!-- All equations shall have all of their variables defined. -->

### 2020-05-07 conversation.Rmd
Article template

@@
 
-<!-- <!-- (why is this story interesting/relevant? Particularly to general audience, why should they care?)  --> -->
+<!-- (why is this story interesting/relevant? Particularly to general audience, why should they care?)  --> 
 <!-- Politicians of all stripes recognise the importance of acting now to prevent a surge in car use and resulting air pollution when the lockdown is gradually released. -->
@@
 -->
+
+ <!-- supporting the case for reclaiming road space for safe, socially distanced key worker travel. -->

### 2020-05-08 exploratory-analysis-leeds.R
Add exploratory data for leeds

@@ -1,0 +1,102 @@
+# Aim: explore data for methods development in Leeds
+
+
+# set-up and load data ----------------------------------------------------
+
+library(sf)
+library(tidyverse)
+library(tmap)
+tmap_mode("view")
+
+# run once to get leeds
+# region_name = "Leeds"
+# rsf = readRDS("rsf.Rds")
+# r = rsf[rsf$region == "Leeds", ]
+# r = r[ukboundaries::leeds, ]
+# saveRDS(r, "rsf_leeds.Rds")
+# piggyback::pb_upload("rsf_leeds.Rds")
+# piggyback::pb_download_url("rsf_leeds.Rds")
+
+u = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rsf_leeds.Rds"
+r_original = readRDS(url(u))
+nrow(r_original)
+plot(r_original$geometry)
+head(r_original)
+names(r_original)
+
+ur = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rtid.csv"
+rtid = readr::read_csv(ur)
+
+# parameters --------------------------------------------------------------
+
+min_pctgov = 0
+min_grouped_pctgov = 100
+min_grouped_length = 500
+city_centre_buffer_radis = 8000
+city_centre = tmaptools::geocode_OSM("leeds", as.sf = TRUE)
+city_centre_buffer = stplanr::geo_buffer(city_centre, dist = 8000)
+
+# preprocess + filter -----------------------------------------------------
+
+rj = inner_join(r_original, rtid)
+summary(as.factor(rj$highway))
+
+# remove motorways
+r = rj %>% 
+  filter(!grepl(pattern = "motorway", x = highway))
+summary(r$length)
+
+# cycleways
+cy = r %>% filter(highway == "cycleway")
+plot(st_geometry(cy))
+
+# simple pctgov and lanes filter ------------------------------------------
+
+r_pct_lanes = r %>% 
+  filter(pctgov > min_pctgov) %>% 
+  filter((lanespsvforward + lanesforward) > 1 |
+      (lanespsvbackward + lanesbackward) > 1
+  )
+
+plot(st_geometry(r_pct_lanes))
+
+# issues: lots of 'bitty' sections, lanes extend way beyond city centre (pct threshold issue?)
+# solution: add new variable with groups: https://gis.stackexchange.com/questions/310462
[…]

### 2020-05-08 get-data-db.R

@@ -18,6 +18,11 @@ for(i in seq_along(f)) {
 r = data.table::fread("roads.csv")
 setwd(old)
 
+region_name = "Leeds"
+summary(sel_region <- r$region == region_name)
+r_region = r[sel_region, ]
+
+  
 # test read-in as geographic object ---------------------------------------
 
 names(r)

### 2020-05-08 README.Rmd
Add article

@@
 rtid = readr::read_csv("~/cyipt/cyipt-server-data/malcolm/roadtypes.csv")
+readr::write_csv(rtid, "rtid.csv")
+piggyback::pb_upload("rtid.csv")
 rsf = left_join(rsf, rtid)

### 2020-05-08 exploratory-analysis-leeds.R

@@ -98,5 +98,5 @@ m2 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
   tm_layout(title = "Group then filter (length > 500, cycling_potential > 100)")
 m2
 # to go into paper
-tmap_arrange(m1, m2)
-
+ma = tmap_arrange(m1, m2)
+tmap_save(.Last.value, "figures/grouping.png", width = 7, height = 4)

### 2020-05-08 article.Rmd
Update article

@@
 ---
-title: "Methods to support cycleways and pedestrianised spaces in response to the Covid-19 pandemic: overview and findings from 10 UK cities"
+title: "Methods to prioritise pop-up active transport infrastructure: overview and findings from 10 UK cities"
 author: "Robin Lovelace"
 output: github_document
+bibliography: references.bib
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)
 ```
+
+```{r, echo=FALSE, eval=FALSE}
+citr::tidy_bib_file("article.Rmd", "~/uaf/allrefs.bib", file = "references.bib")
+```
+
+# RESEARCH QUESTIONS AND HYPOTHESIS
+
+The Covid-19 pandemic has transformed all sectors of the economy, not least transport.
+[Demand for long distance trips has plummetted](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies have gone bust](https://arxiv.org/abs/2004.08460) [@iacus_estimating_2020; @jittrapirom_exploratory_2020].
+Yet short distance travel patterns have also changed greatly and there has been an uptick in cycling in some places [@harrabin_boom_2020].
+There are two main explanations for increased active transport during pandemic-related physical distancing measures: 1) with fewer opportunities to exercise, more people are taking to walking, running and cycling in their local area for mental and physical health; and 2) public transport options have been greatly reduced due to cuts in services and fear of being infected while in enclosed spaces, meaning that walking and particularly cycling may be the only way that workers can reach key destinations such as hospitals.
+The second reason is particularly important given that many 'key workers' are low paid and may have access to a working private automobile.
+From a physical activity perspective this change is welcome: obesity is a powerful predictor of all-cause mortality, including from Covid-19 [@docherty_features_2020].
+From a physical distancing perspective, increased levels of walking and cycling,
+<!-- particularly in densly populated urban areas where pavements and cycleways may be narrow,  -->
+the shift creates pressure on governments to ensure sufficient 'space for social distancing', while enabling increased active mobility for health and travel to key workplaces.
+
+Local and national governments worldwide are still working out how best to respond to these changes and many options are available to ensure that citizens can benefit from outdoor activity while minimizing health risks, ranging from the provision of hand sanitisers to opening-up spaces such as parking lots and golf courses [@freeman_covid19_2020].
+Many local governments have responded by implementing 'pop-up' active transport infrastructure, including:
+
+1. 'filtered permeability' interventions to prevent through traffic
+2. banning cars and to pedestrianise streets, greatly increasing the width of walkable space, New York's 'Open Streets' initiative being a prominent example [@litman_pandemicresilient_2020], and
+3. from the through-flow of motor vehicles
+
+The Scottish government, for example have provided emergency funding "to keep key workers moving" by "reallocating road space to better enable this shift and make it safer for people who choose to walk, cycle or wheel for essential trips or for exercise" [@transportscotland_10_2020].
+
+Hundreds of kilometres of temporary infrastructure being put in place in dozens of cities worldwide, from [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e) to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf), never has there been such a need for affordable, healthy and socially isolated transport that does not create the fumes that emerging evidence links to coronavirus susceptibility.
+With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised.
+As with the medical science, research is needed now.
+Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in now is effective, safe, and placed where it is most needed.
+
+# METHODS AND DATA
+
+```{r load}
+library(sf)
+library(tidyverse)
+library(tmap)
+u = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rsf_leeds.Rds"
+r_original = readRDS(url(u))
+ur = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rtid.csv"
+rtid = readr::read_csv(ur)
+```
+
+```{r parameters}
+min_pctgov = 0
+min_grouped_pctgov = 100
+min_grouped_length = 500
+city_centre_buffer_radis = 8000
+city_centre = tmaptools::geocode_OSM("leeds", as.sf = TRUE)
+city_centre_buffer = stplanr::geo_buffer(city_centre, dist = 8000)
+```
+
+```{r preprocess}
+rj = inner_join(r_original, rtid)
+# remove motorways
+r = rj %>% 
+  filter(!grepl(pattern = "motorway", x = highway))
+cy = r %>% filter(highway == "cycleway")
+```
+
+## Levels of analysis
+
+An important distinction when developing methods for automated analysis of transport networks is the level of analysis.
+
+```{r levels}
+r_pct_lanes = r %>% 
+  filter(pctgov > min_pctgov) %>% 
+  filter((lanespsvforward + lanesforward) > 1 |
+      (lanespsvbackward + lanesbackward) > 1
+  )
+touching_list = st_touches(r_pct_lanes)
+g = igraph::graph.adjlist(touching_list)
+components = igraph::components(g)
+r_pct_lanes$group = components$membership
+r_pct_lanes = r_pct_lanes %>% 
+  group_by(group) %>% 
+  mutate(group_length = sum(length)) %>% 
+  mutate(pctgov_mean = mean(pctgov, na.rm = TRUE)) %>% 
+  filter(pctgov_mean > min_grouped_pctgov)
+r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length)
+r_filter_before_grouping = r %>% 
+  filter(pctgov > min_pctgov) %>% 
+  filter((lanespsvforward + lanesforward) > 1 |
+           (lanespsvbackward + lanesbackward) > 1
+  ) %>% 
+  filter(pctgov > min_grouped_pctgov) %>% 
+  filter(length > 100)
+tmap_mode("plot")
+m1 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
+  tm_shape(r_filter_before_grouping) + tm_lines() +
+  tm_layout(title = "Filter then group (length > 100, cycling_potential > 100)")
+m2 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
+  tm_shape(r_pct_lanes) + tm_lines() +
+  tm_layout(title = "Group then filter (length > 500, cycling_potential > 100)")
+ma = tmap_arrange(m1, m2)
+ma
+```
+
+```{r}
+```
+
+# FINDINGS
 

### 2020-05-09 conversation.Rmd
Update article

@@
 
-Local and national governments worldwide are still working out how best to respond to these changes to lif.
-Hundreds of kilometres of temporary infrastructure being put in place in dozens of cities worldwide, from [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e "") to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/ "") to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf ""), never has there been such a need for affordable, healthy and socially isolated transport that does not create the fumes that emerging evidence links to coronavirus susceptibility.
+Local and national governments worldwide are still working out how best to respond to these changes.
+Hundreds of kilometres of temporary infrastructure being put in place in dozens of cities worldwide, from [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e) to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf), never has there been such a need for affordable, healthy and socially isolated transport that does not create the fumes that emerging evidence links to coronavirus susceptibility.
 With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised.

### 2020-05-09 cities/bristol.Rmd
Tidy-up

# Bristol

```{r bristol}
region = las_top %>% filter(name == "Bristol")
region = sf::st_make_valid(region)
roads = rsf %>% filter(region == "Bristol", pctgov > 100)
roads = st_intersection(roads, region)

bristol_centre = tmaptools::geocode_OSM("Bristol", as.sf = TRUE)
bristol_buff = stplanr::geo_buffer(bristol_centre, dist = 1500)
roads_top = roads[bristol_buff, ] %>%
  filter(length > 300) %>%
  filter(name != "Temple Way Underpass") %>%
  filter(!name == "Newfoundland Street", !name == "") %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)

tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = c(-2.5879, 51.454, 13)) + 
  tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```

### 2020-05-09 cities/leeds.Rmd
Tidy-up

# Leeds

```{r leeds}
leeds_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
leeds_buff = stplanr::geo_buffer(leeds_centre, dist = 4000)
roads = rsf %>% filter(region == "Leeds", pctgov > 100)
region_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
roads = st_intersection(roads, region_buff)

roads_top = roads[leeds_buff, ] %>%
  filter(length > 300) %>%
  filter(name != "Broadway") %>%
  filter(name != "Wellington Street") %>% 
  arrange(desc(pctgov)) %>%
  slice(1:10)
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = c(-1.549, 53.805, 13)) + 
  tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```

### 2020-05-09 cities/leicester.Rmd
Tidy-up

--- 2020-05-06 sheffield.Rmd
@@
-# Sheffield
+# Leicester
 
-```{r sheffield}
-roads = rsf %>% filter(region == "Sheffield", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Sheffield", as.sf = TRUE)
+```{r leicester}
+roads = rsf %>% filter(region == "Leicester", pctgov > 100)
+region_centre = tmaptools::geocode_OSM("Leicester", as.sf = TRUE)
 region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
@@
   filter(length > 200) %>%
-  filter(!name == "Tinsley Roundabout")  %>%
-  filter(!name == "Meadowhead Roundabout")  %>%
   arrange(desc(pctgov)) %>%

### 2020-05-09 cities/liverpool.Rmd
Tidy-up

# Liverpool

```{r liverpool}
roads = rsf %>% filter(region == "Liverpool", pctgov > 100)
region = las_top %>% filter(name == "Liverpool")
roads = st_intersection(roads, region)
# region_centre = tmaptools::geocode_OSM("Liverpool", as.sf = TRUE)
# region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
# tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
#   tm_view(set.view = 10) + tm_scale_bar() +
#   tm_shape(region) + tm_borders()

roads_top = roads %>%
  filter(length > 200) %>%
  filter(name != "Queensway (Mersey Tunnel)") %>%
  filter(name != "Queensway") %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = c(-2.95, 53.4084, 13)) + 
  tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```

### 2020-05-09 cities/manchester.Rmd
Tidy-up

# Manchester

```{r manchester}
roads = rsf %>% filter(region == "Manchester", pctgov > 100)
region_centre = tmaptools::geocode_OSM("Manchester", as.sf = TRUE)
region_buff = stplanr::geo_buffer(region_centre, dist = 7000)
roads = st_intersection(roads, region_buff)

roads_top = roads %>%
  filter(length > 300) %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = 12) + tm_scale_bar() +
  tm_basemap(server = s)
roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```

### 2020-05-09 README.Rmd
Update location of child .Rmd files
Fix minor issues with README

@@
 These cities represent around 1/4 of the population of England.
-Welsh and Scottish cities with high cycling potential such as Cardiff and Edinbugh were not included in the analysis because the CyIPT does not currently have data outside of England, although we could extend the methods to cover all UK cities at some point.
+Welsh and Scottish cities with high cycling potential such as Cardiff and Edinburgh were not included in the analysis because the CyIPT does not currently have data outside of England, although we could extend the methods to cover all UK cities at some point.
 
@@
 
-The cycling potential of the 'top 10' streets is presented in terms of the 'Government Target' scenario, which represents a doubling in cycling compared with 2011 levels.
+The cycling potential of the top 10 streets is calculated based on the 'Government Target' scenario in the Propensity to Cycle Tool, which represents a doubling in cycling compared with 2011 levels.
 London is close to meeting this target already.
@@
 
-The results are not a definitive list of places where pop-up cycleways should be prioritised but a 'starter for 10' highlighting roads that good candidates for 'pop-up' active transport infrastructure through reallocation of road space, as has happened in [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e).
+The results are not a definitive list of places where pop-up cycleways should be prioritised but a 'starter for 10' highlighting roads that may be good candidates for 'pop-up' active transport infrastructure.
+There are many types of pop-up infrastructure, but focus of this project is through reallocation lanes of traffic, as planned for [Park Lane and other wide roads in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e).
+<!-- It seems this scheme is not in operation yet from https://www.tfljamcams.net/ -->
+
 The results highlight roads that have cycling potential and at least one spare lane, meaning a 2 lanes in one direction.
-There will be many road sections that would benefit from interventions not shown in the maps below: for example roads with only one lane in each direction could be made oneway temporarily, creating a spare lane for cycleways or extra pavement width.
-Prevent through-flow in residential areas, as is happening in [Lewisham](https://853.london/2020/05/05/lewisham-council-asks-which-streets-it-should-open-up-to-walkers-and-cyclists/) is another option that can complement road reallocation.
+There will be many road sections that would benefit from interventions not shown in the maps below: roads with only one lane in each direction could be made oneway temporarily, creating a spare lane for cycleways or extra pavement width.
+Preventing through-flow in residential areas, as is happening in [Lewisham](https://853.london/2020/05/05/lewisham-council-asks-which-streets-it-should-open-up-to-walkers-and-cyclists/) and [Salford City Council](https://salfordliveablestreets.commonplace.is/) is another option that can complement road reallocation.
 
@@
 
-```{r, child="london.Rmd"}
+```{r, child="cities/london.Rmd"}
 
@@
 
-```{r, child="birmingham.Rmd"}
+```{r, child="cities/birmingham.Rmd"}
 
@@
 
-```{r, child="manchester.Rmd"}
+```{r, child="cities/manchester.Rmd"}
 
@@
 
-```{r, child="leeds.Rmd"}
+```{r, child="cities/leeds.Rmd"}
 
@@
 
-```{r, child="liverpool.Rmd"}
+```{r, child="cities/liverpool.Rmd"}
 
@@
 
-```{r, child="bristol.Rmd"}
+```{r, child="cities/bristol.Rmd"}
 
@@
 
-```{r, child="leicester.Rmd"}
+```{r, child="cities/leicester.Rmd"}
 
@@
 
-```{r, child="sheffield.Rmd"}
+```{r, child="cities/sheffield.Rmd"}
 
@@
 
-```{r, child="newcastle.Rmd"}
+```{r, child="cities/newcastle.Rmd"}
 
@@
 
-```{r, child="cambridge.Rmd"}
+```{r, child="cities/cambridge.Rmd"}
 

### 2020-05-09 conversation.Rmd
Conversation piece

@@
 
-The Covid-19 pandemic has transformed all sectors of the economy, not least transport.
+Around the world, as nations prepare to emerge from months of Covid-19 induced lockdown, life cannot simply return to previous norms.
+The pandemic has transformed all sectors of the economy, not least transport.
 Long distance travel has been dramatically altered, as [demand plummets](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies go bust](https://arxiv.org/abs/2004.08460).
@@
 Forced to stay close to home, many people have started walking and cycling more.
-From a health and well-being perspective, this may be a more important shift, especially in the long term as coronavirus measures are gradually eased.
+From a health and well-being perspective, this may be a more important shift, especially in the long term as lockdown measures are gradually eased.
 
-Local and national governments worldwide are still working out how best to respond to these changes.
-Hundreds of kilometres of temporary infrastructure being put in place in dozens of cities worldwide, from [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e) to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf), never has there been such a need for affordable, healthy and socially isolated transport that does not create the fumes that emerging evidence links to coronavirus susceptibility.
+In the coming weeks and months, active travel will play a vital role in enabling the reopening of many countries.
+As travel increases, we must ensure that safe, socially distanced walking and cycling remains possible.
+Maintaining social distancing on public transport could severely reduce the capacity of rail and bus networks, and there is a risk that this may lead to a rapid rise in road traffic.
+Without careful management, increased congestion could deter others from the streets, especially as [ record bicycle sales ] suggest that many people have taken advantage of lockdown conditions to take up or return to cycling.
+These people may have less experience in dealing with high volumes of road traffic.
+
+Active travel has extensive [ social and environmental benefits ]. Physical activity is associated with XX health conditions….. mental health…..isolation...
+
+Evidence is beginning to emerge of [ potential links ] between air pollution and Covid-19 lethality, not least because many health conditions such as asthma and xxx are [ worsened by poor air quality ].
+Indeed, it is estimated that global improvements in air quality in recent months have led to [ xx thousand fewer deaths ] from conditions such as these.
+This makes it all the more important to restrict any rise in road traffic and to encourage more beneficial means of travel.
+
+Local and national governments worldwide are still working out how best to respond to these challenges.
+On Saturday 9th May, British Transport Secretary Grant Shapps announced a £2 billion investment in measures to promote walking and cycling.
+This follows a £10 million commitment from the Scottish Government and action from Wales.
+Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf).
+Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.
+
+In terms of practical steps to encourage walking and cycling, a range of simple interventions can be undertaken, initially on a temporary basis.
+Roads can be closed to motor vehicles, or physical barriers can be used to enable [ filtered permeability ], in which access for residents is retained but modal filters enable routes for walking and cycling that are faster and more direct than the routes for motor vehicles.
+This is well established in [ the London Borough of Hackney ] where attractive barriers such as planters have been used to good effect.
+Another option involves keeping the roads open while reallocating excess roadspace to widen pavements or introduce pop-up cycleways.
+
+These measures can all be undertaken on a temporary basis, but if action is prioritised in places where there is a *long-term need* for cycle improvements, there is a greater chance that people will demand they become permanent.
+This is what happened with the experimental cycleway in Torrington Place in London, where a 2 way road for motor vehicles has been converted into a thriving 2 way cycleway, enabling social distancing and zero emissions transport for keyworkers in central London.
+[Paris plans].
+
 With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised.
 As with the medical science, research is needed now.
-Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in now is effective, safe, and placed where it is most needed.
+Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in is effective, safe, and placed where it is most needed.
+Agile research can respond to rapidly changing conditions, as government priorities shift and different sectors of the economy reopen.
+
+Building on the previous projects such as the [ Propensity to Cycle Tool ] and the [ Cycling Infrastructure Prioritisation Toolkit ], researchers from the University of Leeds have investigated where roadspace could be reallocated for pop-up cycleways.
+They have mapped out the top locations for intervention in ten major cities across England, with the intention of rolling out these methods across the country in the near future.
+
+The methods have focused on road capacity and cycling potential, identifying roads where a ‘spare lane’ could be repurposed for cycling.
+Cycle potential is assessed based on travel to work data and estimates of cycling uptake under the UK Government’s national target of doubling cycling levels by 2025, with the length of road segments also being used to filter the top schemes.
+
+The use of open data and publicly accessible models is a key aspect of the research.
+Given the global nature of the challenges we face, open and collaborative research is vital so that methods can be applied by as wide a user base as possible.
+The potential for international application can be seen for example in [ WHO report ] modelling cycling uptake in cities in low income countries.
+Within the UK, [ 50+ Local Authorities ] are now using the PCT to inform their strategic cycle network plans.
+
+Maps (see below) show the recommendations for where to introduce pop-up cycleways.
+With the situation rapidly developing, some of the roads shown here, such as [ Princess Street in Manchester ], have already been identified by Local Authorities as targets for intervention.
 

### 2020-05-09 conversation.Rmd
edits

@@
 Without careful management, increased congestion could deter others from the streets, especially as [ record bicycle sales ] suggest that many people have taken advantage of lockdown conditions to take up or return to cycling.
-These people may have less experience in dealing with high volumes of road traffic.
+Some of these cyclists may have less confidence in dealing with high volumes of road traffic.
 
@@
 On Saturday 9th May, British Transport Secretary Grant Shapps announced a £2 billion investment in measures to promote walking and cycling.
-This follows a £10 million commitment from the Scottish Government and action from Wales.
+This follows a £10 million commitment from the Scottish Government and action announced in Wales.
 Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf).
@@
 Agile research can respond to rapidly changing conditions, as government priorities shift and different sectors of the economy reopen.
+Today the greatest need may be for medical workers to safely access [hospitals], tomorrow it could be access to schools or retail centres.
 
@@
 
-The methods have focused on road capacity and cycling potential, identifying roads where a ‘spare lane’ could be repurposed for cycling.
-Cycle potential is assessed based on travel to work data and estimates of cycling uptake under the UK Government’s national target of doubling cycling levels by 2025, with the length of road segments also being used to filter the top schemes.
+The methods have focused on road capacity and cycling potential, identifying roads where a ‘spare lane’ could be re-purposed for cycling.
+Cycle potential is assessed based on travel to work data and cycling uptake scenarios under the UK Government’s national target of doubling cycling levels by 2025, with the length of road segments also being used to filter the top schemes.
 
-The use of open data and publicly accessible models is a key aspect of the research.
+The use of open data and publicly accessible models is a key aspect of this research.
 Given the global nature of the challenges we face, open and collaborative research is vital so that methods can be applied by as wide a user base as possible.
@@
 Maps (see below) show the recommendations for where to introduce pop-up cycleways.
-With the situation rapidly developing, some of the roads shown here, such as [ Princess Street in Manchester ], have already been identified by Local Authorities as targets for intervention.
+With the situation rapidly developing, some of the roads shown here, such as [Princess Street in Manchester](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175), have already been identified by Local Authorities as targets for intervention.
 

### 2020-05-10 article.R
Add test code for #14

@@ -1,0 +1,252 @@
+## ----setup, include=FALSE--------------------------------------------------------------------
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)
+
+
+## ---- echo=FALSE, eval=FALSE-----------------------------------------------------------------
+## citr::tidy_bib_file("article.Rmd", "~/uaf/allrefs.bib", file = "ref.bib")
+## wordcountaddin::word_count("article.Rmd")
+
+
+## ----commonplace, fig.cap="Screenshot from the website salfordliveablestreets.commonplace.is to support local responses to the Covid-19 pandemic, including the prioritisation of pop-up active transport infrastructure."----
+knitr::include_graphics("https://user-images.githubusercontent.com/1825120/81451234-ed82d200-917b-11ea-977d-fff1665378c5.png")
+
+
+## ----load------------------------------------------------------------------------------------
+library(sf)
+library(tidyverse)
+library(tmap)
+if(!file.exists("rsf_leeds.Rds")) {
+  u = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rsf_leeds.Rds"
+  download.file(u, "rsf_leeds.Rds")
+}
+r_original = readRDS("rsf_leeds.Rds")
+if(!file.exists("rtid.csv")) {
+  ur = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rtid.csv"
+  download.file(ur, "rtid.csv")
+}
+rtid = readr::read_csv("rtid.csv")
+# uh = "http://media.nhschoices.nhs.uk/data/foi/Hospital.csv"
+# download.file(uh, "uh.csv")
+h = readr::read_delim("uh-edited.csv", delim = "|")
+# nrow(h) # 1221 hospitals
+h_clean = h %>% 
+  mutate(Longitude = as.numeric(Longitude)) %>% 
+  mutate(Latitude = as.numeric(Latitude)) %>% 
+  filter(!is.na(Longitude)) %>% 
+  filter(!is.na(Latitude))
+hsf = h_clean %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)
+region = ukboundaries::leeds
+tmap_mode("plot")
+
+
+## ----parameters------------------------------------------------------------------------------
+regexclude = "welling"
+min_cycling_potential = 0
[…]

### 2020-05-10 code/article-methods-cambridge-test.R

@@ -1,0 +1,202 @@
+## ----load------------------------------------------------------------------------------------
+library(sf)
+library(tidyverse)
+library(tmap)
+
+if(!exists("region_name")) region_name = "Cambridge"
+
+r_all = readRDS("rsf.Rds")
+rtid = readr::read_csv("rtid.csv")
+hsf = readRDS("hsf.Rds")
+regions = readRDS("~/cyipt/cyipt-bigdata/boundaries/local_authority/local_authority.Rds")
+region = regions %>% filter(lad16nmw == region_name)
+r_original = r_all %>% filter(region == region_name)
+r_original = r_original[st_transform(region, st_crs(r_original)), ]
+
+## ----parameters------------------------------------------------------------------------------
+# regexclude = "welling"
+min_cycling_potential = 0
+min_grouped_cycling_potential = 100
+min_grouped_length = 100
+city_centre_buffer_radius_small = 5000
+city_centre_buffer_radius = 8000
+city_centre_buffer_radius_large = 10000
+key_destination_buffer_radius = 5000
+city_centre = tmaptools::geocode_OSM(region_name, as.sf = TRUE)
+city_centre_buffer_small = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius_small)
+city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
+city_centre_buffer_large = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius_large)
+h_leeds = hsf[city_centre_buffer, ]
+h_leeds_buffer = stplanr::geo_buffer(h_leeds, dist = key_destination_buffer_radius)
+city_key_buffer = st_union(city_centre_buffer, st_union(h_leeds_buffer))
+
+
+## ----preprocess------------------------------------------------------------------------------
+# remove motorways
+r_original$highway_type = r_original$highway
+r_original$highway_type = gsub(pattern = "_link", replacement = "", r_original$highway)
+highway_table = table(r_original$highway)
+highway_rare = highway_table[highway_table < nrow(r_original) / 100]
[…]

### 2020-05-11 conversation.Rmd

@@
 
-In the coming weeks and months, active travel will play a vital role in enabling the reopening of many countries.
+Active travel is already [recommended by the World Health Organisation](https://who.canto.global/pdfviewer/viewer/viewer.html?share=share%2Calbum%2CU6GDM&column=document&id=q3v02qdsh114763ffe3apo4s2p&suffix=pdf) as a safe mode of transport, and in the coming weeks and months it will play a vital role in enabling the reopening of many countries.
 As travel increases, we must ensure that safe, socially distanced walking and cycling remains possible.
 Maintaining social distancing on public transport could severely reduce the capacity of rail and bus networks, and there is a risk that this may lead to a rapid rise in road traffic.
-Without careful management, increased congestion could deter others from the streets, especially as [ record bicycle sales ] suggest that many people have taken advantage of lockdown conditions to take up or return to cycling.
-Some of these cyclists may have less confidence in dealing with high volumes of road traffic.
+[ record bicycle sales ] suggest that many people have taken advantage of lockdown conditions to take up or return to cycling, but without careful management, increased traffic congestion could again deter many would-be cyclists from the streets.
 
@@
 This is what happened with the experimental cycleway in Torrington Place in London, where a 2 way road for motor vehicles has been converted into a thriving 2 way cycleway, enabling social distancing and zero emissions transport for keyworkers in central London.
-[Paris plans].
+In Paris, measures aim to contribute towards Mayor Anne Hidalgo's long-term [Plan Velo](https://www.curbed.com/2020/1/15/21065343/bike-paris-cycling-anne-hidalgo).
 
@@
 The methods have focused on road capacity and cycling potential, identifying roads where a ‘spare lane’ could be re-purposed for cycling.
-Cycle potential is assessed based on travel to work data and cycling uptake scenarios under the UK Government’s national target of doubling cycling levels by 2025, with the length of road segments also being used to filter the top schemes.
+Cycle potential is assessed based on census data for travel to work and cycling uptake scenarios under the UK Government’s national target of doubling cycling levels by 2025.
+The length of road segments is also used to filter the top schemes.
 

### 2020-05-11 conversation.Rmd
Tweak title + fix typos

@@
 ---
-title: "Pop-up cycleways and pedestrian spaces: a peak into post Covid-19 transport futures"
+title: "Pop-up cycleways: a peak into post COVID-19 transport futures"
 ---
 
-Around the world, as nations prepare to emerge from months of Covid-19 induced lockdown, life cannot simply return to previous norms.
+Around the world, as nations prepare to emerge from months of COVID-19 induced lockdown, life cannot simply return to previous norms.
 The pandemic has transformed all sectors of the economy, not least transport.
@@
 
-Evidence is beginning to emerge of [ potential links ] between air pollution and Covid-19 lethality, not least because many health conditions such as asthma and xxx are [ worsened by poor air quality ].
+Evidence is beginning to emerge of [ potential links ] between air pollution and COVID-19 lethality, not least because many health conditions such as asthma and xxx are [ worsened by poor air quality ].
 Indeed, it is estimated that global improvements in air quality in recent months have led to [ xx thousand fewer deaths ] from conditions such as these.
@@
 
-<!-- The Covid-19 pandemic has transformed all sectors of the economy. -->
+<!-- The COVID-19 pandemic has transformed all sectors of the economy. -->
 <!-- Transport patterns have been dramatically altered, with much discussion of plummeting demand for long distance travel. -->
@@
 <!-- Politicians of all stripes recognise the importance of acting now to prevent a surge in car use and resulting air pollution when the lockdown is gradually released. -->
-<!-- Sadiq Kahn has advocated re-allocating road space for walking and cycling, and Boris Johnson has [talked](https://twitter.com/SomethingELLIS/status/1258031387334053888) of a post Covid-19 "new golden age for cycling".
+<!-- Sadiq Kahn has advocated re-allocating road space for walking and cycling, and Boris Johnson has [talked](https://twitter.com/SomethingELLIS/status/1258031387334053888) of a post COVID-19 "new golden age for cycling".
 -->
@@
 -->
-<!-- Polls showing people do not want to 'return to normal' also suggest that the Covid-19 crisis represents an historic opportunity to readjust transport cities in favour of health. -->
+<!-- Polls showing people do not want to 'return to normal' also suggest that the COVID-19 crisis represents an historic opportunity to readjust transport cities in favour of health. -->
 <!-- If pop-up cycleways go where there is *long term need* there is a greater chance that people will demand they become permanent, as happened with the experimental cycleway in Torrington Place in London, where a 2 way road for motor vehicles has been converted into a thriving 2 way cycleway, enabling social distancing and zero emissions transport for keyworkers in central London.
@@
 
-<!-- Results for Manchester shown below, which highlights Priness Street, one of the roads where [space for active travel will be increased](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175) as part of the post Covid-19 recovery package in Greater Manchester.
+<!-- Results for Manchester shown below, which highlights Priness Street, one of the roads where [space for active travel will be increased](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175) as part of the post COVID-19 recovery package in Greater Manchester.
 -->

### 2020-05-11 code/article-methods-cambridge-test.R

@@ -2,8 +2,13 @@
 library(sf)
 library(tidyverse)
 library(tmap)
-
-if(!exists("region_name")) region_name = "Cambridge"
+s = c(
+  "Esri.WorldGrayCanvas",
+  "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
+  "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
+)
+if(!exists("region_name"))
+  region_name = "Leeds"
 
 r_all = readRDS("rsf.Rds")
 rtid = readr::read_csv("rtid.csv")
@@ -11,13 +16,13 @@ hsf = readRDS("hsf.Rds")
 regions = readRDS("~/cyipt/cyipt-bigdata/boundaries/local_authority/local_authority.Rds")
 region = regions %>% filter(lad16nmw == region_name)
 r_original = r_all %>% filter(region == region_name)
-r_original = r_original[st_transform(region, st_crs(r_original)), ]
+# r_original = r_original[st_transform(region, st_crs(r_original)), ]
 
 ## ----parameters------------------------------------------------------------------------------
 # regexclude = "welling"
 min_cycling_potential = 0
 min_grouped_cycling_potential = 100
-min_grouped_length = 100
+min_grouped_length = 500
 city_centre_buffer_radius_small = 5000
 city_centre_buffer_radius = 8000
 city_centre_buffer_radius_large = 10000
@@ -174,10 +179,11 @@ r_pct_top_n = r_pct_top %>%
 
 ## ----res, fig.cap="Results, showing road segments with a spare lane (light blue) and road groups with a minium threshold length, 1km in this case (dark blue). The top 10 road groups are labelled."----
 tmap_mode("view")
-tm_shape(r_pct_no_overlap) + tm_lines(col = "lightblue", lwd = 4, alpha = 0.6) +
-  tm_shape(r_pct_top) + tm_lines(col = "blue", lwd = 4, alpha = 0.6) +
+tm_shape(r_pct_no_overlap) + tm_lines(col = "turquoise", lwd = 6, alpha = 0.6) +
+  tm_shape(r_pct_top) + tm_lines(col = "blue", lwd = 6, alpha = 0.6) +
   tm_shape(r_pct_top_n) + tm_text("name") +
-  tm_shape(h_leeds) + tm_dots(size = 0.1)
+  tm_shape(h_leeds) + tm_dots(size = 0.1) +
[…]

### 2020-05-11 conversation.Rmd
Update conversation article - work in progress

@@
 
-Around the world, as nations prepare to emerge from months of COVID-19 induced lockdown, life cannot simply return to previous norms.
-The pandemic has transformed all sectors of the economy, not least transport.
-Long distance travel has been dramatically altered, as [demand plummets](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies go bust](https://arxiv.org/abs/2004.08460).
-Yet short distance travel patterns have also changed greatly.
-Forced to stay close to home, many people have started walking and cycling more.
-From a health and well-being perspective, this may be a more important shift, especially in the long term as lockdown measures are gradually eased.
+Around the world, as nations prepare to emerge from months of COVID-19 induced lockdown, life cannot simply return to pre-pandemic normal.
+The pandemic has transformed all sectors of the economy, not least transport, with much attention on [plummeting levels of long distance travel](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies going bust](https://arxiv.org/abs/2004.08460).
+Yet short distance travel patterns have also changed greatly and, perhaps, forever.
+This may be a more important shift from a health perspective, especially in the long term as lockdown measures are gradually eased.
 
-Active travel is already [recommended by the World Health Organisation](https://who.canto.global/pdfviewer/viewer/viewer.html?share=share%2Calbum%2CU6GDM&column=document&id=q3v02qdsh114763ffe3apo4s2p&suffix=pdf) as a safe mode of transport, and in the coming weeks and months it will play a vital role in enabling the reopening of many countries.
-As travel increases, we must ensure that safe, socially distanced walking and cycling remains possible.
-Maintaining social distancing on public transport could severely reduce the capacity of rail and bus networks, and there is a risk that this may lead to a rapid rise in road traffic.
-[ record bicycle sales ] suggest that many people have taken advantage of lockdown conditions to take up or return to cycling, but without careful management, increased traffic congestion could again deter many would-be cyclists from the streets.
+Active travel is already [recommended by the World Health Organisation](https://who.canto.global/pdfviewer/viewer/viewer.html?share=share%2Calbum%2CU6GDM&column=document&id=q3v02qdsh114763ffe3apo4s2p&suffix=pdf) during the outbreak.
+With health and scientific evidence at the top of government priorities, walking and cycling will play a vital role in enabling the reopening of many countries.
+There are two main reasons why walking and cycling levels can be expected to grow during the pandemic and its aftermath.
+First, forced to stay close to home, many people have started walking and cycling more for health reasons.
+Second, public transport options have been greatly reduced in many [countries](https://doi.org/10.1126/science.abb6105), voluntary use has fallen, and services that do run will run at a fraction of their intentended capacity.
+As UK Transport Minister Grant Shapps [said](https://www.gov.uk/government/speeches/transport-secretarys-statement-on-coronavirus-covid-19-9-may-2020) on 9th March, "once you take into account the 2 metre social distancing rule, there would only be effective capacity for one in ten passengers on many parts of the network."
+This second reason is particularly important given that many 'key workers' who keep economies and vital services moving are low paid, reliant on public transport, and with limited capacity to buy and run private automobiles.
 
-Active travel has extensive [ social and environmental benefits ]. Physical activity is associated with XX health conditions….. mental health…..isolation...
-
-Evidence is beginning to emerge of [ potential links ] between air pollution and COVID-19 lethality, not least because many health conditions such as asthma and xxx are [ worsened by poor air quality ].
+Furthermore, evidence is beginning to emerge of [ potential links ] between air pollution and COVID-19 lethality, not least because many health conditions such as asthma and xxx are [ worsened by poor air quality ].
 Indeed, it is estimated that global improvements in air quality in recent months have led to [ xx thousand fewer deaths ] from conditions such as these.
@@
 
-Local and national governments worldwide are still working out how best to respond to these challenges.
+To some extend the authorities are responding to bottom-up demand, with [sales of bicycles have gone through the roof](https://www.bbc.co.uk/news/business-52564351) suggesting that there is a huge health opportunity.
+There is a [window of opportunity](link to Forbes article) that could be missed if roads become busy and dangerous again before new active transport habits become embedded.
+
+<!-- Active travel has extensive [ social and environmental benefits ]. Physical activity is associated with XX health conditions….. mental health…..isolation... -->
+In response to this emerging evidence, local and national transport authorities have been scrambling to ensure that walking and cycling levels can grow, while social distancing guidance is observed.
+A range of interventions has been [proposed](link to academic article from USA - rl).
 On Saturday 9th May, British Transport Secretary Grant Shapps announced a £2 billion investment in measures to promote walking and cycling.
@@
 This is well established in [ the London Borough of Hackney ] where attractive barriers such as planters have been used to good effect.
-Another option involves keeping the roads open while reallocating excess roadspace to widen pavements or introduce pop-up cycleways.
+Another option involves keeping the roads open while reallocating excess road space to widen pavements or introduce pop-up cycleways.
 
@@
 
-Building on the previous projects such as the [ Propensity to Cycle Tool ] and the [ Cycling Infrastructure Prioritisation Toolkit ], researchers from the University of Leeds have investigated where roadspace could be reallocated for pop-up cycleways.
+Building on the previous projects such as the [ Propensity to Cycle Tool ] and the [ Cycling Infrastructure Prioritisation Toolkit ], researchers from the University of Leeds have investigated where road space could be reallocated for pop-up cycleways.
 They have mapped out the top locations for intervention in ten major cities across England, with the intention of rolling out these methods across the country in the near future.

### 2020-05-11 conversation.Rmd
Finish draft article

@@
 In response to this emerging evidence, local and national transport authorities have been scrambling to ensure that walking and cycling levels can grow, while social distancing guidance is observed.
-A range of interventions has been [proposed](link to academic article from USA - rl).
-On Saturday 9th May, British Transport Secretary Grant Shapps announced a £2 billion investment in measures to promote walking and cycling.
-This follows a £10 million commitment from the Scottish Government and action announced in Wales.
-Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf).
+<!-- A range of interventions has been [proposed](link to academic article from USA - rl) but it is clear that leadership, national guidance, funding and evidence on *where to invest* is needed to 'lock-in' the benefits of the nascent active transport revolution.
+-->
+In the UK, £2 billion investment in measures to promote walking and cycling has been announced, following a £10 million package from the Scottish Government, action announced in Wales and decisive action by leaders in many other local and national authorities.
+The UK package includes £250 million for "emergency interventions to make cycling and walking safer" as [Shapps put it](https://www.gov.uk/government/speeches/transport-secretarys-statement-on-coronavirus-covid-19-9-may-2020), including pop-up bike lanes, wider pavements and car-free streets.
+
+Many pop-up interventions have already been documented internationally and in the UK and there is much to learn much from these experiences already.
+There are three broad categories of pop-up interventions...
+Geographically, cities with pop-up infrastructure range from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf).
 Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.

### 2020-05-11 conversation.Rmd
Minor tweaks
more references

--- 2020-05-11 conversation.Rmd
@@
 Maintaining social distancing on public transport could severely reduce the capacity of rail and bus networks, and there is a risk that this may lead to a rapid rise in road traffic.
-[ record bicycle sales ] suggest that many people have taken advantage of lockdown conditions to take up or return to cycling, but without careful management, increased traffic congestion could again deter many would-be cyclists from the streets.
+[Record bicycle sales](https://www.bbc.co.uk/news/business-52564351) suggest that many people have taken advantage of lockdown conditions to take up or return to cycling, but without careful management, increased traffic congestion could again deter many would-be cyclists from the streets.
 
-Active travel has extensive [ social and environmental benefits ]. Physical activity is associated with XX health conditions….. mental health…..isolation...
-
-Evidence is beginning to emerge of [ potential links ] between air pollution and COVID-19 lethality, not least because many health conditions such as asthma and xxx are [ worsened by poor air quality ].
-Indeed, it is estimated that global improvements in air quality in recent months have led to [ xx thousand fewer deaths ] from conditions such as these.
+Active travel has extensive [social and environmental benefits](https://www.bma.org.uk/what-we-do/population-health/preventing-ill-health/steps-to-increase-physical-activity-levels-in-the-uk).
+Meeting recommended levels of physical activity can help to prevent or manage 20 chronic health conditions, such as cancer, stroke and type 2 diabetes.
+It can manage or reduce the risk of mental health conditions such as depression and anxiety, and aid with maintaining a healthy weight.
+Evidence is beginning to emerge of potential links between air pollution and COVID-19 [infection](https://www.ncbi.nlm.nih.gov/pubmed/32315904) and [lethality](https://www.sciencedirect.com/science/article/pii/S0269749120320601), and many health conditions are [associated with poor air quality](https://www.ncbi.nlm.nih.gov/pubmed/30419237).
+Indeed, across Europe it is estimated that improvements in air quality in recent months have led to [11,000 fewer deaths](https://energyandcleanair.org/wp/wp-content/uploads/2020/04/CREA-Europe-COVID-impacts.pdf) from conditions such as these.
 This makes it all the more important to restrict any rise in road traffic and to encourage more beneficial means of travel.
@@
 Local and national governments worldwide are still working out how best to respond to these challenges.
-On Saturday 9th May, British Transport Secretary Grant Shapps announced a £2 billion investment in measures to promote walking and cycling.
+On Saturday 9th May, British Transport Secretary Grant Shapps announced a [£2 billion investment](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) in measures to promote walking and cycling.
 This follows a £10 million commitment from the Scottish Government and action announced in Wales.
-Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf).
+Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan](https://www.theguardian.com/world/2020/apr/21/milan-seeks-to-prevent-post-crisis-return-of-traffic-pollution) to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf).
 Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.
@@
 In terms of practical steps to encourage walking and cycling, a range of simple interventions can be undertaken, initially on a temporary basis.
-Roads can be closed to motor vehicles, or physical barriers can be used to enable [ filtered permeability ], in which access for residents is retained but modal filters enable routes for walking and cycling that are faster and more direct than the routes for motor vehicles.
+Roads can be closed to motor vehicles, or physical barriers can be used to enable [filtered permeability](https://cyclehighways.eu/design-and-build/infrastructure/filtered-permeability.html), in which access for residents is retained but modal filters enable routes for walking and cycling that are faster and more direct than the routes for motor vehicles.
 This is well established in [ the London Borough of Hackney ] where attractive barriers such as planters have been used to good effect.

### 2020-05-11 conversation.Rmd
Add tweet

--- 2020-05-11 conversation.Rmd
@@
 ---
-title: "Pop-up cycleways: a peak into post COVID-19 transport futures"
+title: "Pop-up cycleways provide a peak into post COVID-19 transport futures"
 ---
@@
 Many pop-up interventions have already been documented internationally and in the UK and there is much to learn much from these experiences already.
-There are three broad categories of pop-up interventions...
 Geographically, cities with pop-up infrastructure range from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf).
 Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.
+
+These interventions can be classified into three broad groups, as outlined in a [paper](add link) that we have written outlining methods for prioritising pop-up active transport infrastructure.
+
+First...
+
+An example of the first type of intervention is illustrated in the image below, from Salford, Manchester
+
+[![](https://pbs.twimg.com/media/EXZ1t2EXgAELrdO?format=jpg&name=large)](https://twitter.com/CatrionaSwanson/status/1258322956595453952)
+
+<tweet url="https://twitter.com/CatrionaSwanson/status/1258322956595453952" />
+
+Second...
+Third...
 

### 2020-05-12 conversation.Rmd
Update articles

--- 2020-05-11 conversation.Rmd
@@
 ---
-title: "Pop-up cycleways: a peak into post COVID-19 transport futures"
+title: "Pop-up cycleways provide a peak into post COVID-19 transport futures"
 ---
 
-Around the world, as nations prepare to emerge from months of COVID-19 induced lockdown, life cannot simply return to previous norms.
+As nations prepare to emerge from months of COVID-19 induced lockdowns, it is becoming clear life cannot simply return to normal.
 The pandemic has transformed all sectors of the economy, not least transport.
-Long distance travel has been dramatically altered, as [demand plummets](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies go bust](https://arxiv.org/abs/2004.08460).
-Yet short distance travel patterns have also changed greatly.
-Forced to stay close to home, many people have started walking and cycling more.
-From a health and well-being perspective, this may be a more important shift, especially in the long term as lockdown measures are gradually eased.
 
-Active travel is already [recommended by the World Health Organisation](https://who.canto.global/pdfviewer/viewer/viewer.html?share=share%2Calbum%2CU6GDM&column=document&id=q3v02qdsh114763ffe3apo4s2p&suffix=pdf) as a safe mode of transport, and in the coming weeks and months it will play a vital role in enabling the reopening of many countries.
-As travel increases, we must ensure that safe, socially distanced walking and cycling remains possible.
-Maintaining social distancing on public transport could severely reduce the capacity of rail and bus networks, and there is a risk that this may lead to a rapid rise in road traffic.
-[Record bicycle sales](https://www.bbc.co.uk/news/business-52564351) suggest that many people have taken advantage of lockdown conditions to take up or return to cycling, but without careful management, increased traffic congestion could again deter many would-be cyclists from the streets.
+In the coming weeks and months, walking and cycling will play a vital role in enabling the reopening of many countries, as continued social distancing measures will mean that public transport options are likely to run at a fraction of their intended capacity.
+There is a window of opportunity here that could be missed if roads become busy and dangerous again before new active transport habits become embedded.
 
-Active travel has extensive [social and environmental benefits](https://www.bma.org.uk/what-we-do/population-health/preventing-ill-health/steps-to-increase-physical-activity-levels-in-the-uk).
-Meeting recommended levels of physical activity can help to prevent or manage 20 chronic health conditions, such as cancer, stroke and type 2 diabetes.
-It can manage or reduce the risk of mental health conditions such as depression and anxiety, and aid with maintaining a healthy weight.
-Evidence is beginning to emerge of potential links between air pollution and COVID-19 [infection](https://www.ncbi.nlm.nih.gov/pubmed/32315904) and [lethality](https://www.sciencedirect.com/science/article/pii/S0269749120320601), and many health conditions are [associated with poor air quality](https://www.ncbi.nlm.nih.gov/pubmed/30419237).
-Indeed, across Europe it is estimated that improvements in air quality in recent months have led to [11,000 fewer deaths](https://energyandcleanair.org/wp/wp-content/uploads/2020/04/CREA-Europe-COVID-impacts.pdf) from conditions such as these.
-This makes it all the more important to restrict any rise in road traffic and to encourage more beneficial means of travel.
+Local and national governments are scrambling to act.
+In the UK, £2 billion investment in measures to promote walking and cycling has been announced.
+£250 million of this is for "emergency interventions to make cycling and walking safer" as [Transport Secretary Grant Shapps put it](https://www.gov.uk/government/speeches/transport-secretarys-statement-on-coronavirus-covid-19-9-may-2020).
+New [guidance from the Department for Transport](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities/traffic-management-act-2004-network-management-in-response-to-covid-19) advocates "measures to reallocate road space to people walking and cycling".
 
-Local and national governments worldwide are still working out how best to respond to these challenges.
-On Saturday 9th May, British Transport Secretary Grant Shapps announced a [£2 billion investment](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) in measures to promote walking and cycling.
-This follows a £10 million commitment from the Scottish Government and action announced in Wales.
-Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan](https://www.theguardian.com/world/2020/apr/21/milan-seeks-to-prevent-post-crisis-return-of-traffic-pollution) to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf).
-Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.
+<tc-read-more id="137909"></tc-read-more>
 
-In terms of practical steps to encourage walking and cycling, a range of simple interventions can be undertaken, initially on a temporary basis.
-Roads can be closed to motor vehicles, or physical barriers can be used to enable [filtered permeability](https://cyclehighways.eu/design-and-build/infrastructure/filtered-permeability.html), in which access for residents is retained but modal filters enable routes for walking and cycling that are faster and more direct than the routes for motor vehicles.
-This is well established in [ the London Borough of Hackney ] where attractive barriers such as planters have been used to good effect.
-Another option involves keeping the roads open while reallocating excess roadspace to widen pavements or introduce pop-up cycleways.
+Local authorities urgently need answers on how this funding should be spent, and where road space reallocation should be prioritised.
+New cyclinginfrastructure which is developed based on rigorous analysis of the best available data, in combination with local knowledge and community engagement, are more likely to be effective than plans based on intuition alone.
+Recognising the importance of a strong and actionable evidence-base, and the window of opportunity, we used our experience developing tools that are already being used by local authorities, notably the multi-institution PCT and the CyIPT, to develop methods and outputs taylored for the new transport planning landscape.
 
-These measures can all be undertaken on a temporary basis, but if action is prioritised in places where there is a *long-term need* for cycle improvements, there is a greater chance that people will demand they become permanent.
-This is what happened with the experimental cycleway in Torrington Place in London, where a 2 way road for motor vehicles has been converted into a thriving 2 way cycleway, enabling social distancing and zero emissions transport for keyworkers in central London.
-In Paris, measures aim to contribute towards Mayor Anne Hidalgo's long-term [Plan Velo](https://www.curbed.com/2020/1/15/21065343/bike-paris-cycling-anne-hidalgo).
+<image id="334100" align="centre" caption="Screenshot from the website salfordliveablestreets.commonplace.is to support local responses to the COVID-19 pandemic, including the prioritisation of pop-up active transport infrastructure." />
 
-With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised.
-As with the medical science, research is needed now.
-Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in is effective, safe, and placed where it is most needed.
-Agile research can respond to rapidly changing conditions, as government priorities shift and different sectors of the economy reopen.
-Today the greatest need may be for medical workers to safely access [hospitals], tomorrow it could be access to schools or retail centres.
+In a pre-print [paper](https://osf.io/7wjb6) on methods to prioritise pop-up road space reallocation schemes, we have classified pop-up interventions into three categories.
+Firstly, in streets that are wide or have multiple lanes, roadspace can be reallocated to widen pavements or introduce pop-up cycleways, such as outside a hospital in [Leicester](https://www.leicestermercury.co.uk/news/leicester-news/light-lockdown-traffic-leicester-allows-4083625).
+Secondly, physical barriers can be used to make routes for walking and cycling [faster and more direct](https://cyclehighways.eu/design-and-build/infrastructure/filtered-permeability.html) than the routes for motor vehicles.
+This can be seen in Tower Hamlets, London, where simple barriers like planters are used to good effect.
+Thirdly, roads can be closed altogether to motor vehicles, as is happening in [Deansgate, Manchester](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175).
 
-Building on the previous projects such as the [ Propensity to Cycle Tool ] and the [ Cycling Infrastructure Prioritisation Toolkit ], researchers from the University of Leeds have investigated where roadspace could be reallocated for pop-up cycleways.
-They have mapped out the top locations for intervention in ten major cities across England, with the intention of rolling out these methods across the country in the near future.
-
-The methods have focused on road capacity and cycling potential, identifying roads where a ‘spare lane’ could be re-purposed for cycling.
-Cycle potential is assessed based on census data for travel to work and cycling uptake scenarios under the UK Government’s national target of doubling cycling levels by 2025.
+The first of these three options is the most simple to implement,  with relatively few implications for traffic flow.
+We have mapped out **link** the top locations for this reallocation of road space in ten major cities across England, with the intention of rolling out these methods across the country in the near future.
+Having identified roads where a spare lane could be re-purposed for cycling, we have assessed the cycle potential of these streets based on census data for travel to work and cycling uptake scenarios under the UK Government’s national target of doubling cycling levels by 2025.
 The length of road segments is also used to filter the top schemes.
@@
 Given the global nature of the challenges we face, open and collaborative research is vital so that methods can be applied by as wide a user base as possible.
-The potential for international application can be seen for example in [ WHO report ] modelling cycling uptake in cities in low income countries.
-Within the UK, [ 50+ Local Authorities ] are now using the PCT to inform their strategic cycle network plans.
+The potential for international application can be seen for example in [research for the World Health Organisation](https://atfutures.github.io/upthat/articles/adaptation.html) modelling cycling uptake in cities in low income countries.
+Within the UK, [50+ Local Authorities](https://npct.github.io/pct-shiny/regions_www/www/static/03d_other_reports/2019-use-of-pct-report.pdf) are now using the PCT to inform their strategic cycle network plans.
 
-Maps (see below) show the recommendations for where to introduce pop-up cycleways.
-With the situation rapidly developing, some of the roads shown here, such as [Princess Street in Manchester](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175), have already been identified by Local Authorities as targets for intervention.
+Results from our [methods](https://osf.io/7wjb6) show that many cities have 'spare lanes' under the current levels of traffic in city centres and near key worker destinations such as hospitals.
+These results could support the prioritisation of vital resources dedicated to pop-up infrastructure.
+With the situation rapidly developing, some of the roads shown here, such as [Kirkstall Road](https://www.transportxtra.com/publications/local-transport-today/news/65382/leeds-to-pilot-enhanced-cycling-safety-u) (in the top 10 results our methods generated for Leeds), have already been identified by Local Authorities as targets for intervention.
+As the recovery continues, we hope our methods will support cities to take the opportunities that the crisis presents for active travel.
+
+<image id="334113" align="centre" caption="Roads in Leeds with at least one spare lane that could be reallocate to provide space for active travel, and locations of hospitals" />
+
+<tweet url="https://twitter.com/laura_laker/status/1257660564098162689" />
+
+With the rush to act, there is a great need for evidence of where new interventions should be prioritised.
+As with the medical science, research is needed now.
+Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in is effective, safe, and placed where it is most needed.
+Agile research can respond to rapidly changing conditions, as government priorities shift and different sectors of the economy reopen.
+Measures can be undertaken on a temporary basis, but if action is prioritised in places where there is a long-term need for cycle improvements, there is a greater chance that they become permanent.
+In Paris, Covid-19 related measures aim to contribute towards Mayor Anne Hidalgo's long-term [Plan Velo](https://www.curbed.com/2020/1/15/21065343/bike-paris-cycling-anne-hidalgo).
 

### 2020-05-12 conversation.Rmd
Update conversation

@@
 ---
+As cities across the world prepare to emerge from lockdown, continued social distancing measures will mean that public transport options are likely to run at a fraction of their intended capacity.
+This is a crucial moment for urban cycling.
 
-As nations prepare to emerge from months of COVID-19 induced lockdowns, it is becoming clear life cannot simply return to normal.
-The pandemic has transformed all sectors of the economy, not least transport.
+There is a window of opportunity to change urban travel.
+Acting fast, [before roads become busy again](https://www.forbes.com/sites/carltonreid/2020/05/07/car-traffic-rise-of-11-shows-urgent-need-for-pop-ups-to-protect-cyclists-and-pedestrians/amp/), will enable new and healthy active transport habits to become embedded long-term.
 
-In the coming weeks and months, walking and cycling will play a vital role in enabling the reopening of many countries, as continued social distancing measures will mean that public transport options are likely to run at a fraction of their intended capacity.
-There is a window of opportunity here that could be missed if roads become busy and dangerous again before new active transport habits become embedded.
+Our research at the University of Leeds is investigating how best to make rapid changes to urban cycling infrastructure.
+We have developed [methods](https://osf.io/7wjb6) to identify roads where there is space for pop-up infrastructure, and where it is most needed.
 
-Local and national governments are scrambling to act.
-In the UK, £2 billion investment in measures to promote walking and cycling has been announced.
-£250 million of this is for "emergency interventions to make cycling and walking safer" as [Transport Secretary Grant Shapps put it](https://www.gov.uk/government/speeches/transport-secretarys-statement-on-coronavirus-covid-19-9-may-2020).
-New [guidance from the Department for Transport](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities/traffic-management-act-2004-network-management-in-response-to-covid-19) advocates "measures to reallocate road space to people walking and cycling".
+The UK government has announced £2 billion of investment in measures to promote walking and cycling in England over the next 5 years.
+£250 million has been allocated for "emergency interventions to make cycling and walking safer" as [Transport Secretary Grant Shapps put it](https://www.gov.uk/government/speeches/transport-secretarys-statement-on-coronavirus-covid-19-9-may-2020), with similar commitments being made by [Scottish](https://www.transport.gov.scot/news/10-million-to-support-pop-up-active-travel-infrastructure/) and [Welsh](https://gov.wales/councils-challenged-change-wales-transport-environment-during-coronavirus-lockdown) governments.
 
@@
 
-Local authorities urgently need answers on how this funding should be spent, and where road space reallocation should be prioritised.
-New cyclinginfrastructure which is developed based on rigorous analysis of the best available data, in combination with local knowledge and community engagement, are more likely to be effective than plans based on intuition alone.
-Recognising the importance of a strong and actionable evidence-base, and the window of opportunity, we used our experience developing tools that are already being used by local authorities, notably the multi-institution PCT and the CyIPT, to develop methods and outputs taylored for the new transport planning landscape.
+Local authorities urgently need to decide how this funding should be spent, and where the reallocation of road space should be prioritised.
+New cycling infrastructure is more likely to be effective when it is developed based on analysis of the best available data, in combination with local knowledge and community engagement.
 
-<image id="334100" align="centre" caption="Screenshot from the website salfordliveablestreets.commonplace.is to support local responses to the COVID-19 pandemic, including the prioritisation of pop-up active transport infrastructure." />
+<tweet url="https://twitter.com/laura_laker/status/1257660564098162689" />
 
-In a pre-print [paper](https://osf.io/7wjb6) on methods to prioritise pop-up road space reallocation schemes, we have classified pop-up interventions into three categories.
-Firstly, in streets that are wide or have multiple lanes, roadspace can be reallocated to widen pavements or introduce pop-up cycleways, such as outside a hospital in [Leicester](https://www.leicestermercury.co.uk/news/leicester-news/light-lockdown-traffic-leicester-allows-4083625).
-Secondly, physical barriers can be used to make routes for walking and cycling [faster and more direct](https://cyclehighways.eu/design-and-build/infrastructure/filtered-permeability.html) than the routes for motor vehicles.
-This can be seen in Tower Hamlets, London, where simple barriers like planters are used to good effect.
-Thirdly, roads can be closed altogether to motor vehicles, as is happening in [Deansgate, Manchester](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175).
+If action is prioritised in places where there is a long-term need for cycle improvements, there is a greater chance that these pop-up developments can become permanent.
+In Paris, COVID-19 related measures aim to contribute towards Mayor Anne Hidalgo's long-term [Plan Velo](https://www.curbed.com/2020/1/15/21065343/bike-paris-cycling-anne-hidalgo).
 
-The first of these three options is the most simple to implement,  with relatively few implications for traffic flow.
-We have mapped out **link** the top locations for this reallocation of road space in ten major cities across England, with the intention of rolling out these methods across the country in the near future.
-Having identified roads where a spare lane could be re-purposed for cycling, we have assessed the cycle potential of these streets based on census data for travel to work and cycling uptake scenarios under the UK Government’s national target of doubling cycling levels by 2025.
-The length of road segments is also used to filter the top schemes.
+We have experience of building tools and software to help highlight roads that may be suitable candidates for pop-up infrastructure.
+The [Propensity to Cycle Tool](https://www.pct.bike/), built in collaboration with other universities and funded by the Department for Transport, helps local authorities identify [where cycle networks](https://npct.github.io/pct-shiny/regions_www/www/static/03d_other_reports/2019-use-of-pct-report.pdf) can be established.
+
+Now, we have used this knowledge to generate results that can support investment in pop-up infrastructure.
+As with the PCT, the visualised and interpreted using interactive maps.
+The accessible and open nature of the results encourages community engagement and online debate, based on the evidence.
+
+<tweet url="https://twitter.com/robinlovelace/status/1258411769682366466" />
+
+In streets that are wide or have multiple lanes, road space can be reallocated to widen pavements or introduce pop-up cycleways, such as outside a hospital in [Leicester](https://www.leicestermercury.co.uk/news/leicester-news/light-lockdown-traffic-leicester-allows-4083625).
+This is simple to implement,  with relatively few implications for traffic flow.
+
+[Publicly available and interactive preliminary results](https://cyipt.github.io/popupCycleways/) for 10 cities highlight the potential of the methods and provide an opportunity for people with local knowledge to feedback, via ['issues' on the collaboration platform GitHub](https://github.com/cyipt/popupCycleways/issues) (requires a [GitHub](https://github.com/) account).
+Our work has already been [used](https://www.cyclinguk.org/press-release/100-miles-pop-cycle-lanes-could-enable-millions-cycle-safely-post-lockdown) to support cycling advocates to push pop-up infrastructure where it is most needed.
+An [interactive map of local interventions nationwide](https://www.cyclinguk.org/covid-19-safe-space-social-distancing) highlights what is possible and provides a basis for engaging with local authorities to encourage swift and effective action to encourage walking and cycling.
+Our research will contribute to the effort by strengthening the case for pop-up cycleways and highlighting available road space and locations with high long-term cycling potential.
+
+The results are based on two main inputs: the Propensity to Cycle Tools's analysis of Census data under the UK Government’s [national target](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy) of doubling cycling levels by 2025, and data on the characteristics of road sections from the CyIPT project.
+Roads with spare lanes were defined as roads on which there are at least two lanes going in the same direction.
+The methods, based on open data, open source software and reproducible code, could scale nationwide and to other cities worldwide.
+Further refinements of the methods could include the prioritisation of roads to 'key destinations' such as hospitals (with hospitals in Leeds shown in the map below) and the inclusion of other relevant variables such as road width and safety statistics.
+
+<image id="334113" align="centre" caption="Roads in Leeds with at least one spare lane that could be reallocated to provide space for active travel, and locations of hospitals" />
 
@@
 The potential for international application can be seen for example in [research for the World Health Organisation](https://atfutures.github.io/upthat/articles/adaptation.html) modelling cycling uptake in cities in low income countries.
-Within the UK, [50+ Local Authorities](https://npct.github.io/pct-shiny/regions_www/www/static/03d_other_reports/2019-use-of-pct-report.pdf) are now using the PCT to inform their strategic cycle network plans.
+Within the UK, [over 50 Local Authorities](https://npct.github.io/pct-shiny/regions_www/www/static/03d_other_reports/2019-use-of-pct-report.pdf) are now using the PCT to inform their strategic cycle network plans.
 
-Results from our [methods](https://osf.io/7wjb6) show that many cities have 'spare lanes' under the current levels of traffic in city centres and near key worker destinations such as hospitals.
-These results could support the prioritisation of vital resources dedicated to pop-up infrastructure.
-With the situation rapidly developing, some of the roads shown here, such as [Kirkstall Road](https://www.transportxtra.com/publications/local-transport-today/news/65382/leeds-to-pilot-enhanced-cycling-safety-u) (in the top 10 results our methods generated for Leeds), have already been identified by Local Authorities as targets for intervention.
-As the recovery continues, we hope our methods will support cities to take the opportunities that the crisis presents for active travel.
-
-<image id="334113" align="centre" caption="Roads in Leeds with at least one spare lane that could be reallocate to provide space for active travel, and locations of hospitals" />
-
-<tweet url="https://twitter.com/laura_laker/status/1257660564098162689" />
-
-With the rush to act, there is a great need for evidence of where new interventions should be prioritised.
-As with the medical science, research is needed now.
-Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in is effective, safe, and placed where it is most needed.
-Agile research can respond to rapidly changing conditions, as government priorities shift and different sectors of the economy reopen.
-Measures can be undertaken on a temporary basis, but if action is prioritised in places where there is a long-term need for cycle improvements, there is a greater chance that they become permanent.
-In Paris, Covid-19 related measures aim to contribute towards Mayor Anne Hidalgo's long-term [Plan Velo](https://www.curbed.com/2020/1/15/21065343/bike-paris-cycling-anne-hidalgo).
+An important [finding](https://osf.io/7wjb6) is that many cities have excess road space, under reduced traffic conditions, on routes leading to key destinations.
+Identifying such roads, like [Kirkstall Road](https://www.openstreetmap.org/way/7991370#map=15/53.8049/-1.5718), highlighted by the results shown in the map above and already part of plans by [Leeds City Council](https://news.leeds.gov.uk/news/new-measures-to-support-social-distancing-for-people-walking-and-cycling-in-leeds) to become a [pilot pop-up cycleway](https://www.transportxtra.com/publications/local-transport-today/news/65382/leeds-to-pilot-enhanced-cycling-safety-u), could be the key to effective investment in pop-up infrastructure.
+The evidence generated by our methods, in combination with community engagement, can ensure that transport interventions made during the crisis are of maximum benefit now and in the post pandemic world.
 

### 2020-05-12 code/article-methods-cambridge-test.R
Add input parameters

@@ -11,6 +11,7 @@ if(!exists("region_name"))
   region_name = "Leeds"
 
 r_all = readRDS("rsf.Rds")
+region_names = unique(rsf$region)
 rtid = readr::read_csv("rtid.csv")
 hsf = readRDS("hsf.Rds")
 regions = readRDS("~/cyipt/cyipt-bigdata/boundaries/local_authority/local_authority.Rds")
@@ -27,6 +28,20 @@ city_centre_buffer_radius_small = 5000
 city_centre_buffer_radius = 8000
 city_centre_buffer_radius_large = 10000
 key_destination_buffer_radius = 5000
+
+parameters_df = data.frame(
+  name = region_name,
+  min_cycling_potential = 0,
+  min_grouped_cycling_potential = 500,
+  min_grouped_length = 500,
+  city_centre_buffer_radius = 8000,
+  key_destination_buffer_radius = 5000
+)
+
+parameters_df = parameters_df[rep(1, length(region_names)), ]
+parameters_df$name = region_names
+write_csv(parameters_df, "input-data/parameters.csv")
+
 city_centre = tmaptools::geocode_OSM(region_name, as.sf = TRUE)
 city_centre_buffer_small = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius_small)
 city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)

### 2020-05-13 README.Rmd
Updated results - leeds + london

@@
 piggyback::pb_download("las_top.Rds")
-pct_las = sf::read_sf("https://github.com/cyipt/tempCycleways/releases/download/0.1/pct_la_summaries.geojson")
-las_top = sf::read_sf("https://github.com/cyipt/tempCycleways/releases/download/0.1/las_top.geojson")
+pct_las = sf::read_sf("https://github.com/cyipt/popupCycleways/releases/download/0.1/pct_la_summaries.geojson")
+las_top = sf::read_sf("https://github.com/cyipt/popupCycleways/releases/download/0.1/las_top.geojson")
 # las_top = readRDS("las_top.Rds")
@@
 
-See [london.html](https://cyipt.github.io/tempCycleways/london.html) for the interactive version of the map above.
+See [london.html](https://cyipt.github.io/popupCycleways/london.html) for the interactive version of the map above.
 
@@
 
-See [birmingham.html](https://cyipt.github.io/tempCycleways/birmingham.html) for the interactive version of the map above.
+See [birmingham.html](https://cyipt.github.io/popupCycleways/birmingham.html) for the interactive version of the map above.
 
@@
 
-See [manchester.html](https://cyipt.github.io/tempCycleways/manchester.html) for the interactive version of the map above.
+See [manchester.html](https://cyipt.github.io/popupCycleways/manchester.html) for the interactive version of the map above.
 
@@
 
-See [leeds.html](https://cyipt.github.io/tempCycleways/leeds.html) for the interactive version of the map above.
+See [leeds.html](https://cyipt.github.io/popupCycleways/leeds.html) for the interactive version of the map above.
 
@@
 
-See [liverpool.html](https://cyipt.github.io/tempCycleways/liverpool.html) for the interactive version of the map above.
+See [liverpool.html](https://cyipt.github.io/popupCycleways/liverpool.html) for the interactive version of the map above.
 
@@
 
-See [bristol.html](https://cyipt.github.io/tempCycleways/bristol.html) for the interactive version of the map above.
+See [bristol.html](https://cyipt.github.io/popupCycleways/bristol.html) for the interactive version of the map above.
 
@@
 
-See [leicester.html](https://cyipt.github.io/tempCycleways/leicester.html) for the interactive version of the map above.
+See [leicester.html](https://cyipt.github.io/popupCycleways/leicester.html) for the interactive version of the map above.
 
@@
 
-See [sheffield.html](https://cyipt.github.io/tempCycleways/sheffield.html) for the interactive version of the map above.
+See [sheffield.html](https://cyipt.github.io/popupCycleways/sheffield.html) for the interactive version of the map above.
 
@@
 
-See [newcastle.html](https://cyipt.github.io/tempCycleways/newcastle.html) for the interactive version of the map above.
+See [newcastle.html](https://cyipt.github.io/popupCycleways/newcastle.html) for the interactive version of the map above.
 
@@
 
-See [cambridge.html](https://cyipt.github.io/tempCycleways/cambridge.html) for the interactive version of the map above.
+See [cambridge.html](https://cyipt.github.io/popupCycleways/cambridge.html) for the interactive version of the map above.
 
@@
 # deploy
-git clone git@github.com:cyipt/tempCycleways
-cd tempCycleways
+git clone git@github.com:cyipt/popupCycleways
+cd popupCycleways
 git checkout gh-pages
@@
 echo '*.html' >> .gitignore
-echo 'tempCycleways/' >> .gitignore
+echo 'popupCycleways/' >> .gitignore
 pandoc -o README.html ../README.md

### 2020-05-13 cities/leeds.Rmd
Updated results - leeds + london

@@
 ```{r leeds}
-leeds_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
-leeds_buff = stplanr::geo_buffer(leeds_centre, dist = 4000)
-roads = rsf %>% filter(region == "Leeds", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-roads = st_intersection(roads, region_buff)
-
-roads_top = roads[leeds_buff, ] %>%
-  filter(length > 300) %>%
-  filter(name != "Broadway") %>%
-  filter(name != "Wellington Street") %>% 
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = c(-1.549, 53.805, 13)) + 
-  tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
+setwd("..")
+region_name = "Leeds"
+source("code/build.R")
+m
+knitr::kable(res_table, caption = "The top 10 candidate roads for space reallocation for pop-up active transport infrastructure according to methods developed for the project.", digits = 0)
 ```

### 2020-05-13 code/article-methods-1.R

@@ -16,12 +16,12 @@ library(sf)
 library(tidyverse)
 library(tmap)
 if(!file.exists("rsf_leeds.Rds")) {
-  u = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rsf_leeds.Rds"
+  u = "https://github.com/cyipt/popupCycleways/releases/download/0.1/rsf_leeds.Rds"
   download.file(u, "rsf_leeds.Rds")
 }
 r_original = readRDS("rsf_leeds.Rds")
 if(!file.exists("rtid.csv")) {
-  ur = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rtid.csv"
+  ur = "https://github.com/cyipt/popupCycleways/releases/download/0.1/rtid.csv"
   download.file(ur, "rtid.csv")
 }
 rtid = readr::read_csv("rtid.csv")

### 2020-05-13 code/article-methods-cambridge-test.R

@@ -29,13 +29,14 @@ city_centre_buffer_radius = 8000
 city_centre_buffer_radius_large = 10000
 key_destination_buffer_radius = 5000
 
-parameters_df = data.frame(
+parameters_df = data.frame(stringsAsFactors = FALSE,
   name = region_name,
   min_cycling_potential = 0,
-  min_grouped_cycling_potential = 500,
+  min_grouped_cycling_potential = 100,
   min_grouped_length = 500,
   city_centre_buffer_radius = 8000,
-  key_destination_buffer_radius = 5000
+  key_destination_buffer_radius = 5000,
+  regexclude = "roundabout"
 )
 
 parameters_df = parameters_df[rep(1, length(region_names)), ]

### 2020-05-13 code/exploratory-analysis-leeds.R

@@ -17,14 +17,14 @@ tmap_mode("view")
 # piggyback::pb_upload("rsf_leeds.Rds")
 # piggyback::pb_download_url("rsf_leeds.Rds")
 
-u = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rsf_leeds.Rds"
+u = "https://github.com/cyipt/popupCycleways/releases/download/0.1/rsf_leeds.Rds"
 r_original = readRDS(url(u))
 nrow(r_original)
 plot(r_original$geometry)
 head(r_original)
 names(r_original)
 
-ur = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rtid.csv"
+ur = "https://github.com/cyipt/popupCycleways/releases/download/0.1/rtid.csv"
 rtid = readr::read_csv(ur)
 
 # parameters --------------------------------------------------------------

### 2020-05-13 README.Rmd
Update results, close #19

@@
 ```{r, eval=FALSE}
+params = list(
+  message = FALSE,
+  warning = FALSE,
+  echo = FALSE,
+  collapse = TRUE,
+  comment = "#>"
+  )
 
-rmarkdown::render("london.Rmd") # check
+old_wd = setwd("cities")
+rmarkdown::render("leeds.Rmd") # check
 rmarkdown_files = paste0(tolower(las_top$name), ".Rmd")
-lapply(rmarkdown_files, rmarkdown::render)
+lapply(rmarkdown_files, rmarkdown::render, params = params)
+setwd(old_wd)
 webshot::install_phantomjs()
@@
 git checkout gh-pages
-git rm -rf .
-git commit --allow-empty -m 'Initial gh-pages commit'
+# git rm -rf .
+# git commit --allow-empty -m 'Initial gh-pages commit'
 # git push origin gh-pages
@@
 # cp -Rv ../README* . # it's 100 MB!
-cp -Rv ../*.html . # it's 100 MB!
+git status
+cp -Rv ../cities/*.html . # it's 100 MB!
 git status
 ls -hal
-mv README.html index.html
+pandoc -o index.html ../README.md
 firefox index.html
 git status
-git commit -am 'Update interactive maps'
+git commit -am 'Update results'
 git push origin gh-pages
 cd ..
-echo '*.html' >> .gitignore
-echo 'popupCycleways/' >> .gitignore
-pandoc -o README.html ../README.md
 ```

### 2020-05-13 cities/bristol.Rmd
Update results, close #19

@@
 
-```{r bristol}
-region = las_top %>% filter(name == "Bristol")
-region = sf::st_make_valid(region)
-roads = rsf %>% filter(region == "Bristol", pctgov > 100)
-roads = st_intersection(roads, region)
-
-bristol_centre = tmaptools::geocode_OSM("Bristol", as.sf = TRUE)
-bristol_buff = stplanr::geo_buffer(bristol_centre, dist = 1500)
-roads_top = roads[bristol_buff, ] %>%
-  filter(length > 300) %>%
-  filter(name != "Temple Way Underpass") %>%
-  filter(!name == "Newfoundland Street", !name == "") %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = c(-2.5879, 51.454, 13)) + 
-  tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
+```{r Bristol}
+setwd("..")
+region_name = "Bristol"
+source("code/build.R")
+m
+knitr::kable(res_table, caption = "The top 10 candidate roads for space reallocation for pop-up active transport infrastructure according to methods developed for the project.", digits = 0)
 ```

### 2020-05-13 cities/leicester.Rmd
Update results, close #19

@@
 
-```{r leicester}
-roads = rsf %>% filter(region == "Leicester", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Leicester", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-roads = st_intersection(roads, region_buff)
-
-roads_top = roads %>%
-  filter(length > 200) %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
+```{r Leicester}
+setwd("..")
+region_name = "Leicester"
+source("code/build.R")
+m
+knitr::kable(res_table, caption = "The top 10 candidate roads for space reallocation for pop-up active transport infrastructure according to methods developed for the project.", digits = 0)
 ```

### 2020-05-13 cities/liverpool.Rmd
Update results, close #19

@@
 
-```{r liverpool}
-roads = rsf %>% filter(region == "Liverpool", pctgov > 100)
-region = las_top %>% filter(name == "Liverpool")
-roads = st_intersection(roads, region)
-# region_centre = tmaptools::geocode_OSM("Liverpool", as.sf = TRUE)
-# region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
-# tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
-#   tm_view(set.view = 10) + tm_scale_bar() +
-#   tm_shape(region) + tm_borders()
-
-roads_top = roads %>%
-  filter(length > 200) %>%
-  filter(name != "Queensway (Mersey Tunnel)") %>%
-  filter(name != "Queensway") %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = c(-2.95, 53.4084, 13)) + 
-  tm_scale_bar() +
-  tm_basemap(server = s)
-
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
+```{r Liverpool}
+setwd("..")
+region_name = "Liverpool"
+source("code/build.R")
+m
+knitr::kable(res_table, caption = "The top 10 candidate roads for space reallocation for pop-up active transport infrastructure according to methods developed for the project.", digits = 0)
 ```

### 2020-05-13 cities/manchester.Rmd
Update results, close #19

@@
 
-```{r manchester}
-roads = rsf %>% filter(region == "Manchester", pctgov > 100)
-region_centre = tmaptools::geocode_OSM("Manchester", as.sf = TRUE)
-region_buff = stplanr::geo_buffer(region_centre, dist = 7000)
-roads = st_intersection(roads, region_buff)
-
-roads_top = roads %>%
-  filter(length > 300) %>%
-  arrange(desc(pctgov)) %>%
-  slice(1:10)
-tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
-  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
-  tm_text("name") +
-  tm_view(set.view = 12) + tm_scale_bar() +
-  tm_basemap(server = s)
-roads_top %>%
-  st_drop_geometry() %>%
-  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
-  knitr::kable(digits = 0)
+```{r Manchester}
+setwd("..")
+region_name = "Manchester"
+source("code/build.R")
+m
+knitr::kable(res_table, caption = "The top 10 candidate roads for space reallocation for pop-up active transport infrastructure according to methods developed for the project.", digits = 0)
 ```

### 2020-05-13 code/build.R

@@ -1,0 +1,190 @@
+# aim: output results highlighting and ranking roads that have 'spare lanes'
+
+# global parameters ----------------------------------------------------------
+library(sf)
+library(tidyverse)
+library(tmap)
+if(!exists("parameters")) {
+message("Loading global parameters")  
+s = c(
+  "Esri.WorldGrayCanvas",
+  "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
+  "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
+)
+parameters = read_csv("input-data/parameters.csv")
+
+# read-in national data ---------------------------------------------------
+r_all = readRDS("rsf.Rds")
+region_names = unique(r_all$region)
+rtid = readr::read_csv("rtid.csv")
+hsf = readRDS("hsf.Rds")
+regions = readRDS("~/cyipt/cyipt-bigdata/boundaries/local_authority/local_authority.Rds")
+# r_original = r_original[st_transform(region, st_crs(r_original)), ]
+}
+
+# local parameters --------------------------------------------------------
+# i = 1
+if(!exists("region_name"))
+  region_name = "Leeds"
+r_original = r_all %>% filter(region == region_name) %>% 
+  select(name, ref, highway, maxspeed, pctgov, width, Existing, length, rtid, idGlobal) 
+i = which(parameters$name %in% region_name)
+p = parameters[i, ]
+list2env(p, envir = .GlobalEnv)
+
+
+# buffers -----------------------------------------------------------------
+# for(i in seq(length(region_names))) {
+region = regions %>% filter(lad16nmw == region_name)
+city_centre = tmaptools::geocode_OSM(region_name, as.sf = TRUE)
+city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
+h_leeds = hsf[city_centre_buffer, ]
+h_leeds_buffer = stplanr::geo_buffer(h_leeds, dist = key_destination_buffer_radius)
+city_key_buffer = st_union(city_centre_buffer, st_union(h_leeds_buffer))
+
+## ----preprocess------------------------------------------------------------------------------
+# remove motorways
[…]

### 2020-05-13 README.Rmd
Update readme, include new liverpool map

@@
 
-in the context of increased demand for cycling to keyworker workplaces.
+in the context of increased demand for cycling to key worker workplaces.
+
+See this [pre-print paper](https://osf.io/7wjb6) for a generalised summary of the methods. This document describes the work in the context of UK planning priorities.
 

### 2020-05-13 README.Rmd
Update build code

@@
 ```{r, eval=FALSE}
-params = list(
-  message = FALSE,
-  warning = FALSE,
-  echo = FALSE,
-  collapse = TRUE,
-  comment = "#>"
-  )
+old_wd = setwd("cities")
+rmarkdown::render("sheffield.Rmd") # check
+rmarkdown_files = paste0(tolower(las_top$name), ".Rmd")
+for(i in rmarkdown_files) {
+  rmarkdown::render(i)
+}
+# lapply(rmarkdown_files, rmarkdown::render)
+# parallel::mclapply(rmarkdown_files, rmarkdown::render, mc.cores = 5)
+getwd()
+browseURL("liverpool.html")
+setwd(old_wd)
 
-old_wd = setwd("cities")
-rmarkdown::render("leeds.Rmd") # check
-rmarkdown_files = paste0(tolower(las_top$name), ".Rmd")
-lapply(rmarkdown_files, rmarkdown::render, params = params)
-setwd(old_wd)
 webshot::install_phantomjs()
@@
 ls -hal
-pandoc -o index.html ../README.md
-firefox index.html
+# firefox index.html
 git status

### 2020-05-13 cities/bristol.Rmd
Update build code

@@
 region_name = "Bristol"
-source("code/build.R")
+source("code/build.R", local = TRUE)
 m

### 2020-05-13 cities/leeds.Rmd
Update build code

@@
 region_name = "Leeds"
-source("code/build.R")
+source("code/build.R", local = TRUE)
 m

### 2020-05-13 cities/leicester.Rmd
Update build code

@@
 region_name = "Leicester"
-source("code/build.R")
+source("code/build.R", local = TRUE)
 m

### 2020-05-13 cities/liverpool.Rmd
Update build code

@@
 region_name = "Liverpool"
-source("code/build.R")
+source("code/build.R", local = TRUE)
 m

### 2020-05-13 cities/manchester.Rmd
Update build code

@@
 region_name = "Manchester"
-source("code/build.R")
+source("code/build.R", local = TRUE)
 m

### 2020-05-13 code/build.R

@@ -24,8 +24,8 @@ regions = readRDS("~/cyipt/cyipt-bigdata/boundaries/local_authority/local_author
 
 # local parameters --------------------------------------------------------
 # i = 1
-if(!exists("region_name"))
-  region_name = "Leeds"
+# if(!exists("region_name"))
+#   region_name = "Leeds"
 r_original = r_all %>% filter(region == region_name) %>% 
   select(name, ref, highway, maxspeed, pctgov, width, Existing, length, rtid, idGlobal) 
 i = which(parameters$name %in% region_name)
@@ -38,9 +38,9 @@ list2env(p, envir = .GlobalEnv)
 region = regions %>% filter(lad16nmw == region_name)
 city_centre = tmaptools::geocode_OSM(region_name, as.sf = TRUE)
 city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
-h_leeds = hsf[city_centre_buffer, ]
-h_leeds_buffer = stplanr::geo_buffer(h_leeds, dist = key_destination_buffer_radius)
-city_key_buffer = st_union(city_centre_buffer, st_union(h_leeds_buffer))
+h_city = hsf[city_centre_buffer, ]
+h_city_buffer = stplanr::geo_buffer(h_city, dist = key_destination_buffer_radius)
+city_key_buffer = st_union(city_centre_buffer, st_union(h_city_buffer))
 
 ## ----preprocess------------------------------------------------------------------------------
 # remove motorways
@@ -87,7 +87,7 @@ t1 = rj %>%
   # select(name, highway_type, maxspeed, cycling_potential, width) %>%
   table1::table1(~ highway_type + cycling_potential + width + n_lanes | maxspeed, data = ., )
 
-## ----hospitals, fig.cap="Overview map of input data, showing the main highway types and location of hospitals in Leeds"----
+## ----hospitals, fig.cap="Overview map of input data, showing the main highway types and location of hospitals in city"----
 m1 = r_main_region %>%
   sample_n(1000) %>% 
   mutate(`Highway type` = highway_type) %>% 
@@ -96,7 +96,7 @@ m1 = r_main_region %>%
   tm_lines(col = "Highway type", palette = c("green", "black", "blue", "grey"),
[…]

### 2020-05-14 README.Rmd
Add uas boundary code

@@
 # general analysis of data
+# piggyback::pb_upload("rsf.Rds", "cyipt/cyipt-phase-1-data")
 # piggyback::pb_download("rsf.Rds", "cyipt/cyipt-phase-1-data")

### 2020-05-14 code/boundaries.R

@@ -1,0 +1,11 @@
+library(tidyverse)
+
+counties = ukboundaries::duraz("https://borders.ukdataservice.ac.uk/ukborders/easy_download/prebuilt/kml/England_ct_2001_clipped.zip")
+counties = sf::read_sf("england_ct_2001_clipped.kml")
+mapview::mapview(counties)
+uas_uk = sf::read_sf("https://opendata.arcgis.com/datasets/6638c31a8e9842f98a037748f72258ed_3.kml?outSR=%7B%22latestWkid%22%3A27700%2C%22wkid%22%3A27700%7D")
+uas_uk
+head(uas_uk$ctyua17cd)
+uas_en = uas_uk %>% filter(str_detect(string = ctyua17cd, pattern = "E"))
+nrow(uas_en)
+mapview::mapview(uas_en)

### 2020-05-14 code/boundaries.R
Add analysis of bham for #22

@@ -4,8 +4,13 @@ counties = ukboundaries::duraz("https://borders.ukdataservice.ac.uk/ukborders/ea
 counties = sf::read_sf("england_ct_2001_clipped.kml")
 mapview::mapview(counties)
 uas_uk = sf::read_sf("https://opendata.arcgis.com/datasets/6638c31a8e9842f98a037748f72258ed_3.kml?outSR=%7B%22latestWkid%22%3A27700%2C%22wkid%22%3A27700%7D")
+uas_uk = uas_uk %>% select(matches("cty"))
+plot(uas_uk)
+sf::write_sf(uas_uk, "~/hd/data/uk/uas_uk.gpkg")
 uas_uk
 head(uas_uk$ctyua17cd)
 uas_en = uas_uk %>% filter(str_detect(string = ctyua17cd, pattern = "E"))
 nrow(uas_en)
+saveRDS(uas_en, "uas_en.Rds")
+piggyback::pb_upload("uas_en.Rds")
 mapview::mapview(uas_en)

### 2020-05-14 code/build.R

@@ -19,6 +19,8 @@ region_names = unique(r_all$region)
 rtid = readr::read_csv("rtid.csv")
 hsf = readRDS("hsf.Rds")
 regions = readRDS("~/cyipt/cyipt-bigdata/boundaries/local_authority/local_authority.Rds")
+nrow(regions)
+mapview::mapview(regions)
 # r_original = r_original[st_transform(region, st_crs(r_original)), ]
 }
 

### 2020-05-14 code/explore-existing-popup-integration-birmingham.R

@@ -1,0 +1,23 @@
+# aim: explore existing infrastructure in relation to pop-up analysis
+library(sf)
+library(tidyverse)
+library(tmap)
+
+country_name = "England"
+region_name = "Birmingham"
+
+library(geofabrik)
+gfz = geofabrik::geofabrik_zones
+e = gfz %>% 
+  filter(name == country_name)
+f = geofabrik::gf_filename(name = country_name)
+download.file(url = e$pbf_url, destfile = f)
+cycleways_en = geofabrik::read_pbf(dsn = f, key = "highway", value = "cycleway")
+plot(sf::st_geometry(cycleways_en))
+
+uas_en = readRDS("uas_en.Rds")
+
+region = uas_en %>% filter(ctyua17nm == region_name)
+mapview::mapview(region)
+cycleways_region = cycleways_en[region, ]
+source("code/build.R")

### 2020-05-14 code/explore-existing-popup-integration-birmingham.R
Update cycleway overlay script

@@ -20,4 +20,28 @@ uas_en = readRDS("uas_en.Rds")
 region = uas_en %>% filter(ctyua17nm == region_name)
 mapview::mapview(region)
 cycleways_region = cycleways_en[region, ]
+# cycleways_region_long = cycleways_region %>% 
+#   mutate(length = round(as.numeric(st_length(.)))) %>% 
+#   filter(length > 50)
+# summary(cycleways_region_long$length)
 source("code/build.R")
+m = tm_shape(r_pct_top) + tm_lines(col = "blue", lwd = 6, alpha = 0.6) +
+  tm_shape(r_pct_top_n) + tm_text("name") +
+  tm_shape(cycleways_region) + tm_lines() +
+  tm_basemap(server = s) +
+  tm_scale_bar()
+m
+cycleway_buffer = stplanr::geo_buffer(cycleways_region, dist = 50) %>% sf::st_union()
+r_pct_top_no_cycleways = sf::st_difference(r_pct_top, cycleway_buffer)
+# r_pct_top_no_cycleways = r_pct_top[cycleway_buffer, , op = sf::st_disjoint]
+r_pct_top_no_cycleways = r_pct_top_no_cycleways %>% 
+  st_cast("LINESTRING") %>% 
+  mutate(group_length = round(as.numeric(st_length(.)))) %>% 
+  filter(group_length > 50)
+
+m = tm_shape(r_pct_top_no_cycleways) + tm_lines(col = "blue", lwd = 6, alpha = 0.6) +
+  tm_shape(r_pct_top_n) + tm_text("name") +
+  tm_shape(cycleways_region_long) + tm_lines() +
+  tm_basemap(server = s) +
+  tm_scale_bar()
+m

### 2020-05-18 code/mcr-safe-streets-analysis.Rmd
Add mcr analysis doc

---
title: "Analysis of cycling potential around key corridors in Greater Manchester"
author: "Robin Lovelace"
date: "18/05/2020"
# output: bookdown::html_document2
output: bookdown::word_document2
---

```{r, echo=FALSE, eval=FALSE}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(tmap)
z = pct::get_pct_zones("greater-manchester")
lads = z %>% 
  sf::st_buffer(0.0001) %>% 
  group_by(lad_name) %>% 
  summarise(all = sum(all))
# plot(lads)
l = pct::get_pct_lines("greater-manchester")
sum(z$bicycle) / sum(z$all)
sum(z$dutch_slc) / sum(z$all)
z$walk_2011 = z$foot / z$all * 100
z$cycle_2011 = z$bicycle / z$all * 100
z$pt_2011 = (z$bus + z$train_tube) / z$all * 100
z$cycle_dutch = z$dutch_slc / z$all * 100
```

```{r}
# od analysis
od = pct::get_od()
od$pt = od$train + od$light_rail + od$bus
od_mcr = od %>% filter(geo_code1 %in% z$geo_code)
# sum(od$pt) / sum(od$all)
# 0.159002
# sum(od_mcr$pt) / sum(od_mcr$all)
# 0.1434537
```

```{r}
z$active_2011 = (z$foot + z$bicycle) / z$all * 100
z$active_dutch = (z$foot + z$dutch_slc) / z$all * 100
b = 0:8 * 10
```

# Context

## Public transport

Greater Manchester has high cycling potential and high reliance on public transport (PT) .
Across the region overall, 14.3% of commutes were made by public transport in 2011, although the reliance on public transport varies substantially across the region.
Around a third of trips are made by PT in many parts of Manchester and Salford, as shown in Figure \@ref(fig:pt).
These areas areas that are particularly vulnerable to reduced capacity by public transport also tend to be within cycling distance from the main employment centre in Manchester's city centre.

```{r pt, fig.cap="The percentage of trips made by public transport. Source: 2011 Census"}
tm_shape(z) + tm_fill("pt_2011", palette = "Blues", title = "% PT (2011)") +
  tm_shape(lads) + tm_borders(col = "black") + tm_text("lad_name") +
  tm_scale_bar()
```

## Active transport

The geographic distribution of active transport (walking and cycling) for commuting is shown in Figure \@ref(fig:active).
13% of single mode commutes were made by active modes in 2011.
This has the potential to more than double to 33% for the region overall under the Go Dutch scenario, represented in Figure \@ref(fig:activedutch).

```{r active, fig.cap="Percentage of commutes in Greater Manchester by active transport. Source: 2011 Census"}
tm_shape(z) + tm_fill("active_2011", palette = "Greens", title = "% Active (2011)", breaks = b) +
  tm_shape(lads) + tm_borders(col = "black") + tm_text("lad_name") +
  tm_scale_bar()
# (sum(z$foot) + sum(z$bicycle)) / sum(z$all)
# (sum(z$foot) + sum(z$dutch_slc)) / sum(z$all)
```

The majority of potential growth in active transport comes from cycling due to the increased range of commutes by this mode, around 5 km for most people and further for many, due to high fitness and use of e-bikes.

```{r activedutch, fig.cap="Percentage of commutes in Greater Manchester by active transport under the Go Dutch scenario. Source: Propensity to Cycle Tool"}
tm_shape(z) + tm_fill("active_dutch", palette = "Greens", title = "% Active (Go Dutch)", breaks = b) +
  tm_shape(lads) + tm_borders(col = "black") + tm_text("lad_name") +
  tm_scale_bar()
```

Like surrounding regions, only 2% of single mode commutes in were made by bike according to data from the 2011 Census (this figure will be higher, perhaps doubling since the introduction of pro-cycling policies since 2011 and the uptake of cycling in response to PT capacity restrictions).

Under the Go Dutch scenario, this rises to 22%, compared with 20% in Lancashire and Cheshire, with a geographic distribution shown in Figure \@ref(fig:cycling).
The region has high cycling potential because of the high population density in the region and relatively short average trip distances.

```{r cycling, fig.cap="Mode share of cycling from the 2011 Census and under the Go Dutch scenario in the Propensity to Cycle Tool", out.width="50%", fig.show='hold'}
knitr::include_graphics(c(
  "https://user-images.githubusercontent.com/1825120/82181124-d65d8600-98d9-11ea-8835-98647333ffe2.png",
  "https://user-images.githubusercontent.com/1825120/82181210-fee58000-98d9-11ea-8d9c-cc716556f769.png"
))
```

Data on walking, cycling and PT levels from the 2011 Census for MSOA zones in the region are summarised by local authority in the table below.

```{r}
table1::table1(~ walk_2011 + cycle_2011 + pt_2011 | lad_name, data = z, transpose = TRUE)
```

# Coridor analysis

# Cycling potential on key corridors

### 2020-05-18 code/update-rnet.R
update rnet

@@ -3,26 +3,88 @@ library(stplanr)
 library(tidyverse)
 
 # readRDS(url("https://github.com/npct/pct-outputs-national/blob/master/commute/lsoa/rnet_all.Rds"))
-rnet = pct::get_pct_rnet # LSOA route network
-mapview(rnet["govtarget_slc"])
+rnet = pct::get_pct_rnet(region = "avon") # LSOA route network
+# mapview(rnet["govtarget_slc"])
 
 r_cyipt = readRDS("rsf.Rds")
 r_cyipt = r_cyipt[r_cyipt$region == "Bristol",]
 
-st_crs(rnet)
-st_crs(r_cyipt)
+# st_crs(rnet)
+# st_crs(r_cyipt)
 
 dim(rnet) # 13829 this only has streets which are part of the LSOA route network
 dim(r_cyipt) # 104805 this has every street
 
 rnet_buff = geo_buffer(shp = rnet, dist = 10)
-mapview(rnet_buff["govtarget_slc"])
+# mapview(rnet_buff["govtarget_slc"])
 
-rnet_buff2 = geo_buffer(rnet, dist = 2)
-mapview(rnet_buff2["govtarget_slc"])
+r_cyipt_joined = st_join(r_cyipt, rnet_buff, join = st_within)
 
-rnet_joined = st_join(rnet_buff, r_cyipt, join = st_within)
+dim(r_cyipt_joined) # 111865 
+nrow(r_cyipt_joined)/nrow(r_cyipt)
 
-dim(rnet_joined) # 13829 same as rnet
+# test_all_present = r_cyipt_joined[which(duplicated(r_cyipt_joined$idGlobal) == TRUE) ,]
+# dim(r_cyipt_joined) - dim(test_all_present) # this must equal dim(r_cyipt) if all segments can be joined
+
+dupes = r_cyipt_joined[which(duplicated(r_cyipt_joined$idGlobal) == TRUE | duplicated(r_cyipt_joined$idGlobal, fromLast = TRUE) == TRUE),]
+# dim(dupes) # 10593
+
+# mapview(dupes) + mapview(rnet_buff)
+
+dupes_max = dupes %>% 
+  group_by(idGlobal) %>%
+  summarise(cycling_potential_max = max(govtarget_slc)) %>%
+  st_drop_geometry()
+
+dim(dupes_max) # 3533
+
+##Need to join dupes_max with the rnet. Then join the results of that with the r_cyipt/r_cyipt_joined
+
+
[…]

### 2020-05-18 code/mcr-safe-streets-analysis.Rmd
Update mcr analysis

@@
 output: bookdown::word_document2
+# output: bookdown::pdf_document2
 ---
@@
 ```{r setup, include=FALSE}
+knitr::opts_knit$set(root.dir = "..")
 knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE)
+region_name = "Manchester"
 library(tidyverse)
@@
 z$active_dutch = (z$foot + z$dutch_slc) / z$all * 100
+z$active_ebike = (z$foot + z$ebike_slc) / z$all * 100
 b = 0:8 * 10
@@
 13% of single mode commutes were made by active modes in 2011.
-This has the potential to more than double to 33% for the region overall under the Go Dutch scenario, represented in Figure \@ref(fig:activedutch).
+This has the potential to more than double to 33% for the region overall under the Ebikes scenario, represented in Figure \@ref(fig:activedutch).
 
@@
 
-```{r activedutch, fig.cap="Percentage of commutes in Greater Manchester by active transport under the Go Dutch scenario. Source: Propensity to Cycle Tool"}
-tm_shape(z) + tm_fill("active_dutch", palette = "Greens", title = "% Active (Go Dutch)", breaks = b) +
+```{r activedutch, fig.cap="Percentage of commutes in Greater Manchester by active transport under the Ebikes scenario. Source: Propensity to Cycle Tool"}
+tm_shape(z) + tm_fill("active_dutch", palette = "Greens", title = "% Active (Ebikes)", breaks = b) +
   tm_shape(lads) + tm_borders(col = "black") + tm_text("lad_name") +
@@
 
-Under the Go Dutch scenario, this rises to 22%, compared with 20% in Lancashire and Cheshire, with a geographic distribution shown in Figure \@ref(fig:cycling).
+Under the Ebikes scenario, this rises to 22%, compared with 20% in Lancashire and Cheshire, with a geographic distribution shown in Figure \@ref(fig:cycling).
 The region has high cycling potential because of the high population density in the region and relatively short average trip distances.
 
-```{r cycling, fig.cap="Mode share of cycling from the 2011 Census and under the Go Dutch scenario in the Propensity to Cycle Tool", out.width="50%", fig.show='hold'}
+```{r cycling, fig.cap="Mode share of cycling from the 2011 Census and under the Ebikes scenario in the Propensity to Cycle Tool", out.width="50%", fig.show='hold'}
 knitr::include_graphics(c(
   "https://user-images.githubusercontent.com/1825120/82181124-d65d8600-98d9-11ea-8835-98647333ffe2.png",
-  "https://user-images.githubusercontent.com/1825120/82181210-fee58000-98d9-11ea-8d9c-cc716556f769.png"
+  "https://user-images.githubusercontent.com/1825120/82255290-4e649400-994c-11ea-98c4-267fe936ccdd.png"
 ))
@@
 ```{r}
-table1::table1(~ walk_2011 + cycle_2011 + pt_2011 | lad_name, data = z, transpose = TRUE)
+# table1::table1(~ walk_2011 + cycle_2011 + pt_2011 | lad_name, data = z, transpose = TRUE)
+# knitr::include_graphics("mcr-census-results.png")
+knitr::include_graphics("https://user-images.githubusercontent.com/1825120/82254618-10b33b80-994b-11ea-85ef-5edee3a53ac5.png")
 ```
 
-# Coridor analysis
+# Corridor analysis
+
+We constrained the analysis to 10 key corridors.
+These provide access to the main employment hub from the North, East, South and West, as illustrated in Figure \@ref(fig:corridors).
+
+```{r, fig.cap="Key corridors analysed for the project."}
+# Airport
+# Altrincham
+# Leigh
+# Oldham
+
+city_centre = tmaptools::geocode_OSM(region_name, as.sf = TRUE)
+city_centre_buffer = stplanr::geo_buffer(city_centre, dist = 2000)
+city_centre_buffer8 = stplanr::geo_buffer(city_centre, dist = 8000)
+city_centre_buffer_square = stplanr::geo_bb(scale_factor = 1.1, city_centre_buffer8)
+r_pct_lanes_all = readRDS("mcr_r_pct_lanes_all.Rds")
+# mapview::mapview(r_pct_lanes_all)
+corridor_names = c(
+  "Chester Road", # (Trafford)
+  "A6",
+  "A580", #A6/A580 – East Lancs to Chapel Street (Salford)
+  "A34", # – Anson road/Birchfield Road (Stockport)
+  "A57", # Hyde Road (Tameside)
+  "A62", # A62 – Oldham Road (Oldham)
+  "A56", # A56 – Bury New Road (Bury)
+  "Chorlton Road", # (all Manchester)
+  "A5103" # Princess road (all Manchester)
+)
+princess_road = r_pct_lanes_all %>% filter(name == "Princess Road")
+# table(princess_road$ref) # A5103
+bury = r_pct_lanes_all %>% filter(name == "Bury New Road")
+# mapview::mapview(princess_road)
+
+# load all roads using code in "code/get-data-db.R":
+rsf_mcr = readRDS("mcr_rsf.Rds")
+regexclude_corridors = "A537|A5149|A5102|A5034|A556|5014|5067"
+
+key_corridors_mcr = rsf_mcr %>% 
+  filter(name %in% corridor_names | ref %in% corridor_names) %>% 
+  filter(str_detect(string = ref, pattern = "A")) %>% 
+  filter(!str_detect(string = ref, pattern = regexclude_corridors)) %>% 
+  filter(!is.na(width)) %>% 
+  mutate(cycling_potential = as.numeric(pctebike)) %>% 
+  select(name, ref, cycling_potential, maxspeed, aadt, width, length, idGlobal)
+princess_road = key_corridors_mcr %>% filter(name == "Princess Road")
+
+# Split up A6 and A56
+# table(key_corridors_mcr$name[key_corridors_mcr$ref == "A6"])
+zbs = zonebuilder::zb_zone(x = city_centre, n_circles = 1, n_segments = 4, distance = 50, segment_center = TRUE)
+# plot(zbs$geometry)
+# plot(princess_road, add = T)
+# plot(key_corridors_mcr %>% filter(ref == "A6"), add = T)
+# plot(key_corridors_mcr %>% filter(ref == "A56"), add = T)
+zbs_north = zbs %>% filter(segment_id == 1)
+zbs_east = zbs %>% filter(segment_id == 4)
+# plot(zbs_east)
+a56_north = key_corridors_mcr %>% filter(ref == "A56") %>% .[zbs_north, ]
+a6_east = key_corridors_mcr %>% filter(ref == "A6") %>% .[zbs_east, ]
+a57_east = key_corridors_mcr %>% filter(ref == "A57") %>% .[zbs_east, ]
+
+key_corridors_mcr$ref[key_corridors_mcr$ref == "A56"] = "A56 (South)"
+key_corridors_mcr$ref[key_corridors_mcr$idGlobal %in% a56_north$idGlobal] = "A56 (North)"
+
+key_corridors_mcr$ref[key_corridors_mcr$ref == "A6"] = "A6 (West)"
+key_corridors_mcr$ref[key_corridors_mcr$idGlobal %in% a6_east$idGlobal] = "A6 (East)"#
+
+key_corridors_mcr$ref[key_corridors_mcr$ref == "A57"] = "A57 (West)"
+key_corridors_mcr$ref[key_corridors_mcr$idGlobal %in% a57_east$idGlobal] = "A57 (East)"
+
+key_corridors_mcr_grouped = key_corridors_mcr %>% 
+  # group_by(name, ref) %>%
+  group_by(ref) %>%
+  summarise(
+    name = first(name),
+    cycling_potential = round(mean(cycling_potential)),
+    width = round(mean(width)),
+    length = round(sum(length))
+    ) %>% 
+  ungroup()
+
+m = tm_shape(key_corridors_mcr_grouped) +
+  tm_lines(lwd = "width", scale = 9, alpha = 0.8, col = "ref") +
+  tm_shape(lads) + tm_borders() +
+  tm_scale_bar()
+# m
+```
+
+```{r corridors, fig.cap="Key corridors in Greater Manchester by road reference and direction relative to Manchester city centre."}
+knitr::include_graphics("https://user-images.githubusercontent.com/1825120/82258452-caada600-9951-11ea-8af0-520c8407928a.png")
+```
+
+To further focus the analysis, only the sections within 8 km of Manchester city centre were included.
+The results of clipping key corridors to an 8 km buffer surrounding the city centre are shown in the Figure \@ref(fig:buff8).
+
+```{r buff8, fig.cap="Illustration of the key corridors within an 8 km buffer of Manchester city centre, with the width of road line segments proportional to estimated carriageway width.
+Source: www.cyipt.bike."}
+key_corridors8 = key_corridors_mcr[city_centre_buffer8, ]
+lads8 = sf::st_intersection(lads, city_centre_buffer_square)
+
+m = tm_shape(key_corridors8) +
+  tm_lines(lwd = "width", scale = 9, alpha = 0.8, col = "ref") +
+  tm_shape(lads8) + tm_borders() + tm_text("lad_name") +
+  tm_shape(city_centre_buffer8) + tm_borders() +
+  tm_scale_bar() +
+  tm_layout(legend.outside = TRUE)
+m
+```
 
 # Cycling potential on key corridors
+
+The cycling potential under the E-bike scenario was assessed as follows:
+
+```{r}
+key_corridors_mcr_grouped2 = key_corridors8 %>% 
+  # group_by(name, ref) %>%
+  group_by(ref) %>%
+  summarise(
+    name = first(name),
+    cycling_potential = round(mean(cycling_potential)),
+    width = round(mean(width)),
+    length = round(sum(length))
+    ) %>% 
+  ungroup()
+
+key_corridors_mcr_grouped2 %>% 
+  sf::st_drop_geometry() %>% 
+  group_by(ref) %>% 
+  summarise(
+    `Length (m)` = round(sum(length)),
+    `Cycling potential` = mean(cycling_potential),
+    `Width (m)` = mean(width),
+    `Km/day` = round(`Length (m)` * `Cycling potential` / 1000) * 2
+    ) %>% 
+  arrange(desc(`Km/day`)) %>% 
+  filter(`Length (m)` > 2000) %>% 
+  knitr::kable(caption = "Average cycling potential, width and potential distance cycled per day for commuting (to and from work on a typical work day) for each corridor under the Ebikes scenario.")
+```
+
+Many of the key corridors have spare lanes during times of low traffic, as shown in Figure \@ref(fig:nlanes), which also shows the variability of cycling potential in segments along the different corridors.
+
+```{r nlanes, fig.cap="Illustration of spare lanes on key corridors"}
+# find the r_pct_lanes in the key roads
+key_corridors_buffer = stplanr::geo_buffer(key_corridors8, dist = 20)
+r_pct_lanes_in_corridors = r_pct_lanes_all[key_corridors_buffer, , op = sf::st_within]
+m = tm_shape(key_corridors8) +
+  tm_lines(lwd = "cycling_potential", scale = 9, alpha = 0.8, col = "ref") +
+  tm_shape(r_pct_lanes_in_corridors) + tm_lines() +
+  tm_scale_bar()
+m
+```
+
+```{r}
+l = readRDS("mcr-l-pct.Rds")
+r = readRDS("mcr-r-pct.Rds")
+
+# potential analysis ------------------------------------------------------
+
+r$pt_mode_shift = r$ebike_sip * -1
+# summary(r$pt_mode_shift)
+# nrow(r) # 16363
+# rnet_pt = stplanr::overline(r, "pt_mode_shift")
+# saveRDS(rnet_pt, "mcr-rnet_pt.Rds")
+# mapview::mapview(rnet_pt)
+rnet_pt = readRDS("mcr-rnet_pt.Rds")
+
+# since 2011 there has been a doubling
+
+# high percentage of key workers - stats
+
+```
+
+Cycling potential can also be estimated in terms of the number of new cycling trips that could result from a mode shift to cycling.
+Results of the shift from PT under the E-bikes scenario, which were calculated at the route segment level for the first time in this study, are shown for each key corridor in Figure \@ref(fig:shift):
+
+```{r shift, fig.cap="Cycling potential associated with a PT/cycling mode shift under the Ebikes scenario at the route network level within 500 m of each corridor, under the (unmet) assumption that every mode has equal chance of being replaced by cycling for a given origin-destination pair.
+In practice, the probability of a trip being replaced by cycling is mode and context-dependent, with physical distancing rules making a PT/cycling switch proportionally more likely."}
+key_corridors_buffer_100 = stplanr::geo_buffer(key_corridors8, dist = 500)
+# slow operation:
+rnet_pt_in_corridors = rnet_pt[key_corridors_buffer_100, , op = sf::st_within]
+# saveRDS(rnet_pt_in_corridors, "rnet_pt_in_corridors.Rds")
+# rnet_pt_in_corridors = readRDS("rnet_pt_in_corridors.Rds")
+plot(rnet_pt_in_corridors)
+# mapview::mapview(rnet_pt_in_corridors)
+```
+
+The results from this dataset will be used to assess cycling potential from a PT perspective, with results shown below.
+
+```{r}
+key_corridors_pt = st_join(rnet_pt_in_corridors, key_corridors_buffer["ref"])
+key_corridors_pt$km_day = key_corridors_pt$pt_mode_shift * as.numeric(sf::st_length(key_corridors_pt)) / 1000
+res_pt_corridors = key_corridors_pt %>% 
+  sf::st_drop_geometry() %>% 
+  group_by(ref) %>% 
+  summarise(
+    `Km/day` = sum(km_day)
+  )
+res_pt_corridors %>%
+  filter(!is.na(ref)) %>% 
+  arrange(desc(`Km/day`)) %>% 
+  knitr::kable(digits = 0, caption = "Ranking of corridors according to estimated distance of PT trips that switch to cycling within a 500m buffer of each corridor.")
+# key_corridors_pt = aggregate(rnet_pt_in_corridors, key_corridors_buffer, op = sf::st_within, FUN = mean)
+# key_corridors_pt %>% 
+#   sf::st_drop_geometry() %>% 
+#   arrange(desc(pt_mode_shift)) 
+```

### 2020-05-18 code/pct-manchester.R

@@ -2,6 +2,8 @@ library(tidyverse)
 
 # shown input data
 source("code/build.R")
+saveRDS(r_pct_top, "mcr_r_pct_lanes_top.Rds")
+saveRDS(r_pct_lanes_all, "mcr_r_pct_lanes_all.Rds")
 mapview::mapview(r_pct_lanes_all)
 corridor_names = c(
   "Chester Road", # (Trafford)
@@ -22,6 +24,7 @@ mapview::mapview(princess_road)
 rsf = readRDS("rsf.Rds")
 unique(rsf$region)
 rsf_mcr = rsf %>% filter(region == "Manchester")
+saveRDS(rsf_mcr, "mcr_rsf.Rds")
 rsf_1k = rsf_mcr %>% sample_n(1000)
 mapview::mapview(rsf_1k) # ttwa
 
@@ -41,7 +44,7 @@ tm_shape(key_corridors_mcr) +
   tm_lines(lwd = "width", scale = 9)
 
 saveRDS(key_corridors_mcr, "key_corridors_mcr.Rds")
-piggyback::pb_upload("key_corridors_mcr.Rds")
+# piggyback::pb_upload("key_corridors_mcr.Rds")
 
 # potential analysis ------------------------------------------------------
 
@@ -56,9 +59,11 @@ mapview::mapview(rnet_pt)
 
 # high percentage of key workers - stats
 
-
+library(pct)
 l = pct::get_desire_lines("greater-manchester")
 r = pct::get_pct_routes_fast("greater-manchester")
+saveRDS(r, "mcr-r-pct.Rds")
+saveRDS(l, "mcr-l-pct.Rds")
 r_top_pt = r %>% filter(train_tube > 50)
 mapview::mapview(r_top_pt["train_tube"])
 rnet_top = stplanr::overline(r_top_pt, "train_tube")

### 2020-05-19 code/mcr-safe-streets-analysis.Rmd
Final tweaks to mcr analysis

@@
 # output: bookdown::html_document2
-output: bookdown::word_document2
+# output: bookdown::word_document2
+output: bookdown::github_document2
 # output: bookdown::pdf_document2
@@
 ```{r, echo=FALSE, eval=FALSE}
-
+pagedown::chrome_print("mcr-safe-streets-analysis.html")
 ```
@@
 Around a third of trips are made by PT in many parts of Manchester and Salford, as shown in Figure \@ref(fig:pt).
-These areas areas that are particularly vulnerable to reduced capacity by public transport also tend to be within cycling distance from the main employment centre in Manchester's city centre.
+These areas that are particularly vulnerable to reduced capacity by public transport also tend to be within cycling distance from the main employment hub in Manchester's city centre.
 

### 2020-05-19 code/preprocess.R
Add preprocessing script

@@ -1,0 +1,12 @@
+# Aim: preprocess national roads data for pop-up analysis
+# This is an updated and tidier version of code in get-data-db.R
+
+library(tidyverse)
+
+# The data relies on roads.csv, a 2 GB file generated for the CyIPT project
+r = data.table::fread("roads.csv")
+names(r)
+# time-consuming
+r_sfc_all = sf::st_as_sfc(r$geotext, crs = 4326)
+r_subset_of_data = r %>% select(-geotext)
+rsf = sf::st_sf(r_subset_of_data, geometry = r_sfc_all)

### 2020-05-19 code/preprocess.R
Update preprocessing script

@@ -5,8 +5,53 @@ library(tidyverse)
 
 # The data relies on roads.csv, a 2 GB file generated for the CyIPT project
 r = data.table::fread("roads.csv")
-names(r)
-# time-consuming
+rtid = readr::read_csv("rtid.csv")
+
+# Create geometries
 r_sfc_all = sf::st_as_sfc(r$geotext, crs = 4326)
 r_subset_of_data = r %>% select(-geotext)
 rsf = sf::st_sf(r_subset_of_data, geometry = r_sfc_all)
+names(rsf)
+r_original = rsf %>%
+  select(name, ref, highway, maxspeed, pctgov, pctdutch, pctebike, width, Existing, length, aadt, idGlobal, rtid) 
+
+## ----preprocess------------------------------------------------------------------------------
+r_original$highway_type = r_original$highway
+r_original$highway_type = gsub(pattern = "_link", replacement = "", r_original$highway)
+highway_table = table(r_original$highway_type)
+highway_rare = highway_table[highway_table < nrow(r_original) / 100]
+highway_rare
+# bus_guideway living_street      motorway    pedestrian          road         steps 
+# 137          7792         14062         17898          3379         41168
+highway_remove = names(highway_rare)[!grepl(pattern = "motor|living|pedestrian|bus", x = names(highway_rare))]
+r_cleaned = r_original %>% 
+  filter(!grepl(pattern = "motorway", x = highway)) %>% 
+  mutate(highway_type = case_when(
+    highway_type %in% highway_remove ~ "other",
+    grepl(pattern = "path|track|trunk", highway_type) ~ "other",
+    grepl(pattern = "ped|liv", highway_type) ~ "pedestrian/living_street",
+    TRUE ~ highway_type
+  )) %>% 
+  mutate(maxspeed = case_when(
+    maxspeed <= 20 ~ "20 mph or less",
+    maxspeed > 20 & maxspeed <= 30 ~ "30 mph",
+    maxspeed > 30 ~ "40+ mph",
+  )) %>% 
+  mutate(cycling_potential = as.numeric(pctgov)) 
+# table(r_cleaned$maxspeed)
+rj = inner_join(r_cleaned, rtid) %>% 
+  mutate(lanes_f = abs(lanespsvforward) + abs(lanesforward)) %>% 
[…]

### 2020-05-19 code/update-rnet.R
finished method to update cycling potential

@@ -1,6 +1,7 @@
 library(pct)
 library(stplanr)
 library(tidyverse)
+library(sf)
 
 # readRDS(url("https://github.com/npct/pct-outputs-national/blob/master/commute/lsoa/rnet_all.Rds"))
 rnet = pct::get_pct_rnet(region = "avon") # LSOA route network
@@ -23,68 +24,58 @@ r_cyipt_joined = st_join(r_cyipt, rnet_buff, join = st_within)
 dim(r_cyipt_joined) # 111865 
 nrow(r_cyipt_joined)/nrow(r_cyipt)
 
-# test_all_present = r_cyipt_joined[which(duplicated(r_cyipt_joined$idGlobal) == TRUE) ,]
-# dim(r_cyipt_joined) - dim(test_all_present) # this must equal dim(r_cyipt) if all segments can be joined
 
+# Duplicates are mostly very short sections within junctions, or dual carriageways.
 dupes = r_cyipt_joined[which(duplicated(r_cyipt_joined$idGlobal) == TRUE | duplicated(r_cyipt_joined$idGlobal, fromLast = TRUE) == TRUE),]
-# dim(dupes) # 10593
+dim(dupes) # 10593
 
 # mapview(dupes) + mapview(rnet_buff)
 
+# Max cycling potential is best because on dual carriageways, each carriageway already has its cycle potential half (or less) what it would be if this was a single carriageway road
 dupes_max = dupes %>% 
+  st_drop_geometry() %>% 
   group_by(idGlobal) %>%
-  summarise(cycling_potential_max = max(govtarget_slc)) %>%
-  st_drop_geometry()
+  summarise(cycling_potential_max = max(govtarget_slc)) 
 
 dim(dupes_max) # 3533
 
-##Need to join dupes_max with the rnet. Then join the results of that with the r_cyipt/r_cyipt_joined
-
-
+##Need to join dupes_max with the rnet. Take cycling potential from dupes_max where duplicates exist. Otherwise take it from the new pct dataset, unless there is no buffer segment that the cyipt segment fits inside. In that case take it from cyipt. 
+# Otherwise some segments will be missing. Eg Anchor Road, Cumberland Basin, Gloucester Road Horfields. These go outside the buffer. 
 r_joined = left_join(r_cyipt_joined, dupes_max, by = "idGlobal") %>%
[…]

### 2020-05-19 code/preprocess.R
Complete preprocessing code for now

@@ -37,7 +37,7 @@ r_cleaned = r_original %>%
     maxspeed > 20 & maxspeed <= 30 ~ "30 mph",
     maxspeed > 30 ~ "40+ mph",
   )) %>% 
-  mutate(cycling_potential = as.numeric(pctgov)) 
+  mutate(cycling_potential = as.numeric(pctebike)) 
 # table(r_cleaned$maxspeed)
 rj = inner_join(r_cleaned, rtid) %>% 
   mutate(lanes_f = abs(lanespsvforward) + abs(lanesforward)) %>% 
@@ -50,8 +50,8 @@ rj$n_lanes[rj$n_lanes == "4"] = "4+"
 rj$n_lanes[rj$n_lanes == "5"] = "4+"
 rj$n_lanes[rj$n_lanes == "6"] = "4+"
 cy = r_cleaned %>% filter(highway == "cycleway")
-r_central = rj[city_key_buffer, ]
-r_main = r_central %>% 
-  filter(grepl(pattern = "cycleway|primary|second|tert", highway_type))
-r_main_region = rj %>% 
-  filter(grepl(pattern = "cycleway|primary|second|tert", highway_type))
+
+saveRDS(cy, "cy.Rds")
+saveRDS(rj, "rj.Rds") # 300 MB file
+rj_sample = rj %>% sample_n(1000)
+mapview::mapview(rj_sample) # national coverage

### 2020-05-19 code/build.R
Update build, prepare for national build

@@ -7,10 +7,14 @@ library(tmap)
 if(!exists("parameters")) {
 message("Loading global parameters")  
 s = c(
-  "Esri.WorldGrayCanvas",
-  "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
-  "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
+  `Grey` = "Esri.WorldGrayCanvas",
+  `PCT (Ebike)` = "https://npttile.vs.mythic-beasts.com/commute/v2/ebike/{z}/{x}/{y}.png",
+  `Cycleways` = "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
+  `Satellite` = "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
 )
+# test basemap:
+tmap_mode("view")
+tm_shape(spData::lnd) + tm_borders() + tm_basemap(s, tms = c(FALSE, TRUE, FALSE, FALSE))
 parameters = read_csv("input-data/parameters.csv")
 
 # read-in national data ---------------------------------------------------

### 2020-05-19 code/preprocess.R

@@ -55,3 +55,4 @@ saveRDS(cy, "cy.Rds")
 saveRDS(rj, "rj.Rds") # 300 MB file
 rj_sample = rj %>% sample_n(1000)
 mapview::mapview(rj_sample) # national coverage
+piggyback::pb_upload("rj.Rds", "cyipt/cyipt-phase-1-data")

### 2020-05-19 code/mcr-safe-streets-analysis.Rmd

@@
 # output: bookdown::html_document2
-# output: bookdown::word_document2
-output: bookdown::github_document2
+output: bookdown::word_document2
+# output: bookdown::github_document2
 # output: bookdown::pdf_document2
@@
 
+<!-- Summary route network results from Ebikes scenario, aggregated to the route network level, are shown for Greater Manchester below -->
+
 ```{r}
@@
 rnet_pt = readRDS("mcr-rnet_pt.Rds")
-
+# summary(rnet_pt$pt_mode_shift)
+rnet_pt_high = rnet_pt %>% filter(pt_mode_shift > 20)
+# mapview::mapview(rnet_pt_high)
 # since 2011 there has been a doubling
@@
 
-Cycling potential can also be estimated in terms of the number of new cycling trips that could result from a mode shift to cycling.
 Results of the shift from PT under the E-bikes scenario, which were calculated at the route segment level for the first time in this study, are shown for each key corridor in Figure \@ref(fig:shift):

### 2020-05-19 code/boundaries.R
Update build.R

@@ -181,6 +181,9 @@ cas_uas$included = !is.na(cas_uas$Level_dft)
 mapview::mapview(cas_uas["included"]) +
   mapview::mapview(lads_missing)
 
+regions = cas_uas
+saveRDS(regions, "regions.Rds")
+
 # Suggested solutions, to check:
 # Replace geographic region Cheshire with d regions of Chesire East and west Cheshire
 # Replace geographic region Bedfordshire with d regions of Bedford and west Central Bedfordshire

### 2020-05-19 code/build.R

@@ -18,30 +18,31 @@ tm_shape(spData::lnd) + tm_borders() + tm_basemap(s, tms = c(FALSE, TRUE, FALSE,
 parameters = read_csv("input-data/parameters.csv")
 
 # read-in national data ---------------------------------------------------
-r_all = readRDS("rsf.Rds")
-region_names = unique(r_all$region)
-rtid = readr::read_csv("rtid.csv")
-hsf = readRDS("hsf.Rds")
-regions = readRDS("~/cyipt/cyipt-bigdata/boundaries/local_authority/local_authority.Rds")
+regions = readRDS("regions.Rds")
+rj = readRDS("rj.Rds")
+region_names = regions$Name
+# hospitals:
+# hsf = readRDS("hsf.Rds")
 nrow(regions)
-mapview::mapview(regions)
+# mapview::mapview(regions)
 # r_original = r_original[st_transform(region, st_crs(r_original)), ]
 }
 
 # local parameters --------------------------------------------------------
 # i = 1
-# if(!exists("region_name"))
-#   region_name = "Leeds"
-r_original = r_all %>% filter(region == region_name) %>% 
-  select(name, ref, highway, maxspeed, pctgov, width, Existing, length, rtid, idGlobal) 
+if(!exists("region_name"))
+  region_name = "West Midlands"
+region = regions %>% filter(Name == region_name)
+r_original = rj[region_to_build, ]
 i = which(parameters$name %in% region_name)
+if(length(i) == 0) i = 1
 p = parameters[i, ]
 list2env(p, envir = .GlobalEnv)
 
+is_city = FALSE # Todo: add a new is_city parameter
 
 # buffers -----------------------------------------------------------------
 # for(i in seq(length(region_names))) {
-region = regions %>% filter(lad16nmw == region_name)
 city_centre = tmaptools::geocode_OSM(region_name, as.sf = TRUE)
 city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
 h_city = hsf[city_centre_buffer, ]

### 2020-05-19 code/build.R
Add city specific code

@@ -42,12 +42,13 @@ list2env(p, envir = .GlobalEnv)
 is_city = FALSE # Todo: add a new is_city parameter
 
 # buffers -----------------------------------------------------------------
-# for(i in seq(length(region_names))) {
-city_centre = tmaptools::geocode_OSM(region_name, as.sf = TRUE)
-city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
-h_city = hsf[city_centre_buffer, ]
-h_city_buffer = stplanr::geo_buffer(h_city, dist = key_destination_buffer_radius)
-city_key_buffer = st_union(city_centre_buffer, st_union(h_city_buffer))
+if(is_city) {
+  city_centre = tmaptools::geocode_OSM(region_name, as.sf = TRUE)
+  city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
+  h_city = hsf[city_centre_buffer, ]
+  h_city_buffer = stplanr::geo_buffer(h_city, dist = key_destination_buffer_radius)
+  city_key_buffer = st_union(city_centre_buffer, st_union(h_city_buffer))
+}
 
 ## ----preprocess------------------------------------------------------------------------------
 # remove motorways

### 2020-05-19 code/build.R
Update build for notts

@@ -1,6 +1,7 @@
 # aim: output results highlighting and ranking roads that have 'spare lanes'
 
 # global parameters ----------------------------------------------------------
+library(DescTools)
 library(sf)
 library(tidyverse)
 library(tmap)
@@ -19,10 +20,10 @@ parameters = read_csv("input-data/parameters.csv")
 
 # read-in national data ---------------------------------------------------
 regions = readRDS("regions.Rds")
-rj = readRDS("rj.Rds")
+rj_all = readRDS("rj.Rds")
 region_names = regions$Name
 # hospitals:
-# hsf = readRDS("hsf.Rds")
+hsf = readRDS("hsf.Rds")
 nrow(regions)
 # mapview::mapview(regions)
 # r_original = r_original[st_transform(region, st_crs(r_original)), ]
@@ -31,9 +32,16 @@ nrow(regions)
 # local parameters --------------------------------------------------------
 # i = 1
 if(!exists("region_name"))
-  region_name = "West Midlands"
-region = regions %>% filter(Name == region_name)
-r_original = rj[region_to_build, ]
+  region_name = "Nottingham"
+if(region_name == "Nottingham") {
+  region = regions %>% filter(str_detect(string = Name, pattern = "Nott")) %>% 
+    st_union()
+} else {
+  region = regions %>% filter(Name == region_name)
+}
+# time consuming: ~1 minute
+rj = rj_all[region, ]
+h_city = hsf[region, ]
 i = which(parameters$name %in% region_name)
 if(length(i) == 0) i = 1
 p = parameters[i, ]
@@ -48,70 +56,35 @@ if(is_city) {
   h_city = hsf[city_centre_buffer, ]
   h_city_buffer = stplanr::geo_buffer(h_city, dist = key_destination_buffer_radius)
   city_key_buffer = st_union(city_centre_buffer, st_union(h_city_buffer))
+  r_main_region = rj[city_key_buffer, ]
+} else {
+  city_key_buffer = sf::st_geometry(region)
+  r_main_region = rj
 }
-
-## ----preprocess------------------------------------------------------------------------------
-# remove motorways
-r_original$highway_type = r_original$highway
-r_original$highway_type = gsub(pattern = "_link", replacement = "", r_original$highway)
[…]

### 2020-05-19 code/build.R
Get existing infra into build script

@@ -19,12 +19,15 @@ tm_shape(spData::lnd) + tm_borders() + tm_basemap(s, tms = c(FALSE, TRUE, FALSE,
 parameters = read_csv("input-data/parameters.csv")
 
 # read-in national data ---------------------------------------------------
+# see preprocess.R for data origins
 regions = readRDS("regions.Rds")
 rj_all = readRDS("rj.Rds")
 region_names = regions$Name
 # hospitals:
 hsf = readRDS("hsf.Rds")
 nrow(regions)
+cycleways_en = readRDS("cycleways_en.Rds")
+
 # mapview::mapview(regions)
 # r_original = r_original[st_transform(region, st_crs(r_original)), ]
 }
@@ -105,51 +108,54 @@ group_table = table(r_pct_lanes$group)
 top_groups = tail(sort(group_table), 5)
 r_pct_lanes$graph_group[!r_pct_lanes$graph_group %in% names(top_groups)] = "other"
 
-r_filter_before_grouping = rj %>% 
-  filter(cycling_potential > min_cycling_potential) %>% 
-  filter(lanes_f > 1 | lanes_b > 1) %>% 
-  filter(cycling_potential > min_grouped_cycling_potential) %>% 
-  filter(length > 100)
 tmap_mode("plot")
-m0 = tm_shape(city_key_buffer) + tm_borders(col = "grey") +
-  tm_shape(r_pct_lanes_all) + tm_lines() +
-  tm_layout(title = "Roads on which there are spare lanes.")
-m1 = tm_shape(city_key_buffer) + tm_borders(col = "grey") +
-  tm_shape(r_filter_before_grouping) + tm_lines() +
-  tm_layout(title = "Filter then group:\n(length > 100, cycling_potential > 100)")
 m2 = tm_shape(city_key_buffer) + tm_borders(col = "grey") +
   tm_shape(r_pct_lanes) + tm_lines("graph_group", palette = "Dark2") +
   tm_layout(title = "Group then filter:\n(length > 500, cycling_potential > 100)")
+# m2
 
-## ----Grouping - work in progress---------------------------------------------------------------------------------
-r_pct_lanes$rounded_cycle_potential = RoundTo(r_pct_lanes$cycling_potential, 50)
-
-## Take segments (which are already grouped by the initial igraph list) and group by cycle potential rounded to the nearest 50
-
-agg_var = st_sfc(list(r_pct_lanes$group), list(r_pct_lanes$rounded_cycle_potential))
-r_pct_group1 = r_pct_lanes %>%
[…]

### 2020-05-19 code/explore-existing-popup-integration-birmingham.R

@@ -13,6 +13,7 @@ e = gfz %>%
 f = geofabrik::gf_filename(name = country_name)
 download.file(url = e$pbf_url, destfile = f)
 cycleways_en = geofabrik::read_pbf(dsn = f, key = "highway", value = "cycleway")
+saveRDS(cycleways_en, "cycleways_en.Rds") # 10 MB file 
 plot(sf::st_geometry(cycleways_en))
 
 uas_en = readRDS("uas_en.Rds")

### 2020-05-20 code/build.Rmd
Add build.Rmd - wip

```{r, include=FALSE}
if(!exists("region_name"))
  region_name = "Nottingham"
# set working directory to root dir:

knitr::opts_knit$set(root.dir = "..")
```

```{r}
# aim: output results highlighting and ranking roads that have 'spare lanes'

# global parameters ----------------------------------------------------------
library(DescTools)
library(sf)
library(tidyverse)
library(tmap)
if(!exists("parameters")) {
message("Loading global parameters")  
s = c(
  `Grey` = "Esri.WorldGrayCanvas",
  `PCT (Ebike)` = "https://npttile.vs.mythic-beasts.com/commute/v2/ebike/{z}/{x}/{y}.png",
  `Cycleways` = "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
  `Satellite` = "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
)
# test basemap:
tmap_mode("view")
tm_shape(spData::lnd) + tm_borders() + tm_basemap(s, tms = c(FALSE, TRUE, FALSE, FALSE))
parameters = read_csv("input-data/parameters.csv")

# read-in national data ---------------------------------------------------
# see preprocess.R for data origins
regions = readRDS("regions.Rds")
rj_all = readRDS("rj.Rds")
region_names = regions$Name
# hospitals:
hsf = readRDS("hsf.Rds")
nrow(regions)
cycleways_en = readRDS("cycleways_en.Rds")

# mapview::mapview(regions)
# r_original = r_original[st_transform(region, st_crs(r_original)), ]
}

# local parameters --------------------------------------------------------
# i = 1
if(region_name == "Nottingham") {
  region = regions %>% filter(str_detect(string = Name, pattern = "Nott")) %>% 
    st_union()
} else {
  region = regions %>% filter(Name == region_name)
}
# time consuming: ~1 minute
rj = rj_all[region, ]
h_city = hsf[region, ]
i = which(parameters$name %in% region_name)
if(length(i) == 0) i = 1
p = parameters[i, ]
list2env(p, envir = .GlobalEnv)

is_city = FALSE # Todo: add a new is_city parameter

# buffers -----------------------------------------------------------------
if(is_city) {
  city_centre = tmaptools::geocode_OSM(region_name, as.sf = TRUE)
  city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
  h_city = hsf[city_centre_buffer, ]
  h_city_buffer = stplanr::geo_buffer(h_city, dist = key_destination_buffer_radius)
  city_key_buffer = st_union(city_centre_buffer, st_union(h_city_buffer))
  r_main_region = rj[city_key_buffer, ]
} else {
  city_key_buffer = sf::st_geometry(region)
  r_main_region = rj
}

## ----t1, results='asis'----------------------------------------------------------------------
t1 = rj %>%
  st_drop_geometry() %>%
  # select(name, highway_type, maxspeed, cycling_potential, width) %>%
  table1::table1(~ highway_type + cycling_potential + width + n_lanes | maxspeed, data = .)

## ----hospitals, fig.cap="Overview map of input data, showing the main highway types and location of hospitals in city"----
m1 = r_main_region %>%
  sample_n(1000) %>% 
  mutate(`Highway type` = highway_type) %>% 
  mutate(`Cycling potential` = case_when(cycling_potential < 100 ~ 100, TRUE ~ cycling_potential)) %>% 
  tm_shape() +
  tm_lines(col = "Highway type", palette = c("green", "black", "blue", "grey"),
           lwd = "Cycling potential", scale = 5, lwd.legend = c(100, 200, 500, 1000),
           lwd.legend.labels = c("0 - 100", "100 - 200", "200 - 500", "500+")) +
  tm_shape(h_city) + tm_dots(size = 0.2, col = "ParentName", palette = "Dark2", title = "Hospital group") +
  tm_layout(legend.outside = TRUE)
# + tm_text("OrganisationName")
# m1

## ----levels, fig.height=3, fig.cap="Illustration of the 'group then filter' method to identify long sections with spare lanes *and* high cycling potential"----
r_pct_lanes_all = r_main_region %>% 
  filter(cycling_potential > min_cycling_potential) %>% 
  filter(lanes_f > 1 | lanes_b > 1)
# mapview::mapview(r_pct_lanes)

touching_list = st_touches(r_pct_lanes_all)
g = igraph::graph.adjlist(touching_list)
components = igraph::components(g)
r_pct_lanes_all$group = components$membership
r_pct_lanes = r_pct_lanes_all %>% 
  group_by(group) %>% 
  mutate(group_length = sum(length)) %>% 
  mutate(cycling_potential_mean = weighted.mean(cycling_potential, w = length, na.rm = TRUE)) %>% 
  filter(cycling_potential_mean > min_grouped_cycling_potential)
r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length)

r_pct_lanes$graph_group = r_pct_lanes$group
group_table = table(r_pct_lanes$group)
top_groups = tail(sort(group_table), 5)
r_pct_lanes$graph_group[!r_pct_lanes$graph_group %in% names(top_groups)] = "other"

tmap_mode("plot")
m2 = tm_shape(city_key_buffer) + tm_borders(col = "grey") +
  tm_shape(r_pct_lanes) + tm_lines("graph_group", palette = "Dark2") +
  tm_layout(title = "Group then filter:\n(length > 500, cycling_potential > 100)")
m2

r_pct_grouped = r_pct_lanes %>%
  group_by(name, group) %>%
  summarise(
    group_length = sum(length),
    cycling_potential = round(weighted.mean(cycling_potential_mean, length))
  )
# summary(r_pct_grouped$group_length)
r_pct_top = r_pct_grouped %>%
  filter(group_length > min_grouped_length) %>% 
  filter(cycling_potential > min_grouped_cycling_potential) %>% 
  filter(!grepl(pattern = regexclude, name, ignore.case = TRUE)) %>% 
  mutate(km_day = round(cycling_potential * group_length / 1000))

## ----Grouping - work in progress---------------------------------------------------------------------------------

# commented-out for now (RL)
# r_pct_lanes$rounded_cycle_potential = RoundTo(r_pct_lanes$cycling_potential, 50)
# 
# ## Take segments (which are already grouped by the initial igraph list) and group by cycle potential rounded to the nearest 50
# 
# agg_var = st_sfc(list(r_pct_lanes$group), list(r_pct_lanes$rounded_cycle_potential))
# r_pct_group1 = r_pct_lanes %>%
#   group_by(group, rounded_cycle_potential) %>%
#   select(group, rounded_cycle_potential) %>%
#   st_drop_geometry() %>%
#   aggregate(by = list(r_pct_lanes$group, r_pct_lanes$rounded_cycle_potential), FUN = mean)
# 
# # Now need to separate non-adjacent groups with the same cycle potential
# 
# touching_list2 = st_touches(r_pct_group1)
# g2 = igraph::graph.adjlist(touching_list2)
# components2 = igraph::components(g2)
# r_pct_group1$group2 = components2$membership
# 
# 
# # This one should be shown on the map
# r_pct_grouped = r_pct_group1 %>%
#   group_by(group2) %>%
#   summarise(
#     group_length = sum(length),
#     cycling_potential = round(weighted.mean(cycling_potential, length))
#   )

# Generate lists of top segments ------------------------------------------------------------

# summary(r_pct_grouped$group_length)
r_pct_top = r_pct_grouped %>%
  filter(group_length > min_grouped_length) %>% 
  filter(cycling_potential > min_grouped_cycling_potential) %>% 
  filter(!grepl(pattern = regexclude, name, ignore.case = TRUE)) %>% 
  mutate(km_cycled = round(cycling_potential * group_length / 1000))
# head(r_pct_top$kkm_cycled)

r_pct_lanes_overlap = r_pct_lanes[r_pct_top, , op = sf::st_covered_by] # works
# r_pct_no_overlap = st_difference(r_pct_lanes, r_pct_top) # slow
r_pct_no_overlap = r_pct_lanes %>% 
  filter(!idGlobal %in% r_pct_lanes_overlap$idGlobal)

r_pct_top_n = r_pct_top %>% 
  group_by(name) %>% 
  slice(which.max(km_cycled)) %>% 
  filter(name != "") %>% 
  ungroup() %>% 
  arrange(desc(km_cycled)) %>% 
  slice(1:10)

# get existing infrastructure ---------------------------------------------
cycleways = cycleways_en[region, ]

# note: dist could be a parameter (RL)
cycleway_buffer = stplanr::geo_buffer(cycleways, dist = 50) %>% sf::st_union()
r_pct_top_no_cycleways = sf::st_difference(r_pct_top, cycleway_buffer)
r_pct_top_no_cycleways = r_pct_top_no_cycleways %>% 
  st_cast("LINESTRING") %>% 
  mutate(group_length = round(as.numeric(st_length(.)))) %>% 
  filter(group_length > 50)

## ----res, fig.cap="Results, showing road segments with a spare lane (light blue) and road groups with a minium threshold length, 1km in this case (dark blue). The top 10 road groups are labelled."----
tmap_mode("view")
m =
  tm_shape(r_pct_no_overlap) + tm_lines(col = "turquoise", lwd = 6, alpha = 0.6) +
  tm_shape(r_pct_top_no_cycleways) + tm_lines(col = "blue", lwd = 6, alpha = 0.6) +
  tm_shape(r_pct_top) + tm_lines(col = "red", lwd = 6, alpha = 0.6) +
  tm_shape(cycleways) + tm_lines() +
  tm_shape(r_pct_top_n) + tm_text("name") +
  tm_shape(h_city) + tm_dots(size = 0.1, col = "red", alpha = 0.4) +
  tm_basemap(server = s) +
  tm_scale_bar()
tmap::tmap_save(m, "m.html")

## --------------------------------------------------------------------------------------------
res_table = r_pct_top_n %>% 
  sf::st_drop_geometry() %>% 
  select(name, length = group_length, cycling_potential, km_cycled) 
knitr::kable(res_table, caption = "The top 10 candidate roads for space reallocation for pop-up active transport infrastructure according to methods presented in this paper.", digits = 0)
```

### 2020-05-20 code/build.R
updated cycle potential in build.R

@@ -5,6 +5,8 @@ library(DescTools)
 library(sf)
 library(tidyverse)
 library(tmap)
+library(pct)
+library(stplanr)
 if(!exists("parameters")) {
 message("Loading global parameters")  
 s = c(
@@ -85,9 +87,27 @@ m1 = r_main_region %>%
 # + tm_text("OrganisationName")
 # m1
 
+##Update cycling potential with new PCT figures
+rnet = pct::get_pct_rnet(region = "avon")
+rnet_buff = geo_buffer(shp = rnet, dist = 10)
+r_cyipt_joined = st_join(r_main_region, rnet_buff, join = st_within)
+
+dupes = r_cyipt_joined[which(duplicated(r_cyipt_joined$idGlobal) == TRUE | duplicated(r_cyipt_joined$idGlobal, fromLast = TRUE) == TRUE),]
+dupes_max = dupes %>% 
+  st_drop_geometry() %>% 
+  group_by(idGlobal) %>%
+  summarise(cycling_potential_max = max(govtarget_slc)) 
+
+r_joined = left_join(r_cyipt_joined, dupes_max, by = "idGlobal") %>%
+  mutate(cycling_potential = ifelse(is.na(cycling_potential_max), ifelse(is.na(govnearmkt_slc), pctgov, govnearmkt_slc), cycling_potential_max),
+         cycling_potential_source = ifelse(is.na(cycling_potential_max), ifelse(is.na(govnearmkt_slc), "cyipt", "updated"), "updated_duplicate"))
+
+r_positive = r_joined[which(r_joined$cycling_potential > 0),] %>%
+  select(name:n_lanes, cycling_potential_source) %>%
+  distinct(.keep_all = TRUE) # remove the duplicates
 
 ## ----levels, fig.height=3, fig.cap="Illustration of the 'group then filter' method to identify long sections with spare lanes *and* high cycling potential"----
-r_pct_lanes_all = r_main_region %>% 
+r_pct_lanes_all = r_positive %>% 
   filter(cycling_potential > min_cycling_potential) %>% 
   filter(lanes_f > 1 | lanes_b > 1)
 # mapview::mapview(r_pct_lanes)

### 2020-05-20 code/update-rnet.R

@@ -42,13 +42,11 @@ dim(dupes_max) # 3533
 ##Need to join dupes_max with the rnet. Take cycling potential from dupes_max where duplicates exist. Otherwise take it from the new pct dataset, unless there is no buffer segment that the cyipt segment fits inside. In that case take it from cyipt. 
 # Otherwise some segments will be missing. Eg Anchor Road, Cumberland Basin, Gloucester Road Horfields. These go outside the buffer. 
 r_joined = left_join(r_cyipt_joined, dupes_max, by = "idGlobal") %>%
-  mutate(cycling_potential = ifelse(is.na(cycling_potential_max), ifelse(is.na(govtarget_slc), pctgov, govtarget_slc), cycling_potential_max))
+  mutate(cycling_potential = ifelse(is.na(cycling_potential_max), ifelse(is.na(govtarget_slc), pctgov, govtarget_slc), cycling_potential_max),
+         cycling_potential_source = ifelse(is.na(cycling_potential_max), ifelse(is.na(govtarget_slc), "cyipt", "updated"), "updated_duplicate"))
 dim(r_joined) # 111865 needs to be the same as dim(r_cyipt_joined). Be careful, this includes large numbers of streets that with 0 cycling potential
 
-r_joined = r_joined %>%
-  distinct()
-
-head(r_joined)
+# head(r_joined)
 
 r_positive = r_joined[which(r_joined$cycling_potential > 0),] %>%
   select(idGlobal:widthstatus, Existing, length, cycling_potential) %>%

### 2020-05-20 code/update-rnet.R
school rnet

@@ -7,16 +7,24 @@ library(sf)
 rnet = pct::get_pct_rnet(region = "avon") # LSOA route network
 # mapview(rnet["govtarget_slc"])
 
+##Combine with schools rnet
+rnet_school = get_pct_rnet(region = "avon", purpose = "school")
+rnet_reduce = rnet[,c(1,3)]
+rnet_school_reduce = rnet[,c(1,3)]
+combine = rbind(rnet_reduce, rnet_school_reduce)
+
+rnet_combined = stplanr::overline2(x = combine, attrib = "govtarget_slc")
+
 r_cyipt = readRDS("rsf.Rds")
 r_cyipt = r_cyipt[r_cyipt$region == "Bristol",]
 
 # st_crs(rnet)
 # st_crs(r_cyipt)
 
-dim(rnet) # 13829 this only has streets which are part of the LSOA route network
+dim(rnet_combined) # 13829 this only has streets which are part of the LSOA route network
 dim(r_cyipt) # 104805 this has every street
 
-rnet_buff = geo_buffer(shp = rnet, dist = 10)
+rnet_buff = geo_buffer(shp = rnet_combined, dist = 10)
 # mapview(rnet_buff["govtarget_slc"])
 
 r_cyipt_joined = st_join(r_cyipt, rnet_buff, join = st_within)

### 2020-05-28 code/build.R
Reduce minp_exclude to 0.5

@@ -365,7 +365,7 @@ r_lanes_grouped_in_cycleway = st_intersection(r_lanes_grouped2, cycleway_buffer)
 r_lanes_grouped_in_cycleway = r_lanes_grouped_in_cycleway %>% 
   st_drop_geometry() 
 
-minp_exclude = 0.75
+minp_exclude = 0.5
 r_lanes_joined = left_join(r_lanes_grouped2, r_lanes_grouped_in_cycleway) %>% 
   mutate(km_cycled = round(mean_cycling_potential * group_length / 1000)) 
 r_lanes_joined$proportion_on_cycleway = r_lanes_joined$length_in_cycleway / r_lanes_joined$group_length

### 2020-05-28 code/build.R
Add grouping variable for #42

@@ -329,9 +329,21 @@ rg_long = rg_new2 %>%
 rg_new3 = rg_new2[rg_new2$lastgroup %in% rg_long$lastgroup,]
 
 # mapview::mapview(rg_new3)
+rg_new4 = rg_new3 %>% 
+  group_by(ref, group, ig, name) %>% 
+  mutate(long_named_section = case_when(
+    sum(length) > 5000 & name != "" ~ name,
+    TRUE ~ "Other"
+  )
+  ) %>% 
+  ungroup()
 
-r_lanes_grouped2 = rg_new3 %>% 
-  group_by(ref, group, ig) %>% 
+table(rg_new4$long_named_section)
+summary(as.factor(rg_new4$long_named_section))
+# find group membership of top named roads
+
+r_lanes_grouped2 = rg_new4 %>% 
+  group_by(ref, group, ig, long_named_section) %>% 
   summarise(
     name = ifelse(names(table(name))[which.max(table(name))] != "", 
                   names(table(name))[which.max(table(name))],
@@ -344,13 +356,33 @@ r_lanes_grouped2 = rg_new3 %>%
     spare_lane = sum(length[spare_lane]) > sum(length[!spare_lane])
   ) %>% 
   filter(mean_cycling_potential > min_grouped_cycling_potential | group_length > min_grouped_length) %>%
-  # filter(group_length > min_grouped_length |
-           #  mean_cycling_potential > min_grouped_cycling_potential &
-           # (group_length * 2 > min_grouped_length)) %>% 
   ungroup() %>% 
   mutate(group_id = 1:nrow(.))
 
-# mapview::mapview(r_lanes_grouped2, zcol = "mean_cycling_potential", lwd = 3)
+# max_length = 10000
+# r_lanes_grouped_long = r_lanes_grouped2 %>% 
+#   filter(group_length > max_length) %>% 
+#   filter(ref != "")
+# # rg_new3_long = rg_new3 %>% filter(ref %in% r_lanes_grouped_long$ref)
+# r_lanes_long_segments = rg_new3[r_lanes_grouped_long, , op = st_within]
+# 
+# r_lanes_grouped3 = r_lanes_long_segments %>% 
+#   group_by(ref, name) %>%
+#   summarise(
+#     group_length = round(sum(length)),
+#     mean_cycling_potential = round(weighted.mean(cycling_potential, length, na.rm = TRUE)),
[…]

### 2020-05-28 code/build.R
Update top_n with new long roads group, close #42

@@ -408,12 +408,14 @@ summary(r_lanes_joined$proportion_on_cycleway) # all between 0 and 1
 r_lanes_top = r_lanes_joined %>%
   ungroup() %>% 
   filter(name != "" & ref != "") %>%
-  filter(mean_cycling_potential > 2*min_grouped_cycling_potential) %>% 
+  filter(group_length > min_grouped_length) %>%
+  filter(mean_cycling_potential > min_grouped_cycling_potential) %>% 
   filter(!grepl(pattern = regexclude, name, ignore.case = TRUE)) %>% 
   filter(proportion_on_cycleway < minp_exclude) %>% 
   arrange(desc(km_cycled)) %>% 
   slice(1:20)
 nrow(r_lanes_top)
+r_lanes_top
 
 # classify roads to visualise
 labels = c("Top route", "Spare lane(s)", "Estimated width > 10m")
@@ -497,5 +499,6 @@ m_leaflet = tmap_leaflet(m)
 res_table = r_lanes_top %>% 
   sf::st_drop_geometry() %>% 
   select(name, ref, length = group_length, mean_cycling_potential, km_cycled) 
+res_table
 # knitr::kable(res_table, caption = "The top 10 candidate roads for space reallocation for pop-up active transport infrastructure according to methods presented in this paper.", digits = 0)
 

### 2020-05-28 code/index.Rmd
Prepare website

---
title: "Rapid cycleway prioritisation tool"
author: "University of Leeds"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r pkg, include=FALSE}
library(sf)
library(tidyverse)
library(tmap)
tmap_mode("view")
```

```{r}
regions_dft = readRDS("regions_dft.Rds")
regions_dft_centroids = sf::st_point_on_surface(regions_dft)
regions_dft$html_name = tolower(gsub(pattern = " |, ", replacement = "-", regions_dft$Name))
regions_dft$url = paste0("<a href='", regions_dft$html_name ,"' target='_blank'>", regions_dft$html_name, "</a>")
```

```{r}
pvars = c("Name", "Level")
tm_shape(regions_dft) + tm_polygons("Level", popup.vars = pvars) + tm_shape(regions_dft_centroids) + tm_text("Name", size = 0.7)
```

```{r}
regions_dft %>%
  sf::st_drop_geometry() %>%
  select(Name, Level, url) %>% 
  arrange(Name) %>% 
  DT::datatable(escape = -4)
```

### 2020-05-28 code/boundaries2.R

@@ -107,6 +107,8 @@ regions_within = nmatches == 2
 regions_dft$Name[regions_within]
 # manually:
 regions_within_names = c("Leicester", "Nottingham", "Derby", "Stoke-on-Trent")
-
+regions_dft$url = paste0("<a href='test'>test</a>")
+pvars = c("url", "Name")
+tm_shape(regions_dft) + tm_polygons("Level", popup.vars = pvars) + tm_shape(regions_dft_centroids) + tm_text("Name", size = 0.7)
 
 saveRDS(regions_dft, "regions_dft.Rds")

### 2020-05-28 code/index.Rmd
Add per region landing page

@@
 
+```{r, eval=FALSE}
+# test build works before running for all:
+source("code/build.R")
+```
+
 ```{r}
-regions_dft = readRDS("regions_dft.Rds")
-regions_dft_centroids = sf::st_point_on_surface(regions_dft)
-regions_dft$html_name = tolower(gsub(pattern = " |, ", replacement = "-", regions_dft$Name))
-regions_dft$url = paste0("<a href='", regions_dft$html_name ,"' target='_blank'>", regions_dft$html_name, "</a>")
+regions = readRDS("regions_dft.Rds")
+regions_centroids = sf::st_point_on_surface(regions)
+regions$html_name = tolower(gsub(pattern = " |, ", replacement = "-", regions$Name))
+regions$url = paste0("<a href='", regions$html_name ,"' target='_blank'>", regions$Name, "</a>")
 ```
@@
 pvars = c("Name", "Level")
-tm_shape(regions_dft) + tm_polygons("Level", popup.vars = pvars) + tm_shape(regions_dft_centroids) + tm_text("Name", size = 0.7)
+tm_shape(regions) + tm_polygons("Level", popup.vars = pvars) + tm_shape(regions_centroids) + tm_text("Name", size = 0.7)
 ```
@@
 ```{r}
-regions_dft %>%
+regions %>%
   sf::st_drop_geometry() %>%
@@
 ```
+
+```{r, eval=FALSE}
+
+sample_to_build_regex = "west of|west y|hereford|nott|leic|manc|liv|cam"
+region_names_to_build = regions %>% 
+  filter(grepl(pattern = sample_to_build_regex, x = Name, ignore.case = TRUE)) %>% 
+  pull(Name)
+
+dn = dn # dirctory name
+dir.create(dn)
+# region_names_to_build = c("West Yorkshire", "Nottingham")
+rmarkdown::render("code/index.Rmd", output_dir = dn)
+# region_names_to_build = c("Greater London", "Cambridgeshire")
+
+rn = "West of England"
+for(rn in region_names_to_build) {
+  region = regions %>% filter(Name == rn)
+  t = Sys.time()
+  message("Building for ", rn)
+  d = file.path(dn, region$html_name)
+  dir.create(d)
+  region_name = region$Name
+  # source("code/build.R")
+  rmarkdown::render(input = "code/build.R", output_dir = d, knit_root_dir = ".")
+  plot(sf::st_geometry(r_lanes_top))
+  htmlwidgets::saveWidget(m_leaflet, file.path("~/cyipt/tempCycleways/", d, "m.html"))
+  time_to_run = Sys.time() - t
+  rmarkdown::render("code/region_index.Rmd", output_file = file.path("~/cyipt/tempCycleways/", d, "index.html"))
+  message(round(time_to_run), " minutes to build ", rn)
+}
+```

### 2020-05-28 code/region_index.Rmd
Add per region landing page

--- 2020-05-28 code/index.Rmd
@@
 ---
-title: "Rapid cycleway prioritisation tool"
-author: "University of Leeds"
 output: html_document
+title: '`r region$Name`'
 ---
@@
 ```{r}
-regions_dft = readRDS("regions_dft.Rds")
-regions_dft_centroids = sf::st_point_on_surface(regions_dft)
-regions_dft$html_name = tolower(gsub(pattern = " |, ", replacement = "-", regions_dft$Name))
-regions_dft$url = paste0("<a href='", regions_dft$html_name ,"' target='_blank'>", regions_dft$html_name, "</a>")
+knitr::include_url("m.html")
 ```
 
-```{r}
-pvars = c("Name", "Level")
-tm_shape(regions_dft) + tm_polygons("Level", popup.vars = pvars) + tm_shape(regions_dft_centroids) + tm_text("Name", size = 0.7)
-```
+See full map [here](m.html).
 
 ```{r}
-regions_dft %>%
-  sf::st_drop_geometry() %>%
-  select(Name, Level, url) %>% 
-  arrange(Name) %>% 
-  DT::datatable(escape = -4)
+DT::datatable(res_table)
 ```

### 2020-05-28 code/build.R

@@ -27,7 +27,7 @@ if(!exists("s")) {
   
   # read-in national data ---------------------------------------------------
   # see preprocess.R for data origins
-  regions = readRDS("regions.Rds")
+  regions = readRDS("regions_dft.Rds")
   rj_all = readRDS("rj.Rds")
   region_names = regions$Name
   # hospitals:

### 2020-05-28 code/index.Rmd
Update build script

--- 2020-05-28 code/index.Rmd
@@
 ```{r}
-regions_dft %>%
+dn = "popupCycleways/v0.2" # dirctory name
+built_regions = list.files(dn)
+regions %>%
   sf::st_drop_geometry() %>%
+  filter(html_name %in% built_regions) %>% 
   select(Name, Level, url) %>% 
@@
 ```
+
+```{r, eval=FALSE}
+
+sample_to_build_regex = "west of|west y|west mid|hereford|nott|leic|manc|liv|cam"
+region_names_to_build = regions %>% 
+  filter(grepl(pattern = sample_to_build_regex, x = Name, ignore.case = TRUE)) %>% 
+  pull(Name)
+dir.create(dn)
+# region_names_to_build = c("West Yorkshire", "Nottingham")
+rmarkdown::render("code/index.Rmd", output_dir = dn)
+# region_names_to_build = c("Greater London", "Cambridgeshire")
+regions_within_names = c("Leicester", "Nottingham", "Derby", "Stoke-on-Trent")
+names(regions_within_names) = c(paste0(regions_within_names[1:3], "shire"), "Staffordshire")
+regions_containing_names = names(regions_within_names)
+names(regions_containing_names) = regions_within_names
+
+rn = "West of England"
+for(rn in region_names_to_build) {
+  region = regions %>% filter(Name == rn)
+  if(rn %in% c(regions_within_names)) {
+    region$geometry = sf::st_union(
+      region,
+      regions %>% filter(Name == regions_containing_names[rn])
+      ) %>% sf::st_geometry()
+  }
+  t = Sys.time()
+  message("Building for ", rn)
+  d = file.path(dn, region$html_name)
+  dir.create(d)
+  region_name = region$Name
+  # source("code/build.R")
+  rmarkdown::render(input = "code/build.R", output_dir = d, knit_root_dir = ".")
+  plot(sf::st_geometry(r_lanes_top))
+  htmlwidgets::saveWidget(m_leaflet, file.path("~/cyipt/tempCycleways/", d, "m.html"))
+  time_to_run = Sys.time() - t
+  rmarkdown::render("code/region_index.Rmd", output_file = file.path("~/cyipt/tempCycleways/", d, "index.html"))
+  message(round(time_to_run), " minutes to build ", rn)
+}
+```

### 2020-05-28 code/build.R

@@ -27,7 +27,7 @@ if(!exists("s")) {
   
   # read-in national data ---------------------------------------------------
   # see preprocess.R for data origins
-  regions = readRDS("regions_dft.Rds")
+  if(!exists("regions")) regions = readRDS("regions_dft.Rds")
   rj_all = readRDS("rj.Rds")
   region_names = regions$Name
   # hospitals:

### 2020-05-28 code/index.Rmd
index text

@@
 ```
+
+The aim of this tool is to help reveal the locations that have the best potential for pop-up cycleways.
+It should be used together with local knowledge, to prioritise where to install rapid measures to support safe travel as part of the COVID-19 response.
+We anticipate that this may be used in the development of a range of temporary, experimental and longer term measures to make the streets safer for us all, as new travel patterns and behaviours emerge in the coming weeks and months.
+
+Maps are available for regions covering the whole of England.
+
+We have focused on roads which are sufficiently wide that part of the roadspace could potentially be reallocated to pop-up cycleways without preventing traffic movement.
+In identifying these locations, we have not taken average traffic volumes into account.
+The pandemic has had a severe impact on movement and traffic flows are currently below normal levels.
+Instead, we simply identify streets that have multiple vehicle lanes or a carriageway width of greater than 10m.
+
+Reallocating roadspace is just one of the measures available for transport authorities.
+Other approaches include modal filters, road closures, and the creation of bus, walking and cycling only streets, as well as speed limits and traffic calming measures.
+Used strategically, a combination of these measures can promote active travel through the creation of filtered permeability networks and Low Traffic Neighbourhoods.
+For certain types of area, such as residential zones, this may be the most suitable approach.
+However, roadspace reallocation is naturally suited to larger arterial-type roads and that is the focus of this tool.
+
+We have used several types of data.
+The cycling potential of each road is estimated according the Propensity to Cycle Tool's Government Target scenario, which represents the government target to double cycling levels by 2025.
+Our figures include both travel to work and travel to school, and are based on 2011 Census data.
+Data on road widths and number of lanes is derived from the Cycling Infrastructure Prioritisation Toolkit.
+
+### Map layers
+
+The maps show five layers.
+
+1) Key network:
+
+- we have created this to represent the most important cycle corridors.
+For this layer, we do not take into account the width of the roads, but simply represent them as continuous corridors.
+This underscores the importance of ensuring that cycle routes are continuous and do not 'give up' at awkward junctions.
+To derive this layer, we have selected all road segments with cycle potential in the 85th percentile and above, which also belong to classified trunk roads, A roads or B roads.
+These segments are then joined together to form a continuous network.
+
+2) Existing cycleways (500m +)
+
+- this layer shows existing cycleways.
+This may be helpful to view in connection with the other layers, because it can reveal where gaps in the existing cycle network could be filled by new infrastructure.
+This layer does not include some on-road cycle lanes or cycleways less than 500m in length, but these can be viewed in one of the basemap options (see below for more information about the basemaps).
+
+The following three layers are the central results of our analysis, representing the most promising locations for pop-up cycleways.
+
+3) Top routes
+
+- these are the top routes for pop-up cycleways.
+These are classified by "km cycled", calculated as the length of the route multiplied by the mean cycling potential.
+All of these road segments will either have more than one vehicle lane in either direction, or be at least 10m in width.
+The routes are listed in the table below the map, so you can see more information on them.
+The number of top routes selected is normally 20, with some variation by region.
+
+4) Spare lane(s)
+
+- these are routes that have more than one vehicle lane in either direction and have a cycling potential above the minimum threshold (which varies according to region as it needs to be higher in e.g.
+Greater London than in rural areas).
+
+5) Width >= 10m
+
+- these routes are at least 10m in width and have a cycling potential above the minimum threshold.
+
+Click on any of these layers to see a pop-up box containing more information about the road in question.
+
+### Basemaps
+
+There are also seven basemaps to choose from.
+The default basemap is grey and provides an uncluttered view.
+We then have basemap options showing four different scenarios from the Propensity to Cycle Tool.
+This allows our results to be compared with cycling potentials across the road network.
+The scenarios available show cycling potential for Commuting (government target), Schools (government targets), Commuting (e-bikes), and schools (Go Dutch).
+
+A further basemap option shows all existing cycleways, including on-road cycle lanes. Finally, a satellite image is also available as a basemap.
 

### 2020-05-28 code/index.Rmd
Update index, rename project

@@
 ```
+
+## PRELIMINARY RESULTS
+
+Not quality assured.
+
+Please provide feedback on these preliminary results on this [web form (estimated time to complete: 5 minites)](https://urldefense.proofpoint.com/v2/url?u=https-3A__forms.office.com_Pages_ResponsePage.aspx-3Fid-3DqO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u&d=DwMCaQ&c=troKkvwivNn_CddsvWCHHPiPoFoTgTGIbXJULvYU158&r=DXeEYtdSYDBZN2DBwEgAHMmX5PQ_ob9r2MynFWta2mU&m=mkyUa3RIC9d5LSrHUqk_WdstQR1OQuTWyYLBecHFMm0&s=dD7GVcQbjZ2AcdCObSe-rNKmLLDBhu98ChFupIIrlJA&e=).
+
+## Introduction
 

### 2020-05-28 code/build.R
Tweak parameters based on Cambridge

@@ -220,13 +220,13 @@ key_network = r_key_roads_plus_high_pct %>%
 r_lanes_all = r_main_region %>% 
   filter(cycling_potential > min_cycling_potential) %>% # min_cycling_potential = 0 so this simply selects multilane roads
   mutate(spare_lane = lanes_f > 1 | lanes_b > 1) %>% 
-  filter(spare_lane | width >= 10)
+  filter(spare_lane | width >= 9)
 
 # Optionally remove areas outside key network:
 # r_lanes_all = r_lanes_all[r_key_network_buffer_large, ]
 # mapview::mapview(r_lanes_all)
-
-r_lanes_all_buff = geo_buffer(shp = r_lanes_all, dist = 50)
+buff_dist_large = 100
+r_lanes_all_buff = geo_buffer(shp = r_lanes_all, dist = buff_dist_large)
 touching_list = st_intersects(r_lanes_all_buff)
 g = igraph::graph.adjlist(touching_list)
 components = igraph::components(g)
@@ -236,29 +236,8 @@ r_lanes_all$group = components$membership
 r_linestrings_with_ref = r_lanes_all %>%
   filter(ref != "")
 
-# to optionally create and filter by group characteristics:
-# %>%
-#   group_by(ref, group) %>%
-#   mutate(
-#     group_length = round(sum(length)),
-#     mean_cycling_potential = round(weighted.mean(cycling_potential, length, na.rm = TRUE)),
-#     mean_width = round(weighted.mean(width, length, na.rm = TRUE)),
-#     majority_spare_lane = sum(length[spare_lane]) > sum(length[!spare_lane]),
-#     main_road_name = names(table(name))[which.max(table(name))],
-#     width_status = case_when(
-#       mean_width >= 10 ~ "Width >= 10 m",
-#       spare_lane ~ "Spare lane"
-#       # spare_lane & mean_width > 10 ~ "Spare lane & width > 10 m"
-#     )
-#   ) %>% 
-#   ungroup() %>% 
-# filter(group_length > min_grouped_length) 
-# mapview::mapview(r_linestrings, zcol = "group_length", lwd = 5)
-
 # Roads with no ref -------------------------------------------------------
 
[…]

### 2020-05-28 code/build.R
Reinstate roads with spare lanes

@@ -289,13 +289,14 @@ rg_new = do.call(rbind, rg_list)
 # mapview::mapview(rg_new)
 
 # Only keep groups of sufficient width and cycling potential
-rg_new2= rg_new %>% 
+rg_new2 = rg_new %>% 
   group_by(ig, group, ref) %>%
   mutate(
     mean_width = round(weighted.mean(width, length, na.rm = TRUE)),
-    mean_cycling_potential = round(weighted.mean(cycling_potential, length, na.rm = TRUE))
+    mean_cycling_potential = round(weighted.mean(cycling_potential, length, na.rm = TRUE)),
+    spare_lane = sum(length[spare_lane]) > sum(length[!spare_lane])
     ) %>%
-  filter(mean_width >= 10) %>%
+  filter(mean_width >= 10 | spare_lane) %>%
   filter(mean_cycling_potential >= min_grouped_cycling_potential) %>%
   ungroup()
 # mapview::mapview(rg_new2)

### 2020-05-29 code/index.Rmd
Tweak build

@@
 ```{r}
-dn = "popupCycleways/v0.2" # dirctory name
+dn = "popupCycleways/v0.3" # dirctory name
 built_regions = list.files(dn)
@@
 ```{r, eval=FALSE}
-
-sample_to_build_regex = "west of|west y|west mid|hereford|nott|leic|manc|liv|cam"
-region_names_to_build = regions %>% 
-  filter(grepl(pattern = sample_to_build_regex, x = Name, ignore.case = TRUE)) %>% 
-  pull(Name)
 dir.create(dn)
+region_names_to_build = regions$Name
+# sample_to_build_regex = "west of|west y|west mid|hereford|nott|leic|manc|liv|cam"
+# region_names_to_build = regions %>% 
+#   filter(grepl(pattern = sample_to_build_regex, x = Name, ignore.case = TRUE)) %>% 
+#   pull(Name)
 # region_names_to_build = c("West Yorkshire", "Nottingham")
-rmarkdown::render("code/index.Rmd", output_dir = dn)
 # region_names_to_build = c("Greater London", "Cambridgeshire")
@@
 rn = "West of England"
-for(rn in region_names_to_build) {
+for(rn in region_names_to_build[22:length(region_names_to_build)]) {
   region = regions %>% filter(Name == rn)
@@
 }
+rmarkdown::render("code/index.Rmd", output_dir = dn)
 ```

### 2020-05-29 code/build.R
Use variable min_cycling_potential params, fix #54

@@ -67,8 +67,15 @@ list2env(p, envir = .GlobalEnv)
 is_city = FALSE 
 pct_dist_within = 50 
 r_min_width_highlighted = 10
-min_cycling_potential = 5
-min_grouped_cycling_potential = 30
+# min_cycling_potential = 5
+# min_grouped_cycling_potential = 30
+# min_cycling_potential = quantile(x = rj$cycling_potential, probs = 0.5)
+# min_grouped_cycling_potential = quantile(x = rj$cycling_potential, probs = 0.8)
+# Select cycling potential thresholds based on max potential
+# 1000 for West Yorkshire, 200 for Rutland
+max_cycling_potential_99th = quantile(x = rj$cycling_potential, probs = 0.99)
+min_cycling_potential = round(max_cycling_potential_99th / 250)
+min_grouped_cycling_potential = round(max_cycling_potential_99th / 25)
 n_top_roads = (1 + round(nrow(rj) / 80000)) * 10
 
 # buffers -----------------------------------------------------------------
@@ -335,7 +342,7 @@ r_lanes_grouped2 = rg_new4 %>%
   group_by(ref, group, ig, long_named_section) %>% 
   summarise(
     name = case_when(
-      length(table(name)) > 4 ~ "Unnammed road",
+      length(table(name)) > 4 ~ "Unnamed road",
       names(table(name))[which.max(table(name))] != "" ~
         names(table(name))[which.max(table(name))],
       names(table(ref))[which.max(table(ref))] != "" ~
@@ -376,7 +383,8 @@ summary(r_lanes_joined$proportion_on_cycleway) # all between 0 and 1
 
 # we need to add in all segments within the grey key roads, and usethe combined dataset to pick the top routes 
 r_lanes_top = r_lanes_joined %>%
-  filter(name != "" & ref != "") %>%
+  # filter(name != "Unnamed road" & ref != "") %>%
+  filter(name != "Unnamed road") %>%
   filter(group_length > min_grouped_length) %>%
   filter(mean_cycling_potential > min_grouped_cycling_potential) %>% 
   filter(!grepl(pattern = regexclude, name, ignore.case = TRUE)) %>% 
@@ -428,6 +436,10 @@ top_routes = r_lanes_final %>% filter(Status == labels[1])
 spare_lane_groups = r_lanes_final %>% filter(Status == labels[2])
[…]

### 2020-05-29 code/index.Rmd
Update landing page map

@@
 pvars = c("Name", "Level")
-tm_shape(regions) + tm_polygons("Level", popup.vars = pvars) + tm_shape(regions_centroids) + tm_text("Name", size = 0.7)
+tm_shape(regions) + tm_polygons("Level", popup.vars = pvars) 
+  # + tm_shape(regions_centroids) + tm_text("Name", size = 0.7)
 ```
@@
 rn = "West of England"
-for(rn in region_names_to_build[22:length(region_names_to_build)]) {
+for(rn in region_names_to_build[3:length(region_names_to_build)]) {
   region = regions %>% filter(Name == rn)

### 2020-05-29 code/index.Rmd

@@
 
-3) Top routes
-
-- these are the top routes for pop-up cycleways.
-These are classified by "km cycled", calculated as the length of the route multiplied by the mean cycling potential.
-All of these road segments will either have more than one vehicle lane in either direction, or be at least 10m in width.
-The routes are listed in the table below the map, so you can see more information on them.
-The number of top routes selected is normally 20, with some variation by region.
-
-4) Spare lane(s)
+3) Spare lane(s)
 
@@
 
-5) Width >= 10m
+4) Width >= 10m
 
 - these routes are at least 10m in width and have a cycling potential above the minimum threshold.
+
+5) Top routes
+
+- these are the top routes for pop-up cycleways according to our analysis.
+These roads have either a spare lane or width of at least 10m, and are ranked according to their mean cycling potential.
+For routes that are shorter than 1km in length, we use 'km_cycled' instead of mean cycling potential - this is simply the cycling potential multiplied by the length of the route.
+We used certain eligibility criteria for the top routes, for example, a road which already has an existing cycleway alongside the majority of its length cannot be a top route.
+Top routes are listed in the table below the map, so you can see more information on them.
+The number of top routes selected is normally 20, with some variation by region.
 

### 2020-05-29 code/index.Rmd
Minor updates to build

--- 2020-05-29 code/index.Rmd
@@
 region_names_to_build = regions$Name
+region_names_to_build = regions$Name[!regions$html_name %in% built_regions]
 # sample_to_build_regex = "west of|west y|west mid|hereford|nott|leic|manc|liv|cam"
@@
 rn = "West of England"
-for(rn in region_names_to_build[3:length(region_names_to_build)]) {
+for(rn in region_names_to_build) {
   region = regions %>% filter(Name == rn)
@@
 rmarkdown::render("code/index.Rmd", output_dir = dn)
+browseURL(file.path(dn, "index.html"))
 ```

### 2020-05-29 code/build.R

@@ -441,7 +441,10 @@ if(nrow(spare_lanes) == 0) {
   spare_lanes = rg_new4 %>% top_n(n = 1, wt = width)
 }
 width_10m = r_lanes_final %>% filter(Status == labels[3])
-
+# edge case: there are no wide roads
+if(nrow(spare_lanes) == 0) {
+  spare_lanes = rg_new4 %>% filter(width > 10)
+}
 tmap_mode("view")
 m =
   tm_shape(key_network, name = "Key network") +

### 2020-05-29 code/build.R
Alter text and popups

@@ -39,8 +39,8 @@ if(!exists("s")) {
   # get pct data
   rnet_url = "https://github.com/npct/pct-outputs-national/raw/master/commute/lsoa/rnet_all.Rds"
   rnet_url_school = "https://github.com/npct/pct-outputs-national/raw/master/school/lsoa/rnet_all.Rds"
-  rnet_all = sf::st_as_sf(readRDS(url(rnet_url)))
-  rnet_all_school = sf::st_as_sf(readRDS(url(rnet_url_school)))
+  rnet_all = sf::st_as_sf(read_rds(url(rnet_url)))
+  rnet_all_school = sf::st_as_sf(read_rds(url(rnet_url_school)))
 }
 
 # local parameters --------------------------------------------------------
@@ -453,7 +453,7 @@ m =
            lwd = 3,
            # scale = 5,
            alpha = 1,
-           popup.vars = popup.vars
+           popup.vars = c("name", "ref", "maxspeed", "cycling_potential", "n_lanes")
            ) +
   tm_shape(width_10m, name = labels[2]) +
   tm_lines(legend.col.show = FALSE,
@@ -461,7 +461,8 @@ m =
            lwd = 3,
            # scale = 5,
            alpha = 1,
-           popup.vars = popup.vars) +
+           popup.vars = popup.vars
+           ) +
   tm_shape(top_routes, name = "Top routes") +
   tm_lines(legend.col.show = FALSE,
            col = cols_status[3], 

### 2020-05-29 code/index.Rmd
Alter text

--- 2020-05-29 code/index.Rmd
@@
 These roads have either a spare lane or width of at least 10m, and are ranked according to their mean cycling potential.
-For routes that are shorter than 1km in length, we use 'km_cycled' instead of mean cycling potential - this is simply the cycling potential multiplied by the length of the route.
-We used certain eligibility criteria for the top routes, for example, a road which already has an existing cycleway alongside the majority of its length cannot be a top route.
+For routes that are shorter than 1km in length, we use 'km_cycled' instead of mean cycling potential - this is simply the mean cycling potential multiplied by the length of the route.
 Top routes are listed in the table below the map, so you can see more information on them.
 The number of top routes selected is normally 20, with some variation by region.
+We used certain eligibility criteria for the top routes:
+
+- a road which already has an existing cycleway alongside the majority of its length cannot be a top route. 
+- unnamed roads cannot be a top route 
+- routes less than 500m in length cannot be a top route
 

### 2020-05-29 code/index.Rmd
Add toc

--- 2020-05-29 code/index.Rmd
@@
 title: "Rapid cycleway prioritisation tool"
-author: "University of Leeds"
-output: html_document
+author: "Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds"
+output: 
+  html_document: 
+    toc: yes
 ---
@@
 
-## PRELIMINARY RESULTS
+# NOTE: PRELIMINARY RESULTS
 
@@
 
-## Introduction
+# Introduction
 
@@
 
-### Map layers
+# Map layers
 
-The maps show five layers.
+The maps contain five layers of data that overlay the basemaps.
 
-1) Key network:
+## 1. Key network:
 
-- we have created this to represent the most important cycle corridors.
-For this layer, we do not take into account the width of the roads, but simply represent them as continuous corridors.
-This underscores the importance of ensuring that cycle routes are continuous and do not 'give up' at awkward junctions.
-To derive this layer, we have selected all road segments with cycle potential in the 85th percentile and above, which also belong to classified trunk roads, A roads or B roads.
+This layer is designed to represent a cohesive network of direct routes.
+It shows what a joined-up cycle network could look like if there were no constraints on road space, along some of the most important corridors.
+The layer takes no account spare space and is designed to support consideratioin of alternative options, e.g.
+one way streets, where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
+The layer was developed based on evidence on the importance of active transport network that are continuous and do not 'give up' at awkward junctions.
+The layer was created by selecting all road segments with cycle potential in the 85th percentile and above, which also belong to classified trunk roads, A roads or B roads.
 These segments are then joined together to form a continuous network.
 
-2) Existing cycleways (500m +)
+## 2. Existing cycleways
 
-- this layer shows existing cycleways.
+This layer shows existing cycleways.
 This may be helpful to view in connection with the other layers, because it can reveal where gaps in the existing cycle network could be filled by new infrastructure.
@@
 
-3) Spare lane(s)
+## 3. Spare lane(s)
 
-- these are routes that have more than one vehicle lane in either direction and have a cycling potential above the minimum threshold (which varies according to region as it needs to be higher in e.g.
+These are routes that have more than one vehicle lane in either direction and have a cycling potential above the minimum threshold (which varies according to region as it needs to be higher in e.g.
 Greater London than in rural areas).
 
-4) Width >= 10m
+## Width >= 10m
 
-- these routes are at least 10m in width and have a cycling potential above the minimum threshold.
+These routes are at least 10m in width and have a cycling potential above the minimum threshold.
 
-5) Top routes
+## Top routes
 
-- these are the top routes for pop-up cycleways according to our analysis.
+These are the top routes for pop-up cycleways according to our analysis.
 These roads have either a spare lane or width of at least 10m, and are ranked according to their mean cycling potential.
@@
 
-### Basemaps
+# Basemaps
 

### 2020-05-29 code/index.Rmd

--- 2020-05-29 code/index.Rmd
@@
 
-The maps show five layers.
+The maps show five layers. Click on any of these layers to see a pop-up box containing more information about the road in question.
 
@@
 These roads have either a spare lane or width of at least 10m, and are ranked according to their mean cycling potential.
-For routes that are shorter than 1km in length, we use 'km_cycled' instead of mean cycling potential - this is simply the cycling potential multiplied by the length of the route.
-We used certain eligibility criteria for the top routes, for example, a road which already has an existing cycleway alongside the majority of its length cannot be a top route.
+For routes that are shorter than 1km in length, we use 'km_cycled' instead of mean cycling potential - this is simply the mean cycling potential multiplied by the length of the route.
 Top routes are listed in the table below the map, so you can see more information on them.
 The number of top routes selected is normally 20, with some variation by region.
+We used certain eligibility criteria for the top routes:
 
-Click on any of these layers to see a pop-up box containing more information about the road in question.
+- a road which already has an existing cycleway alongside the majority of its length cannot be a top route. 
+- unnamed roads cannot be a top route 
+- routes less than 500m in length cannot be a top route
 

### 2020-05-29 code/index.Rmd

--- 2020-05-29 code/index.Rmd
@@
 
+Click on any of these layers to see a pop-up box containing more information about the road in question.
+
 ## 1. Key network:
@@
 These roads have either a spare lane or width of at least 10m, and are ranked according to their mean cycling potential. 
-For routes that are shorter than 1km in length, we use 'km_cycled' instead of mean cycling potential - this is simply the cycling potential multiplied by the length of the route.
-We used certain eligibility criteria for the top routes, for example, a road which already has an existing cycleway alongside the majority of its length cannot be a top route.
+For routes that are shorter than 1km in length, we use 'km_cycled' instead of mean cycling potential - this is simply the mean cycling potential multiplied by the length of the route. 
 Top routes are listed in the table below the map, so you can see more information on them. 
@@
 
-Click on any of these layers to see a pop-up box containing more information about the road in question.
+We used certain eligibility criteria for the top routes:
+
+- a road which already has an existing cycleway alongside the majority of its length cannot be a top route. 
+- unnamed roads cannot be a top route 
+- routes less than 500m in length cannot be a top route
 

### 2020-05-29 code/index.Rmd
Update intro

--- 2020-05-29 code/index.Rmd
@@
 
-The aim of this tool is to help reveal the locations that have the best potential for pop-up cycleways.
+The aim of this tool is to help reveal the locations that have the greatest potential for new cycleways.
 It should be used together with local knowledge, to prioritise where to install rapid measures to support safe travel as part of the COVID-19 response.
@@
 
-We have focused on roads which are sufficiently wide that part of the roadspace could potentially be reallocated to pop-up cycleways without preventing traffic movement.
+We have focused on roads which are sufficiently wide that part of the road space could potentially be reallocated to provide additional space for active transport without preventing traffic movement.
 In identifying these locations, we have not taken average traffic volumes into account.
@@
 
-We have used several types of data.
-The cycling potential of each road is estimated according the Propensity to Cycle Tool's Government Target scenario, which represents the government target to double cycling levels by 2025.
-Our figures include both travel to work and travel to school, and are based on 2011 Census data.
-Data on road widths and number of lanes is derived from the Cycling Infrastructure Prioritisation Toolkit.
+The cycling potential of each road is based on data from the Propensity to Cycle Tool (see [www.pct.bike](https://www.pct.bike/) for further details) and represents the number of one way cycle trips to work and school that could be expected when the [Government's aim of doubling cycling is met](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy).
+The scenario assumes cycling grows evenly nationwide accounting for distance and hilliness.
+Our figures include data on travel to work from 2011 Census data and travel to school. Data on road widths and number of lanes is derived from the Cycling Infrastructure Prioritisation Toolkit.
 
@@
 A further basemap option shows all existing cycleways, including on-road cycle lanes. Finally, a satellite image is also available as a basemap.
+
+# Known issues and limitations
+
+Road names ...
 

### 2020-05-29 code/build.R

@@ -400,7 +400,7 @@ r_lanes_top %>% sf::st_drop_geometry()
 
 # classify roads to visualise
 labels = c("Spare lane(s)", "Estimated width > 10m", "Top route")
-cycleways_name = "Existing cycleways (500 m+)"
+cycleways_name = "Existing off road cycleways"
 
 r_lanes_final = r_lanes_joined %>% 
   mutate(

### 2020-05-29 code/index.Rmd
Update landing page

@@
 Our figures include data on travel to work from 2011 Census data and travel to school. Data on road widths and number of lanes is derived from the Cycling Infrastructure Prioritisation Toolkit.
+
+To use the evidence generated for this project:
+
+- Type in the name of the local authority of interest into the search bar above the table below and click on the associated URL.
+- Explore the results in the map (you will find a link to a full screen map) and the table of 'top roads' which can be ranked in various ways with reference to descriptions of the map layers described below.
+- Interpret the results with care based on other sources of information and an understanding of the limitations of the analysis (see 'Known issues and limitations' below).
 
@@
 
-Road names ...
+<!-- We have developed this evidence base rapidly and -->
+
+- The 'top routes' highlighted in the map-based results were produced automatically based on a simple score of continuous candidate road length multiplied by cycling potential.
+For roads with a length of more than 1 km, the cycling potential alone is used to rank roads.
+**The results should be interpreted based on local knowledge and an understanding of the limitations of the data.
+The results are designed to complement not replace other factors such as feedback from the local community and long-term strategic cycle network plans.**
+- The estimates of cycling potential only include school and work travel.
+Accounting for increases in cycling for other trip purposes such as for shopping and leisure would increase cycling potential substantially.
+- The results rely on grouping roads by proximity, name and other variables.
+Currently some roads that are strong candidates for new cycleways may be missing from the results because of how the algorithm works (we are working on this issue).
+- The road network data was taken from the Cycling Infrastructure Prioritisation Toolkit (CyIPT) project which was completed in 2017, so may be out of date (we are looking at ways to update the data).
+- The results are based on the existing road network and does not flag potential routes that could be constructed.
+
+<!-- Like many projects that have been initiated in response to COVID-19 we have developed the analy -->
 

### 2020-05-29 code/index.Rmd
Add link to github

@@
 
-## 1. Key network:
+## 1. Key cycle routes
 
@@
 
-## 3. Spare lane(s)
-
-These are routes that have more than one vehicle lane in either direction and have a cycling potential above the minimum threshold (which varies according to region as it needs to be higher in e.g.
-Greater London than in rural areas).
-
-## Width >= 10m
-
-These routes are at least 10m in width and have a cycling potential above the minimum threshold.
-
-## Top routes
+## 3. Top routes
 
@@
 - routes less than 500m in length cannot be a top route
+
+## 3. Spare lane(s)
+
+These are routes that have more than one vehicle lane in either direction and have a cycling potential above the minimum threshold (which varies according to region as it needs to be higher in e.g.
+Greater London than in rural areas).
+
+## 4. Width >= 10m
+
+These routes are at least 10m in width and have a cycling potential above the minimum threshold.
 
@@
 <!-- Like many projects that have been initiated in response to COVID-19 we have developed the analy -->
+
+If you find any other issues with the results or have suggestions of ways to improve the results, please provide feedback on the [web form](https://urldefense.proofpoint.com/v2/url?u=https-3A__forms.office.com_Pages_ResponsePage.aspx-3Fid-3DqO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u&d=DwMCaQ&c=troKkvwivNn_CddsvWCHHPiPoFoTgTGIbXJULvYU158&r=DXeEYtdSYDBZN2DBwEgAHMmX5PQ_ob9r2MynFWta2mU&m=mkyUa3RIC9d5LSrHUqk_WdstQR1OQuTWyYLBecHFMm0&s=dD7GVcQbjZ2AcdCObSe-rNKmLLDBhu98ChFupIIrlJA&e=).
+If you are interested in the code underlying the project and would like to support the project with technical input, please provide feedback via the [GitHub issue tracker (requires a GitHub account)](https://github.com/cyipt/popupCycleways/issues).
 

### 2020-05-29 code/build.R
Tidy-up tmap code

@@ -423,14 +423,7 @@ summary(factor(r_lanes_final$Status))
 
 # Post-processing and set-up for map --------------------------------------
 
-popup.vars = c("name", "ref", "spare_lane", "Estimated width", "mean_cycling_potential", "length (m)")
-pvars_key = c("ref", "name", "highway_type", "cycling_potential", "n_lanes")
-key_network = key_network[pvars_key]
-
-# cols_status = c("blue", "orange", tmaptools::get_brewer_pal("Accent", n = 3)[3])
-cols_status = c("#B91F48", "#FF7F00", "blue")
 summary(r_lanes_final$Status)
-
 top_routes = r_lanes_final %>% filter(Status == labels[1])
 # idea: show spare lane segments not groups:
 spare_lane_groups = r_lanes_final %>% filter(Status == labels[2])
@@ -446,6 +439,21 @@ if(nrow(spare_lanes) == 0) {
   spare_lanes = rg_new4 %>% filter(width > 10)
 }
 tmap_mode("view")
+
+popup.vars = c(
+  "name",
+  "ref",
+  "spare_lane",
+  "Estimated width",
+  "mean_cycling_potential",
+  "length (m)"
+)
+pvars_key = c("ref", "name", "highway_type", "cycling_potential", "n_lanes")
+key_network = key_network[pvars_key]
+legend_labels = c("Key network", cycleways_name, labels[1], labels[2], labels[3])
+legend_colours = c("darkgrey", "darkgreen", cols_status)
+cols_status = c("#B91F48", "#FF7F00", "blue")
+
 m =
   tm_shape(key_network, name = "Key network") +
   tm_lines(lwd = 5, col = "darkgrey", popup.vars = pvars_key) +
@@ -454,9 +462,9 @@ m =
   tm_lines(legend.col.show = FALSE,
            col = cols_status[1], 
            lwd = 3,
-           # scale = 5,
            alpha = 1,
-           popup.vars = c("name", "ref", "maxspeed", "cycling_potential", "n_lanes")
+           popup.vars = c("name", "ref", "maxspeed", "cycling_potential", "n_lanes"),
+           group = "layers"
            ) +
   tm_shape(width_10m, name = labels[2]) +
   tm_lines(legend.col.show = FALSE,
@@ -464,7 +472,9 @@ m =
            lwd = 3,
            # scale = 5,
            alpha = 1,
-           popup.vars = popup.vars) +
[…]

### 2020-05-29 code/build.R
Remove commented lines

@@ -425,7 +425,7 @@ summary(factor(r_lanes_final$Status))
 
 summary(r_lanes_final$Status)
 top_routes = r_lanes_final %>% filter(Status == labels[1])
-# idea: show spare lane segments not groups:
+# show spare lane segments not groups:
 spare_lane_groups = r_lanes_final %>% filter(Status == labels[2])
 spare_lane_segments = rg_new4[spare_lane_groups, , op = sf::st_within]
 spare_lanes = spare_lane_segments %>% filter(spare_lane)
@@ -470,7 +470,6 @@ m =
   tm_lines(legend.col.show = FALSE,
            col = cols_status[2], 
            lwd = 3,
-           # scale = 5,
            alpha = 1,
            popup.vars = popup.vars,
            group = "layers"
@@ -479,7 +478,6 @@ m =
   tm_lines(legend.col.show = FALSE,
            col = cols_status[3], 
            lwd = 5,
-           # scale = 5,
            alpha = 1,
            popup.vars = popup.vars
   ) +

### 2020-05-29 code/build.R
Close #63

@@ -451,14 +451,14 @@ popup.vars = c(
 pvars_key = c("ref", "name", "highway_type", "cycling_potential", "n_lanes")
 key_network = key_network[pvars_key]
 legend_labels = c("Key network", cycleways_name, labels[1], labels[2], labels[3])
-legend_colours = c("darkgrey", "darkgreen", cols_status)
+legend_colours = c(cols_status, "darkgrey", "darkgreen")
 cols_status = c("#B91F48", "#FF7F00", "blue")
 
 m =
   tm_shape(key_network, name = "Key network") +
   tm_lines(lwd = 5, col = "darkgrey", popup.vars = pvars_key) +
   tm_shape(cycleways, name = cycleways_name) + tm_lines(popup.vars = c("surface", "name", "osm_id"), col = "darkgreen", lwd = 1.3) +
-  tm_shape(spare_lanes, name = labels[1]) +
+  tm_shape(spare_lanes, name = labels[2]) +
   tm_lines(legend.col.show = FALSE,
            col = cols_status[1], 
            lwd = 3,
@@ -466,7 +466,7 @@ m =
            popup.vars = c("name", "ref", "maxspeed", "cycling_potential", "n_lanes"),
            group = "layers"
            ) +
-  tm_shape(width_10m, name = labels[2]) +
+  tm_shape(width_10m, name = labels[3]) +
   tm_lines(legend.col.show = FALSE,
            col = cols_status[2], 
            lwd = 3,

### 2020-05-29 code/build.R
Fix labels (fix #63 mkII)

@@ -450,9 +450,9 @@ popup.vars = c(
 )
 pvars_key = c("ref", "name", "highway_type", "cycling_potential", "n_lanes")
 key_network = key_network[pvars_key]
-legend_labels = c("Key network", cycleways_name, labels[1], labels[2], labels[3])
-legend_colours = c(cols_status, "darkgrey", "darkgreen")
-cols_status = c("#B91F48", "#FF7F00", "blue")
+legend_labels = c("Key network", cycleways_name, labels[3], labels[1], labels[2])
+cols_status = c("blue", "#B91F48", "#FF7F00")
+legend_colours = c("darkgrey", "darkgreen", cols_status)
 
 m =
   tm_shape(key_network, name = "Key network") +
@@ -460,7 +460,7 @@ m =
   tm_shape(cycleways, name = cycleways_name) + tm_lines(popup.vars = c("surface", "name", "osm_id"), col = "darkgreen", lwd = 1.3) +
   tm_shape(spare_lanes, name = labels[2]) +
   tm_lines(legend.col.show = FALSE,
-           col = cols_status[1], 
+           col = cols_status[3], 
            lwd = 3,
            alpha = 1,
            popup.vars = c("name", "ref", "maxspeed", "cycling_potential", "n_lanes"),
@@ -476,7 +476,7 @@ m =
            ) +
   tm_shape(top_routes, name = "Top routes") +
   tm_lines(legend.col.show = FALSE,
-           col = cols_status[3], 
+           col = cols_status[1], 
            lwd = 5,
            alpha = 1,
            popup.vars = popup.vars
@@ -485,7 +485,8 @@ m =
   tm_add_legend(type = "fill", labels = legend_labels, col = legend_colours) +
   tm_scale_bar() 
 # m
-m_leaflet = tmap_leaflet(m) %>% leaflet::hideGroup(c("layers"))
+m_leaflet = tmap_leaflet(m)
+# %>% leaflet::hideGroup(c("layers"))
 # htmlwidgets::saveWidget(m_leaflet, "/tmp/m.html")
 # system("ls -hal /tmp/m.html") # 15 MB for West Yorkshire
 # browseURL("/tmp/m.html")

### 2020-05-29 code/build.R
Move cycleway layer up 2 places

@@ -457,7 +457,6 @@ legend_colours = c("darkgrey", "darkgreen", cols_status)
 m =
   tm_shape(key_network, name = "Key network") +
   tm_lines(lwd = 5, col = "darkgrey", popup.vars = pvars_key) +
-  tm_shape(cycleways, name = cycleways_name) + tm_lines(popup.vars = c("surface", "name", "osm_id"), col = "darkgreen", lwd = 1.3) +
   tm_shape(spare_lanes, name = labels[2]) +
   tm_lines(legend.col.show = FALSE,
            col = cols_status[3], 
@@ -474,6 +473,7 @@ m =
            popup.vars = popup.vars,
            group = "layers"
            ) +
+  tm_shape(cycleways, name = cycleways_name) + tm_lines(popup.vars = c("surface", "name", "osm_id"), col = "darkgreen", lwd = 1.3) +
   tm_shape(top_routes, name = "Top routes") +
   tm_lines(legend.col.show = FALSE,
            col = cols_status[1], 

### 2020-05-29 code/build.R
Untick layers, close #55

@@ -399,7 +399,7 @@ nrow(r_lanes_top)
 r_lanes_top %>% sf::st_drop_geometry()
 
 # classify roads to visualise
-labels = c("Spare lane(s)", "Estimated width > 10m", "Top route")
+labels = c("Top route", "Spare lane(s)", "Estimated width > 10m")
 cycleways_name = "Existing off road cycleways"
 
 r_lanes_final = r_lanes_joined %>% 
@@ -428,15 +428,24 @@ top_routes = r_lanes_final %>% filter(Status == labels[1])
 # show spare lane segments not groups:
 spare_lane_groups = r_lanes_final %>% filter(Status == labels[2])
 spare_lane_segments = rg_new4[spare_lane_groups, , op = sf::st_within]
-spare_lanes = spare_lane_segments %>% filter(spare_lane)
+spare_lanes = spare_lane_segments %>%
+  filter(spare_lane) %>% 
+  mutate(Status = labels[2])
 # edge case: there are no spare lanes
 if(nrow(spare_lanes) == 0) {
   spare_lanes = rg_new4 %>% top_n(n = 1, wt = width)
 }
-width_10m = r_lanes_final %>% filter(Status == labels[3])
+# show wide segments not groups:
+# width_10m = r_lanes_final %>% filter(Status == labels[1])
+wide_lane_groups = r_lanes_final %>% filter(Status == labels[3])
+wide_lane_segments = rg_new4[wide_lane_groups, , op = sf::st_within]
+wide_lanes = wide_lane_segments %>%
+  filter(width >= 10 & !idGlobal %in% spare_lanes$idGlobal) %>% 
+  mutate(Status = labels[3])
+spare_wide_lanes = rbind(wide_lanes, spare_lanes)
 # edge case: there are no wide roads
 if(nrow(spare_lanes) == 0) {
-  spare_lanes = rg_new4 %>% filter(width > 10)
+  spare_lanes = rg_new4 %>% filter(width > 10 )
 }
 tmap_mode("view")
 
@@ -450,7 +459,7 @@ popup.vars = c(
 )
 pvars_key = c("ref", "name", "highway_type", "cycling_potential", "n_lanes")
 key_network = key_network[pvars_key]
-legend_labels = c("Key network", cycleways_name, labels[3], labels[1], labels[2])
+legend_labels = c("Key network", cycleways_name, labels[1], labels[2], labels[3])
 cols_status = c("blue", "#B91F48", "#FF7F00")
[…]

### 2020-05-29 code/build.R
Add boundaries

@@ -28,6 +28,8 @@ if(!exists("s")) {
   # read-in national data ---------------------------------------------------
   # see preprocess.R for data origins
   if(!exists("regions")) regions = readRDS("regions_dft.Rds")
+  lads_all = readRDS("lads.Rds")
+  lads_all_centroids = sf::st_centroid(lads_all)
   rj_all = readRDS("rj.Rds")
   region_names = regions$Name
   # hospitals:
@@ -54,6 +56,8 @@ if(region_name == "Nottingham") {
 } else {
   region = regions %>% filter(Name == region_name)
 }
+lads_centroids = lads_all_centroids[region, ]
+lads = lads_all %>% filter(Name %in% lads_centroids$Name)
 # time consuming: ~1 minute
 rj = rj_all[region, ]
 # h_city = hsf[region, ]
@@ -459,12 +463,13 @@ popup.vars = c(
 )
 pvars_key = c("ref", "name", "highway_type", "cycling_potential", "n_lanes")
 key_network = key_network[pvars_key]
-legend_labels = c("Key network", cycleways_name, labels[1], labels[2], labels[3])
+legend_labels = c("Key cycle routes", cycleways_name, labels[1], labels[2], labels[3])
 cols_status = c("blue", "#B91F48", "#FF7F00")
 legend_colours = c("darkgrey", "darkgreen", cols_status)
 
 m =
-  tm_shape(key_network, name = "Key network") +
+  tm_shape(lads) + tm_borders() +
+  tm_shape(key_network, name = "Key cycle routes") +
   tm_lines(lwd = 5, col = "darkgrey", popup.vars = pvars_key) +
   tm_shape(spare_lanes, name = labels[2]) +
   tm_lines(legend.col.show = FALSE,

### 2020-05-29 code/index.Rmd
Update presentation of results

@@
 region_names_to_build = regions$Name
-region_names_to_build = regions$Name[!regions$html_name %in% built_regions]
+# region_names_to_build = regions$Name[!regions$html_name %in% built_regions]
 # sample_to_build_regex = "west of|west y|west mid|hereford|nott|leic|manc|liv|cam"
@@
 rn = "West of England"
-for(rn in region_names_to_build) {
+for(rn in region_names_to_build[50:nrow(regions)]) {
   region = regions %>% filter(Name == rn)

### 2020-05-29 code/region_index.Rmd
Update presentation of results

@@
 
-```{r}
+The map shows a network of key roads with high cycling potential (grey), existing off road cycleways (and shared space paths) according to OpenStreetMap data (green) and top road sections with high cycling potential and spare lanes or estimated width above 10m (blue).
+To see additional layers showing road sections with 'spare lanes' and width greater than 10 m, click on the layers icon in the top left, just below the +/- buttons on the map.
+
+```{r, out.width="100%"}
 knitr::include_url("m.html")
@@
 
+The table below shows the top `r nrow(res_table)` roads, in terms of cycling potential and spare space criteria outlined on the landing page.
+The length refers to the continuous length of road.
+Cycling potential represents the average *combined number of commute and school* cycle trips that would use the road each morning under the Government Target scenario in the PCT.
+'Length * potential' is the length of the road multiplied by its average cycling potential.
+
 ```{r}

### 2020-05-29 code/build.R

@@ -468,7 +468,7 @@ cols_status = c("blue", "#B91F48", "#FF7F00")
 legend_colours = c("darkgrey", "darkgreen", cols_status)
 
 m =
-  tm_shape(lads) + tm_borders() +
+  tm_shape(lads, name = "Local authority district boundaries") + tm_borders() +
   tm_shape(key_network, name = "Key cycle routes") +
   tm_lines(lwd = 5, col = "darkgrey", popup.vars = pvars_key) +
   tm_shape(spare_lanes, name = labels[2]) +
@@ -508,7 +508,13 @@ m_leaflet = tmap_leaflet(m) %>% leaflet::hideGroup(labels[2:3])
 
 res_table = r_lanes_top %>% 
   sf::st_drop_geometry() %>% 
-  select(name, ref, length = group_length, mean_cycling_potential, km_cycled) 
+  select(
+    Name = name,
+    Reference = ref,
+    `Continuous length (m)` = group_length,
+    `Cycling potential` = mean_cycling_potential,
+    `Length * potential (km)` = km_cycled
+    ) 
 res_table
 # knitr::kable(res_table, caption = "The top 10 candidate roads for space reallocation for pop-up active transport infrastructure according to methods presented in this paper.", digits = 0)
 

### 2020-05-30 code/build.R
Fix edge case where la not in region

@@ -58,6 +58,9 @@ if(region_name == "Nottingham") {
 }
 lads_centroids = lads_all_centroids[region, ]
 lads = lads_all %>% filter(Name %in% lads_centroids$Name)
+if(nrow(lads) == 0) {
+  lads = region
+}
 # time consuming: ~1 minute
 rj = rj_all[region, ]
 # h_city = hsf[region, ]

### 2020-05-30 code/index.Rmd
Fix bug in build for edge case

@@
 rn = "West of England"
-for(rn in region_names_to_build) {
+for(rn in region_names_to_build[38:nrow(regions)]) {
   region = regions %>% filter(Name == rn)

### 2020-05-30 code/build.R

@@ -440,7 +440,9 @@ spare_lanes = spare_lane_segments %>%
   mutate(Status = labels[2])
 # edge case: there are no spare lanes
 if(nrow(spare_lanes) == 0) {
-  spare_lanes = rg_new4 %>% top_n(n = 1, wt = width)
+  spare_lanes = rg_new4 %>%
+    top_n(n = 1, wt = width) %>% 
+    mutate(Status = labels[2]) 
 }
 # show wide segments not groups:
 # width_10m = r_lanes_final %>% filter(Status == labels[1])
@@ -449,11 +451,14 @@ wide_lane_segments = rg_new4[wide_lane_groups, , op = sf::st_within]
 wide_lanes = wide_lane_segments %>%
   filter(width >= 10 & !idGlobal %in% spare_lanes$idGlobal) %>% 
   mutate(Status = labels[3])
-spare_wide_lanes = rbind(wide_lanes, spare_lanes)
-# edge case: there are no wide roads
-if(nrow(spare_lanes) == 0) {
-  spare_lanes = rg_new4 %>% filter(width > 10 )
+# edge case: there are no wide lanes
+if(nrow(wide_lanes) == 0) {
+  wide_lanes = rg_new4 %>%
+    top_n(n = 1, wt = width) %>% 
+    mutate(Status = labels[3]) 
 }
+spare_wide_lanes = rbind(wide_lanes, spare_lanes)
+
 tmap_mode("view")
 
 popup.vars = c(

### 2020-05-30 code/index.Rmd
Remove isles of scilly

--- 2020-05-29 code/index.Rmd
@@
 built_regions = list.files(dn)
+regions = regions %>% 
+  filter(stringr::str_detect(string = Name, "Scilly", negate = TRUE))
 regions %>%
   sf::st_drop_geometry() %>%
-  filter(html_name %in% built_regions) %>% 
+  filter(html_name %in% built_regions) 
   select(Name, Level, url) %>% 

### 2020-06-01 code/index.Rmd
Change name of Key cycle routes to Cohesive network
see #55 for context

command to make these changes:

sd 'Key cycle routes' 'Cohesive network' README.Rmd
README.md
README_files
article.R
article.Rmd
cities/birmingham.Rmd
cities/bristol.Rmd
cities/cambridge.Rmd
cities/leeds.Rmd
cities/leicester.Rmd
cities/liverpool.Rmd
cities/london.Rmd
cities/manchester.Rmd
cities/newcastle.Rmd
cities/sheffield.Rmd
code/article-methods-1.R
code/article-methods-cambridge-test.R
code/boundaries.R
code/boundaries2.R
code/build.R
code/build.R.orig
code/build.Rmd
code/build.spin.R
code/build.spin.Rmd
code/builda

@@
 
-The aim of this tool is to help reveal the locations that have the greatest potential for new cycleways.
-It should be used together with local knowledge, to prioritise where to install rapid measures to support safe travel as part of the COVID-19 response.
-We anticipate that this may be used in the development of a range of temporary, experimental and longer term measures to make the streets safer for us all, as new travel patterns and behaviours emerge in the coming weeks and months.
+This tool helps identify road sections for road space reallocation and other interventions to create new cycleways.
+Used with local knowledge, it provides a strong evidence base to support planning of rapid measures to support safe travel as part of the COVID-19 response.
+The evidence may be used to support a range of temporary, experimental and permanent treatments that will enable shift to walking and cycling in the coming weeks and months, to help overcome capacity constraints on public transport modes and prevent gridlock as the lockdown eases.
 
@@
 
-We have focused on roads which are sufficiently wide that part of the road space could potentially be reallocated to provide additional space for active transport without preventing traffic movement.
-In identifying these locations, we have not taken average traffic volumes into account.
-The pandemic has had a severe impact on movement and traffic flows are currently below normal levels.
-Instead, we simply identify streets that have multiple vehicle lanes or a carriageway width of greater than 10m.
+The focus is on roads the either have a spare lane or that have an estimated width of above 10 m, *and* high cycling potential.
+The 10 m width is an indicator of road conditions and larger widths, of 15+ m, may be required for road space reallocation without other changes such as making roads one way for motor traffic or removing central markings as described in , depending on factors such as on street parking levels.
+The work is based evidence that road space can be reallocated without causing congestion at a time when traffic levels, particularly during the rush hours, are greatly reduced.
 
@@
 
+# How to use the tool
+
 To use the evidence generated for this project:
@@
 - Type in the name of the local authority of interest into the search bar above the table below and click on the associated URL.
-- Explore the results in the map (you will find a link to a full screen map) and the table of 'top roads' which can be ranked in various ways with reference to descriptions of the map layers described below.
+- Explore the results in the map (you will find a link to a full screen map) by panning, zooming and clicking on the layers.
+See additional layers by clicking on the layers icon in the top left of the map.
+- Rank the 'top roads' in various ways (including by length and cycling potential) to support prioritisation of road space reallocation schemes.
 - Interpret the results with care based on other sources of information and an understanding of the limitations of the analysis (see 'Known issues and limitations' below).
+- Refer to the description of the map layers below when exploring the results and complete the [web form](https://urldefense.proofpoint.com/v2/url?u=https-3A__forms.office.com_Pages_ResponsePage.aspx-3Fid-3DqO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u&d=DwMCaQ&c=troKkvwivNn_CddsvWCHHPiPoFoTgTGIbXJULvYU158&r=DXeEYtdSYDBZN2DBwEgAHMmX5PQ_ob9r2MynFWta2mU&m=mkyUa3RIC9d5LSrHUqk_WdstQR1OQuTWyYLBecHFMm0&s=dD7GVcQbjZ2AcdCObSe-rNKmLLDBhu98ChFupIIrlJA&e=) with any feedback.
 
@@
 
-## 3. Spare lane(s)
+## 3. Spare lane(s) (not shown be default)
 
@@
   sf::st_drop_geometry() %>%
-  filter(html_name %in% built_regions) 
+  filter(html_name %in% built_regions) %>% 
   select(Name, Level, url) %>% 

### 2020-06-01 code/boundaries2.R

@@ -112,3 +112,5 @@ pvars = c("url", "Name")
 tm_shape(regions_dft) + tm_polygons("Level", popup.vars = pvars) + tm_shape(regions_dft_centroids) + tm_text("Name", size = 0.7)
 
 saveRDS(regions_dft, "regions_dft.Rds")
+piggyback::pb_upload("regions_dft.Rds", "cyipt/cyipt-phase-1-data")
+piggyback::pb_upload("lads.Rds", "cyipt/cyipt-phase-1-data")

### 2020-06-01 code/index.Rmd
keeping groups

@@
 
-Please provide feedback on these preliminary results on this [web form (estimated time to complete: 5 minites)](https://urldefense.proofpoint.com/v2/url?u=https-3A__forms.office.com_Pages_ResponsePage.aspx-3Fid-3DqO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u&d=DwMCaQ&c=troKkvwivNn_CddsvWCHHPiPoFoTgTGIbXJULvYU158&r=DXeEYtdSYDBZN2DBwEgAHMmX5PQ_ob9r2MynFWta2mU&m=mkyUa3RIC9d5LSrHUqk_WdstQR1OQuTWyYLBecHFMm0&s=dD7GVcQbjZ2AcdCObSe-rNKmLLDBhu98ChFupIIrlJA&e=).
+Please provide feedback on these preliminary results on this [web form (estimated time to complete: 5 minutes)](https://urldefense.proofpoint.com/v2/url?u=https-3A__forms.office.com_Pages_ResponsePage.aspx-3Fid-3DqO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u&d=DwMCaQ&c=troKkvwivNn_CddsvWCHHPiPoFoTgTGIbXJULvYU158&r=DXeEYtdSYDBZN2DBwEgAHMmX5PQ_ob9r2MynFWta2mU&m=mkyUa3RIC9d5LSrHUqk_WdstQR1OQuTWyYLBecHFMm0&s=dD7GVcQbjZ2AcdCObSe-rNKmLLDBhu98ChFupIIrlJA&e=).
 
@@
 
-## 1. Key cycle routes
+## 1. Cohesive network
 
@@
 
-## 3. Spare lane(s) (not shown be default)
+## 3. Spare lane(s) (not shown by default)
 
@@
 
-## 4. Width >= 10m
+## 4. Width >= 10m (not shown by default)
 

### 2020-06-01 code/build.R

@@ -252,41 +252,48 @@ r_linestrings_with_ref = r_lanes_all %>%
 
 # Roads with no ref -------------------------------------------------------
 
+# Remove segments with cycling potential below 30 (to prevent side street segments)
 min_cycling_potential_without_ref = 30
 r_linestrings_without_ref =  r_lanes_all %>%
   filter(ref == "") %>% 
   filter(cycling_potential >= min_cycling_potential_without_ref)
+
+# Put into groups, using a 20m buffer (stricter than for roads with a ref, to prevent groups covering multiple streets)
 r_linestrings_without_ref_buff = geo_buffer(shp = r_linestrings_without_ref, dist = 20)
 touching_list = st_intersects(r_linestrings_without_ref_buff)
 g = igraph::graph.adjlist(touching_list)
 components = igraph::components(g)
-r_linestrings_without_ref$nogroup = components$membership
+r_linestrings_without_ref$group2 = components$membership
 
-# mapview::mapview(r_linestrings_without_ref["nogroup"], lwd = 3)
+# mapview::mapview(r_linestrings_without_ref["group2"], lwd = 3)
 
-# Remove groups with length <300m (20m buffer). (these are groups consisting purely of road segments with no ref) this is stricter than for roads with refs, because otherwise too many short segments are picked up. 
+# Remove groups with length <300m (20m buffer). (these are groups consisting purely of road segments with no ref) this is stricter than for roads with refs, because otherwise too many short segments are picked up. # Remove groups without sufficient width/lanes or cycle potential (the same as for roads with a ref)
 r_linestrings_without_ref2 = r_linestrings_without_ref %>%
-  group_by(nogroup) %>%
+  group_by(group2) %>%
   mutate(
-    group_length = round(sum(length)),
-    mean_cycling_potential = round(weighted.mean(cycling_potential, length, na.rm = TRUE))
+    group2_length = round(sum(length)),
+    mean_cycling_potential = round(weighted.mean(cycling_potential, length, na.rm = TRUE)),
+    mean_width = round(weighted.mean(width, length, na.rm = TRUE)),
+    majority_spare_lane = sum(length[spare_lane]) > sum(length[!spare_lane])
     ) %>% 
[…]

### 2020-06-01 code/build.R
road naming

@@ -379,7 +379,9 @@ r_lanes_grouped2 = rg_new4 %>%
       TRUE ~ case_when(
         names(table(ref))[which.max(table(ref))] != "" ~
           names(table(ref))[which.max(table(ref))],
-        TRUE ~ "Unnamed road")
+        TRUE ~ case_when(
+          names(desc(table(ref)))[2] != "NA" ~ paste("(",names(desc(table(ref)))[2],")"),
+          TRUE ~ "Unnamed road"))
       ), 
     group_length = round(sum(length)),
     mean_cycling_potential = round(weighted.mean(cycling_potential, length, na.rm = TRUE)),

### 2020-06-01 code/index.Rmd
Fix typos in index

@@
 
-Please provide feedback on these preliminary results on this [web form (estimated time to complete: 5 minutes)](https://urldefense.proofpoint.com/v2/url?u=https-3A__forms.office.com_Pages_ResponsePage.aspx-3Fid-3DqO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u&d=DwMCaQ&c=troKkvwivNn_CddsvWCHHPiPoFoTgTGIbXJULvYU158&r=DXeEYtdSYDBZN2DBwEgAHMmX5PQ_ob9r2MynFWta2mU&m=mkyUa3RIC9d5LSrHUqk_WdstQR1OQuTWyYLBecHFMm0&s=dD7GVcQbjZ2AcdCObSe-rNKmLLDBhu98ChFupIIrlJA&e=).
+Please provide feedback on these preliminary results on this [web form (estimated time to complete: 5 minutes)](https://forms.office.com/Pages/ResponsePage.aspx?id=qO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u).
 
@@
 
-The focus is on roads the either have a spare lane or that have an estimated width of above 10 m, *and* high cycling potential.
-The 10 m width is an indicator of road conditions and larger widths, of 15+ m, may be required for road space reallocation without other changes such as making roads one way for motor traffic or removing central markings as described in , depending on factors such as on street parking levels.
-The work is based evidence that road space can be reallocated without causing congestion at a time when traffic levels, particularly during the rush hours, are greatly reduced.
+The focus is on roads that have 'spare space' (defined as those with a spare lane or that have an estimated width of above 10 m) and high cycling potential.
+The 10 m width is an indicator of road conditions and larger widths, of 15+ m, may be required for road space reallocation without other changes such as making roads one way for motor traffic or removing central markings as described in a paper by Stella Shackel and John Parkin ([2014](https://doi.org/10.1016/j.aap.2014.08.015)).
+Which roads to prioritise and the most effective interventions will also depend on factors such as on street parking levels not present in the analysis.
 
@@
 Our figures include data on travel to work from 2011 Census data and travel to school. Data on road widths and number of lanes is derived from the Cycling Infrastructure Prioritisation Toolkit.
+
+Overall the evidence presented in the tool is designed to encourage bold interventions where more space for active modes is most needed.
+The work is based evidence that road space can be reallocated without causing congestion at a time when traffic levels, particularly during the rush hours, are greatly reduced.
 
@@
 
-If you find any other issues with the results or have suggestions of ways to improve the results, please provide feedback on the [web form](https://urldefense.proofpoint.com/v2/url?u=https-3A__forms.office.com_Pages_ResponsePage.aspx-3Fid-3DqO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u&d=DwMCaQ&c=troKkvwivNn_CddsvWCHHPiPoFoTgTGIbXJULvYU158&r=DXeEYtdSYDBZN2DBwEgAHMmX5PQ_ob9r2MynFWta2mU&m=mkyUa3RIC9d5LSrHUqk_WdstQR1OQuTWyYLBecHFMm0&s=dD7GVcQbjZ2AcdCObSe-rNKmLLDBhu98ChFupIIrlJA&e=).
+If you find any other issues with the results or have suggestions of ways to improve the results, please provide feedback on the [web form](https://forms.office.com/Pages/ResponsePage.aspx?id=qO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u).
 If you are interested in the code underlying the project and would like to support the project with technical input, please provide feedback via the [GitHub issue tracker (requires a GitHub account)](https://github.com/cyipt/popupCycleways/issues).

### 2020-06-01 code/build.R
wide segments

@@ -441,7 +441,7 @@ r_lanes_final = r_lanes_joined %>%
     Status = case_when(
       group_id %in% r_lanes_top$group_id ~ labels[1],
       majority_spare_lane ~ labels[2],
-      mean_width >= 10 ~ labels[3]
+      mean_width >= 10 & ! majority_spare_lane ~ labels[3]
     ),
     `Estimated width` = case_when(
       mean_width < 10 ~ "<10 m",
@@ -476,7 +476,8 @@ if(nrow(spare_lanes) == 0) {
 wide_lane_groups = r_lanes_final %>% filter(Status == labels[3])
 wide_lane_segments = rg_new4[wide_lane_groups, , op = sf::st_within]
 wide_lanes = wide_lane_segments %>%
-  filter(width >= 10 & !idGlobal %in% spare_lanes$idGlobal) %>% 
+  # filter(width >= 10 & !idGlobal %in% spare_lanes$idGlobal) %>% 
+  filter(width >= 10) %>% 
   mutate(Status = labels[3])
 # edge case: there are no wide lanes
 if(nrow(wide_lanes) == 0) {

### 2020-06-01 code/build.R
fix spare lanes issue

@@ -473,10 +473,12 @@ if(nrow(spare_lanes) == 0) {
 }
 # show wide segments not groups:
 # width_10m = r_lanes_final %>% filter(Status == labels[1])
-wide_lane_groups = r_lanes_final %>% filter(Status == labels[3])
+wide_lane_groups = r_lanes_final %>% 
+  filter(`Estimated width` != "<10 m") %>%
+  filter(Status != labels[1])
 wide_lane_segments = rg_new4[wide_lane_groups, , op = sf::st_within]
 wide_lanes = wide_lane_segments %>%
-  # filter(width >= 10 & !idGlobal %in% spare_lanes$idGlobal) %>% 
+  filter(!idGlobal %in% spare_lanes$idGlobal) %>% 
   filter(width >= 10) %>% 
   mutate(Status = labels[3])
 # edge case: there are no wide lanes

### 2020-06-01 code/build.R
extra bit

@@ -479,7 +479,7 @@ wide_lane_groups = r_lanes_final %>%
 wide_lane_segments = rg_new4[wide_lane_groups, , op = sf::st_within]
 wide_lanes = wide_lane_segments %>%
   filter(!idGlobal %in% spare_lanes$idGlobal) %>% 
-  filter(width >= 10) %>% 
+  filter(width >= 10 | spare_lane == TRUE) %>% 
   mutate(Status = labels[3])
 # edge case: there are no wide lanes
 if(nrow(wide_lanes) == 0) {

### 2020-06-02 code/wales.R
Add code for wales

@@ -1,0 +1,55 @@
+# Aim: find spare lanes in Wales
+
+remotes::install_github("itsleeds/geofabrik")
+library(geofabrik)
+library(tidyverse)
+library(tmap)
+tmap_mode("view")
+city_centre = tmaptools::geocode_OSM("cardiff", as.sf = TRUE)
+mapview::mapview(city_centre)
+
+osm_data = geofabrik::get_geofabrik(city_centre)
+nrow(osm_data) # 700,000+ filas
+summary(osm_data$lanes)
+table(osm_data$lanes)
+osm_spare_lanes = osm_data %>%
+  mutate(n_lanes = as.numeric(lanes)) %>%
+  filter(n_lanes >= 2)
+osm_cycleways = osm_data %>%
+  filter(highway == "cycleway")
+nrow(osm_spare_lanes)
+
+mapview::mapview(osm_spare_lanes)
+buffer_distance = 20 # km
+sf::st_crs(city_centre)
+# city_centre_geographic = sf::st_buffer(city_centre, 1) # incorrect
+# mapview::mapview(city_centre_geographic)
+# city_buffer = stplanr::geo_buffer(city_centre, buffer_distance * 1000) # fails
+city_buffer = city_centre %>%
+  sf::st_transform(crs = 5361) %>%
+  sf::st_buffer(dist = buffer_distance * 1000) %>%
+  sf::st_transform(4326)
+osm_spare_city = sf::st_intersection(osm_spare_lanes, city_buffer)
+osm_cycleways_city = sf::st_intersection(osm_cycleways, city_buffer) %>%
+  mutate(n_lanes = 1) %>%
+  select(name, highway, n_lanes, maxspeed)
+osm_spare_clean = osm_spare_city %>%
+  select(name, highway, n_lanes, maxspeed, oneway) %>%
+  filter((n_lanes >= 2 & oneway == "yes") | (n_lanes >= 3)) %>% 
+  filter(highway != "residential") %>% 
+  filter(highway != "unclassified") %>% 
+  filter(highway != "motorway") %>% 
+  
+
+names(osm_cycleways_city)
+names(osm_spare_clean)
+osm_popup = rbind(osm_cycleways_city, osm_spare_clean)
+s = c(
+  "Esri.WorldGrayCanvas",
+  "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
[…]

### 2020-06-02 code/preprocess.R
Close #66

@@ -55,4 +55,5 @@ saveRDS(cy, "cy.Rds")
 saveRDS(rj, "rj.Rds") # 300 MB file
 rj_sample = rj %>% sample_n(1000)
 mapview::mapview(rj_sample) # national coverage
+system("ls -hal *.Rds")
 piggyback::pb_upload("rj.Rds", "cyipt/cyipt-phase-1-data")

### 2020-06-02 code/wales.R

@@ -32,13 +32,13 @@ city_buffer = city_centre %>%
 osm_spare_city = sf::st_intersection(osm_spare_lanes, city_buffer)
 osm_cycleways_city = sf::st_intersection(osm_cycleways, city_buffer) %>%
   mutate(n_lanes = 1) %>%
-  select(name, highway, n_lanes, maxspeed)
+  select(name, highway, n_lanes, maxspeed, oneway)
 osm_spare_clean = osm_spare_city %>%
   select(name, highway, n_lanes, maxspeed, oneway) %>%
   filter((n_lanes >= 2 & oneway == "yes") | (n_lanes >= 3)) %>% 
   filter(highway != "residential") %>% 
   filter(highway != "unclassified") %>% 
-  filter(highway != "motorway") %>% 
+  filter(highway != "motorway") 
   
 
 names(osm_cycleways_city)
@@ -53,3 +53,4 @@ tm_shape(osm_popup) +
   tm_lines(col = "highway", lwd = "n_lanes", scale = 5) +
   tm_scale_bar() +
   tm_basemap(server = s)
+

### 2020-06-02 code/build.R
crack it

@@ -359,7 +359,10 @@ rg_new3 = r_lanes %>%
 # If that's not possible, split using 50m buffer, if each section is > min_grouped_length (500m). Or identify the longest gap within the group
 
 # If that's not possible, split at midpoint(s), or use cycling potential
+short_roads = rg_new3 %>%
+  filter(group2_length <= 2000)
 
+rg_new4 = rbind()
 
 # create a new group to capture long continuous sections with the same name
 min_length_named_road = 1000

### 2020-06-02 code/chelt.R

@@ -54,7 +54,7 @@ lg_new2 = lg_new %>%
     ) %>%
   st_transform(27700)
 
-groups(lg_new2)
+# groups(lg_new2)
 
 # Find the centroid of each section
 lg_c = sf::st_centroid(lg_new2)
@@ -64,7 +64,7 @@ lg_c = sf::st_centroid(lg_new2)
 
 # If the shortest section is < min_grouped_length (500m), join it together with the section with the closest centroid. Recalculate the length and centroid of this new section. 
 # Keep going until all sections are >500m (since this is the minimum length required for top routes) 
-summary(lg_new2$group3_length)
+# summary(lg_new2$group3_length)
 while (min(lg_c$group3_length) < min_grouped_length) {
   shortest = NULL
   nearest = NULL
@@ -99,22 +99,28 @@ while (min(lg_c$group3_length) < min_grouped_length) {
   if(length(mindist) > 1) mindist = mindist[1]
   
   lg_new2$group3_length[lg_new2$group3 == near$group3[mindist]] =
-    sum(shortest_segment$group3_length, nearest_l$group3_length)
-  nearest_l$geometry = sf::st_union(
-    shortest_segment$geometry,
-    sf::st_cast(nearest_l$geometry, "LINESTRING")
+    sum(shortest_segment$group3_length, lg_new2$group3_length[lg_new2$group3 == near$group3[mindist]])
+  # lg_new2$geometry[lg_new2$group3 == near$group3[mindist]] =
+  #   sf::st_union(
+  #   shortest_segment$geometry,
+  #   sf::st_cast(lg_new2$geometry[lg_new2$group3 == near$group3[mindist]], "LINESTRING")
+  #   )
+  lg_new2$geometry[lg_new2$group3 == near$group3[mindist]] =
+    sf::st_union(
+    shortest_segment$geometry, lg_new2$geometry[lg_new2$group3 == near$group3[mindist]]
     )
   
+  lg_c[lg_c$group3 == near$group3[mindist],] =
+    sf::st_centroid(lg_new2[lg_new2$group3 == near$group3[mindist],])
+  
   lg_new2 = lg_new2 %>%
-    filter(group3 != shortest_segment$group3, group3 != nearest_l$group3) %>%
-    rbind(new_joined)
+    filter(group3 != shortest_segment$group3) 
   lg_c = lg_c %>%
[…]

### 2020-06-03 report.Rmd
Update report

---
title: "Data analysis to prioritise new cycleways in response to COVID-19"
author: Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds
output: bookdown::pdf_document2
bibliography: report-refs.bib
---

```{r, eval=FALSE, echo=FALSE}
citr::tidy_bib_file("~/uaf/allrefs.bib", rmd_file = "report.Rmd", file = "report-refs.bib")
citr::tidy_bib_file("article.Rmd", "~/uaf/allrefs.bib", file = "ref.bib")
browseURL("report.pdf")
```

```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
```

# Introduction

New statutory guidance and funding for active transport infrastructure was announced by the Secretary of State for Transport on May 9^th^, to support the UK's response to the COVID-19 crisis by enabling healthy and safe travel at a time of reduced capacity on public transport networks.^[
https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities/traffic-management-act-2004-network-management-in-response-to-covid-19
]

The statutory guidance on network management in response to COVID-19 outlines a range of measures that can support a safe and healthy restart, including the creation of new cycleways, reducing speed limits, widening of footways and closing roads to motor traffic via 'modal filters'.
**This report describes the rapid cycleway prioritisation tool project, which provides aims to ensure evidence-based and effective investment in first of these measures: the creation of new cycleways in existing carriageway space reclaimed from motor traffic.**
Light segregation infrastructure such as flexible plastic wands (illustrated in Figure 1) and bollards can deliver a step change in provision for cycling, when implemented in combination with other measures such reduced speed limits on key corridors (as has happened on the A33 in Southampton and the A56 in Manchester) and when placed where they are most needed.
**The purpose of the tool is to help identify such key corridors, ensuring that funds from the Emergency Active Travel Fund change the status quo of city transport systems in favour of active travel.**

<!-- 17:29:01	 From Gill M (Manchester cyclist) : Trafford is planning to reduce speeds on Bridgewater Way element of A56 when the pop up lane goes in there “OneTrafford tweeted that work is continuing to add pop up cycle lanes along Cornbrook Road and Warwick Road in Stretford too, with temporary 30mph speed limit imposed along Bridgewater Way.” Source (https://www.messengernewspapers.co.uk/news/18489349.cyclists-praise-pop-up-cycle-lanes-along-a56-trafford/) -->

<!-- https://twitter.com/doddyuk/status/1267146449227255808 -->

```{r park-lane, fig.cap="New space for cycling created by road space reallocation in Park Lane, London.", out.width="60%"}
# source: Hillingdon and Uxbridge Times
# https://www.hillingdontimes.co.uk/news/national/18450461.one-londons-prestigious-roads-gets-cycle-lane-due-pandemic/
knitr::include_graphics("figures/cycleway-image.jpg")
```

The £250m released by the Emergency Active Travel Fund will be spent in three broad ways:

- The first tranche: £45m released on 1^st^ June, is for swift interventions that must be implemented by the end of June
- The second tranche: £180m will be released later in the summer for more permanent measures
- £25m will be invested in cycle repair schemes

The emphasis of the evidence presented in this report is providing a strong evidence base to prioritisation of schemes eligible for the second tranche.
However, the evidence could also be used to support identification of schemes for the first tranche, such as reallocation of road space on arterial routes, where decisions have not yet been made.

Given the importance of acting quickly, and the unique opportunity that the crisis poses to improve physical activity levels in the long term, it is vital that local authorities have the evidence they need.
Without a nationally consistent and systematic evidence base, there is a risk that some already stretched local authorities will rely on ad-hoc methods to allocate funding.
The rapid cycleway prioritisation tool can also support authorities to make bold plans by highlighting the potential of key strategic corridors.

# Methods

The approach was designed to identify roads and road segments where investment in new cycleways should be prioritised using a transparent methodology and nationally consistent datasets, alongside local authority expertise and local knowledge.
Strong and actionable datasets already exist to support strategic cycle network planning, notably the Propensity to Cycle Tool (PCT), a national planning support system [@lovelace_propensity_2017].
The PCT is already used by local authorities and contains key data showing not only current levels of cycling but also scenarios of growth, notably the Government Target scenario of doubling the number of cycling stages made by 2025 [@departmentfortransport_cycling_2017], at area, route and route network levels, for commuting and school travel (as of May 2020) [@goodman_scenarios_2019].
A limitation of the PCT, however, is that it takes little account of existing infrastructure: variables such as road width and even the presence/absence of cycle infrastructure currently are not included in the results.

To address this limitation, the work set out below will use national road network data which includes geographic attributes and cycle infrastructure from OpenStreetMap, plus carriageway width data derived from analysis of Ordnance Survey data.
This road data was developed for the Cycling Infrastructure Prioritisation Tool (CyIPT) project, funded as a prototype project by a DfT Transport-technology research innovation grants (T-TRIG).
The results of CyIPT, while not publicly accessible, are well known among local authority transport planners and are reported to be of great use to practitioners in scheme prioritisation, according to a DfT funded Discovery project on tools to support investment in active transport.^[
Planning Tools for cycle networks and infrastructure prioritisation, unpublished.
]

The rapid cycleway prioritisation tool is based on levels of cycling potential *and* road attribute data, with an emphasis on road space reallocation.
The method was parameterised to highlight roads that have sufficient space to allow for new cycleways without affecting the navigable network for motorised modes [@lovelace_methods_2020].
A key benefit of this method is that it allows continuous road sections to be identified and ranked, supporting prioritisation of investment, based not only on cycling potential, but also based on continuous road length and estimated cycle usage measured in km per day.
The tool highlights continuous road sections with high cycling potential in the context of existing infrastructure on the basis that new infrastructure should maximise cycle network cohesion and continuity [@parkin_designing_2018].

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE, out.width = "100%", eval = FALSE)
```

## Attribute filtering and grouping

At a time of reduced travel, fewer lanes dedicated to motor traffic are needed.
Based on this observation, we defined roads with a 'spare lane' as those on which there is more than one lane in either direction.
This definition assumes no reduction in mobility for motor vehicles (making two-way lanes one-way is another option not explored in this analysis).

To identify road sections on which there is a spare lane we developed a simple algorithm that takes the OSM variable [`lanes`](https://wiki.openstreetmap.org/wiki/Key:lanes) if it is present and, if not, derives the number from the highway type and presence/absence of bus lanes.
All segments defined as having a spare lane using this method is shown in Figure \@ref(fig:levels) (left).
In future, this methodology could be enhanced to take into account the effect of lane widths, which increase the effective available space, and the presence of vehicle parking bays, which reduce it.
The result of filtering by distance and cycling potential before and after grouping using graph membership of touching roads is shown in \@ref(fig:levels) (middle and right, respectively).
Grouping linked roads before filtering results in a more cohesive network.

<!-- Note this could be a function in an R package.. -->
<!-- see https://github.com/cyipt/cyipt/blob/82248b2f99e388fac314d34ec5aa49bb90a737a3/scripts/prep_data/clean_osm.R#L349 -->

<!-- reallocated road space in particular -->
<!-- An important distinction when developing methods for automated analysis of transport networks is the level of analysis. -->

```{r levels, fig.height=3, fig.cap="Illustration of the 'group then filter' method to identify long sections with spare lanes *and* high cycling potential"}
r_pct_lanes_all = r_central %>% 
  filter(cycling_potential > min_cycling_potential) %>% 
  filter(lanes_f > 1 | lanes_b > 1)
# mapview::mapview(r_pct_lanes)

touching_list = st_touches(r_pct_lanes_all)
g = igraph::graph.adjlist(touching_list)
components = igraph::components(g)
r_pct_lanes_all$group = components$membership
r_pct_lanes = r_pct_lanes_all %>% 
  group_by(group) %>% 
  mutate(group_length = sum(length)) %>% 
  mutate(cycling_potential_mean = weighted.mean(cycling_potential, w = length, na.rm = TRUE)) %>% 
  filter(cycling_potential_mean > min_grouped_cycling_potential)
r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length)

r_pct_lanes$graph_group = r_pct_lanes$group
group_table = table(r_pct_lanes$group)
top_groups = tail(sort(group_table), 5)
r_pct_lanes$graph_group[!r_pct_lanes$graph_group %in% names(top_groups)] = "other"

r_filter_before_grouping = rj %>% 
  filter(cycling_potential > min_cycling_potential) %>% 
  filter(lanes_f > 1 | lanes_b > 1) %>% 
  filter(cycling_potential > min_grouped_cycling_potential) %>% 
  filter(length > 100)
tmap_mode("plot")
m0 = tm_shape(city_key_buffer) + tm_borders(col = "grey") +
  tm_shape(r_pct_lanes_all) + tm_lines() +
  tm_layout(title = "Roads on which there are spare lanes.")
m1 = tm_shape(city_key_buffer) + tm_borders(col = "grey") +
  tm_shape(r_filter_before_grouping) + tm_lines() +
  tm_layout(title = "Filter then group:\n(length > 100, cycling_potential > 100)")
m2 = tm_shape(city_key_buffer) + tm_borders(col = "grey") +
  tm_shape(r_pct_lanes) + tm_lines("graph_group", palette = "Dark2") +
  tm_layout(title = "Group then filter:\n(length > 500, cycling_potential > 100)")
# todo show group membership with colours
ma = tmap_arrange(m0, m1, m2, nrow = 1)
ma
```

```{r joining}
r_pct_grouped = r_pct_lanes %>%
  group_by(name, group) %>%
  summarise(
    group_length = sum(length),
    cycling_potential = round(weighted.mean(cycling_potential_mean, length))
  )
# summary(r_pct_grouped$group_length)
r_pct_top = r_pct_grouped %>%
  filter(group_length > min_grouped_length) %>% 
  filter(cycling_potential > min_grouped_cycling_potential) %>% 
  filter(!grepl(pattern = regexclude, name, ignore.case = TRUE)) %>% 
  mutate(km_day = round(cycling_potential * group_length / 1000))
# head(r_pct_top$km_day)

r_pct_lanes_overlap = r_pct_lanes[r_pct_top, , op = sf::st_covered_by] # works
# r_pct_no_overlap = st_difference(r_pct_lanes, r_pct_top) # slow
r_pct_no_overlap = r_pct_lanes %>% 
  filter(!idGlobal %in% r_pct_lanes_overlap$idGlobal)

r_pct_top_n = r_pct_top %>% 
  group_by(name) %>% 
  slice(which.max(km_day)) %>% 
  filter(name != "") %>% 
  ungroup() %>% 
  top_n(10, wt = km_day) 
# aim: make the labels go on the longest sections
# %>% 
#   st_cast("LINESTRING") %>% 
#   mutate(length_section = as.numeric(st_length(.))) %>% 
#   group_by(name) %>% 
#   slice(which.max(length_section))
# r_pct_top_n$name
# summary(r_pct_grouped$group_length)
```

# Limitations and known issues

The approach is not without limitations.
Its reliance on data rather than community engagement represents a rather top-down approach to transport planning.
To overcome this issue, future work could seek to incorporate the results such as those presented above into a participatory map of the type shown in Figure 1.

# Next steps

# References

### 2020-06-03 code/build.R
Close #57, for now

@@ -252,7 +252,7 @@ r_linestrings_with_ref = r_lanes_all %>%
 
 # Roads with no ref -------------------------------------------------------
 
-min_cycling_potential_without_ref = 30
+min_cycling_potential_without_ref = min_cycling_potential * 5
 r_linestrings_without_ref =  r_lanes_all %>%
   filter(ref == "") %>% 
   filter(cycling_potential >= min_cycling_potential_without_ref)
@@ -275,7 +275,9 @@ r_linestrings_without_ref2 = r_linestrings_without_ref %>%
   filter(mean_cycling_potential > min_grouped_cycling_potential) %>%  # this is currently 50
   ungroup() %>% 
   select(-nogroup, -group_length, -mean_cycling_potential)
-# mapview::(no_ref_grouped["mean_cycling_potential"])
+# mapview::mapview(no_ref_grouped["mean_cycling_potential"])
+
+
 
 r_lanes = rbind(r_linestrings_with_ref, r_linestrings_without_ref2)
 # mapview::mapview(r_lanes)
@@ -313,7 +315,7 @@ rg_new2 = rg_new %>%
   filter(mean_width >= 10 | majority_spare_lane) %>%
   filter(mean_cycling_potential >= min_grouped_cycling_potential) %>%
   ungroup()
-# mapview::mapview(rg_new2)
+# mapview::mapview(rg_new2, zcol = "mean_cycling_potential")
 
 rg_buff = geo_buffer(shp = rg_new2, dist = buff_dist_large)
 touching_list = st_intersects(rg_buff)
@@ -330,7 +332,7 @@ rg_new3 = rg_new2 %>%
 
 # mapview::mapview(rg_new3)
 # create a new group to capture long continuous sections with the same name
-min_length_named_road = 1000
+min_length_named_road = 800
 rg_new4 = rg_new3 %>% 
   group_by(ref, group, ig, name) %>% 
   mutate(long_named_section = case_when(
@@ -341,7 +343,6 @@ rg_new4 = rg_new3 %>%
   ungroup()
 # Only one group (meaning no impact on results) in many regions:
 table(rg_new4$long_named_section)
-
 
 
 # find group membership of top named roads
@@ -392,6 +393,7 @@ summary(r_lanes_joined$proportion_on_cycleway) # all between 0 and 1
 r_lanes_top = r_lanes_joined %>%
   # filter(name != "Unnamed road" & ref != "") %>%
[…]

### 2020-06-03 code/build.R
Increase the number of top roads being shown

@@ -83,7 +83,7 @@ r_min_width_highlighted = 10
 max_cycling_potential_99th = quantile(x = rj$cycling_potential, probs = 0.99)
 min_cycling_potential = round(max_cycling_potential_99th / 250)
 min_grouped_cycling_potential = round(max_cycling_potential_99th / 25)
-n_top_roads = (1 + round(nrow(rj) / 80000)) * 10
+n_top_roads = (1 + round(nrow(rj) / 60000)) * 10
 
 # buffers -----------------------------------------------------------------
 if(is_city) {

### 2020-06-03 code/chelt.R
segment level

@@ -46,27 +46,34 @@ lg_new$il = NULL
 lg_new2 = lg_new %>% 
   group_by(group3) %>%
   mutate(
-    group3_length = sum(length)
+    group3_length = sum(length),
+    new_group3 = group3
   ) %>%
-  summarise(
-    group2 = first(group2),
-    group3_length = mean(group3_length)
-    ) %>%
+  # summarise(
+  #   group2 = first(group2),
+  #   group3_length = mean(group3_length),
+  #   new_group3 = first(new_group3)
+  #   ) %>%
   st_transform(27700)
 
 # groups(lg_new2)
 
 # Find the centroid of each section
-lg_c = sf::st_centroid(lg_new2)
+lg_c = lg_new2 %>%
+  group_by(new_group3) %>%
+  summarise(group2 = first(group2),
+            group3_length = mean(group3_length)) %>%
+  select(group2, group3_length, new_group3) %>%
+  sf::st_centroid() 
 
 
-# why are West Street group2 vanishing? "4 120 A38" all joined groups seem to be disappearing see mapview(lg_new2)
 
 # If the shortest section is < min_grouped_length (500m), join it together with the section with the closest centroid. Recalculate the length and centroid of this new section. 
 # Keep going until all sections are >500m (since this is the minimum length required for top routes) 
 # summary(lg_new2$group3_length)
 while (min(lg_c$group3_length) < min_grouped_length) {
-  shortest = NULL
+  shortest_segment = NULL
+  shortest_centroid = NULL
   nearest = NULL
   near = NULL
   mindist = NULL
@@ -82,7 +89,7 @@ while (min(lg_c$group3_length) < min_grouped_length) {
   if(dim(shortest_centroid)[1] > 1) shortest = shortest[1,]
   near = lg_c %>% 
     filter(group2 == shortest_centroid$group2) %>% 
-    filter(group3 != shortest_centroid$group3) # must remove the point itself from this group
+    filter(new_group3 != shortest_centroid$new_group3) # must remove the point itself from this group
   
   distances = NULL
   
@@ -98,25 +105,33 @@ while (min(lg_c$group3_length) < min_grouped_length) {
   mindist = which(distances == min(distances))
[…]

### 2020-06-03 code/index.Rmd
Revert to Key cycle network
Close #71, fix #72

@@
 ```
-
-# NOTE: PRELIMINARY RESULTS
-
-Not quality assured.
-
-Please provide feedback on these preliminary results on this [web form (estimated time to complete: 5 minutes)](https://forms.office.com/Pages/ResponsePage.aspx?id=qO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u).
 
@@
 
-## 1. Cohesive network
+## 1. Key cycle routes
 
@@
 ```{r}
-dn = "popupCycleways/v0.3" # dirctory name
+dn = "popupCycleways/v1" # dirctory name
 built_regions = list.files(dn)
@@
   sf::st_drop_geometry() %>%
-  filter(html_name %in% built_regions) %>% 
+  # filter(html_name %in% built_regions) %>% 
   select(Name, Level, url) %>% 
@@
 rn = "West of England"
-for(rn in region_names_to_build[50:nrow(regions)]) {
+for(rn in region_names_to_build) {
   region = regions %>% filter(Name == rn)

### 2020-06-03 code/build.R

@@ -52,11 +52,12 @@ if(!exists("region_name")) {
 }
 if(region_name == "Nottingham") {
   region = regions %>% filter(str_detect(string = Name, pattern = "Nott")) %>% 
+    st_geometry() %>% 
     st_union()
 } else {
   region = regions %>% filter(Name == region_name)
 }
-lads_centroids = lads_all_centroids[region, ]
+lads_centroids = lads_all_centroids[region$geometry, ]
 lads = lads_all %>% filter(Name %in% lads_centroids$Name)
 if(nrow(lads) == 0) {
   lads = region
@@ -83,6 +84,7 @@ r_min_width_highlighted = 10
 max_cycling_potential_99th = quantile(x = rj$cycling_potential, probs = 0.99)
 min_cycling_potential = round(max_cycling_potential_99th / 250)
 min_grouped_cycling_potential = round(max_cycling_potential_99th / 25)
+min_top_cycling_potential = min_grouped_cycling_potential + 10
 n_top_roads = (1 + round(nrow(rj) / 60000)) * 10
 
 # buffers -----------------------------------------------------------------
@@ -119,7 +121,7 @@ rnet_school_done = rnet_all_school[reg_school_logical, ]
 
 # Combine the two
 combine = rbind(rnet_done, rnet_school_done) # change to rbindlist
-rnet_combined = stplanr::overline2(x = sf::st_cast(combine, "LINESTRING"), attrib = "govtarget_slc")
+rnet_combined = stplanr::overline2(sf::st_cast(combine, "LINESTRING"), attrib = "govtarget_slc")
 
 # Link the updated cycle potential to the road widths ---------------------
 
@@ -225,7 +227,14 @@ group_table = sort(table(r_key_roads_plus_high_pct$group), decreasing = TRUE)
 key_network = r_key_roads_plus_high_pct %>%
   group_by(group) %>% 
   mutate(group_length = sum(length)) %>% 
-  filter(group_length > 5 * min_grouped_length)
+  filter(group_length > 5 * min_grouped_length) %>% 
+  mutate(
+    `Estimated width` = case_when(
+      width < 10 ~ "<10 m",
+      width >= 10 & width < 15 ~ "10-15 m",
+      width >= 15 ~ ">15 m"
+    )
+  ) 
 # mapview::mapview(key_network)
 # tm_shape(key_network) + tm_lines(lwd = "mean_width", scale = 7, col = "lightsalmon2")
 
[…]

### 2020-06-03 code/build.R
Fix #73 - thanks @joeytalbot

@@ -84,7 +84,7 @@ r_min_width_highlighted = 10
 max_cycling_potential_99th = quantile(x = rj$cycling_potential, probs = 0.99)
 min_cycling_potential = round(max_cycling_potential_99th / 250)
 min_grouped_cycling_potential = round(max_cycling_potential_99th / 25)
-min_top_cycling_potential = min_grouped_cycling_potential + 10
+# min_top_cycling_potential = min_grouped_cycling_potential + 10
 n_top_roads = (1 + round(nrow(rj) / 60000)) * 10
 
 # buffers -----------------------------------------------------------------
@@ -341,22 +341,43 @@ rg_new3 = rg_new2 %>%
 
 # mapview::mapview(rg_new3)
 # create a new group to capture long continuous sections with the same name
-min_length_named_road = 800
-rg_new4 = rg_new3 %>% 
-  group_by(ref, group, ig, name) %>% 
+min_length_named_road = min_grouped_length
+rg_new4 = rg_new3 %>%
+  group_by(ref, group, ig, name) %>%
   mutate(long_named_section = case_when(
     sum(length) > min_length_named_road & name != "" ~ name,
     TRUE ~ "Other"
   )
-  ) %>% 
+  ) %>%
   ungroup()
-# Only one group (meaning no impact on results) in many regions:
 table(rg_new4$long_named_section)
+# new approach
 
+# Split into sections by road name, and split these into contiguous sections using buff_dist_large (100m).
+lgs = unique(rg_new4$long_named_section)
+
+i = "Melbourn bypass"
+
+# create per name groups
+long_list = lapply(lgs, FUN = function(i) {
+  lg = rg_new4 %>% filter(long_named_section == i)
+  # mapview::mapview(lg)
+  l_buff = lg %>%
+    st_transform(27700) %>%
+    st_buffer(buff_dist_large) %>%
+    st_transform(4326)
+  touching_list = st_intersects(l_buff)
+  g = igraph::graph.adjlist(touching_list)
+  components = igraph::components(g)
+  lg$long_named_group = components$membership
+  lg
+})
+
+lg_new = do.call(rbind, long_list)
 
 # find group membership of top named roads
-r_lanes_grouped2 = rg_new4 %>% 
-  group_by(ref, group, ig, long_named_section) %>% 
+r_lanes_grouped2 = lg_new %>% 
[…]

### 2020-06-03 code/build.R
keep all roads

@@ -375,8 +375,13 @@ long_list = lapply(lgs, FUN = function(i) {
 
 lg_new = do.call(rbind, long_list)
 
+other_roads = rg_new4[rg_new4$long_named_section == "Other", ]
+other_roads$long_named_group = NA
+rejoined = rbind(lg_new, other_roads)
+
+
 # find group membership of top named roads
-r_lanes_grouped2 = lg_new %>% 
+r_lanes_grouped2 = rejoined %>%
   group_by(ref, group, ig, long_named_section, long_named_group) %>% 
   summarise(
     name = case_when(
@@ -438,18 +443,18 @@ nrow(r_lanes_top)
 r_lanes_top %>% sf::st_drop_geometry()
 
 # remove any disconnected bits
-rg_buff = geo_buffer(shp = r_lanes_top, dist = buff_dist_large)
-touching_list = st_intersects(rg_buff)
-g = igraph::graph.adjlist(touching_list)
-components = igraph::components(g)
-rg_new2$lastgroup = components$membership
-
-# Only keep segments which are part of a wider group (including roads with different refs/names) of >500m length (100m buffer)
-rg_new3 = rg_new2 %>% 
-  group_by(lastgroup) %>%
-  mutate(last_length = round(sum(length))) %>%
-  filter(last_length >= min_grouped_length) %>% 
-  ungroup()
+# rg_buff = geo_buffer(shp = r_lanes_top, dist = buff_dist_large)
+# touching_list = st_intersects(rg_buff)
+# g = igraph::graph.adjlist(touching_list)
+# components = igraph::components(g)
+# rg_new2$lastgroup = components$membership
+# 
+# # Only keep segments which are part of a wider group (including roads with different refs/names) of >500m length (100m buffer)
+# rg_new3 = rg_new2 %>% 
+#   group_by(lastgroup) %>%
+#   mutate(last_length = round(sum(length))) %>%
+#   filter(last_length >= min_grouped_length) %>% 
+#   ungroup()
 
 # classify roads to visualise
 labels = c("Top route", "Spare lane(s)", "Estimated width > 10m")

### 2020-06-03 code/build.R
update with changes

@@ -261,34 +261,80 @@ r_linestrings_with_ref = r_lanes_all %>%
 
 # Roads with no ref -------------------------------------------------------
 
-min_cycling_potential_without_ref = min_cycling_potential * 5
+# Remove segments with cycling potential below 30 (to prevent side street segments)
+min_cycling_potential_without_ref = 30
 r_linestrings_without_ref =  r_lanes_all %>%
   filter(ref == "") %>% 
   filter(cycling_potential >= min_cycling_potential_without_ref)
+
+# Put into groups, using a 20m buffer (stricter than for roads with a ref, to prevent groups covering multiple streets)
 r_linestrings_without_ref_buff = geo_buffer(shp = r_linestrings_without_ref, dist = 20)
 touching_list = st_intersects(r_linestrings_without_ref_buff)
 g = igraph::graph.adjlist(touching_list)
 components = igraph::components(g)
-r_linestrings_without_ref$nogroup = components$membership
+r_linestrings_without_ref$group2 = components$membership
 
-# mapview::mapview(r_linestrings_without_ref["nogroup"], lwd = 3)
+# mapview::mapview(r_linestrings_without_ref["group2"], lwd = 3)
 
-# Remove groups with length <300m (20m buffer). (these are groups consisting purely of road segments with no ref) this is stricter than for roads with refs, because otherwise too many short segments are picked up. 
+# Remove groups with length <300m (20m buffer). (these are groups consisting purely of road segments with no ref) this is stricter than for roads with refs, because otherwise too many short segments are picked up. # Remove groups without sufficient width/lanes or cycle potential (the same as for roads with a ref)
 r_linestrings_without_ref2 = r_linestrings_without_ref %>%
-  group_by(nogroup) %>%
+  group_by(group2) %>%
   mutate(
-    group_length = round(sum(length)),
-    mean_cycling_potential = round(weighted.mean(cycling_potential, length, na.rm = TRUE))
-    ) %>% 
-  filter(group_length >= 300) %>%
-  filter(mean_cycling_potential > min_grouped_cycling_potential) %>%  # this is currently 50
-  ungroup() %>% 
-  select(-nogroup, -group_length, -mean_cycling_potential)
[…]

### 2020-06-03 code/index.Rmd

@@
 # region_names_to_build = regions$Name[!regions$html_name %in% built_regions]
-# sample_to_build_regex = "west of|west y|west mid|hereford|nott|leic|manc|liv|cam"
-# region_names_to_build = regions %>% 
-#   filter(grepl(pattern = sample_to_build_regex, x = Name, ignore.case = TRUE)) %>% 
-#   pull(Name)
+sample_to_build_regex = "west of|west y|west mid|hereford|nott|leic|manc|liv|cam|bright|suff"
+region_names_to_build = regions %>%
+  filter(grepl(pattern = sample_to_build_regex, x = Name, ignore.case = TRUE)) %>%
+  pull(Name)
 # region_names_to_build = c("West Yorkshire", "Nottingham")

### 2020-06-03 report.Rmd
Close #71, fix #72
Update report

@@
 
-The £250m released by the Emergency Active Travel Fund will be spent in three broad ways:
+The £250m released by the Emergency Active Travel Fund will be spent in three broad [ways](https://www.transportxtra.com/publications/local-transport-today/news/65661/dft-releases-first-45m-of-250m-fund/):
 
@@
 The PCT is already used by local authorities and contains key data showing not only current levels of cycling but also scenarios of growth, notably the Government Target scenario of doubling the number of cycling stages made by 2025 [@departmentfortransport_cycling_2017], at area, route and route network levels, for commuting and school travel (as of May 2020) [@goodman_scenarios_2019].
-A limitation of the PCT, however, is that it takes little account of existing infrastructure: variables such as road width and even the presence/absence of cycle infrastructure currently are not included in the results.
 
-To address this limitation, the work set out below will use national road network data which includes geographic attributes and cycle infrastructure from OpenStreetMap, plus carriageway width data derived from analysis of Ordnance Survey data.
-This road data was developed for the Cycling Infrastructure Prioritisation Tool (CyIPT) project, funded as a prototype project by a DfT Transport-technology research innovation grants (T-TRIG).
-The results of CyIPT, while not publicly accessible, are well known among local authority transport planners and are reported to be of great use to practitioners in scheme prioritisation, according to a DfT funded Discovery project on tools to support investment in active transport.^[
-Planning Tools for cycle networks and infrastructure prioritisation, unpublished.
-]
+A limitation of the PCT is that it takes little account of existing infrastructure: variables such as road width and even the presence/absence of cycle infrastructure currently are not included in the results.
+The rapid cycleway prioritisation tool overcomes this issu by using road network data which includes geographic attributes and cycle infrastructure from OpenStreetMap.
+Carriageway width data were taken from the Cycling Infrastructure Prioritisation Tool (CyIPT) project, funded as a prototype project by a DfT Transport-technology research innovation grants (T-TRIG).
 
@@
 The method was parameterised to highlight roads that have sufficient space to allow for new cycleways without affecting the navigable network for motorised modes [@lovelace_methods_2020].
-A key benefit of this method is that it allows continuous road sections to be identified and ranked, supporting prioritisation of investment, based not only on cycling potential, but also based on continuous road length and estimated cycle usage measured in km per day.
+A benefit of this method is that it allows continuous road sections to be identified and ranked, supporting prioritisation of investment.
+Roads can be ranked based not only on cycling potential, but also based on continuous road length and estimated cycle usage measured in km per day.
 The tool highlights continuous road sections with high cycling potential in the context of existing infrastructure on the basis that new infrastructure should maximise cycle network cohesion and continuity [@parkin_designing_2018].
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE, out.width = "100%", eval = FALSE)
-```
 
@@
 To identify road sections on which there is a spare lane we developed a simple algorithm that takes the OSM variable [`lanes`](https://wiki.openstreetmap.org/wiki/Key:lanes) if it is present and, if not, derives the number from the highway type and presence/absence of bus lanes.
-All segments defined as having a spare lane using this method is shown in Figure \@ref(fig:levels) (left).
-In future, this methodology could be enhanced to take into account the effect of lane widths, which increase the effective available space, and the presence of vehicle parking bays, which reduce it.
-The result of filtering by distance and cycling potential before and after grouping using graph membership of touching roads is shown in \@ref(fig:levels) (middle and right, respectively).
-Grouping linked roads before filtering results in a more cohesive network.
+All segments defined as having a spare lane using this method are shown in Figure \@ref(fig:levels) (left).
+We also included estimates of road width, only flagging roads with an average estimated width of more than 10m (the absolute minumum that should be considered for cycleways without additional changes).
+In future, this methodology could be enhanced to include verge and pavement width estimates, which could increase the effective available space for new cycleways, and the presence of vehicle parking bays, and on street parking which reduce it.
+The result of filtering by distance and cycling potential before and after grouping using graph membership of touching roads is shown in Figure \@ref(fig:levels) (middle and right, respectively), which shows that grouping linked roads before filtering results in a more cohesive network.
 
-<!-- Note this could be a function in an R package.. -->
-<!-- see https://github.com/cyipt/cyipt/blob/82248b2f99e388fac314d34ec5aa49bb90a737a3/scripts/prep_data/clean_osm.R#L349 -->
-
-<!-- reallocated road space in particular -->
-<!-- An important distinction when developing methods for automated analysis of transport networks is the level of analysis. -->
-
-```{r levels, fig.height=3, fig.cap="Illustration of the 'group then filter' method to identify long sections with spare lanes *and* high cycling potential"}
-r_pct_lanes_all = r_central %>% 
-  filter(cycling_potential > min_cycling_potential) %>% 
-  filter(lanes_f > 1 | lanes_b > 1)
-# mapview::mapview(r_pct_lanes)
-
-touching_list = st_touches(r_pct_lanes_all)
-g = igraph::graph.adjlist(touching_list)
-components = igraph::components(g)
-r_pct_lanes_all$group = components$membership
-r_pct_lanes = r_pct_lanes_all %>% 
-  group_by(group) %>% 
-  mutate(group_length = sum(length)) %>% 
-  mutate(cycling_potential_mean = weighted.mean(cycling_potential, w = length, na.rm = TRUE)) %>% 
-  filter(cycling_potential_mean > min_grouped_cycling_potential)
-r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length)
-
-r_pct_lanes$graph_group = r_pct_lanes$group
-group_table = table(r_pct_lanes$group)
-top_groups = tail(sort(group_table), 5)
-r_pct_lanes$graph_group[!r_pct_lanes$graph_group %in% names(top_groups)] = "other"
-
-r_filter_before_grouping = rj %>% 
-  filter(cycling_potential > min_cycling_potential) %>% 
-  filter(lanes_f > 1 | lanes_b > 1) %>% 
-  filter(cycling_potential > min_grouped_cycling_potential) %>% 
-  filter(length > 100)
-tmap_mode("plot")
-m0 = tm_shape(city_key_buffer) + tm_borders(col = "grey") +
-  tm_shape(r_pct_lanes_all) + tm_lines() +
-  tm_layout(title = "Roads on which there are spare lanes.")
-m1 = tm_shape(city_key_buffer) + tm_borders(col = "grey") +
-  tm_shape(r_filter_before_grouping) + tm_lines() +
-  tm_layout(title = "Filter then group:\n(length > 100, cycling_potential > 100)")
-m2 = tm_shape(city_key_buffer) + tm_borders(col = "grey") +
-  tm_shape(r_pct_lanes) + tm_lines("graph_group", palette = "Dark2") +
-  tm_layout(title = "Group then filter:\n(length > 500, cycling_potential > 100)")
-# todo show group membership with colours
-ma = tmap_arrange(m0, m1, m2, nrow = 1)
-ma
+```{r levels, out.width="90%", fig.cap="Illustration of the 'group then filter' method to identify long sections with spare lanes *and* high cycling potential"}
+knitr::include_graphics("figures/filter-then-group.png")
 ```
 
-```{r joining}
-r_pct_grouped = r_pct_lanes %>%
-  group_by(name, group) %>%
-  summarise(
-    group_length = sum(length),
-    cycling_potential = round(weighted.mean(cycling_potential_mean, length))
-  )
-# summary(r_pct_grouped$group_length)
-r_pct_top = r_pct_grouped %>%
-  filter(group_length > min_grouped_length) %>% 
-  filter(cycling_potential > min_grouped_cycling_potential) %>% 
-  filter(!grepl(pattern = regexclude, name, ignore.case = TRUE)) %>% 
-  mutate(km_day = round(cycling_potential * group_length / 1000))
-# head(r_pct_top$km_day)
+## Feedback from stakeholders
 
-r_pct_lanes_overlap = r_pct_lanes[r_pct_top, , op = sf::st_covered_by] # works
-# r_pct_no_overlap = st_difference(r_pct_lanes, r_pct_top) # slow
-r_pct_no_overlap = r_pct_lanes %>% 
-  filter(!idGlobal %in% r_pct_lanes_overlap$idGlobal)
+We elicited feedback from stakeholders via the Basecamp forum for local authority transport planners.
+As a result of the feedback we made a few adjustments to the results:
 
-r_pct_top_n = r_pct_top %>% 
-  group_by(name) %>% 
-  slice(which.max(km_day)) %>% 
-  filter(name != "") %>% 
-  ungroup() %>% 
-  top_n(10, wt = km_day) 
-# aim: make the labels go on the longest sections
-# %>% 
-#   st_cast("LINESTRING") %>% 
-#   mutate(length_section = as.numeric(st_length(.))) %>% 
-#   group_by(name) %>% 
-#   slice(which.max(length_section))
-# r_pct_top_n$name
-# summary(r_pct_grouped$group_length)
-```
+- We refined the road identification method, to highlight more central road sections and fewer long sections far from high density areas, in response to feedback suggesting that some of the road sections went too far into the countryside.
+- We simplified the legend of the basemap, so that the layers were presented in time order: Existing cycleway (now); Top ranked new cycleways (near term); Cohesive network (long term).
+- We requested data from Highways England to flag the recommendations on the Strategic Road Network (SRN)
 

### 2020-06-03 code/build.R
Update legend, close #76

@@ -489,8 +489,8 @@ nrow(r_lanes_top)
 r_lanes_top %>% sf::st_drop_geometry()
 
 # classify roads to visualise
-labels = c("Top route", "Spare lane(s)", "Estimated width > 10m")
-cycleways_name = "Existing off road cycleways"
+labels = c("Top ranked new cycleways", "Spare lane(s)", "Estimated width > 10m")
+cycleways_name = "Existing cycleways"
 
 r_lanes_final = r_lanes_joined %>% 
   mutate(
@@ -555,9 +555,9 @@ popup.vars = c(
 )
 pvars_key = c("ref", "name", "highway_type", "cycling_potential", "n_lanes", "Estimated width")
 key_network_final = key_network[pvars_key]
-legend_labels = c("Cohesive network", cycleways_name, labels[1], labels[2], labels[3])
+legend_labels = c(cycleways_name, labels[1], "Cohesive network", labels[2], labels[3])
 cols_status = c("blue", "#B91F48", "#FF7F00")
-legend_colours = c("darkgrey", "darkgreen", cols_status)
+legend_colours = c("darkgreen", cols_status[1], "darkgrey", cols_status[2:3])
 
 m =
   tm_shape(lads, name = "Local authority district boundaries") + tm_borders() +

### 2020-06-03 code/build.R
Close #77

@@ -57,6 +57,10 @@ if(region_name == "Nottingham") {
 } else {
   region = regions %>% filter(Name == region_name)
 }
+if(as.numeric(sf::st_area(region)) < 5000000000) {
+  region = stplanr::geo_buffer(region, dist = 1000)
+}
+
 lads_centroids = lads_all_centroids[region$geometry, ]
 lads = lads_all %>% filter(Name %in% lads_centroids$Name)
 if(nrow(lads) == 0) {

### 2020-06-03 code/index.Rmd
Remove pop-up terminology

@@
 
-The maps contain five layers of data that overlay the basemaps.
+The maps contain five layers (two of which are hidden by default) that provide the results of our analysis, overlaying the basemaps.
 
@@
 
-The following three layers are the central results of our analysis, representing the most promising locations for pop-up cycleways.
+The following three layers are the central results of our analysis, representing the most promising locations for new cycleways.
 
@@
 
-These are the top routes for pop-up cycleways according to our analysis. 
+These are the 'top n' road sections in terms of cycling potential with 'spare space' defined above.
+They may be suitable candidates for road space reallocation to create new cycleways. 
 These roads have either a spare lane or width of at least 10m, and are ranked according to their mean cycling potential. 

### 2020-06-03 code/index.Rmd
Update description of three main layers

@@
 The maps contain five layers (two of which are hidden by default) that provide the results of our analysis, overlaying the basemaps. 
+Clicking on the lines on the map for any of these layers will reveal a pop-up box containing more information about the road or segment of interest.
 
-Click on any of these layers to see a pop-up box containing more information about the road in question.
+The three main layers are arranged in order of planning timescales:
+
+1.
+The **Existing cycleways** layer provides an approximation of where cycling infrastructure exists **in the recent past** and is intended to help identify gaps in the network that could be filled with new interventions.
+2.
+The **Top ranked new cycleways** layer is the central result of the analysis, providing a list of roads that have high cycling potential, a minimum threshold length, and spare space according to our definition.
+These may be strong candidates for reallocation of road space **immediately or in the near future** to create new cycleways on strategic corridors under the first and second tranches of funding.
+3.
+The **Cohesive network** layer is intended to show what a joined-up cycle network could look like, if road space was not a constraint on creating new infrastructure.
+Composed of roads that have high cycling potential on some or all of their length, the layer is designed to guide **long term planning** towards a continuous and coherent network.
+
+These main layers, and two other layers that informed the results but which are not shown be default, are described in detail below.
 
@@
 This layer does not include some on-road cycle lanes or cycleways less than 500m in length, but these can be viewed in one of the basemap options (see below for more information about the basemaps).
+The definition of 'cycleway' includes shared use paths but excludes cycle lanes painted on roads in this layer.
 
-The following three layers are the central results of our analysis, representing the most promising locations for new cycleways.
-
-## 3. Top routes
+## 3. Top ranked new cycleways
 
@@
 For routes that are shorter than 1km in length, we use 'km_cycled' instead of mean cycling potential - this is simply the mean cycling potential multiplied by the length of the route. 
-Top routes are listed in the table below the map, so you can see more information on them. 
-The number of top routes selected is normally 20, with some variation by region.
+Top ranked new cycleways are listed in the table below the map, so you can see more information on them. 
+The number of Top ranked new cycleways selected is normally 20, with some variation by region.
 
-We used certain eligibility criteria for the top routes:
+We used certain eligibility criteria for the Top ranked new cycleways:
 
@@
 
-- The 'top routes' highlighted in the map-based results were produced automatically based on a simple score of continuous candidate road length multiplied by cycling potential.
+- The 'Top ranked new cycleways' highlighted in the map-based results were produced automatically based on a simple score of continuous candidate road length multiplied by cycling potential.
 For roads with a length of more than 1 km, the cycling potential alone is used to rank roads.

### 2020-06-03 code/index.Rmd
Update layer descriptions

@@
 It shows what a joined-up cycle network could look like if there were no constraints on road space, along some of the most important corridors.
+
 The layer takes no account spare space and is designed to support consideratioin of alternative options, e.g.
 one way streets, where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
-The layer was developed based on evidence on the importance of active transport network that are continuous and do not 'give up' at awkward junctions.
-The layer was created by selecting all road segments with cycle potential in the 85th percentile and above, which also belong to classified trunk roads, A roads or B roads.
+The cohesive network layer was created based on evidence that cohesive cycle networks, that are continuous and do not 'give up' at awkward junctions, are a key component of plans to get people cycling.
+
+The layer was generated by first identifying the roads (classified by reference number) on which there is most cycling potential accross the city and joining them together.
+Specifically we first selected all road segments with cycle potential in the 85th percentile and above, which also belong to classified trunk roads, A roads or B roads.
 These segments are then joined together to form a continuous network.

### 2020-06-03 code/index.Rmd
Add references

@@
     toc: yes
+bibliography: 
+  # - index.bib
+  - ../refs.bib
 ---
+
+```{r, eval=FALSE, echo=FALSE}
+citr::tidy_bib_file("~/uaf/allrefs.bib", rmd_file = "code/index.Rmd", file = "code/index.bib")
+```
 
@@
 
+```{r bibs, engine='bash', eval=FALSE}
+bibtool
+ls *.bib
+bibtool refs.bib report-refs.bib code/index.bib > refs.bib
+```
+
 # Introduction
 
-This tool helps identify road sections for road space reallocation and other interventions to create new cycleways.
+The rapid cycleway prioritisation tool helps identify road sections for road space reallocation and other interventions to create new cycleways.
 Used with local knowledge, it provides a strong evidence base to support planning of rapid measures to support safe travel as part of the COVID-19 response.
-The evidence may be used to support a range of temporary, experimental and permanent treatments that will enable shift to walking and cycling in the coming weeks and months, to help overcome capacity constraints on public transport modes and prevent gridlock as the lockdown eases.
+In the context of COVID-19, big decisions need to be made **rapidly**.
 
-Maps are available for regions covering the whole of England.
+The rapid cycleway prioritisation tool aims to help produce evidence-based answers the question: **where to reallocate road space on arterial routes away from motor traffic and towards active transport?**
+The evidence can be used to support a range of temporary, experimental and permanent treatments that will enable shift to walking and cycling in the coming weeks and months.
+The focus of the tool is road space reallocation on strategic corridors, a key way in which transport systems can be transformed to overcome capacity constraints on public transport modes and prevent gridlock as the lockdown eases.
 
-The focus is on roads that have 'spare space' (defined as those with a spare lane or that have an estimated width of above 10 m) and high cycling potential.
-The 10 m width is an indicator of road conditions and larger widths, of 15+ m, may be required for road space reallocation without other changes such as making roads one way for motor traffic or removing central markings as described in a paper by Stella Shackel and John Parkin ([2014](https://doi.org/10.1016/j.aap.2014.08.015)).
+Maps are currently available for regions covering the whole of England.
+
+The focus is on roads that have 'spare space' (defined as those with a spare lane or that have an estimated width of above 10 m) and high cycling potential based on the [Propensity to Cycle Tool](https://www.pct.bike/) [@lovelace_propensity_2017].
+The 10 m width is an indicator of road conditions and larger widths, of 15+ m, may be required for road space reallocation without other changes such as making roads one way for motor traffic or removing [central markings](https://doi.org/10.1016/j.aap.2014.08.015) [@shackel_influence_2014].
 Which roads to prioritise and the most effective interventions will also depend on factors such as on street parking levels not present in the analysis.
@@
 
-## 1. Cohesive network
+## 1. Existing cycleways
 
-This layer is designed to represent a cohesive network of direct routes.
-It shows what a joined-up cycle network could look like if there were no constraints on road space, along some of the most important corridors.
-
-The layer takes no account spare space and is designed to support consideratioin of alternative options, e.g.
-one way streets, where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
-The cohesive network layer was created based on evidence that cohesive cycle networks, that are continuous and do not 'give up' at awkward junctions, are a key component of plans to get people cycling.
-
-The layer was generated by first identifying the roads (classified by reference number) on which there is most cycling potential accross the city and joining them together.
-Specifically we first selected all road segments with cycle potential in the 85th percentile and above, which also belong to classified trunk roads, A roads or B roads.
-These segments are then joined together to form a continuous network.
-
-## 2. Existing cycleways
-
-This layer shows existing cycleways.
+On the basis that plans should build on and extend existing provision, this layer shows existing cycleways.
 This may be helpful to view in connection with the other layers, because it can reveal where gaps in the existing cycle network could be filled by new infrastructure.
@@
 
-## 3. Top ranked new cycleways
+## 2. Top ranked new cycleways
 
@@
 - routes less than 500m in length cannot be a top route
+
+## 3. Cohesive network
+
+This layer is designed to represent a cohesive network of direct routes.
+It shows what a joined-up cycle network could look like if there were no constraints on road space, along some of the most important corridors.
+
+The layer takes no account spare space and is designed to support consideration of alternative options, e.g.
+one way streets, where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
+The cohesive network layer was created based on evidence that cohesive cycle networks, that are continuous and do not 'give up' at awkward junctions, are a key component of plans to get people cycling.
+
+The layer was generated by first identifying the roads (classified by reference number) on which there is most cycling potential accross the city and joining them together.
+Specifically we first selected all road segments with cycle potential in the 85th percentile and above, which also belong to classified trunk roads, A roads or B roads.
+These segments are then joined together to form a continuous network.
 
@@
 ```
+
+# References

### 2020-06-03 report.Rmd
Add references

@@
 output: bookdown::pdf_document2
-bibliography: report-refs.bib
+bibliography: refs.bib
 ---
@@
 ```{r, eval=FALSE, echo=FALSE}
-citr::tidy_bib_file("~/uaf/allrefs.bib", rmd_file = "report.Rmd", file = "report-refs.bib")
-citr::tidy_bib_file("article.Rmd", "~/uaf/allrefs.bib", file = "ref.bib")
+citr::tidy_bib_file("~/uaf/allrefs.bib", rmd_file = "report.Rmd", file = "refs.bib")
+citr::tidy_bib_file("~/uaf/allrefs.bib", rmd_file = "code/index.Rmd", file = "refs.bib")
+citr::tidy_bib_file("article.Rmd", "~/uaf/allrefs.bib", file = "refs.bib")
 browseURL("report.pdf")

### 2020-06-03 article.R

@@ -3,7 +3,7 @@ knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TR
 
 
 ## ---- echo=FALSE, eval=FALSE-----------------------------------------------------------------
-## citr::tidy_bib_file("article.Rmd", "~/uaf/allrefs.bib", file = "ref.bib")
+## citr::tidy_bib_file("article.Rmd", "~/uaf/allrefs.bib", file = "refs.bib")
 ## wordcountaddin::word_count("article.Rmd")
 
 

### 2020-06-03 code/index.Rmd
Update index

@@
 The rapid cycleway prioritisation tool aims to help produce evidence-based answers the question: **where to reallocate road space on arterial routes away from motor traffic and towards active transport?**
-The evidence can be used to support a range of temporary, experimental and permanent treatments that will enable shift to walking and cycling in the coming weeks and months.
 The focus of the tool is road space reallocation on strategic corridors, a key way in which transport systems can be transformed to overcome capacity constraints on public transport modes and prevent gridlock as the lockdown eases.
 
-Maps are currently available for regions covering the whole of England.
+Reallocating road space on strategic corridors is just one of the measures available for transport authorities.
+Evidence provided by the tool can also support a range of temporary, experimental and permanent treatments that will enable shift to walking and cycling in the coming weeks and months.
+New [statutory guidance on network management in response to COVID-19](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities/traffic-management-act-2004-network-management-in-response-to-covid-19) outlines a range of measures, including the creation of new cycleways, reducing speed limits, widening of footways and closing roads to motor traffic via 'modal filters'.
+Road space reallocation is suited to larger arterial roads and can have a large impact quickly, especially when moving large numbers of people away from public transport and capacity is a priority.
+Furthermore, low but gradually increasing levels of peak hour motor traffic mean that there is a window of opportunity to reallocate lanes without causing congestion.
+
+<!-- # Input data -->
 
@@
 Which roads to prioritise and the most effective interventions will also depend on factors such as on street parking levels not present in the analysis.
-
-Reallocating roadspace is just one of the measures available for transport authorities.
-Other approaches include modal filters, road closures, and the creation of bus, walking and cycling only streets, as well as speed limits and traffic calming measures.
-Used strategically, a combination of these measures can promote active travel through the creation of filtered permeability networks and Low Traffic Neighbourhoods.
-For certain types of area, such as residential zones, this may be the most suitable approach.
-However, roadspace reallocation is naturally suited to larger arterial-type roads and that is the focus of this tool.
 
@@
 Overall the evidence presented in the tool is designed to encourage bold interventions where more space for active modes is most needed.
-The work is based evidence that road space can be reallocated without causing congestion at a time when traffic levels, particularly during the rush hours, are greatly reduced.
+Currently the results are only available for England.
+<!-- The work is based evidence that road space can be reallocated without causing congestion at a time when traffic levels, particularly during the rush hours, are greatly reduced. -->
 

### 2020-06-03 code/index.Rmd
Close #79

@@
 
-The rapid cycleway prioritisation tool helps identify road sections for road space reallocation and other interventions to create new cycleways.
+The [rapid cycleway prioritisation tool](https://www.cyipt.bike/rapid/) helps identify road sections for road space reallocation and other interventions to create new cycleways.
 Used with local knowledge, it provides a strong evidence base to support planning of rapid measures to support safe travel as part of the COVID-19 response.

### 2020-06-03 code/region_index.Rmd
Close #79

@@
 
-The map shows a network of key roads with high cycling potential (grey), existing off road cycleways (and shared space paths) according to OpenStreetMap data (green) and top road sections with high cycling potential and spare lanes or estimated width above 10m (blue).
+The map shows three key layers:
+
+1.
+The **Existing cycleways** layer provides an approximation of where cycling infrastructure exists **in the recent past** and is intended to help identify gaps in the network that could be filled with new interventions.
+2.
+The **Top ranked new cycleways** layer is the central result of the analysis, providing a list of roads that have high cycling potential, a minimum threshold length, and spare space according to our definition.
+These may be strong candidates for reallocation of road space **immediately or in the near future** to create new cycleways on strategic corridors under the first and second tranches of funding.
+3.
+The **Cohesive network** layer is intended to show what a joined-up cycle network could look like, if road space was not a constraint on creating new infrastructure.
+Composed of roads that have high cycling potential on some or all of their length, the layer is designed to guide **long term planning** towards a continuous and coherent network.
+
 To see additional layers showing road sections with 'spare lanes' and width greater than 10 m, click on the layers icon in the top left, just below the +/- buttons on the map.
@@
 
-See full map [here](m.html).
+See full map [here](m.html), and a more description of the layers and the methodology at [cyipt.bike/rapid](https://www.cyipt.bike/rapid/)
 

### 2020-06-03 code/build.R
Update build

@@ -57,7 +57,7 @@ if(region_name == "Nottingham") {
 } else {
   region = regions %>% filter(Name == region_name)
 }
-if(as.numeric(sf::st_area(region)) < 5000000000) {
+if(as.numeric(sf::st_area(region)) < 3000000000) {
   region = stplanr::geo_buffer(region, dist = 1000)
 }
 
@@ -293,7 +293,7 @@ r_linestrings_without_ref2 = r_linestrings_without_ref %>%
   filter(mean_cycling_potential > min_grouped_cycling_potential) %>%  # this varies by region
   filter(mean_width >= 10 | majority_spare_lane) %>%
   ungroup() 
-# mapview::(no_ref_grouped["mean_cycling_potential"])
+# mapview::mapview(r_linestrings_without_ref2, zcol = "mean_cycling_potential")
 
 # Roads with a ref --------------------------------------------------------
 
@@ -339,7 +339,7 @@ rg_new2 = rg_new %>%
 
 # Now rejoin the roads with no ref together with the roads with a ref
 r_lanes = rbind(rg_new2, r_linestrings_without_ref2)
-# mapview::mapview(r_lanes)
+# mapview::mapview(r_lanes, zcol = "group2")
 
 gs = unique(r_lanes$ref)
 # i = g[2]
@@ -389,7 +389,7 @@ rg_new3 = rg_new2 %>%
   filter(last_length >= min_grouped_length) %>% 
   ungroup()
 
-# mapview::mapview(rg_new3)
+# mapview::mapview(rg_new3, zcol = "lastgroup")
 # create a new group to capture long continuous sections with the same name
 min_length_named_road = min_grouped_length
 rg_new4 = rg_new3 %>%
@@ -400,6 +400,7 @@ rg_new4 = rg_new3 %>%
   )
   ) %>%
   ungroup()
+# mapview::mapview(rg_new4, zcol = "long_named_section")
 table(rg_new4$long_named_section)
 # new approach
 
@@ -425,13 +426,12 @@ long_list = lapply(lgs, FUN = function(i) {
 
 lg_new = do.call(rbind, long_list)
 
-other_roads = rg_new4[rg_new4$long_named_section == "Other", ]
-other_roads$long_named_group = NA
-rejoined = rbind(lg_new, other_roads)
-
+# other_roads = rg_new4[rg_new4$long_named_section == "Other", ]
+# other_roads$long_named_group = NA
+# rejoined = rbind(lg_new, other_roads)
 
[…]

### 2020-06-03 code/index.Rmd
Save geojson for #78

@@
   rmarkdown::render("code/region_index.Rmd", output_file = file.path("~/cyipt/tempCycleways/", d, "index.html"))
+  sf::write_sf(cycleways, file.path(d, "cycleways.geojson"))
+  sf::write_sf(top_routes, file.path(d, "top_routes.geojson"))
+  sf::write_sf(key_network_final, file.path(d, "cohesive_network.geojson"))
+  sf::write_sf(spare_lanes_final, file.path(d, "spare_lanes.geojson"))
+  sf::write_sf(wide_lanes_final, file.path(d, "wide_lanes.geojson"))
   message(round(time_to_run), " minutes to build ", rn)

### 2020-06-03 code/build.R

@@ -557,8 +557,11 @@ popup.vars = c(
   "mean_cycling_potential",
   "length (m)"
 )
-pvars_key = c("ref", "name", "highway_type", "cycling_potential", "n_lanes", "Estimated width")
+pvars_key = c("ref", "name", "highway_type", "cycling_potential", "n_lanes", "Estimated width", "maxspeed")
 key_network_final = key_network[pvars_key]
+pvars_spare = c("name", "ref", "maxspeed", "cycling_potential", "n_lanes")
+spare_lanes_final = spare_lanes[pvars_spare]
+wide_lanes_final = wide_lanes[pvars_spare]
 legend_labels = c(cycleways_name, labels[1], "Cohesive network", labels[2], labels[3])
 cols_status = c("blue", "#B91F48", "#FF7F00")
 legend_colours = c("darkgreen", cols_status[1], "darkgrey", cols_status[2:3])
@@ -567,20 +570,20 @@ m =
   tm_shape(lads, name = "Local authority district boundaries") + tm_borders() +
   tm_shape(key_network_final, name = "Cohesive network") +
   tm_lines(lwd = 5, col = "darkgrey", popup.vars = pvars_key) +
-  tm_shape(spare_lanes, name = labels[2]) +
+  tm_shape(spare_lanes_final, name = labels[2]) +
   tm_lines(legend.col.show = FALSE,
            col = cols_status[2], 
            lwd = 3,
            alpha = 1,
-           popup.vars = c("name", "ref", "maxspeed", "cycling_potential", "n_lanes"),
+           popup.vars = pvars_spare,
            group = labels[2]
            ) +
-  tm_shape(wide_lanes, name = labels[2]) +
+  tm_shape(wide_lanes_final, name = labels[2]) +
   tm_lines(legend.col.show = FALSE,
            col = cols_status[3], 
            lwd = 3,
            alpha = 1,
-           popup.vars = c("name", "ref", "maxspeed", "cycling_potential", "n_lanes"),
+           popup.vars = pvars_spare,
            group = labels[3]
   ) +
   tm_shape(cycleways, name = cycleways_name) + tm_lines(popup.vars = c("surface", "name", "osm_id"), col = "darkgreen", lwd = 1.3) +

### 2020-06-04 report.Rmd
report

@@
 Light segregation infrastructure such as flexible plastic wands (illustrated in Figure 1) and bollards can deliver a step change in provision for cycling, when implemented in combination with other measures such reduced speed limits on key corridors (as has happened on the A33 in Southampton and the A56 in Manchester) and when placed where they are most needed.
-**The purpose of the tool is to help identify such key corridors, ensuring that funds from the Emergency Active Travel Fund change the status quo of city transport systems in favour of active travel.**
+**The purpose of the tool is to help identify such key corridors, ensuring that funds from the Emergency Active Travel Fund change the status quo of urban and rural transport systems in favour of active travel.**
 
@@
 Strong and actionable datasets already exist to support strategic cycle network planning, notably the Propensity to Cycle Tool (PCT), a national planning support system [@lovelace_propensity_2017].
-The PCT is already used by local authorities and contains key data showing not only current levels of cycling but also scenarios of growth, notably the Government Target scenario of doubling the number of cycling stages made by 2025 [@departmentfortransport_cycling_2017], at area, route and route network levels, for commuting and school travel (as of May 2020) [@goodman_scenarios_2019].
+The PCT is already used by local authorities and contains key data showing not only current levels of cycling but also scenarios of growth, notably the Government Target scenario of doubling the number of cycling stages made by 2025 [@departmentfortransport_cycling_2017].
+The PCT maps cycle uptake at area, route and route network levels, for commuting and school travel (as of May 2020) [@goodman_scenarios_2019].
 
 A limitation of the PCT is that it takes little account of existing infrastructure: variables such as road width and even the presence/absence of cycle infrastructure currently are not included in the results.
-The rapid cycleway prioritisation tool overcomes this issu by using road network data which includes geographic attributes and cycle infrastructure from OpenStreetMap.
+The rapid cycleway prioritisation tool overcomes this issue by using road network data which includes geographic attributes and cycle infrastructure from OpenStreetMap.
 Carriageway width data were taken from the Cycling Infrastructure Prioritisation Tool (CyIPT) project, funded as a prototype project by a DfT Transport-technology research innovation grants (T-TRIG).
@@
 
+## Selection of top routes
+
 ## Feedback from stakeholders
@@
 
+# Outputs
+
+The results of this research are publically accessible at www.cyipt.bike/rapid. 
+Maps are available for regions covering the whole of England, at the level of Combined Authorities, Unitary Authorities, or Upper Tier Authorities. 
+Using the methods outlined above, we have derived a set of outputs which are represented as layers in these maps.
+
+## Output 1: Cohesive network
+
+This represents a cohesive network of cycle routes along the most important road corridors in each region.
+It allows visualisation of what a joined-up cycle network would look like, without considering road capacity constraints.
+
+To identify this network, we select road segments on the trunk road, A road and B road network that have cycling potential in the 85th percentile or above, for their region. 
+We then join together these segments using a 1000m buffer, to form a coherent network.
+This is thus based on the cycling potential of roads, without reference to the road width or number of lanes.
+
+This provides the ability to identify roads that may be key priorities for cycle infrastructure, but which are missing from the other results, either because of limitations with our road width or number of lanes input data, or because the roads are simply too narrow to show up in the other results.
+Interventions may still be possible on narrow roads, for example the removal of central line markings has been shown to reduce average vehicle speeds (ref). 
+As an example of cycle infrastructure on a narrow road, South Road in Lancaster is recorded on OSM as having road width <10m, yet protected cycleways are currently being constructed there (image). 
+In other cases, it may be possible to convert a two-way road to one-way working and re-allocate the space previously occupied by one of the traffic lanes to bi-directional cycleways.
+(Torrington Place image)
+
+Finally the cohesive network can help identify gaps in existing and/or proposed cycle infrastructure.
+Gappy networks are off-putting to many cyclists, and busy or complex road junctions in particular can create barriers to cycle movement (ref).
+
+## Output 2: Existing cycleways
+
+## Output 3: Top routes
+
+## Output 4: Roads with spare lanes and/or estimated width of at least 10m
+
 # Limitations and known issues
@@
 Its reliance on data rather than community engagement represents a rather top-down approach to transport planning.
-To overcome this issue, future work could seek to incorporate the results such as those presented above into a participatory map of the type shown in Figure 1.
+To overcome this issue, future work could seek to incorporate the results such as those presented above into a participatory map of the type shown in Figure x (add widenmypath fig).
+
+As stated above, road space reallocation is the only form of intervention we have explored. 
+This should be investigated in combination with other interventions. 
+Strategic use of modal filters, road closures, traffic calming, and conversions to one-way working can be used to prevent 'rat-running', enhance filtered permeability, and create Low Traffic Neighoubourhoods (image).
+
+We have used crowd-sourced open data for road width and number of lanes, which does not give comprehensive coverage of these road attributes in all parts of the country. 
+The 'cohesive network' layer mitigates the impact of any inaccuracies here.
+
+Our current methodology only selects roads that have the same name for a continuous stretch of at least 500m, as candidates for top routes.
+Roads that don't meet this criteria can still appear in the 'spare lane' and 'estimated width > 10m' layers.
+
+Vehicular traffic volumes may restrict the potential for road space reallocation, and the presence of on-road car parking may prevent the use of particular alignments.
+While we do not account for these limitations, the reduction in traffic due to COVID-19 has brought about both a unique opportunity and an urgent necessity to change travel patterns and behaviours in profound ways.
 

### 2020-06-04 code/index.Rmd
Add code to update map (wip)

@@
 
+```{r, eval=FALSE}
+# rebuild results based on pre-processed data (wip)
+for(rn in region_names_to_build) {
+  region = regions %>% filter(Name == rn)
+  if(rn %in% c(regions_within_names)) {
+    region$geometry = sf::st_union(
+      region,
+      regions %>% filter(Name == regions_containing_names[rn])
+      ) %>% sf::st_geometry()
+  }
+  t = Sys.time()
+  message("Building for ", rn)
+  d = file.path(dn, region$html_name)
+  dir.create(d)
+  region_name = region$Name
+  # source("code/build.R")
+  rmarkdown::render(input = "code/build.R", output_dir = d, knit_root_dir = ".")
+  plot(sf::st_geometry(r_lanes_top))
+  htmlwidgets::saveWidget(m_leaflet, file.path("~/cyipt/tempCycleways/", d, "m.html"))
+  time_to_run = Sys.time() - t
+  rmarkdown::render("code/region_index.Rmd", output_file = file.path("~/cyipt/tempCycleways/", d, "index.html"))
+  sf::write_sf(cycleways, file.path(d, "cycleways.geojson"))
+  sf::write_sf(top_routes, file.path(d, "top_routes.geojson"))
+  sf::write_sf(key_network_final, file.path(d, "cohesive_network.geojson"))
+  sf::write_sf(spare_lanes_final, file.path(d, "spare_lanes.geojson"))
+  sf::write_sf(wide_lanes_final, file.path(d, "wide_lanes.geojson"))
+  message(round(time_to_run), " minutes to build ", rn)
+}
+m =
+  tm_shape(lads, name = "Local authority district boundaries") + tm_borders() +
+  tm_shape(key_network_final, name = "Cohesive network") +
+  tm_lines(lwd = 5, col = "darkgrey", popup.vars = pvars_key) +
+  tm_shape(spare_lanes_final, name = labels[2]) +
+  tm_lines(legend.col.show = FALSE,
+           col = cols_status[2], 
+           lwd = 3,
+           alpha = 1,
+           popup.vars = pvars_spare,
+           group = labels[2]
+           ) +
+  tm_shape(wide_lanes_final, name = labels[2]) +
+  tm_lines(legend.col.show = FALSE,
+           col = cols_status[3], 
+           lwd = 3,
+           alpha = 1,
+           popup.vars = pvars_spare,
+           group = labels[3]
+  ) +
+  tm_shape(cycleways, name = cycleways_name) + tm_lines(popup.vars = c("surface", "name", "osm_id"), col = "darkgreen", lwd = 1.3) +
+  tm_shape(top_routes, name = "Top routes") +
+  tm_lines(legend.col.show = FALSE,
+           col = cols_status[1], 
+           lwd = 5,
+           alpha = 1,
+           popup.vars = popup.vars
+  ) +
+  tm_basemap(server = s, tms = tms) +
+  tm_add_legend(type = "fill", labels = legend_labels[1:3], col = legend_colours[1:3]) +
+  tm_add_legend(type = "fill", labels = labels[2], col = legend_colours[4], group = labels[2]) +
+  tm_add_legend(type = "fill", labels = labels[3], col = legend_colours[5], group = labels[3]) +
+  tm_scale_bar() 
+# m
+m_leaflet = tmap_leaflet(m) %>% leaflet::hideGroup(labels[2:3])
+```
+
 # References

### 2020-06-04 code/build.R
Close #82

@@ -43,6 +43,7 @@ if(!exists("s")) {
   rnet_url_school = "https://github.com/npct/pct-outputs-national/raw/master/school/lsoa/rnet_all.Rds"
   rnet_all = sf::st_as_sf(readRDS(url(rnet_url)))
   rnet_all_school = sf::st_as_sf(readRDS(url(rnet_url_school)))
+  srn_names_df = readr::read_csv("input-data/srn_names_df.csv")
 }
 
 # local parameters --------------------------------------------------------
@@ -445,7 +446,8 @@ r_lanes_grouped2 = lg_new %>%
     group_length = round(sum(length)),
     mean_cycling_potential = round(weighted.mean(cycling_potential, length, na.rm = TRUE)),
     mean_width = round(weighted.mean(width, length, na.rm = TRUE)),
-    majority_spare_lane = sum(length[spare_lane]) > sum(length[!spare_lane])
+    majority_spare_lane = sum(length[spare_lane]) > sum(length[!spare_lane]),
+    speed_limit = names(which.max(table(maxspeed)))
   ) %>% 
   filter(mean_cycling_potential > min_grouped_cycling_potential | group_length > min_grouped_length) %>%
   ungroup() %>% 
@@ -485,12 +487,12 @@ r_lanes_top = r_lanes_joined %>%
   filter(proportion_on_cycleway < minp_exclude) %>% 
   mutate(
     length_up_to_1km = if_else(group_length > 1000, true = 1000, false = group_length),
-    km_cycled_1km = length_up_to_1km * mean_cycling_potential
+    km_cycled_1km = length_up_to_1km * mean_cycling_potential,
+    srn = name %in% srn_names_df$roa_number
     ) %>% 
   arrange(desc(km_cycled_1km)) %>% 
-  slice(1:n_top_roads)
+  slice(1:n_top_roads) 
 nrow(r_lanes_top)
-r_lanes_top %>% sf::st_drop_geometry()
 
 # classify roads to visualise
 labels = c("Top ranked new cycleways", "Spare lane(s)", "Estimated width > 10m")
@@ -503,13 +505,13 @@ r_lanes_final = r_lanes_joined %>%
       majority_spare_lane ~ labels[2],
       mean_width >= 10 ~ labels[3]
     ),
-    `Estimated width` = case_when(
+    `Estimated width (m)` = case_when(
       mean_width < 10 ~ "<10 m",
       mean_width >= 10 & mean_width < 15 ~ "10-15 m",
[…]

### 2020-06-04 code/build.R
Turn off srn

@@ -624,7 +624,7 @@ res_table = r_lanes_top %>%
     `Length (m)` = group_length,
     `Cycling potential` = mean_cycling_potential,
     `Length * potential (km)` = km_cycled,
-    `SRN` = srn,
+    # `SRN` = srn,
     `Speed limit` = speed_limit
     ) 
 res_table

### 2020-06-04 report.Rmd

@@
 Carriageway width data were taken from the Cycling Infrastructure Prioritisation Tool (CyIPT) project, funded as a prototype project by a DfT Transport-technology research innovation grants (T-TRIG).
+This uses road data derived from OpenStreetMap.org (OSM)
 
@@
 
-## Attribute filtering and grouping
+## Road attribute data
 
@@
 Based on this observation, we defined roads with a 'spare lane' as those on which there is more than one lane in either direction.
-This definition assumes no reduction in mobility for motor vehicles (making two-way lanes one-way is another option not explored in this analysis).
+This definition assumes no reduction in mobility for motor vehicles (making two-way streets one-way is another option not explored in this analysis).
 
@@
 All segments defined as having a spare lane using this method are shown in Figure \@ref(fig:levels) (left).
-We also included estimates of road width, only flagging roads with an average estimated width of more than 10m (the absolute minumum that should be considered for cycleways without additional changes).
+We also included estimates of road width, only flagging roads with an average estimated width of more than 10m (the minimum that should be considered for cycleways without additional changes).
 In future, this methodology could be enhanced to include verge and pavement width estimates, which could increase the effective available space for new cycleways, and the presence of vehicle parking bays, and on street parking which reduce it.
-The result of filtering by distance and cycling potential before and after grouping using graph membership of touching roads is shown in Figure \@ref(fig:levels) (middle and right, respectively), which shows that grouping linked roads before filtering results in a more cohesive network.
+
+## Grouping and filtering of road segments
+
+For our three core outputs of near-term cycleway prioritisation, we follow a series of grouping and filtering steps, to identify roads that have both high cycle potential and available road space. 
+Segments are grouped by their road reference number, with stricter filtering criteria for roads with no reference number, to avoid a mass of short road segments on side streets. 
+We use a 100m buffer to further group these segments, as shown in Figure \@ref(fig:levels) (middle and right, respectively), which shows that grouping nearby road segments before filtering results in a more cohesive network.
+
+We then remove groups which do not have either an estimated mean road width of at least 10m or a spare lane for the majority of their length. 
+Groups with mean cycling potential below a minimum threshold are also removed. This threshold varies according to cycling levels in each region. 
+Finally, any segments not part of a wider group of > 500m length (using a 100m buffer) are removed.
 
@@
 ## Selection of top routes
+
+From the results of the previous steps, we can now select the top routes for cycling. 
+Only roads that have the same name for a continuous stretch of at least 500m are eligible for selection as a top route. 
+Roads cannot be candidates for top routes if there is an existing cycleway along 75% or more of their length. 
+We selected such a high value because the presence of a cycleway in OSM does not guarantee that this is a high quality, continuous route.
+Top routes are ranked according to their cycling potential.
+
+Roads not selected as top routes are shown in the layers named 'spare lane' and 'estimated width >10m.'
+
+## Identification of cohesive cycle network
+
+The cohesive network is derived using a different method from the other layers. 
+To identify this network, we select road segments on the trunk road, A road and B road network that have cycling potential in the 85th percentile or above, for their region. 
+We then join together these segments using a 1000m buffer, to form a coherent network.
+This is thus based on the cycling potential of roads, without reference to the road width or number of lanes.
 
@@
 
-The results of this research are publically accessible at www.cyipt.bike/rapid. 
+The results of this research are publicly accessible at www.cyipt.bike/rapid. 
 Maps are available for regions covering the whole of England, at the level of Combined Authorities, Unitary Authorities, or Upper Tier Authorities. 
@@
 
-## Output 1: Cohesive network
+## Existing cycleways
 
-This represents a cohesive network of cycle routes along the most important road corridors in each region.
+The first layer shows existing off-carriageway cycleways.
+It is important to plan new infrastructure in reference to that which already exists.
+
+## Near term: Top priorities for road space reallocation
+
+Three layers show the roads we have identified as candidates for road space reallocation.
+These roads have both high cycling potential and available space.
+They also meet criteria for length and continuity.
+
+Named roads with the highest cycling potential are ranked and selected as top routes.
+For the remaining roads, we identify whether they have a spare lane for the majority of their length, or an estimated mean width >10m.
+
+## Long term: Cohesive network
+
+The final layer represents a cohesive network of cycle routes along the most important road corridors in each region.
 It allows visualisation of what a joined-up cycle network would look like, without considering road capacity constraints.
 
-To identify this network, we select road segments on the trunk road, A road and B road network that have cycling potential in the 85th percentile or above, for their region. 
-We then join together these segments using a 1000m buffer, to form a coherent network.
-This is thus based on the cycling potential of roads, without reference to the road width or number of lanes.
-
-This provides the ability to identify roads that may be key priorities for cycle infrastructure, but which are missing from the other results, either because of limitations with our road width or number of lanes input data, or because the roads are simply too narrow to show up in the other results.
+This provides the ability to identify roads that may be key priorities today for cycle infrastructure, but which are missing from the other results, either due to the limitations of our road attribute data, or because the roads are simply too narrow to show up in the other results.
 Interventions may still be possible on narrow roads, for example the removal of central line markings has been shown to reduce average vehicle speeds (ref). 
@@
 
-## Output 2: Existing cycleways
-
-## Output 3: Top routes
-
-## Output 4: Roads with spare lanes and/or estimated width of at least 10m
-
 # Limitations and known issues
@@
 
-As stated above, road space reallocation is the only form of intervention we have explored. 
-This should be investigated in combination with other interventions. 
-Strategic use of modal filters, road closures, traffic calming, and conversions to one-way working can be used to prevent 'rat-running', enhance filtered permeability, and create Low Traffic Neighoubourhoods (image).
+As stated above, we focus on road space reallocation, but this is best pursued as part of a wider multi-intervention strategy. 
+In urban and residential areas, other interventions such as strategic use of modal filters, road closures, traffic calming, and conversions to one-way working can be used to prevent 'rat-running', enhance filtered permeability, and create Low Traffic Neighbourhoods (image).
+For longer journeys, park-and-cycle schemes can solve 'last-mile' problems and cut congestion.
 
@@
 
-Vehicular traffic volumes may restrict the potential for road space reallocation, and the presence of on-road car parking may prevent the use of particular alignments.
+Vehicular traffic volumes may restrict the potential for road space reallocation, and the presence of on-street car parking may prevent the use of particular alignments. 
 While we do not account for these limitations, the reduction in traffic due to COVID-19 has brought about both a unique opportunity and an urgent necessity to change travel patterns and behaviours in profound ways.
+As an example of the changes are happening right now, Glasgow has recently suspended around a third of its on-street parking spaces, to widen pavements and increase space for social distancing.
 
@@
 
+This tool can be used together with existing knowledge to identify locations for investment in new cycleways.
+
+Further consideration should be given to community consultation and to other factors such as materials, design, and engineering constraints.
+Physical protection of cycleways will improve their safety, but permeable barriers (such as flexible plastic wands) are preferred to solid barriers that may obstruct pedestrians from crossing the road and encourage speeding.
+Interventions will be most effective where they form part of a wider transport policy and where experimental measures that prove successful can be adopted long-term.
+
 # References

### 2020-06-04 code/srn.R
Add figures

@@ -1,0 +1,27 @@
+# aim: get srn data
+srn_posts = readr::read_csv("https://datamillnorth.org/download/strategic-road-network-marker-posts/58ebbe5f-bfb2-4a01-bf30-64587f7a51c4/Gazetteer_All_Mposts_only.csv")
+srn_posts
+head(srn_posts$TN)
+
+library(sf)
+u = "http://data.dft.gov.uk/road-traffic/major-roads-link-network2018.zip"
+download.file(u, "major-roads-link-network2018.zip")
+unzip("major-roads-link-network2018.zip")
+mrn = sf::read_sf("2018-MRDB-minimal.shp")
+plot(mrn)
+head(mrn$CP_Number)
+
+srn = sf::read_sf("~/hd/data/uk/srn/HAPMS_network_20160929.shp")
+sel_unique = duplicated(srn$geometry)
+summary(sel_unique)
+plot(srn$geometry) 
+head(srn$roa_number)
+mapview::mapview(srn)
+srn_names_df = srn %>% 
+  sf::st_drop_geometry() %>% 
+  group_by(roa_number) %>% 
+  summarise(
+    length_km = sum(sec_length)
+  )
+readr::write_csv(srn_names_df, "input-data/srn_names_df.csv")
+summary(str_detect(string = srn_names_df$roa_number, pattern = "A"))

### 2020-06-04 code/index.Rmd
Revert to original index

@@
 
-The [rapid cycleway prioritisation tool](https://www.cyipt.bike/rapid/) helps identify road sections for road space reallocation and other interventions to create new cycleways.
+The [rapid cycleway prioritisation tool](https://www.cyipt.bike/rapid/) helps identifys road sections for road space reallocation and other interventions to create new cycleways.
 Used with local knowledge, it provides a strong evidence base to support planning of rapid measures to support safe travel as part of the COVID-19 response.

### 2020-06-04 code/index.Rmd
Update intro sentences

--- 2020-06-04 code/index.Rmd
@@
 
-The [rapid cycleway prioritisation tool](https://www.cyipt.bike/rapid/) helps identify road sections for road space reallocation and other interventions to create new cycleways.
+The [rapid cycleway prioritisation tool](https://www.cyipt.bike/rapid/) helps identify promising locations for road space reallocation in England to create new cycleways.
 Used with local knowledge, it provides a strong evidence base to support planning of rapid measures to support safe travel as part of the COVID-19 response.
@@
 
-The rapid cycleway prioritisation tool aims to help produce evidence-based answers the question: **where to reallocate road space on arterial routes away from motor traffic and towards active transport?**
+The rapid cycleway prioritisation tool aims to help produce evidence-based answers the question: **where to reallocate road space on strategic corridors away from motor traffic to enable uptake of cycling?**
 The focus of the tool is road space reallocation on strategic corridors, a key way in which transport systems can be transformed to overcome capacity constraints on public transport modes and prevent gridlock as the lockdown eases.

### 2020-06-05 code/index.Rmd
Emphasise range of other measures

@@
 Reallocating road space on strategic corridors is just one of the measures available for transport authorities.
-Evidence provided by the tool can also support a range of temporary, experimental and permanent treatments that will enable shift to walking and cycling in the coming weeks and months.
+Evidence provided by the tool can also support a range of temporary, experimental and permanent treatments that will enable rapid and safe uptake of cycling.
 New [statutory guidance on network management in response to COVID-19](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities/traffic-management-act-2004-network-management-in-response-to-covid-19) outlines a range of measures, including the creation of new cycleways, reducing speed limits, widening of footways and closing roads to motor traffic via 'modal filters'.
-Road space reallocation is suited to larger arterial roads and can have a large impact quickly, especially when moving large numbers of people away from public transport and capacity is a priority.
-Furthermore, low but gradually increasing levels of peak hour motor traffic mean that there is a window of opportunity to reallocate lanes without causing congestion.
+Road space reallocation is suited to larger roads and can have a large impact quickly, especially when moving large numbers of people away from public transport and capacity is a priority.
+Furthermore, low but gradually increasing levels of peak hour motor traffic mean that there is a window of opportunity to reallocate lanes without causing congestion in the short term.
 
@@
 
-The focus is on roads that have 'spare space' (defined as those with a spare lane or that have an estimated width of above 10 m) and high cycling potential based on the [Propensity to Cycle Tool](https://www.pct.bike/) [@lovelace_propensity_2017].
-The 10 m width is an indicator of road conditions and larger widths, of 15+ m, may be required for road space reallocation without other changes such as making roads one way for motor traffic or removing [central markings](https://doi.org/10.1016/j.aap.2014.08.015) [@shackel_influence_2014].
-Which roads to prioritise and the most effective interventions will also depend on factors such as on street parking levels not present in the analysis.
+The tool identifies roads that have 'spare space', defined as roads with a spare lane or a width greater than 10 metres and high cycling potential, based on scenarios of mode shift to cycling for work and school in the [Propensity to Cycle Tool](https://www.pct.bike/).
+The 10 m width is an indicator of where it may be possible to introduce new light segregation, depending on levels of on street parking and other factors, without altering the navigable network for motor vehicles.
+On roads that are around 10 m wide or less, other interventions such as 'modal filters' which prevent through traffic for motor vehicles, making roads one way for motor traffic (as implemented successfully with an Experimental Traffic Regulation Order in [Torrington Place](https://www.camden.gov.uk/torrington-tavistock-consultation), London, for example) or removing [central markings](https://doi.org/10.1016/j.aap.2014.08.015) [@shackel_influence_2014] may be required.^[
+The road width required for new cycleways depends on a range of factors including the presence of on street parking. 
+When on street parking is not an issue new cycleways supported by light segregation measures may be possible on narrower roads, as illustrated by the new trial cycleway on [South Road in Lancaster](https://www.facebook.com/photo/?fbid=10158393513440797&set=gm.1409892779205719&__cft__[0]=AZWl60wQjK9GqKMN_1t25YvwumbO3srWVvigT0H8NpP6n_8oaqPUKCLOW-u5ubg7Lg4agxSS7qNwVzEDCsBZYegxRVBGgHky-AkNLR-DJYyytsJpvi16VVEG8aofuaq9-S1ajmZA24d922GG8aapZ1LWv1dJ6N5ROym1TVK9rdB7mQ&__tn__=EH-R), which involved re-allocating carraigeway space on both sides of a ~9 m wide two way road using plastic wands.
+A photo of this intervention can be seen [here](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/lancashire-lane.jpeg).
+]
+Which roads to prioritise and the most effective interventions will also depend on factors such as on street parking levels, safety considerations and symbolic and aesthetic considerations not present in the analysis.
 
-The cycling potential of each road is based on data from the Propensity to Cycle Tool (see [www.pct.bike](https://www.pct.bike/) for further details) and represents the number of one way cycle trips to work and school that could be expected when the [Government's aim of doubling cycling is met](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy).
-The scenario assumes cycling grows evenly nationwide accounting for distance and hilliness.
-Our figures include data on travel to work from 2011 Census data and travel to school. Data on road widths and number of lanes is derived from the Cycling Infrastructure Prioritisation Toolkit.
+The cycling potential of each road is based on the Government Target (equality) scenario from the Propensity to Cycle Tool. 
+The numbers represents the number of one way cycle trips to work and school that could be expected when the [Government's aim of doubling cycling is met](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy).
+Our figures include data on travel to work from 2011 Census data and travel to school, assuming cycling grows evenly nationwide accounting for distance and hilliness [@lovelace_propensity_2017; @goodman_scenarios_2019]
+Data on road widths and number of lanes is derived from the Cycling Infrastructure Prioritisation Toolkit (see [cyipt.bike](https://www.cyipt.bike/)).
 

### 2020-06-05 code/index.Rmd
Update index, finished

@@
 The 10 m width is an indicator of where it may be possible to introduce new light segregation, depending on levels of on street parking and other factors, without altering the navigable network for motor vehicles.
-On roads that are around 10 m wide or less, other interventions such as 'modal filters' which prevent through traffic for motor vehicles, making roads one way for motor traffic (as implemented successfully with an Experimental Traffic Regulation Order in [Torrington Place](https://www.camden.gov.uk/torrington-tavistock-consultation), London, for example) or removing [central markings](https://doi.org/10.1016/j.aap.2014.08.015) [@shackel_influence_2014] may be required.^[
+On roads that are around 10 m wide or less, not highlighted by the tool, other interventions may be appropriate.
+Options include 'modal filters' which prevent through traffic for motor vehicles, making roads one way for motor traffic (as implemented successfully with an Experimental Traffic Regulation Order in [Torrington Place](https://www.camden.gov.uk/torrington-tavistock-consultation), London, for example) or removing [central markings](https://doi.org/10.1016/j.aap.2014.08.015) [@shackel_influence_2014] may be required.^[
 The road width required for new cycleways depends on a range of factors including the presence of on street parking. 
@@
 ]
-Which roads to prioritise and the most effective interventions will also depend on factors such as on street parking levels, safety considerations and symbolic and aesthetic considerations not present in the analysis.
+Which roads to prioritise for road space reallocation and the most effective designs will also depend on factors such as on street parking levels, safety considerations and symbolic and aesthetic considerations not present in the analysis.
 
@@
 
-Overall the evidence presented in the tool is designed to encourage bold interventions where more space for active modes is most needed.
-Currently the results are only available for England.
-<!-- The work is based evidence that road space can be reallocated without causing congestion at a time when traffic levels, particularly during the rush hours, are greatly reduced. -->
-
 # How to use the tool
@@
 
-- Type in the name of the local authority of interest into the search bar above the table below and click on the associated URL.
+- Type in the name of the authority of interest into the search bar above the table below and click on the associated URL.
 - Explore the results in the map (you will find a link to a full screen map) by panning, zooming and clicking on the layers.
 See additional layers by clicking on the layers icon in the top left of the map.
-- Rank the 'top roads' in various ways (including by length and cycling potential) to support prioritisation of road space reallocation schemes.
+- Rank the 'top roads' in various ways (including by length and cycling potential) in the table below, to explore options highlighted by the tool to help prioritise road space reallocation schemes.
 - Interpret the results with care based on other sources of information and an understanding of the limitations of the analysis (see 'Known issues and limitations' below).
-- Refer to the description of the map layers below when exploring the results and complete the [web form](https://urldefense.proofpoint.com/v2/url?u=https-3A__forms.office.com_Pages_ResponsePage.aspx-3Fid-3DqO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u&d=DwMCaQ&c=troKkvwivNn_CddsvWCHHPiPoFoTgTGIbXJULvYU158&r=DXeEYtdSYDBZN2DBwEgAHMmX5PQ_ob9r2MynFWta2mU&m=mkyUa3RIC9d5LSrHUqk_WdstQR1OQuTWyYLBecHFMm0&s=dD7GVcQbjZ2AcdCObSe-rNKmLLDBhu98ChFupIIrlJA&e=) with any feedback.
+- Refer to the description of the map layers below when exploring the results.
 
@@
 1.
-The **Existing cycleways** layer provides an approximation of where cycling infrastructure exists **in the recent past** and is intended to help identify gaps in the network that could be filled with new interventions.
+The **Existing cycleways** layer provides an approximation of where cycling infrastructure exists **currently** (based on OpenStreetMap data downloaded in May 2020) and is intended to help identify gaps in the network that could be filled with new interventions.
 2.
 The **Top ranked new cycleways** layer is the central result of the analysis, providing a list of roads that have high cycling potential, a minimum threshold length, and spare space according to our definition.
-These may be strong candidates for reallocation of road space **immediately or in the near future** to create new cycleways on strategic corridors under the first and second tranches of funding.
+These may be strong candidates for reallocation of road space **immediately or in the near future** to create new cycleways on strategic corridors under the Emergency Active Travel Fund.
 3.
-The **Cohesive network** layer is intended to show what a joined-up cycle network could look like, if road space was not a constraint on creating new infrastructure.
-Composed of roads that have high cycling potential on some or all of their length, the layer is designed to guide **long term planning** towards a continuous and coherent network.
+The **Cohesive network** layer is intended to show what a joined-up cycle network could look like if we were to consider new cycleways by either closing roads to motorised traffic or creating one-way systems.
+Composed of roads that have high cycling potential on some or all of their length, the layer is designed to guide **long term planning**, alongside pre-existing plans such LCWIPs [@departmentfortransport_cycling_2017], towards a continuous and coherent network.
 
@@
 This may be helpful to view in connection with the other layers, because it can reveal where gaps in the existing cycle network could be filled by new infrastructure.
-This layer does not include some on-road cycle lanes or cycleways less than 500m in length, but these can be viewed in one of the basemap options (see below for more information about the basemaps).
+<!-- This layer does not include some on-road cycle lanes or cycleways less than 500m in length, but these can be viewed in one of the basemap options (see below for more information about the basemaps).
+-->
 The definition of 'cycleway' includes shared use paths but excludes cycle lanes painted on roads in this layer.
@@
 They may be suitable candidates for road space reallocation to create new cycleways. 
-These roads have either a spare lane or width of at least 10m, and are ranked according to their mean cycling potential. 
-For routes that are shorter than 1km in length, we use 'km_cycled' instead of mean cycling potential - this is simply the mean cycling potential multiplied by the length of the route. 
+These roads have either a spare lane or an average width greater than 10 m, and are ranked according to their mean cycling potential. 
+<!-- For routes that are shorter than 1km in length, we use 'km_cycled' instead of mean cycling potential - this is simply the mean cycling potential multiplied by the length of the route.  -->
 Top ranked new cycleways are listed in the table below the map, so you can see more information on them. 
-The number of Top ranked new cycleways selected is normally 20, with some variation by region.
+The number of Top ranked new cycleways varies depending on the size of the region.
 
@@
 
+Some road sections highlighted by the tool may not currently be suitable for new cycleways.
+Road sections that are on strategic and major road networks (SRN and MRN respectively) and which had a high speed limit (40 mph+) before the COVID-19 crises can be in the top routes.
+To highlight such roads, the speed limits of road sections are listed in the table of the top ranked new cycleway locations to inform prioritisation and speed reduction interventions.
+
 ## 3. Cohesive network
 
-This layer is designed to represent a cohesive network of direct routes.
+This layer is represents what a joined-up cycle network could look like if new cycleways were placed along corridors of high potential, by either closing roads to motorised traffic or creating one-way systems.
 It shows what a joined-up cycle network could look like if there were no constraints on road space, along some of the most important corridors.
 
-The layer takes no account spare space and is designed to support consideration of alternative options, e.g.
-one way streets, where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
+The cohesive network layer supports consideration of **long term** plans and alternatives to road space reallocation.
+Road sections in the cohesive network but not in the Top ranked new cycleway layer may be suitable candidates for one way streets, where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
 The cohesive network layer was created based on evidence that cohesive cycle networks, that are continuous and do not 'give up' at awkward junctions, are a key component of plans to get people cycling.
 
-The layer was generated by first identifying the roads (classified by reference number) on which there is most cycling potential accross the city and joining them together.
+The layer was generated by first identifying the roads (classified by reference number) on which there is most cycling potential across the city and joining them together.
 Specifically we first selected all road segments with cycle potential in the 85th percentile and above, which also belong to classified trunk roads, A roads or B roads.
@@
 Greater London than in rural areas).
+This layer can be used to identify where there may be spare lanes on the network that could be reallocated to cycling (or pedestrianisation) beyond those highlighted in the Top ranked new cycleways layer.
 
@@
 These routes are at least 10m in width and have a cycling potential above the minimum threshold. 
+Like the Spare lane(s) layer, this layer is intended to highlight road sections on which increasing space for cycling and walking by reallocating road space may be a viable option.
 
@@
 This allows our results to be compared with cycling potentials across the road network.
-The scenarios available show cycling potential for Commuting (government target), Schools (government targets), Commuting (e-bikes), and schools (Go Dutch).
+The scenarios available show cycling potential for Commuting and travel to Schools under the Government Target scenario.
 
@@
 
-- The 'Top ranked new cycleways' highlighted in the map-based results were produced automatically based on a simple score of continuous candidate road length multiplied by cycling potential.
-For roads with a length of more than 1 km, the cycling potential alone is used to rank roads.
+- The 'Top ranked new cycleways' highlighted in the map-based results were produced automatically based on a simple score combined cycling potential for school and commute trips.
 **The results should be interpreted based on local knowledge and an understanding of the limitations of the data.
@@
 Accounting for increases in cycling for other trip purposes such as for shopping and leisure would increase cycling potential substantially.
-- The results rely on grouping roads by proximity, name and other variables.
-Currently some roads that are strong candidates for new cycleways may be missing from the results because of how the algorithm works (we are working on this issue).
-- The road network data was taken from the Cycling Infrastructure Prioritisation Toolkit (CyIPT) project which was completed in 2017, so may be out of date (we are looking at ways to update the data).
-- The results are based on the existing road network and does not flag potential routes that could be constructed.
+- The road network data was taken from the Cycling Infrastructure Prioritisation Toolkit (CyIPT) project which was completed in 2017, so may be out of date.
+- The results are based on the existing road network and does not flag potential routes that could be constructed, e.g. along old railway tracks or across open areas.
+- The results can highlight major roads that have a speed limit above 40mph that may currently be unsuitable for new cycleways created with light segregation.
+Roads not controlled by local authorities on the SRN can also be highlighted (although motorways are excluded); we do not alter the results based on who is in control of the roads.
+- Other than the Cohesive network layer, which highlights what a joined-up cycle network could look like if roads could be made one way to motor traffic and opened-up to cycling via modal filters, the results do not provide evidence on the full range of interventions.
+Identifying locations for filtered permeability interventions in residential areas, for example, requires data on residential networks not contained in this analysis, for example.
 
@@
 
-If you find any other issues with the results or have suggestions of ways to improve the results, please provide feedback on the [web form](https://forms.office.com/Pages/ResponsePage.aspx?id=qO3qvR3IzkWGPlIypTW3y9orQ9urif9IgBaXiRS9eEpUM0JCTDRFTTEwSDRVMUs0VVhVWEdGRDA3RS4u).
 If you are interested in the code underlying the project and would like to support the project with technical input, please provide feedback via the [GitHub issue tracker (requires a GitHub account)](https://github.com/cyipt/popupCycleways/issues).

### 2020-06-05 code/index.Rmd
Update build + ranking

@@
   sf::write_sf(wide_lanes_final, file.path(d, "wide_lanes.geojson"))
+  sf::write_sf(region, file.path(d, "region.geojson"))
+  sf::write_sf(lads, file.path(d, "lads.geojson"))
   message(round(time_to_run), " minutes to build ", rn)

### 2020-06-05 code/region_index.Rmd
Update build + ranking

--- 2020-05-29 code/region_index.Rmd
@@
 
-The map shows a network of key roads with high cycling potential (grey), existing off road cycleways (and shared space paths) according to OpenStreetMap data (green) and top road sections with high cycling potential and spare lanes or estimated width above 10m (blue).
+The map shows three key layers:
+
+1.
+The **Existing cycleways** layer provides an approximation of where cycling infrastructure exists **currently** (based on OpenStreetMap data downloaded in May 2020) and is intended to help identify gaps in the network that could be filled with new interventions.
+2.
+The **Top ranked new cycleways** layer is the central result of the analysis, providing a list of roads that have high cycling potential, a minimum threshold length, and spare space according to our definition.
+These may be strong candidates for reallocation of road space **immediately or in the near future** to create new cycleways on strategic corridors under the Emergency Active Travel Fund.
+3.
+The **Cohesive network** layer is intended to show what a joined-up cycle network could look like if we were to consider new cycleways by either closing roads to motorised traffic or creating one-way systems.
+Composed of roads that have high cycling potential on some or all of their length, the layer is designed to guide **long term planning**, alongside pre-existing plans.
+
 To see additional layers showing road sections with 'spare lanes' and width greater than 10 m, click on the layers icon in the top left, just below the +/- buttons on the map.

### 2020-06-05 code/build.R

@@ -490,7 +490,7 @@ r_lanes_top = r_lanes_joined %>%
     km_cycled_1km = length_up_to_1km * mean_cycling_potential,
     srn = name %in% srn_names_df$roa_number
     ) %>% 
-  arrange(desc(km_cycled_1km)) %>% 
+  arrange(desc(mean_cycling_potential)) %>% 
   slice(1:n_top_roads) 
 nrow(r_lanes_top)
 

### 2020-06-05 code/index.Rmd
Update index text

--- 2020-06-05 code/index.Rmd
@@
 Used with local knowledge, it provides a strong evidence base to support planning of rapid measures to support safe travel as part of the COVID-19 response.
-In the context of COVID-19, big decisions need to be made **rapidly**.
+The rational for the tool is that, in the context of COVID-19, big decisions need to be made quickly.
+The tool aims to support fast yet informed decision making a 'first pass' to help focus planning attention on road sections that are most likely to be suitable for new schemes.
 
@@
   rmarkdown::render("code/region_index.Rmd", output_file = file.path("~/cyipt/tempCycleways/", d, "index.html"))
+  fg = list.files(path = d, pattern = ".geojson", full.names = TRUE)
+  file.remove(fg)
   sf::write_sf(cycleways, file.path(d, "cycleways.geojson"))

### 2020-06-05 code/build.R

@@ -505,13 +505,13 @@ r_lanes_final = r_lanes_joined %>%
       majority_spare_lane ~ labels[2],
       mean_width >= 10 ~ labels[3]
     ),
-    `Estimated width (m)` = case_when(
+    `Estimated width` = case_when(
       mean_width < 10 ~ "<10 m",
       mean_width >= 10 & mean_width < 15 ~ "10-15 m",
       mean_width >= 15 ~ ">15 m"
     )
   ) %>% 
-  select(name, ref, Status, mean_cycling_potential, spare_lane = majority_spare_lane, `Estimated width (m)`, `length (m)` = group_length, group_id, speed_limit)
+  select(name, ref, Status, mean_cycling_potential, spare_lane = majority_spare_lane, `Estimated width`, `length (m)` = group_length, group_id, speed_limit)
 r_lanes_final$Status = factor(r_lanes_final$Status, levels = c(labels[1], labels[2], labels[3]))
 
 table(r_lanes_final$name)
@@ -526,7 +526,7 @@ top_routes = r_lanes_final %>% filter(Status == labels[1])
 spare_lane_groups = r_lanes_final %>% filter(Status == labels[2])
 route_segments_final = rg_new4 %>% 
   mutate(
-    `Estimated width (m)` = case_when(
+    `Estimated width` = case_when(
       mean_width < 10 ~ "<10 m",
       mean_width >= 10 & mean_width < 15 ~ "10-15 m",
       mean_width >= 15 ~ ">15 m"
@@ -563,7 +563,7 @@ pvars_top = c(
   "name",
   "ref",
   "spare_lane",
-  "Estimated width (m)",
+  "Estimated width",
   "mean_cycling_potential",
   "speed_limit",
   "length (m)"

### 2020-06-05 code/index.Rmd
Update text, typo fixes for r

@@
 New [statutory guidance on network management in response to COVID-19](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities/traffic-management-act-2004-network-management-in-response-to-covid-19) outlines a range of measures, including the creation of new cycleways, reducing speed limits, widening of footways and closing roads to motor traffic via 'modal filters'.
-Road space reallocation is suited to larger roads and can have a large impact quickly, especially when moving large numbers of people away from public transport and capacity is a priority.
+Road space reallocation is suited to larger roads and can have a large impact quickly, especially where moving large numbers of people away from public transport and capacity is a priority.
 Furthermore, low but gradually increasing levels of peak hour motor traffic mean that there is a window of opportunity to reallocate lanes without causing congestion in the short term.
@@
 The cycling potential of each road is based on the Government Target (equality) scenario from the Propensity to Cycle Tool. 
-The numbers represents the number of one way cycle trips to work and school that could be expected when the [Government's aim of doubling cycling is met](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy).
+The figures given represent the numbers of one way cycle trips to work and school that could be expected on each road, in a scenario where the [ Government’s aim of an overall doubling of cycling trips has been met](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy).
 Our figures include data on travel to work from 2011 Census data and travel to school, assuming cycling grows evenly nationwide accounting for distance and hilliness [@lovelace_propensity_2017; @goodman_scenarios_2019]
@@
 -->
-The definition of 'cycleway' includes shared use paths but excludes cycle lanes painted on roads in this layer.
+The definition of ‘cycleway’ for the purpose of this layer includes shared use paths but excludes cycle lanes painted on roads.
 
@@
 
-These are the 'top n' road sections in terms of cycling potential with 'spare space' defined above.
+These are the 'top n' road sections in terms of cycling potential with 'spare space' as defined above.
 They may be suitable candidates for road space reallocation to create new cycleways. 

### 2020-06-05 code/index.Rmd
Fix more typos in index, thanks to kg

@@
 bibliography: 
-  # - index.bib
+  - index.bib
   - ../refs.bib
@@
 
-These main layers, and two other layers that informed the results but which are not shown be default, are described in detail below.
+These main layers, and two other layers that informed the results but which are not shown by default, are described in detail below.
 
@@
 
-We used certain eligibility criteria for the Top ranked new cycleways:
+We used the following criteria for the Top ranked new cycleways:
 
-- a road which already has an existing cycleway alongside the majority of its length cannot be a top route. 
-- unnamed roads cannot be a top route 
-- routes less than 500m in length cannot be a top route
+- A road which already has an existing cycleway alongside the majority of its length cannot be a top route.
+- Unnamed roads cannot be a top route.
+- Routes less than 500m in length cannot be a top route.
 
 Some road sections highlighted by the tool may not currently be suitable for new cycleways.
-Road sections that are on strategic and major road networks (SRN and MRN respectively) and which had a high speed limit (40 mph+) before the COVID-19 crises can be in the top routes.
+Road sections that are on strategic and major road networks (SRN and MRN respectively) and which had a high speed limit (40 mph+) before the COVID-19 crisis can be in the top routes.
 To highlight such roads, the speed limits of road sections are listed in the table of the top ranked new cycleway locations to inform prioritisation and speed reduction interventions.
@@
 
-This layer is represents what a joined-up cycle network could look like if new cycleways were placed along corridors of high potential, by either closing roads to motorised traffic or creating one-way systems.
+This layer represents what a joined-up cycle network could look like if new cycleways were placed along corridors of high potential, by either closing roads to motorised traffic or creating one-way systems.
 It shows what a joined-up cycle network could look like if there were no constraints on road space, along some of the most important corridors.
@@
 Road sections in the cohesive network but not in the Top ranked new cycleway layer may be suitable candidates for one way streets, where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
-The cohesive network layer was created based on evidence that cohesive cycle networks, that are continuous and do not 'give up' at awkward junctions, are a key component of plans to get people cycling.
+The cohesive network layer was created to highlight the need for cohesive cycle network that and do not 'give up' at awkward junctions.
+Cohesive networks are a key to getting people cycling [@buehler_bikeway_2016].
 
@@
 - The road network data was taken from the Cycling Infrastructure Prioritisation Toolkit (CyIPT) project which was completed in 2017, so may be out of date.
-- The results are based on the existing road network and does not flag potential routes that could be constructed, e.g. along old railway tracks or across open areas.
+- The results are based on the existing road network. They do not flag potential routes that could be constructed, e.g. along old railway tracks or across open areas.
 - The results can highlight major roads that have a speed limit above 40mph that may currently be unsuitable for new cycleways created with light segregation.
@@
 - Other than the Cohesive network layer, which highlights what a joined-up cycle network could look like if roads could be made one way to motor traffic and opened-up to cycling via modal filters, the results do not provide evidence on the full range of interventions.
-Identifying locations for filtered permeability interventions in residential areas, for example, requires data on residential networks not contained in this analysis, for example.
+Identifying locations for filtered permeability interventions in residential areas, for example, requires data on residential networks not contained in this analysis.
 

### 2020-06-05 code/index.Rmd
shorten landing page text

@@
 
-The rapid cycleway prioritisation tool aims to help produce evidence-based answers the question: **where to reallocate road space on strategic corridors away from motor traffic to enable uptake of cycling?**
+The rapid cycleway prioritisation tool aims to help produce evidence-based answers the question: **where to reallocate road space on strategic corridors to enable uptake of cycling?**
 The focus of the tool is road space reallocation on strategic corridors, a key way in which transport systems can be transformed to overcome capacity constraints on public transport modes and prevent gridlock as the lockdown eases.

### 2020-06-05 report.Rmd
Update report

@@
 author: Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds
-output: bookdown::pdf_document2
+output:
+  bookdown::pdf_document2:
+    toc: true
+    toc_depth: 1
 bibliography: refs.bib
@@
 citr::tidy_bib_file("~/uaf/allrefs.bib", rmd_file = "code/index.Rmd", file = "refs.bib")
-citr::tidy_bib_file("article.Rmd", "~/uaf/allrefs.bib", file = "refs.bib")
+citr::tidy_bib_file("article.Rmd", "~/uaf/allrefs.bib", file = "article-refs.bib")
 browseURL("report.pdf")
@@
 knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
+```
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+```{r park-lane, fig.cap="New space for cycling created by road space reallocation in Park Lane, London.", out.width="80%"}
+# source: Hillingdon and Uxbridge Times
+# https://www.hillingdontimes.co.uk/news/national/18450461.one-londons-prestigious-roads-gets-cycle-lane-due-pandemic/
+knitr::include_graphics(c(
+  "figures/cycleway-image.jpg"
+  ))
 ```
@@
 
-```{r park-lane, fig.cap="New space for cycling created by road space reallocation in Park Lane, London.", out.width="60%"}
-# source: Hillingdon and Uxbridge Times
-# https://www.hillingdontimes.co.uk/news/national/18450461.one-londons-prestigious-roads-gets-cycle-lane-due-pandemic/
-knitr::include_graphics("figures/cycleway-image.jpg")
-```
-
 The £250m released by the Emergency Active Travel Fund will be spent in three broad [ways](https://www.transportxtra.com/publications/local-transport-today/news/65661/dft-releases-first-45m-of-250m-fund/):
@@
 The rapid cycleway prioritisation tool can also support authorities to make bold plans by highlighting the potential of key strategic corridors.
+
+\newpage
 
@@
 - We simplified the legend of the basemap, so that the layers were presented in time order: Existing cycleway (now); Top ranked new cycleways (near term); Cohesive network (long term).
-- We requested data from Highways England to flag the recommendations on the Strategic Road Network (SRN)
+- We requested data from Highways England to flag the recommendations on the Strategic Road Network (SRN) and trialed adding a column to the results.
+Having tested the results we decided to add a column showing current speed limits on road sections instead, highlighting safety issues around potential routes and encouraging consideration speed limit reductions along strategic corridors with high cycling potential.
 

### 2020-06-05 code/index.Rmd
Update description of cycleway data

@@
 The definition of ‘cycleway’ for the purpose of this layer includes shared use paths but excludes cycle lanes painted on roads.
+The quality of the cycle paths is not reported in the data and the presence of a cycleway cannot guarantee that there is adequate provision, because some cycleways included in this dataset are of poor quality and/or are on paths shared by pedestrians that have limited capacity for high cycle traffic while observing physical distancing guidelines.
+Also, because of the variable time lag between new cycleways being built and their appearance on OSM, not all existing cycleways are included in this layer (although most established cycleways in most areas are).
+The data was accessed from OpenStreetMap in May 2020 so is up-to-date in many areas.
 

### 2020-06-05 report.Rmd
Update description of cycleway data

@@
 - We requested data from Highways England to flag the recommendations on the Strategic Road Network (SRN) and trialed adding a column to the results.
-Having tested the results we decided to add a column showing current speed limits on road sections instead, highlighting safety issues around potential routes and encouraging consideration speed limit reductions along strategic corridors with high cycling potential.
+Having tested the results the great majority of regions did not contain routes on the SRN.
+We decided to add a column showing current speed limits on road sections instead, highlighting safety issues around potential routes and encouraging consideration of speed limit reductions along strategic corridors with high cycling potential.
+- We updated the description and framing of the tool to clarify the limitations of the data and its intended use.
 
@@
 Using the methods outlined above, we have derived a set of outputs which are represented as layers in these maps.
+The layers are presented in chronological order to make reading the maps more intuitive, indicating that the data should be used to identify ways to fill key gaps in the existing network in the short term but also plan for a cohesive network in the long term, providing a long-term 'direction of travel' in line with Local Cycling and Walking Plans (LCWIPs), which provide a strategic vision of joined-up networks.
 
@@
 
-The first layer shows existing off-carriageway cycleways.
-It is important to plan new infrastructure in reference to that which already exists.
+The first layer shows existing off-carriageway cycleways are segregated from motor traffic (it does not include painted cycle lanes).
+The quality of the cycle paths is not reported in the data and the presence of a cycleway cannot guarantee that there is adequate provision, because some cycleways included in this dataset are of poor quality and/or are on paths shared by pedestrians that have limited capacity for high cycle traffic while observing physical distancing guidelines.
+Also, because of the variable time lag between new cycleways being built and their appearance on OSM, not all existing cycleways are included in this layer (although most established cycleways in most areas are).
+The data was accessed from OpenStreetMap in May 2020 so is up-to-date in many areas.
+
+The layer highlights the importance of planning new infrastructure in reference to existing networks or routes, which have not yet been joined up to form a cohesive network in most cities.
+New interventions, including new cycleways created by reallocating space on strategic corridors with sufficient space, identified in the next layer, should be prioritised based on their ability to improve cycle network cohesion.
 

### 2020-06-05 code/index.Rmd
change basemaps

--- 2020-06-05 code/index.Rmd
@@
 3.
-The **Cohesive network** layer is intended to show what a joined-up cycle network could look like if we were to consider new cycleways by either closing roads to motorised traffic or creating one-way systems.
+The **Cohesive network** layer is intended to show what a joined-up cycle network could look like if we were to consider using a wider range of interventions to design new cycleways, such as closing roads to motorised traffic or creating one-way systems.
 Composed of roads that have high cycling potential on some or all of their length, the layer is designed to guide **long term planning**, alongside pre-existing plans such LCWIPs [@departmentfortransport_cycling_2017], towards a continuous and coherent network.
@@
 
-There are also seven basemaps to choose from.
-The default basemap is grey and provides an uncluttered view.
-We then have basemap options showing four different scenarios from the Propensity to Cycle Tool.
-This allows our results to be compared with cycling potentials across the road network.
-The scenarios available show cycling potential for Commuting and travel to Schools under the Government Target scenario.
-
-A further basemap option shows all existing cycleways, including on-road cycle lanes. Finally, a satellite image is also available as a basemap.
+There are also five basemaps to choose from. The default basemap is grey and provides an uncluttered view. 
+The second basemap option is a thematic map which shows all existing cycleways, including on-road cycle lanes. 
+We then have two basemap options showing cycling potential for travel to work and travel to school from the Propensity to Cycle Tool. 
+This allows our results to be compared with cycling potentials across the wider road network. 
+These both follow the Government Target scenario; this is the same scenario as is used to derive the map layers described above. 
+Finally, a satellite image is also available as a basemap.
 

### 2020-06-05 code/build.R

@@ -11,14 +11,12 @@ if(!exists("s")) {
   message("Loading global parameters")  
   s = c(
     `Grey basemap` = "Esri.WorldGrayCanvas",
+    `OSM existing cycle provision (CyclOSM)` = "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
     `PCT commuting, Government Target` = "https://npttile.vs.mythic-beasts.com/commute/v2/govtarget/{z}/{x}/{y}.png",
     `PCT schools, Government Target` = "https://npttile.vs.mythic-beasts.com/school/v2/govtarget/{z}/{x}/{y}.png",
-    `PCT commuting, Ebikes, ` = "https://npttile.vs.mythic-beasts.com/commute/v2/ebike/{z}/{x}/{y}.png",
-    `PCT schools, Go Dutch, ` = "https://npttile.vs.mythic-beasts.com/school/v2/dutch/{z}/{x}/{y}.png",
-    `Existing cycleways and cycle lanes` = "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
     `Satellite image` = "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
   )
-  tms = c(FALSE, TRUE, TRUE, TRUE, TRUE, FALSE, FALSE)
+  tms = c(FALSE, FALSE, TRUE, TRUE, FALSE)
   # test basemap:
   tmap_mode("view")
   if(!exists("parameters")) {

### 2020-06-05 code/build.R
simpler regions

@@ -49,13 +49,9 @@ if(!exists("s")) {
 if(!exists("region_name")) {
   region_name = "West Yorkshire"
 }
-if(region_name == "Nottingham") {
-  region = regions %>% filter(str_detect(string = Name, pattern = "Nott")) %>% 
-    st_geometry() %>% 
-    st_union()
-} else {
-  region = regions %>% filter(Name == region_name)
-}
+
+region = regions %>% filter(Name == region_name)
+
 if(as.numeric(sf::st_area(region)) < 3000000000) {
   region = stplanr::geo_buffer(region, dist = 1000)
 }

### 2020-06-05 report.Rmd
Complete report

@@
 author: Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds
-output:
-  bookdown::pdf_document2:
-    toc: true
-    toc_depth: 1
-bibliography: refs.bib
+output: bookdown::word_document2
+# output:
+#   bookdown::pdf_document2:
+#     toc: true
+#     toc_depth: 1
+bibliography:
+ - refs.bib
+ - report-refs.bib
 ---
@@
 ```{r, eval=FALSE, echo=FALSE}
-citr::tidy_bib_file("~/uaf/allrefs.bib", rmd_file = "report.Rmd", file = "refs.bib")
-citr::tidy_bib_file("~/uaf/allrefs.bib", rmd_file = "code/index.Rmd", file = "refs.bib")
+citr::tidy_bib_file("~/uaf/allrefs.bib", rmd_file = "report.Rmd", file = "report-refs.bib")
 citr::tidy_bib_file("article.Rmd", "~/uaf/allrefs.bib", file = "article-refs.bib")
@@
 Carriageway width data were taken from the Cycling Infrastructure Prioritisation Tool (CyIPT) project, funded as a prototype project by a DfT Transport-technology research innovation grants (T-TRIG).
-This uses road data derived from OpenStreetMap.org (OSM)
+This uses road data derived from OpenStreetMap.org (OSM) [@barrington-leigh_world_2017].
 
@@
 
-The layer highlights the importance of planning new infrastructure in reference to existing networks or routes, which have not yet been joined up to form a cohesive network in most cities.
+The Existing cycleways layer highlights the importance of planning new infrastructure in reference to existing networks or routes, which have not yet been joined up to form a cohesive network in most cities.
 New interventions, including new cycleways created by reallocating space on strategic corridors with sufficient space, identified in the next layer, should be prioritised based on their ability to improve cycle network cohesion.
@@
 Its reliance on data rather than community engagement represents a rather top-down approach to transport planning.
-To overcome this issue, future work could seek to incorporate the results such as those presented above into a participatory map of the type shown in Figure x (add widenmypath fig).
+To overcome this issue, future work could seek to incorporate the results such as those presented above into a participatory map of the type shown in Figure \@ref(fig:widenmypath).
 
-As stated above, we focus on road space reallocation, but this is best pursued as part of a wider multi-intervention strategy. 
+```{r widenmypath, out.width="100%", fig.cap="widenmypath.com, an example of participatory tools to support prioritisation of new active transport infrastructure, including new cycleways.
+The open data provided by the project could be overlaid to encourage community feedback on promising options highlighted by our tool."}
+knitr::include_graphics("figures/widenmypath1.png")
+```
+
+The datasets used also have limitations.
+As with any dataset, the OpenStreetMap (OSM) data on which the estimates of road width and number of lanes are based is not a perfect representation of the real world.
+OSM data is contributed by the community and continuously improving so represents a future-proof and easy-to-update solution [@barrington-leigh_world_2017].
+To the best of our knowledge OSM is also the richest national source of combined cycleway *and* road network data in the UK (and most countries) that is open access and free from licensing restrictions.^[
+We are in discussion with the Ordnance Survey and the Department for Transport about ways to include some of their additional datasets in future transport planning tools.
+]
+However, it is important to understand the limitations of OSM data when using the tool.
+OSM data is better in some places than others, the geometry is not always accurate, and the data model makes some simplifying assumptions (shared use paths without high quality surfaces can be labelled as cycleways, for example).
+
+An strength of the tool is it's focus on one aspect of the wider planning to support focussed attention on road space reallocation on strategic corridors.
+This could also be seen as a limitation: road space reallocation should be pursued as part of a wider multi-intervention strategy, and our tool should be used alongside other sources of evidence of other types of intervention: it has a specific scope.
 In urban and residential areas, other interventions such as strategic use of modal filters, road closures, traffic calming, and conversions to one-way working can be used to prevent 'rat-running', enhance filtered permeability, and create Low Traffic Neighbourhoods (image).
 For longer journeys, park-and-cycle schemes can solve 'last-mile' problems and cut congestion.
+Other tools could be developed to support such broader interventions, perhaps partly building on our work (the methods of which are open and modular and therefore extendible).
 
-We have used crowd-sourced open data for road width and number of lanes, which does not give comprehensive coverage of these road attributes in all parts of the country. 
-The 'cohesive network' layer mitigates the impact of any inaccuracies here.
+OSM data on does not give comprehensive coverage of road width and number of lane attributes in all parts of the country. 
+The Cohesive network layer was developed to mitigate the impact of the disjointed nature of roads with spare space and the data underlying the estimation of where there is (and is not) spare space.
 
-Our current methodology only selects roads that have the same name for a continuous stretch of at least 500m, as candidates for top routes.
-Roads that don't meet this criteria can still appear in the 'spare lane' and 'estimated width > 10m' layers.
+The methodology only selects roads that have the same name for a continuous stretch of at least 500m, as candidates for top routes.
+To mitigate this limitation, roads that don't meet this criteria can still appear in the 'spare lane' and 'estimated width > 10m' layers.
 
@@
 
+The tool cannot account for such nuanced policy responses or, currently, distinguish between design options.
+Physical protection of cycleways will improve their, but permeable barriers (such as flexible plastic wands) are preferred to solid barriers that may obstruct pedestrians from crossing the road and encourage speeding.
+Interventions will be most effective where they form part of a wider transport policy and where experimental measures that prove successful can be adopted long-term.
+The tool is designed to inform a 'first pass' in the planning process around new emergency cycleways but deeper consideration is needed when developing strategic plans and specific designs.
+
 # Next steps
 
-This tool can be used together with existing knowledge to identify locations for investment in new cycleways.
+This tool has been developed rapidly to support rapid decision making, and there are many promising potential future directions of travel, including:
 
-Further consideration should be given to community consultation and to other factors such as materials, design, and engineering constraints.
-Physical protection of cycleways will improve their safety, but permeable barriers (such as flexible plastic wands) are preferred to solid barriers that may obstruct pedestrians from crossing the road and encourage speeding.
-Interventions will be most effective where they form part of a wider transport policy and where experimental measures that prove successful can be adopted long-term.
+- Integration with tools for community consultation.
+- Use of new datasets such as traffic volumes (available as 'AADT' counts on many major roads as open data), road casualty data and more exotic datasets such as passively collected movement data from smartphone applications (pending data availability).
+- Inclusion of 'supply side' considerations such as materials (e.g. wands and orcas), specific design options (as demonstrated with the original CyIPT project), and engineering constraints. 
+- Inclusion of additional trip purposes in estimates of cycling potential, that may require updates to or extensions of the Propensity to Cycle Tool.
+
+Perhaps the most promising future direction of travel is towards more flexible tools for transport planning.
+By demonstrating how existing datasets can be combined to create new results that are useful, the project has shown that, like infrastructure planning itself, an 'agile' approach to transport planning tools can yield useful and perhaps unexpectedly successfull results that support local knowledge and transport planning expertise.
+We hope to take lessons learned during the project forward into future, more strategic, transport planning tool research and development work.
 

### 2020-06-05 code/build.R
make maps less bitty

@@ -538,9 +538,12 @@ if(nrow(spare_lanes) == 0) {
 }
 # show wide segments not groups:
 # width_10m = r_lanes_final %>% filter(Status == labels[1])
-wide_lane_groups = r_lanes_final %>% filter(Status == labels[3])
+wide_lane_groups = r_lanes_final %>% 
+  filter(`Estimated width` != "<10 m") %>%
+  filter(Status != labels[1])
 wide_lane_segments = route_segments_final[wide_lane_groups, , op = sf::st_within]
 wide_lanes = wide_lane_segments %>%
+  filter(!idGlobal %in% spare_lanes$idGlobal) %>% 
   filter(width >= 10 & !idGlobal %in% spare_lanes$idGlobal) %>% 
   mutate(Status = labels[3])
 # edge case: there are no wide lanes

### 2020-06-05 report.Rmd
Update image for @mvl22

@@
 author: Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds
-output: bookdown::word_document2
-# output:
-#   bookdown::pdf_document2:
-#     toc: true
-#     toc_depth: 1
+# output: bookdown::word_document2
+output:
+  bookdown::pdf_document2:
+    toc: true
+    toc_depth: 1
 bibliography:
@@
 The open data provided by the project could be overlaid to encourage community feedback on promising options highlighted by our tool."}
-knitr::include_graphics("figures/widenmypath1.png")
+knitr::include_graphics("figures/widenmypath-bristol.png")
 ```

### 2020-06-05 code/index.Rmd
Fix typo on landing page
Fix another typo in index

@@
 The [rapid cycleway prioritisation tool](https://www.cyipt.bike/rapid/) helps identify promising locations for road space reallocation in England to create new cycleways.
-Used with local knowledge, it provides a strong evidence base to support planning of rapid measures to support safe travel as part of the COVID-19 response.
+Used with local knowledge, it provides an evidence base to support planning of rapid measures to support safe travel as part of the COVID-19 response.
 The rational for the tool is that, in the context of COVID-19, big decisions need to be made quickly.
-The tool aims to support fast yet informed decision making a 'first pass' to help focus planning attention on road sections that are most likely to be suitable for new schemes.
+The tool aims to support fast yet informed decision making, by providing a 'first pass' assessment of the existing road network to help focus planning attention on options that are likely to have sufficient space for new infrastructure in places where there is high latent demand.
 
@@
 
-## 3. Spare lane(s) (not shown by default)
+## 4. Spare lane(s) (not shown by default)
 
@@
 
-## 4. Width >= 10m (not shown by default)
+## 5. Width >= 10m (not shown by default)
 

### 2020-06-05 code/index.Rmd
Update index

--- 2020-06-05 code/index.Rmd
@@
 
-The [rapid cycleway prioritisation tool](https://www.cyipt.bike/rapid/) helps identify promising locations for road space reallocation in England to create new cycleways.
-Used with local knowledge, it provides a strong evidence base to support planning of rapid measures to support safe travel as part of the COVID-19 response.
-The rational for the tool is that, in the context of COVID-19, big decisions need to be made quickly.
-The tool aims to support fast yet informed decision making a 'first pass' to help focus planning attention on road sections that are most likely to be suitable for new schemes.
+The [rapid cycleway prioritisation tool](https://www.cyipt.bike/rapid/) helps identify promising locations for road space reallocation in England.
+The tool provides an evidence base to support investment in cycling infrastructure, in response to new pressures placed on transport and public health systems by the COVID-19 crisis, at local and regional levels.
+The rational for the tool is that, in the context of COVID-19, big decisions around where to prioritise investment on the road network need to be made quickly.
+The tool provides a 'first pass' assessment of the existing road network to help focus planning attention on options that are likely to have sufficient space for new infrastructure in places where there is high latent demand.
 

### 2020-06-08 code/index.Rmd
Shorten intro

@@
 library(tidyverse)
+library(dplyr)
 library(tmap)
@@
 The rational for the tool is that, in the context of COVID-19, big decisions around where to prioritise investment on the road network need to be made quickly.
-The tool provides a 'first pass' assessment of the existing road network to help focus planning attention on options that are likely to have sufficient space for new infrastructure in places where there is high latent demand.
+The tool provides a ‘first pass’ assessment of promising locations for new cycleways which have high latent demand for cycling and where road space can be reallocated whilst maintaining two-way traffic flows for motorists.
 
@@
 ```{r}
-dn = "popupCycleways/v1" # dirctory name
+dn = "popupCycleways/v1.1" # dirctory name
 built_regions = list.files(dn)

### 2020-06-09 report.Rmd
Update title

@@
 ---
-title: "Data analysis to prioritise new cycleways in response to COVID-19"
+title: "Rapid cycleway prioritisation tool"
+subtitle: "Technical specification"
 author: Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds

### 2020-06-09 report.Rmd
Update report intro

@@
 The statutory guidance on network management in response to COVID-19 outlines a range of measures that can support a safe and healthy restart, including the creation of new cycleways, reducing speed limits, widening of footways and closing roads to motor traffic via 'modal filters'.
-**This report describes the rapid cycleway prioritisation tool project, which provides aims to ensure evidence-based and effective investment in first of these measures: the creation of new cycleways in existing carriageway space reclaimed from motor traffic.**
-Light segregation infrastructure such as flexible plastic wands (illustrated in Figure 1) and bollards can deliver a step change in provision for cycling, when implemented in combination with other measures such reduced speed limits on key corridors (as has happened on the A33 in Southampton and the A56 in Manchester) and when placed where they are most needed.
-**The purpose of the tool is to help identify such key corridors, ensuring that funds from the Emergency Active Travel Fund change the status quo of urban and rural transport systems in favour of active travel.**
+**This report describes the rapid cycleway prioritisation tool project, which provides aims help authorities effectively prioritise locations of new cycleways created by reallocating road space.**
+<!-- Light segregation infrastructure such as flexible plastic wands (illustrated in Figure 1) and bollards can deliver a step change in provision for cycling, when implemented in combination with other measures such reduced speed limits on key corridors (as has happened on the A33 in Southampton and the A56 in Manchester) and when placed where they are most needed.
+-->
 
@@
 <!-- https://twitter.com/doddyuk/status/1267146449227255808 -->
+<!-- The purpose of the tool is to help identify such key corridors, ensuring that funds from the Emergency Active Travel Fund change the status quo of urban and rural transport systems in favour of active travel.
+-->
 
-The £250m released by the Emergency Active Travel Fund will be spent in three broad [ways](https://www.transportxtra.com/publications/local-transport-today/news/65661/dft-releases-first-45m-of-250m-fund/):
+The Emergency Active Travel Fund allocated £225m for new infrastructure in two [tranches](https://www.transportxtra.com/publications/local-transport-today/news/65661/dft-releases-first-45m-of-250m-fund/):
 
-- The first tranche: £45m released on 1^st^ June, is for swift interventions that must be implemented by the end of June
-- The second tranche: £180m will be released later in the summer for more permanent measures
-- £25m will be invested in cycle repair schemes
+- The first tranche: £45m allocated for swift interventions for which applications were submitted by authorities on 5^th^ June.
+- The second tranche: £180m will be released later in the summer for permanent measures
 
-The emphasis of the evidence presented in this report is providing a strong evidence base to prioritisation of schemes eligible for the second tranche.
-However, the evidence could also be used to support identification of schemes for the first tranche, such as reallocation of road space on arterial routes, where decisions have not yet been made.
-
+The Rapid Cycleway Prioritisation Tool was designed to provide evidence in support of schemes eligible for the second tranche of funding.
 Given the importance of acting quickly, and the unique opportunity that the crisis poses to improve physical activity levels in the long term, it is vital that local authorities have the evidence they need.

### 2020-06-09 report.Rmd
Link to web page

@@
 
-The Rapid Cycleway Prioritisation Tool was designed to provide evidence in support of schemes eligible for the second tranche of funding.
+The [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/) was designed to provide evidence in support of schemes eligible for the second tranche of funding.
 Given the importance of acting quickly, and the unique opportunity that the crisis poses to improve physical activity levels in the long term, it is vital that local authorities have the evidence they need.
 Without a nationally consistent and systematic evidence base, there is a risk that some already stretched local authorities will rely on ad-hoc methods to allocate funding.
-The rapid cycleway prioritisation tool can also support authorities to make bold plans by highlighting the potential of key strategic corridors.
+The tool can also support authorities to make bold plans by highlighting the potential of key strategic corridors.
 

### 2020-06-09 report.Rmd
Local authority -> authority

@@
 
-The approach was designed to identify roads and road segments where investment in new cycleways should be prioritised using a transparent methodology and nationally consistent datasets, alongside local authority expertise and local knowledge.
+The approach was designed to identify roads and road segments where investment in new cycleways should be prioritised using a transparent methodology and nationally consistent datasets, alongside planning authority expertise and local knowledge.
 Strong and actionable datasets already exist to support strategic cycle network planning, notably the Propensity to Cycle Tool (PCT), a national planning support system [@lovelace_propensity_2017].
@@
 
-We elicited feedback from stakeholders via the Basecamp forum for local authority transport planners.
+We elicited feedback from stakeholders via the Basecamp forum for authority transport planners.
 As a result of the feedback we made a few adjustments to the results:

### 2020-06-09 code/chelt.R
minor

@@ -26,8 +26,7 @@ long_list = lapply(lgs, FUN = function(i) {
   # mapview::mapview(lg)
   l_buff = lg %>%
     st_transform(27700) %>%
-    st_buffer(buff_dist_large) %>%
-    st_transform(4326)
+    st_buffer(buff_dist_large) 
   touching_list = st_intersects(l_buff)
   g = igraph::graph.adjlist(touching_list)
   components = igraph::components(g)
@@ -47,7 +46,7 @@ lg_new2 = lg_new %>%
   group_by(group3) %>%
   mutate(
     group3_length = sum(length),
-    new_group3 = group3
+    new_group3 = as.factor(group3)
   ) %>%
   # summarise(
   #   group2 = first(group2),
@@ -77,7 +76,6 @@ while (min(lg_c$group3_length) < min_grouped_length) {
   nearest = NULL
   near = NULL
   mindist = NULL
-  shortest_l = NULL
   nearest_l = NULL
   new_l = NULL
   new_geom = NULL
@@ -85,8 +83,8 @@ while (min(lg_c$group3_length) < min_grouped_length) {
   new_c = NULL
   
   shortest_centroid = lg_c[which.min(lg_c$group3_length), ]
-  shortest_segment = lg_new2[which.min(lg_c$group3_length), ]
-  if(dim(shortest_centroid)[1] > 1) shortest = shortest[1,]
+  shortest_segment = lg_new2[which.min(lg_new2$group3_length), ]
+  if(dim(shortest_centroid)[1] > 1) shortest_centroid = shortest_centroid[1,]
   near = lg_c %>% 
     filter(group2 == shortest_centroid$group2) %>% 
     filter(new_group3 != shortest_centroid$new_group3) # must remove the point itself from this group

### 2020-06-09 report.Rmd
Update methods part 1

@@
 The approach was designed to identify roads and road segments where investment in new cycleways should be prioritised using a transparent methodology and nationally consistent datasets, alongside planning authority expertise and local knowledge.
-Strong and actionable datasets already exist to support strategic cycle network planning, notably the Propensity to Cycle Tool (PCT), a national planning support system [@lovelace_propensity_2017].
-The PCT is already used by local authorities and contains key data showing not only current levels of cycling but also scenarios of growth, notably the Government Target scenario of doubling the number of cycling stages made by 2025 [@departmentfortransport_cycling_2017].
+Strong and actionable datasets already exist to support strategic cycle network planning, notably the Propensity to Cycle Tool (PCT), a national planning support system.
+The PCT is already used by local authorities and contains key data showing not only current levels of cycling but also scenarios of growth in usage of cycling for everyday trips (for commuting and school travel).
+The scenarios include Government Target, which represents a doubling the number of cycling stages made by 2025 [@departmentfortransport_cycling_2017].
 The PCT maps cycle uptake at area, route and route network levels, for commuting and school travel (as of May 2020) [@goodman_scenarios_2019].
@@
 
-The rapid cycleway prioritisation tool is based on levels of cycling potential *and* road attribute data, with an emphasis on road space reallocation.
-The method was parameterised to highlight roads that have sufficient space to allow for new cycleways without affecting the navigable network for motorised modes [@lovelace_methods_2020].
-A benefit of this method is that it allows continuous road sections to be identified and ranked, supporting prioritisation of investment.
-Roads can be ranked based not only on cycling potential, but also based on continuous road length and estimated cycle usage measured in km per day.
-The tool highlights continuous road sections with high cycling potential in the context of existing infrastructure on the basis that new infrastructure should maximise cycle network cohesion and continuity [@parkin_designing_2018].
+The rapid cycleway prioritisation tool identifies roads which can accommodate new cycleways whilst maintaining 2-way motorized traffic.
+Of these roads, it identifies priority locations for new cycleways that are expected to receive most usage under the 'Government Target' scenario.
+The amount of usage under modelled scenarios of cycling uptake, based on evidence of the percentage of trips cycled in relation to the distance and hilliness of the trip, is known as 'cycling potential' [@lovelace_propensity_2017].
+
+The tool identifies continuous road sections and ranks them according to their cycling potential.
+Roads can also be ranked by length, on the basis that interventions on long, continuous road sections will tend to have a large impact on the overall transport network, and the estimated amount of cycling in units of 'km per day', which is simply cycling potential multiplied by section length.
+The aim is to provide evidence that maximises the impact of new interventions in creating a joined-up and continuous, a vital component of cycling strategies [@parkin_designing_2018].
 

### 2020-06-09 report.Rmd
Complete road attribute section

@@
 
-To identify road sections on which there is a spare lane we developed a simple algorithm that takes the OSM variable [`lanes`](https://wiki.openstreetmap.org/wiki/Key:lanes) if it is present and, if not, derives the number from the highway type and presence/absence of bus lanes.
-All segments defined as having a spare lane using this method are shown in Figure \@ref(fig:levels) (left).
-We also included estimates of road width, only flagging roads with an average estimated width of more than 10m (the minimum that should be considered for cycleways without additional changes).
-In future, this methodology could be enhanced to include verge and pavement width estimates, which could increase the effective available space for new cycleways, and the presence of vehicle parking bays, and on street parking which reduce it.
+The tool highlights roads that have sufficient space for new infrastructure, by which we mean either having more than one lane in either direction *or* an average estimated width of more than 10 metres.
+There are three main reasons for this focus on roads with 'spare space':
+
+- It is possible to rapidly implement new cycleways on wide roads through a range of interventions, including light segregation along one or both sides of the road or point closures.
+There is a window of opportunity to get people cycling and the creation of cycleways by road space reallocation on wide roads with high cycling potential is one of the fastest and most effective ways to maximise mode switch to cycling during the 'restart'.
+- Wide roads tend to be direct, continuous and located along strategic corridors, where cycling potential (including cycling trips replacing trips previously made by public transport) is likely to be highest.
+- The cycleways resulting from space reallocation on such roads are likely to be sufficiently wide, continuous and direct to enable large numbers of people to switch to cycling while observing physical distancing guidelines, potentially reducing pressure on public transport on strategic corridors.
+
+<!-- To identify road sections on which there is a spare lane we developed a simple algorithm that takes the OSM variable [`lanes`](https://wiki.openstreetmap.org/wiki/Key:lanes) if it is present and, if not, derives the number from the highway type and presence/absence of bus lanes.
+-->
+We defined roads with 'spare lanes' as those on which there was at least 2 lanes (including bus lanes) going in the same direction, based on data from [OpenStreetMap](https://www.openstreetmap.org).
+We also included estimates of road width, only flagging roads with an average estimated width of more than 10 m using data that had previously been generated for the Cycling Infrastructure Prioritisation Tool.
+10 m was selected because it is roughly the minimum width required for new cycleways created by road space reallocation measures, although there are examples of roads that are less than 10 m on which new cycleways have been successfully created in response to COVID-19.^[
+[South Road](https://www.lancasterguardian.co.uk/news/uk-news/mixed-reactions-new-lancaster-pop-cycle-lanes-busy-city-centre-road-2875909#gsc.tab=0) in Lancaster, for example, has an average width of around 8 m, yet protected cycleways in either direction have been created by installation of wands on either side.
+]
+The tool does not account for space currently occupied by on street parking, verges and pavements, which could provide additional space for cycling interventions.
+
+<!-- Usingcould increase the effective available space for new cycleways, and the presence of vehicle parking bays, and on street parking which reduce it. -->
+
+<!-- All segments defined as having a spare lane using this method are shown in Figure \@ref(fig:levels) (left). -->
 
@@
 Interventions may still be possible on narrow roads, for example the removal of central line markings has been shown to reduce average vehicle speeds (ref). 
-As an example of cycle infrastructure on a narrow road, South Road in Lancaster is recorded on OSM as having road width <10m, yet protected cycleways are currently being constructed there (image). 
 In other cases, it may be possible to convert a two-way road to one-way working and re-allocate the space previously occupied by one of the traffic lanes to bi-directional cycleways.

### 2020-06-09 report.Rmd
Update top routes section

@@
 
-For our three core outputs of near-term cycleway prioritisation, we follow a series of grouping and filtering steps, to identify roads that have both high cycle potential and available road space. 
-Segments are grouped by their road reference number, with stricter filtering criteria for roads with no reference number, to avoid a mass of short road segments on side streets. 
-We use a 100m buffer to further group these segments, as shown in Figure \@ref(fig:levels) (middle and right, respectively), which shows that grouping nearby road segments before filtering results in a more cohesive network.
+<!-- For our three core outputs of near-term cycleway prioritisation, we follow a series of grouping and filtering steps, to identify roads that have both high cycle potential and available road space.
+-->
+<!-- Segments are grouped by their road reference number, with stricter filtering criteria for roads with no reference number, to avoid a mass of short road segments on side streets.  -->
+<!-- We use a 100m buffer to further group these segments, as shown in Figure \@ref(fig:levels) (middle and right, respectively), which shows that grouping nearby road segments before filtering results in a more cohesive network.
+-->
 
-We then remove groups which do not have either an estimated mean road width of at least 10m or a spare lane for the majority of their length. 
-Groups with mean cycling potential below a minimum threshold are also removed. This threshold varies according to cycling levels in each region. 
-Finally, any segments not part of a wider group of > 500m length (using a 100m buffer) are removed.
+<!-- We then remove groups which do not have either an estimated mean road width of at least 10m or a spare lane for the majority of their length.  -->
+<!-- Groups with mean cycling potential below a minimum threshold are also removed. This threshold varies according to cycling levels in each region.  -->
+<!-- Finally, any segments not part of a wider group of > 500m length (using a 100m buffer) are removed. -->
 
 ```{r levels, out.width="90%", fig.cap="Illustration of the 'group then filter' method to identify long sections with spare lanes *and* high cycling potential"}
-knitr::include_graphics("figures/filter-then-group.png")
+# knitr::include_graphics("figures/filter-then-group.png")
 ```
 
-## Selection of top routes
+## Top ranked new cycleways
 
-From the results of the previous steps, we can now select the top routes for cycling. 
-Only roads that have the same name for a continuous stretch of at least 500m are eligible for selection as a top route. 
-Roads cannot be candidates for top routes if there is an existing cycleway along 75% or more of their length. 
-We selected such a high value because the presence of a cycleway in OSM does not guarantee that this is a high quality, continuous route.
-Top routes are ranked according to their cycling potential.
+To identify wide and continuous roads with high cycling potential we first excluded roads with low cycling potential from the analysis.
 
-Roads not selected as top routes are shown in the layers named 'spare lane' and 'estimated width >10m.'
+Roads with high cycling potential but insufficient width were excluded from the 'Top ranked new cycleways' layer but were included in the 'Cohesive network' layer, described in the next section and on the tools main web page at [www.cyipt.bike/rapid](https://www.cyipt.bike/rapid/#map-layers).
 
-## Identification of cohesive cycle network
+Only roads with a continuous length above a threshold distance (500 m) identified as candidates for inclusion in the Top ranked new cycleways layer. 
+Roads with an existing cycleway along 75% or more of their length were excluded (we selected such a high value because the presence of a cycleway in OSM does not guarantee that this is a high quality, continuous route).
+Roads were ranked in terms of their cycling potential and highest ranking roads were included in the roads.
+The number of roads identified as Top ranked new cycleways varies depending on the size of the authority, ranging from 4 in the [Isle of Wight](https://www.cyipt.bike/rapid/isle-of-wight/) to 80 in [Greater London](https://www.cyipt.bike/rapid/greater-london/).
 
-The cohesive network is derived using a different method from the other layers. 
+Roads that met the spare space criteria but which were not selected as top routes were kept and can be shown in the layers named 'spare lane' and 'estimated width >10m', which are hidden by default to simplify the visualisation of results and to focus attention on the most important results, including the Cohesive network.
+
+## Identification of a cohesive network
+
+The Cohesive network layer was derived using a different method from the other layers. 
 To identify this network, we select road segments on the trunk road, A road and B road network that have cycling potential in the 85th percentile or above, for their region. 

### 2020-06-09 report.Rmd

@@
 
-Roads that met the spare space criteria but which were not selected as top routes were kept and can be shown in the layers named 'spare lane' and 'estimated width >10m', which are hidden by default to simplify the visualisation of results and to focus attention on the most important results, including the Cohesive network.
-
-## Identification of a cohesive network
+## Cohesive network
 
 The Cohesive network layer was derived using a different method from the other layers. 
-To identify this network, we select road segments on the trunk road, A road and B road network that have cycling potential in the 85th percentile or above, for their region. 
-We then join together these segments using a 1000m buffer, to form a coherent network.
-This is thus based on the cycling potential of roads, without reference to the road width or number of lanes.
+To identify this network, we selected 'through traffic' roads, primarily A roads and B roads, that had high cycling potential (in the 85th percentile or above for all roads in the authority) on some or all of their length.
+We then join together these segments together to form a coherent network.
+<!-- using a 1000m buffer -->
+The resulting layer shows what a joined up cycle network could look like if new cycleways were placed along corridors of high potential, by either closing roads to motorised traffic or creating one-way systems.
+
+## Roads with spare space
+
+Roads that met the spare space criteria but which were not selected as top routes were kept and can be shown in the layers named 'spare lane' and 'estimated width >10m'.
+These results are hidden by default to simplify the visualisation of results and to focus attention on the most important results, including the Cohesive network.
+They represent roads on which there is more than one lane in one direction, and roads with an estimated carriageway width of 10 m or more, respectively.
+They can be visualised in the map by clicking on the layers button in the top lefthand corner of the map.
 
@@
 
-- We refined the road identification method, to highlight more central road sections and fewer long sections far from high density areas, in response to feedback suggesting that some of the road sections went too far into the countryside.
-- We simplified the legend of the basemap, so that the layers were presented in time order: Existing cycleway (now); Top ranked new cycleways (near term); Cohesive network (long term).
-- We requested data from Highways England to flag the recommendations on the Strategic Road Network (SRN) and trialed adding a column to the results.
-Having tested the results the great majority of regions did not contain routes on the SRN.
-We decided to add a column showing current speed limits on road sections instead, highlighting safety issues around potential routes and encouraging consideration of speed limit reductions along strategic corridors with high cycling potential.
+- We refined the road identification method, to highlight more roads close to city centres, in response to feedback suggesting that some of the road sections went too far into the countryside.
+- We changed the legend of the basemap, so that the layers were presented in time order: Existing cycleway (now); Top ranked new cycleways (near term); Cohesive network (long term).
+- We added a column showing current speed limits on road sections, highlighting safety issues around potential routes and encouraging consideration of speed limit reductions along strategic corridors with high cycling potential.
 - We updated the description and framing of the tool to clarify the limitations of the data and its intended use.
@@
 Named roads with the highest cycling potential are ranked and selected as top routes.
-For the remaining roads, we identify whether they have a spare lane for the majority of their length, or an estimated mean width >10m.
+For the remaining roads, we identified whether they have a spare lane for the majority of their length, or an estimated mean width >10m.
+The resulting layers, are hidden by default, were used to help identify where 'connecting' cycleway sections' that have spare space but do not meet the criteria for inclusion in the 'top routes' layer that could help form part of a cohesive network.
 
@@
 This provides the ability to identify roads that may be key priorities today for cycle infrastructure, but which are missing from the other results, either due to the limitations of our road attribute data, or because the roads are simply too narrow to show up in the other results.
-Interventions may still be possible on narrow roads, for example the removal of central line markings has been shown to reduce average vehicle speeds (ref). 
-In other cases, it may be possible to convert a two-way road to one-way working and re-allocate the space previously occupied by one of the traffic lanes to bi-directional cycleways.
-(Torrington Place image)
+Interventions may still be possible on narrow roads, for example the removal of central line markings has been shown to reduce average vehicle speeds [@shackel_influence_2014]. 
+In other cases, it may be possible to convert a two-way road to one-way working and re-allocate the space previously occupied by one of the traffic lanes to bi-directional cycleways, as happened in a successful Experimental Traffic Regulation Order (ETRO) intervention in Torrington Place, [Cambden](https://www.camden.gov.uk/documents/20142/18667687/14-1+-+Licensed+Taxi+Drivers%27+Association+%28LTDA%29.pdf/20bc7ead-5ccc-13ed-66f4-e713eee4e103).
 
-Finally the cohesive network can help identify gaps in existing and/or proposed cycle infrastructure. 
-Gappy networks are off-putting to many cyclists, and busy or complex road junctions in particular can create barriers to cycle movement (ref).
+Finally, the cohesive network can help identify gaps in existing and/or proposed cycle infrastructure. 
+'Gappy' networks are off-putting to many cyclists, and busy or complex road junctions in particular can create barriers to cycle movement [@buehler_bikeway_2016].
 
@@
 The approach is not without limitations.
-Its reliance on data rather than community engagement represents a rather top-down approach to transport planning.
-To overcome this issue, future work could seek to incorporate the results such as those presented above into a participatory map of the type shown in Figure \@ref(fig:widenmypath).
+Its reliance on data rather than community engagement represents a rather top-down approach to transport planning, meaning that the results should be sense-checked by people with local knowledge and feedback from stakeholders where timescales permit.
+
+Participatory maps are one way to crowd source such stakeholder feedback, as shown in Figure \@ref(fig:widenmypath).
+The data from the tool could also feed into such participatory tools.
 
@@
 As with any dataset, the OpenStreetMap (OSM) data on which the estimates of road width and number of lanes are based is not a perfect representation of the real world.
-OSM data is contributed by the community and continuously improving so represents a future-proof and easy-to-update solution [@barrington-leigh_world_2017].
+OSM data is contributed by the OSM community and continuously improving so represents a future-proof and easy-to-update solution [@barrington-leigh_world_2017].
 To the best of our knowledge OSM is also the richest national source of combined cycleway *and* road network data in the UK (and most countries) that is open access and free from licensing restrictions.^[
@@
 
-An strength of the tool is it's focus on one aspect of the wider planning to support focussed attention on road space reallocation on strategic corridors.
-This could also be seen as a limitation: road space reallocation should be pursued as part of a wider multi-intervention strategy, and our tool should be used alongside other sources of evidence of other types of intervention: it has a specific scope.
-In urban and residential areas, other interventions such as strategic use of modal filters, road closures, traffic calming, and conversions to one-way working can be used to prevent 'rat-running', enhance filtered permeability, and create Low Traffic Neighbourhoods (image).
+It is important to remember that the Cycleway Prioritisation Tool focuses on only one aspect of the wider planning process.
+It focusses attention on the important question of road space reallocation on strategic corridors with high cycling potential.
+This focus on strategic corridors has advantages but could also be seen as a limitation: road space reallocation is only one part of the range of interventions needed to get people cycling.
+
+The tool should be used alongside other sources of evidence and alongside plans for other types of intervention, outlined in new [statutory guidance](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities).
+In urban and residential areas, other interventions include use of modal filters, point closures of roads, traffic calming, and making roads oneway for motor traffic.
+Future work could help prioritise modal filters to prevent 'rat-running', enhance filtered permeability, and create Low Traffic Neighbourhoods.
 For longer journeys, park-and-cycle schemes can solve 'last-mile' problems and cut congestion.
@@
 
-OSM data on does not give comprehensive coverage of road width and number of lane attributes in all parts of the country. 
-The Cohesive network layer was developed to mitigate the impact of the disjointed nature of roads with spare space and the data underlying the estimation of where there is (and is not) spare space.
-
-The methodology only selects roads that have the same name for a continuous stretch of at least 500m, as candidates for top routes.
+As stated above, OpenStreetMap data on does not give comprehensive coverage in all parts of the country. 
+The Cohesive network layer was developed to mitigate the disjointed nature of roads with spare space, but this layer does not include potential new routes outside the current road network.
+The methodology only selects roads that have the same name for a continuous stretch of at least 500m, as candidates for top routes, meaning that some roads will be excluded from the 'top routes' result due to name changes (although most continuous road sections have consistent names).
 To mitigate this limitation, roads that don't meet this criteria can still appear in the 'spare lane' and 'estimated width > 10m' layers.
@@
 Vehicular traffic volumes may restrict the potential for road space reallocation, and the presence of on-street car parking may prevent the use of particular alignments. 
-While we do not account for these limitations, the reduction in traffic due to COVID-19 has brought about both a unique opportunity and an urgent necessity to change travel patterns and behaviours in profound ways.
+While we do not account for this, the reduction in traffic due to COVID-19 has brought about both a unique opportunity and an urgent necessity to change travel patterns and behaviours in profound ways.
 As an example of the changes are happening right now, Glasgow has recently suspended around a third of its on-street parking spaces, to widen pavements and increase space for social distancing.
@@
 The tool cannot account for such nuanced policy responses or, currently, distinguish between design options.
-Physical protection of cycleways will improve their, but permeable barriers (such as flexible plastic wands) are preferred to solid barriers that may obstruct pedestrians from crossing the road and encourage speeding.
-Interventions will be most effective where they form part of a wider transport policy and where experimental measures that prove successful can be adopted long-term.
-The tool is designed to inform a 'first pass' in the planning process around new emergency cycleways but deeper consideration is needed when developing strategic plans and specific designs.
+After a network of routes has been identified using a range of sources of evidence, including this tool, a number of other policy and design decisions need to be made, that are beyond the scope of this tool.
+Overall, interventions will be most effective where they form part of a wider transport policy and where measures can be adopted long-term.
+
+The coming 'restart' will lead to long lasting changes in transport systems.
+Whether these changes include a worsening of air pollution, traffic and road traffic casualty rates, or uptake of active transport is an open question that depends effective and timely interventions.
+The Rapid Cycleway Prioritisation Tool provides an evidence base in support of effective and timely interventions that will form part of broader strategies to respond to COVID-19.
 

### 2020-06-09 report.Rmd
Add additional reference file

@@
  - report-refs.bib
+ - code/index.bib
 ---

### 2020-06-09 code/index.Rmd
Fix #89

@@
 ---
-title: "Rapid cycleway prioritisation tool"
+title: "Rapid Cycleway Prioritisation Tool"
 author: "Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds"
@@
 
-The [rapid cycleway prioritisation tool](https://www.cyipt.bike/rapid/) helps identify promising locations for road space reallocation in England.
+The [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/) helps identify promising locations for road space reallocation in England.
 The tool provides an evidence base to support investment in cycling infrastructure, in response to new pressures placed on transport and public health systems by the COVID-19 crisis, at local and regional levels.
@@
 
-The rapid cycleway prioritisation tool aims to help produce evidence-based answers the question: **where to reallocate road space on strategic corridors to enable uptake of cycling?**
+The Rapid Cycleway Prioritisation Tool aims to help produce evidence-based answers the question: **where to reallocate road space on strategic corridors to enable uptake of cycling?**
 The focus of the tool is road space reallocation on strategic corridors, a key way in which transport systems can be transformed to overcome capacity constraints on public transport modes and prevent gridlock as the lockdown eases.

### 2020-06-09 report.Rmd
Fix #89

--- 2020-06-09 report.Rmd
@@
 ---
-title: "Rapid cycleway prioritisation tool"
+title: "Rapid Cycleway Prioritisation Tool"
 subtitle: "Technical specification"
@@
 The statutory guidance on network management in response to COVID-19 outlines a range of measures that can support a safe and healthy restart, including the creation of new cycleways, reducing speed limits, widening of footways and closing roads to motor traffic via 'modal filters'.
-**This report describes the rapid cycleway prioritisation tool project, which provides aims help authorities effectively prioritise locations of new cycleways created by reallocating road space.**
+**This report describes the Rapid Cycleway Prioritisation Tool project, which provides aims help authorities effectively prioritise locations of new cycleways created by reallocating road space.**
 <!-- Light segregation infrastructure such as flexible plastic wands (illustrated in Figure 1) and bollards can deliver a step change in provision for cycling, when implemented in combination with other measures such reduced speed limits on key corridors (as has happened on the A33 in Southampton and the A56 in Manchester) and when placed where they are most needed.
@@
 A limitation of the PCT is that it takes little account of existing infrastructure: variables such as road width and even the presence/absence of cycle infrastructure currently are not included in the results.
-The rapid cycleway prioritisation tool overcomes this issue by using road network data which includes geographic attributes and cycle infrastructure from OpenStreetMap.
+The Rapid Cycleway Prioritisation Tool overcomes this issue by using road network data which includes geographic attributes and cycle infrastructure from OpenStreetMap.
 Carriageway width data were taken from the Cycling Infrastructure Prioritisation Tool (CyIPT) project, funded as a prototype project by a DfT Transport-technology research innovation grants (T-TRIG).
@@
 
-The rapid cycleway prioritisation tool identifies roads which can accommodate new cycleways whilst maintaining 2-way motorized traffic.
+The Rapid Cycleway Prioritisation Tool identifies roads which can accommodate new cycleways whilst maintaining 2-way motorized traffic.
 Of these roads, it identifies priority locations for new cycleways that are expected to receive most usage under the 'Government Target' scenario.

### 2020-06-09 code/index.Rmd
Add Use the tool section

@@
 
+# Use the tool
+
+To use the tool, click on the region of interest below and click on the link ([www.cyipt.bike/rapid/west-yorkshire](https://www.cyipt.bike/rapid/west-yorkshire/) for West Yorkshire, for example).
+Explore the results in the new page that appears with reference to the guidance on this page and the [Rapid Cycleway Prioritisation Tool report](https://github.com/cyipt/popupCycleways/releases/download/0.3/report.pdf).
+
 ```{r}

### 2020-06-09 code/index.Rmd
Close #70

@@
 ```{r}
-pvars = c("Name", "Level")
-tm_shape(regions) + tm_polygons("Level", popup.vars = pvars) 
+# pvars = c("Name", "Level")
+# tm_shape(regions) + tm_polygons("Level", popup.vars = pvars) 
   # + tm_shape(regions_centroids) + tm_text("Name", size = 0.7)
+library(leaflet)
+leaflet(data = (regions)) %>% 
+  addTiles(urlTemplate = "//b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png") %>% 
+  addPolygons(popup = ~url, fillOpacity = 0, weight = 1) 
+  # addLabelOnlyMarkers(data = regions_centroids, label = ~as.character(Name), 
+  #                     labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T)
+  #                     )
 ```

### 2020-06-09 code/index.Rmd
Add link to osm
Update basemap, finalise #70

--- 2020-06-09 code/index.Rmd
@@
 If you are interested in the code underlying the project and would like to support the project with technical input, please provide feedback via the [GitHub issue tracker (requires a GitHub account)](https://github.com/cyipt/popupCycleways/issues).
+To see and edit the data in OpenStreetMap on which the results are based, see [osm.org](https://www.openstreetmap.org/#map=7/52.486/-2.428&layers=C) (requires an OSM login to [edit the data](https://www.cyclestreets.net/edit/tutorial/)).
 
@@
 ```{r}
-pvars = c("Name", "Level")
-tm_shape(regions) + tm_polygons("Level", popup.vars = pvars) 
-  # + tm_shape(regions_centroids) + tm_text("Name", size = 0.7)
+s = c(
+    `Grey basemap` = "CartoDB.Positron",
+    `Coloured basemap` = "Esri.WorldTopoMap",
+    `OSM existing cycle provision (CyclOSM)` = "https://b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png",
+    `PCT commuting, Government Target` = "https://npttile.vs.mythic-beasts.com/commute/v2/govtarget/{z}/{x}/{y}.png",
+    `PCT schools, Government Target` = "https://npttile.vs.mythic-beasts.com/school/v2/govtarget/{z}/{x}/{y}.png",
+    `Satellite image` = "https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}'"
+  )
+tms = c(FALSE, FALSE, FALSE, TRUE, TRUE, FALSE)
+# pvars = c("Name", "Level")
+Regions = regions %>% 
+  mutate(Level = gsub(pattern = "UA", replacement = "Unitary authority", x = Level))
+m_regions = tm_shape(Regions) + tm_polygons("Level", popup.vars = FALSE, alpha = 0.3) +
+  tm_basemap(server = s, tms = tms)
+# + tm_shape(regions_centroids)
+# + tm_text("Name", size = 0.7)
+tmap_leaflet(m_regions) %>% 
+  addPolygons(data = regions, popup = ~url, fillOpacity = 0, weight = 1) 
+# library(leaflet)
+# leaflet(data = (regions)) %>% 
+#   addTiles(urlTemplate = "//b.tile-cyclosm.openstreetmap.fr/cyclosm/{z}/{x}/{y}.png") %>% 
+#   addPolygons(popup = ~url, fillOpacity = 0, weight = 1) 
+  # addLabelOnlyMarkers(data = regions_centroids, label = ~as.character(Name), 
+  #                     labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T)
+  #                     )
 ```

### 2020-06-09 code/index.Rmd
Update output directory for #92

@@
 ```{r}
-dn = "popupCycleways/v1.1" # dirctory name
+dn = "popupCycleways/release" # dirctory name
 built_regions = list.files(dn)

### 2020-06-09 report.Rmd
edits
More report edits

@@
 author: Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds
-# output: bookdown::word_document2
-output:
-  bookdown::pdf_document2:
-    toc: true
-    toc_depth: 1
+output: bookdown::word_document2
+# output:
+#   bookdown::pdf_document2:
+#     toc: true
+#     toc_depth: 1
 bibliography:
@@
 The statutory guidance on network management in response to COVID-19 outlines a range of measures that can support a safe and healthy restart, including the creation of new cycleways, reducing speed limits, widening of footways and closing roads to motor traffic via 'modal filters'.
-**This report describes the Rapid Cycleway Prioritisation Tool project, which provides aims help authorities effectively prioritise locations of new cycleways created by reallocating road space.**
+The focus of this report is a tool that provides evidence in support of the first of these measures.
+**The Rapid Cycleway Prioritisation Tool aims provide evidence on where to prioritise cycleways created by reallocating road space.**
 <!-- Light segregation infrastructure such as flexible plastic wands (illustrated in Figure 1) and bollards can deliver a step change in provision for cycling, when implemented in combination with other measures such reduced speed limits on key corridors (as has happened on the A33 in Southampton and the A56 in Manchester) and when placed where they are most needed.
@@
 
-The [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/) was designed to provide evidence in support of schemes eligible for the second tranche of funding.
-Given the importance of acting quickly, and the unique opportunity that the crisis poses to improve physical activity levels in the long term, it is vital that local authorities have the evidence they need.
-Without a nationally consistent and systematic evidence base, there is a risk that some already stretched local authorities will rely on ad-hoc methods to allocate funding.
-The tool can also support authorities to make bold plans by highlighting the potential of key strategic corridors.
+The [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/), supported by the Department for Transport and Sustrans, was designed to provide evidence in support of schemes eligible for the second tranche of funding.
+The tool provides a national evidence base that, used alongside other sources of evidence, and help ensure that the funding is spent effectively, to maximise cycling uptake in the coming and months and years of the post COVID-19 world.
 
@@
 Roads that met the spare space criteria but which were not selected as top routes were kept and can be shown in the layers named 'spare lane' and 'estimated width >10m'.
+This includes roads with high cycling potential but which were not selected as top routes because they do not keep the same name for a stretch of at least 500 m in length.
 These results are hidden by default to simplify the visualisation of results and to focus attention on the most important results, including the Cohesive network.
@@
 The Cohesive network layer was developed to mitigate the disjointed nature of roads with spare space, but this layer does not include potential new routes outside the current road network.
+
 The methodology only selects roads that have the same name for a continuous stretch of at least 500m, as candidates for top routes, meaning that some roads will be excluded from the 'top routes' result due to name changes (although most continuous road sections have consistent names).

### 2020-06-09 report.Rmd
Complete report

@@
 The scenarios include Government Target, which represents a doubling the number of cycling stages made by 2025 [@departmentfortransport_cycling_2017].
-The PCT maps cycle uptake at area, route and route network levels, for commuting and school travel (as of May 2020) [@goodman_scenarios_2019].
+The PCT maps cycle uptake at area, route and route network levels, for commuting and (since May 2019) school travel [@goodman_scenarios_2019].
 
-A limitation of the PCT is that it takes little account of existing infrastructure: variables such as road width and even the presence/absence of cycle infrastructure currently are not included in the results.
-The Rapid Cycleway Prioritisation Tool overcomes this issue by using road network data which includes geographic attributes and cycle infrastructure from OpenStreetMap.
+The PCT takes little account of existing infrastructure: variables such as road width and even the presence/absence of cycle infrastructure are not included in the results, with the focus being on where there is most latent demand for cycling.
+The Rapid Cycleway Prioritisation Tool overcomes this issue from the perspective of planners who need to prioritise scheme candidates from hundreds of potential options.
+It does this by using road network data which includes geographic attributes and cycle infrastructure from OpenStreetMap [@barrington-leigh_world_2017].
 Carriageway width data were taken from the Cycling Infrastructure Prioritisation Tool (CyIPT) project, funded as a prototype project by a DfT Transport-technology research innovation grants (T-TRIG).
-This uses road data derived from OpenStreetMap.org (OSM) [@barrington-leigh_world_2017].
 
-The Rapid Cycleway Prioritisation Tool identifies roads which can accommodate new cycleways whilst maintaining 2-way motorized traffic.
-Of these roads, it identifies priority locations for new cycleways that are expected to receive most usage under the 'Government Target' scenario.
-The amount of usage under modelled scenarios of cycling uptake, based on evidence of the percentage of trips cycled in relation to the distance and hilliness of the trip, is known as 'cycling potential' [@lovelace_propensity_2017].
+Among the wide range of locations where new cycleways could be implemented, the Rapid Cycleway Prioritisation Tool identifies roads which can accommodate new cycleways whilst maintaining 2-way motorized traffic.
+Of these roads, it identifies priority locations for new cycleways that have the highest 'cycling potential' under the 'Government Target' scenario in the PCT [@lovelace_propensity_2017].
+By cycling potential we mean the number of oneway trips that could be expected if cycling levels double, based on the distances and hilliness of routes between [origin-destination pairs](https://docs.ropensci.org/stplanr/articles/stplanr-od.html).
 
-The tool identifies continuous road sections and ranks them according to their cycling potential.
-Roads can also be ranked by length, on the basis that interventions on long, continuous road sections will tend to have a large impact on the overall transport network, and the estimated amount of cycling in units of 'km per day', which is simply cycling potential multiplied by section length.
-The aim is to provide evidence that maximises the impact of new interventions in creating a joined-up and continuous, a vital component of cycling strategies [@parkin_designing_2018].
+The main output of the tool is ranked lists of continuous road sections to support the prioritisation process.
+Roads can also be ranked by cycling potential, length, speed limit and presence of spare lanes.
+Length was included on the basis that interventions on long, continuous road sections will tend to have a large impact on the overall transport network.
+Road sections can also be ranked by estimated amount of cycling in units of 'km per day', which is simply cycling potential multiplied by section length, as illustrated in Figure \@ref(fig:rank) (see Figure \ref(fig:mapout)).
+
+```{r rank, fig.cap="Illustration of the tabular output of the Rapid Cycleway Prioritisation Tool for West Yorkshire."}
+knitr::include_graphics("figures/Screenshot_2020-06-09 West Yorkshire.png")
+```
+
+The rationale was to provide interactive evidence that supports scheme prioritisation based on a range of variables.
+Length, for example, is important for creating a joined-up and continuous, a vital component of cycling strategies [@parkin_designing_2018].
 
@@
 The results of this research are publicly accessible at www.cyipt.bike/rapid. 
-Maps are available for regions covering the whole of England, at the level of Combined Authorities, Unitary Authorities, or Upper Tier Authorities. 
-Using the methods outlined above, we have derived a set of outputs which are represented as layers in these maps.
-The layers are presented in chronological order to make reading the maps more intuitive, indicating that the data should be used to identify ways to fill key gaps in the existing network in the short term but also plan for a cohesive network in the long term, providing a long-term 'direction of travel' in line with Local Cycling and Walking Plans (LCWIPs), which provide a strategic vision of joined-up networks.
+The main output is a list of candidate roads that meet the 'spare space' criteria outlined above, presented in tabular form in Figure \@ref(fig:rank) above.
+From a planning perspective the most useful output is an interactive map of a type shown in Figure \@ref(fig:mapout).
+Results are available for regions covering the whole of England, at the level of Combined Authorities, Unitary Authorities, or Upper Tier Authorities.
+
+```{r mapout, fig.cap="The map output of the Rapid Cycleway Prioritisation Tool for West Yorkshire, zoomed in on Leeds"}
+knitr::include_graphics("figures/mapout.png")
+```
+
+In addition to a list of highlighted road sections, a number of outputs were generated for each authority.
+These are represented as layers in the interactive map, and ordered chronologically, with the aim of making the results more intuitive.
+For the immediate task of joining-up existing cycleways, for example, the Existing cycleways layer may be of most use.
+The Cohesive network provides a long-term 'direction of travel' that could be used to support a longer term strategic vision, in support of Local Cycling and Walking Plans (LCWIPs).
+These are other layers are described below.
 
@@
 'Gappy' networks are off-putting to many cyclists, and busy or complex road junctions in particular can create barriers to cycle movement [@buehler_bikeway_2016].
+
+## Road space layers
+
+In addition to the layers mentioned above, two other layers are available in the map: Spare lane(s) and Estimated width > 10 m layers.
+These layers, which are hidden by default, do not meet the cycling potential or length criteria needed for inclusion in the top ranked routes.
+They were included in the map to highlight routes where 'connecting' infrastructure joining existing infrastructure or routes to destinations not picked up in the PCT, may be appropriate.
 
@@
 Perhaps the most promising future direction of travel is towards more flexible tools for transport planning.
-By demonstrating how existing datasets can be combined to create new results that are useful, the project has shown that, like infrastructure planning itself, an 'agile' approach to transport planning tools can yield useful and perhaps unexpectedly successfull results that support local knowledge and transport planning expertise.
+By demonstrating how existing datasets can be combined to create new results that are useful, the project has shown that, like infrastructure planning itself, an 'agile' approach to transport planning tools can yield useful and perhaps unexpectedly successful results that support local knowledge and transport planning expertise.
 We hope to take lessons learned during the project forward into future, more strategic, transport planning tool research and development work.

### 2020-06-09 code/index.Rmd
Data downloads

@@
   rmarkdown::render("code/region_index.Rmd", output_file = file.path("~/cyipt/tempCycleways/", d, "index.html"))
+  # browseURL(file.path("~/cyipt/tempCycleways/", d, "index.html"))
   sf::write_sf(cycleways, file.path(d, "cycleways.geojson"))

### 2020-06-09 code/region_index.Rmd
Update description of cycleway data
Data downloads

--- 2020-05-29 code/region_index.Rmd
@@
 
-The map shows a network of key roads with high cycling potential (grey), existing off road cycleways (and shared space paths) according to OpenStreetMap data (green) and top road sections with high cycling potential and spare lanes or estimated width above 10m (blue).
-To see additional layers showing road sections with 'spare lanes' and width greater than 10 m, click on the layers icon in the top left, just below the +/- buttons on the map.
+The map shows three key layers (see [www.cyipt.bike/rapid](https://www.cyipt.bike/rapid/#map-layers) for a detailed description of these layers):
+
+1.
+The **Existing cycleways** layer provides an approximation of where cycling infrastructure exists **currently** (based on OpenStreetMap data downloaded in May 2020) and is intended to help identify gaps in the existing network that could be filled with new interventions (data stored as [cycleways.geojson](cycleways.geojson)).
+2.
+The **Top ranked new cycleways** layer is the central result of the analysis, providing a list of roads that have high cycling potential, a minimum threshold length, and spare space according to our definition.
+These may be strong candidates for reallocation of road space **immediately or in the near future** to create new cycleways on strategic corridors under the Emergency Active Travel Fund (data stored as [top_routes.geojson](top_routes.geojson)).
+3.
+The **Cohesive network** layer is intended to show what a joined-up cycle network could look like if we were to consider new cycleways by either closing roads to motorised traffic or creating one-way systems.
+Composed of roads that have high cycling potential on some or all of their length, the layer is designed to guide **long term planning**, alongside pre-existing plans (data stored as [cohesive_network.geojson](cohesive_network.geojson)).
+
+To see additional layers showing road sections with 'spare lanes' (data stored as [spare_lanes.geojson](spare_lanes.geojson)) and width an estimated width of greater than 10 m (data stored as [wide_lanes.geojson](wide_lanes.geojson)), click on the layers icon in the top left, just below the +/- buttons on the map.
 
@@
 
-See full map [here](m.html).
+See full map [here](m.html), and a more description of the layers and the methodology at [cyipt.bike/rapid](https://www.cyipt.bike/rapid/).
 

### 2020-06-10 code/build.R
working at last

@@ -362,28 +362,44 @@ rg_new3 = r_lanes %>%
 short_roads = rg_new3 %>%
   filter(group2_length <= 2000)
 
-rg_new4 = rbind()
+short_roads$group3 = NA
+short_roads$group3_length = NA
+short_roads$new_group3 = NA
+
+rg_new4 = rbind(short_roads, lg_new2)
 
 # create a new group to capture long continuous sections with the same name
-min_length_named_road = 1000
-rg_new4 = rg_new3 %>% 
-  group_by(ref, group, group2, name) %>% 
-  mutate(long_named_section = case_when(
-    sum(length) > min_length_named_road & name != "" ~ name,
-    TRUE ~ "Other"
-  )
-  ) %>% 
-  ungroup()
-# Only one group (meaning no impact on results) in many regions:
-table(rg_new4$long_named_section)
+# min_length_named_road = 1000
+# rg_new4 = rg_new3 %>% 
+#   group_by(ref, group, group2, name) %>% 
+#   mutate(long_named_section = case_when(
+#     sum(length) > min_length_named_road & name != "" ~ name,
+#     TRUE ~ "Other"
+#   )
+#   ) %>% 
+#   ungroup()
+# # Only one group (meaning no impact on results) in many regions:
+# table(rg_new4$long_named_section)
 
 
 # Name the groups and calculate stats -------------------------------------
-
+# new_names = rg_new4 %>%
+#   group_by(ref, group, group2, new_group3, name) %>%
+#   mutate(length = sum(length)) %>%
+#   filter(length == max(length)) 
+# 
+# rg_new4$new_name = rg_new4
+#   
+# 
+# new_name = dalby %>%
+#   group_by(name) %>%
+#   mutate(name_length = sum(length)) %>%
+#   filter(name_length == max(name_length)) 
+# new_name$name
 
 # find group membership of top named roads
 r_lanes_grouped2 = rg_new4 %>% 
-  group_by(ref, group, group2, long_named_section) %>% 
+  group_by(ref, group, group2, new_group3) %>% 
   summarise(
     name = case_when(
       length(table(name)) == 1 & names(table(name))[1] != "" ~ 

### 2020-06-10 code/chelt.R

@@ -83,8 +83,10 @@ while (min(lg_c$group3_length) < min_grouped_length) {
   new_c = NULL
   
   shortest_centroid = lg_c[which.min(lg_c$group3_length), ]
-  shortest_segment = lg_new2[which.min(lg_new2$group3_length), ]
   if(dim(shortest_centroid)[1] > 1) shortest_centroid = shortest_centroid[1,]
+  
+  shortest_segment = lg_new2[which(lg_new2$new_group3 == shortest_centroid$new_group3), ]
+  
   near = lg_c %>% 
     filter(group2 == shortest_centroid$group2) %>% 
     filter(new_group3 != shortest_centroid$new_group3) # must remove the point itself from this group
@@ -106,28 +108,16 @@ while (min(lg_c$group3_length) < min_grouped_length) {
   lg_new2$group3_length[(lg_new2$new_group3 == near$new_group3[mindist]) | lg_new2$new_group3 == shortest_segment$new_group3] =
     sum(mean(shortest_segment$group3_length), mean(lg_new2$group3_length[lg_new2$new_group3 == near$new_group3[mindist]]))
 
-  lg_new2$geometry[(lg_new2$new_group3 == near$new_group3[mindist]) | lg_new2$new_group3 == shortest_segment$new_group3] =
-    sf::st_union(
-    shortest_segment$geometry, lg_new2$geometry[lg_new2$new_group3 == near$new_group3[mindist]]
-    )
-
-  lg_new2$new_group3[lg_new2$new_group3 == shortest_segment$new_group3] = 
-    lg_new2$new_group3[lg_new2$new_group3 == near$new_group3[mindist]]
+  lg_new2$new_group3[lg_new2$new_group3 == shortest_segment$new_group3] = near$new_group3[mindist]
+  
+  # adapt = lg_new2[lg_new2$new_group3 == near$new_group3[mindist],]
   
   lg_c[(lg_c$new_group3 == near$new_group3[mindist]),] =
     lg_new2 %>% filter(new_group3 == near$new_group3[mindist]) %>%
     group_by(group2, group3_length, new_group3) %>%
     summarise() %>%
     sf::st_centroid()
-  
-  # lg_c[lg_c$new_group3 == shortest_centroid$new_group3, ] =
-  #   lg_new2 %>% filter(new_group3 == near$new_group3[mindist]) %>%
[…]

### 2020-06-10 code/index.Rmd
Close #93

@@
 
-The [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/) helps identify promising locations for road space reallocation in England.
+The [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/) is a research project and web application supported by the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) and [Sustrans](https://www.sustrans.org.uk/) to help identify promising locations for road space reallocation in England.
 The tool provides an evidence base to support investment in cycling infrastructure, in response to new pressures placed on transport and public health systems by the COVID-19 crisis, at local and regional levels.
@@
 The Rapid Cycleway Prioritisation Tool aims to help produce evidence-based answers the question: **where to reallocate road space on strategic corridors to enable uptake of cycling?**
-The focus of the tool is road space reallocation on strategic corridors, a key way in which transport systems can be transformed to overcome capacity constraints on public transport modes and prevent gridlock as the lockdown eases.
+The focus of the tool is road space reallocation on strategic corridors, a key way in which transport systems can be transformed to overcome capacity constraints on public transport modes and prevent gridlock during the phased 'restart'.
 
@@
 
+# Credits
+
+Many thanks to Dr Malcolm Morgan ([University of Leeds](http://www.leeds.ac.uk/)) and Martin Lucas-Smith ([CycleStreets.net](https://www.cyclestreets.net/)) for providing substantial input into methodological and web development aspects of the project respectively.
+
+Thanks to Jeremy Clarke, Rabina Nawaz, John Sweetman and Richard Mace from the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) for vital input into the tool's design.
+Thanks to Kevin McCann from [Sustrans](https://www.sustrans.org.uk/) for guidance that improved the tool from a policy perspective.
+Thanks to Roger Geffen and others from [CyclingUK](https://www.cyclinguk.org/) for input into prototype versions of the tool.
+Thanks to developers of open source software, especially to [contributors](https://www.r-project.org/contributors.html) to the R language in which most of the data analysis was conducted.
+Thanks also to all the contributors to OpenStreetMap who provided open data on which much of the analysis was based, allowing use to publish the results as open data under the conditions of the Open Data Commons Open Database License.
+Data (c) [OpenStreetMap](https://www.openstreetmap.org) contributors, ODbL 1.0.
+https://www.openstreetmap.org/copyright .
+
 # References

### 2020-06-10 code/index.Rmd
Update credits

@@
 ```{r}
+library(leaflet)
 s = c(
@@
 Thanks to developers of open source software, especially to [contributors](https://www.r-project.org/contributors.html) to the R language in which most of the data analysis was conducted.
-Thanks also to all the contributors to OpenStreetMap who provided open data on which much of the analysis was based, allowing use to publish the results as open data under the conditions of the Open Data Commons Open Database License.
-Data (c) [OpenStreetMap](https://www.openstreetmap.org) contributors, ODbL 1.0.
-https://www.openstreetmap.org/copyright .
+Thanks also to all the contributors to OpenStreetMap who provided open data on which much of the analysis was based, allowing publication of the results as open data under the conditions of the Open Data Commons Open Database License ([ODbL 1.0](https://www.openstreetmap.org/copyright)).
+Data (c) [OpenStreetMap](https://www.openstreetmap.org) contributors.
 

### 2020-06-10 code/build.R
hospitals

@@ -32,7 +32,7 @@ if(!exists("s")) {
   rj_all = readRDS("rj.Rds")
   region_names = regions$Name
   # hospitals:
-  hsf = readRDS("hsf.Rds")
+  hsf = readRDS("acute_hospitals.Rds")
   nrow(regions)
   cycleways_en = readRDS("cycleways_en.Rds")
   tm_shape(regions) + tm_polygons(alpha = 0.1) + tm_basemap(s, tms = tms) # check basemaps

### 2020-06-10 report.Rmd
Mention dual carriageway limitations
Report updates

@@
 browseURL("report.pdf")
+browseURL("report.docx")
 ```
@@
 To mitigate this limitation, roads that don't meet this criteria can still appear in the 'spare lane' and 'estimated width > 10m' layers. 
+Dual carriageways lengths are currently counted twice, meaning a dual carriageway section of 260 m may be included in the results.
 
@@
 The coming 'restart' will lead to long lasting changes in transport systems.
-Whether these changes include a worsening of air pollution, traffic and road traffic casualty rates, or uptake of active transport is an open question that depends effective and timely interventions.
-The Rapid Cycleway Prioritisation Tool provides an evidence base in support of effective and timely interventions that will form part of broader strategies to respond to COVID-19.
+The changes could include a worsening of air pollution, congestion and road traffic casualty rates.
+Alternatively, the restart could involve continued uptake of active transport.
+The Rapid Cycleway Prioritisation Tool provides an evidence base in support of effective and timely interventions that can help ensure a healthy, safe restart to the transport system that future proofs it for the demands the 21^st^ Century.
 
@@
 
+# Credits
+
+Many thanks to Dr Malcolm Morgan ([University of Leeds](http://www.leeds.ac.uk/)) and Martin Lucas-Smith ([CycleStreets.net](https://www.cyclestreets.net/)) for providing substantial input into methodological and web development aspects of the project respectively.
+
+Thanks to Jeremy Clarke, Rabina Nawaz, John Sweetman and Richard Mace from the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) for vital input into the tool's design.
+Thanks to Kevin McCann from [Sustrans](https://www.sustrans.org.uk/) for guidance that improved the tool from a policy perspective.
+Thanks to Roger Geffen and others from [CyclingUK](https://www.cyclinguk.org/) for input into prototype versions of the tool.
+Thanks to developers of open source software, especially to [contributors](https://www.r-project.org/contributors.html) to the R language in which most of the data analysis was conducted.
+Thanks also to all the contributors to OpenStreetMap who provided open data on which much of the analysis was based, allowing publication of the results as open data under the conditions of the Open Data Commons Open Database License ([ODbL 1.0](https://www.openstreetmap.org/copyright)).
+Data (c) [OpenStreetMap](https://www.openstreetmap.org) contributors.
+
 # References

### 2020-06-10 report.Rmd
PDF version

@@
 ```{r setup2, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
+knitr::opts_chunk$set(echo = FALSE, fig.align = "center", out.width = "100%")
 ```
@@
 ```{r rank, fig.cap="Illustration of the tabular output of the Rapid Cycleway Prioritisation Tool for West Yorkshire."}
-knitr::include_graphics("figures/Screenshot_2020-06-09 West Yorkshire.png")
+knitr::include_graphics("figures/wyca.png")
 ```

### 2020-06-11 code/index.Rmd
Final tweaks to landing page

@@
   mutate(Level = gsub(pattern = "UA", replacement = "Unitary authority", x = Level))
-m_regions = tm_shape(Regions) + tm_polygons("Level", popup.vars = FALSE, alpha = 0.3) +
+m_regions = tm_shape(Regions) + tm_polygons("Level", popup.vars = FALSE, alpha = 0.5) +
   tm_basemap(server = s, tms = tms)
@@
 
-# Credits
+# Acknowledgments
 

### 2020-06-11 code/index.Rmd
spelling

@@
 The tool provides an evidence base to support investment in cycling infrastructure, in response to new pressures placed on transport and public health systems by the COVID-19 crisis, at local and regional levels.
-The rational for the tool is that, in the context of COVID-19, big decisions around where to prioritise investment on the road network need to be made quickly.
+The rationale for the tool is that, in the context of COVID-19, big decisions around where to prioritise investment on the road network need to be made quickly.
 The tool provides a ‘first pass’ assessment of promising locations for new cycleways which have high latent demand for cycling and where road space can be reallocated whilst maintaining two-way traffic flows for motorists.

### 2020-06-11 code/index.Rmd

@@
 Thanks to Roger Geffen and others from [CyclingUK](https://www.cyclinguk.org/) for input into prototype versions of the tool.
-Thanks to developers of open source software, especially to [contributors](https://www.r-project.org/contributors.html) to the R language in which most of the data analysis was conducted.
+
+Thanks to developers of open source software, especially to developers of the [R language](https://www.r-project.org/contributors.html) and the packages [sf](https://github.com/r-spatial/sf) and [tmap](https://github.com/mtennekes/tmap) which underpinned the analysis.
 Thanks also to all the contributors to OpenStreetMap who provided open data on which much of the analysis was based, allowing publication of the results as open data under the conditions of the Open Data Commons Open Database License ([ODbL 1.0](https://www.openstreetmap.org/copyright)).

### 2020-06-11 report.Rmd
Credits -> Acknowledgements

@@
 10 m was selected because it is roughly the minimum width required for new cycleways created by road space reallocation measures, although there are examples of roads that are less than 10 m on which new cycleways have been successfully created in response to COVID-19.^[
-[South Road](https://www.lancasterguardian.co.uk/news/uk-news/mixed-reactions-new-lancaster-pop-cycle-lanes-busy-city-centre-road-2875909#gsc.tab=0) in Lancaster, for example, has an average width of around 8 m, yet protected cycleways in either direction have been created by installation of wands on either side.
+[South Road](https://www.lancasterguardian.co.uk/news/uk-news/mixed-reactions-new-lancaster-pop-cycle-lanes-busy-city-centre-road-2875909#gsc.tab=0) in Lancaster, for example, has an average width of around 9 m, yet protected cycleways in either direction have been created by installation of wands on either side.
 ]

### 2020-06-11 code/index.Rmd
Typo fix II

@@
 The tool provides an evidence base to support investment in cycling infrastructure, in response to new pressures placed on transport and public health systems by the COVID-19 crisis, at local and regional levels.
-The rationale for the tool is that, in the context of COVID-19, big decisions around where to prioritise investment on the road network need to be made quickly.
+The rational for the tool is that, in the context of COVID-19, big decisions around where to prioritise investment on the road network need to be made quickly.
 The tool provides a ‘first pass’ assessment of promising locations for new cycleways which have high latent demand for cycling and where road space can be reallocated whilst maintaining two-way traffic flows for motorists.

### 2020-06-11 code/index.Rmd
cohesive network text

@@
 
-This layer represents what a joined-up cycle network could look like if new cycleways were placed along corridors of high potential, by either closing roads to motorised traffic or creating one-way systems.
+This layer represents what a joined-up cycle network could look like if new cycleways were placed along the key corridors of high potential. 
+Unlike the 'top ranked new cycleways' layer, the cohesive network comprises all of the major high cycle potential corridors, including sections where the roads are narrower. 
 It shows what a joined-up cycle network could look like if there were no constraints on road space, along some of the most important corridors.
 
-The cohesive network layer supports consideration of **long term** plans and alternatives to road space reallocation.
-Road sections in the cohesive network but not in the Top ranked new cycleway layer may be suitable candidates for one way streets, where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
-The cohesive network layer was created to highlight the need for cohesive cycle network that and do not 'give up' at awkward junctions.
+The cohesive network layer includes sections where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
+It was created to highlight the need for a cohesive cycle network that does not 'give up' at awkward junctions.
 Cohesive networks are a key to getting people cycling [@buehler_bikeway_2016].
 
-The layer was generated by first identifying the roads (classified by reference number) on which there is most cycling potential across the city and joining them together.
+This layer supports consideration of **long term** plans and alternatives to road space reallocation.
+Some sections of the cohesive network may still be suitable candidates for road-space reallocation (for example, [in Lancaster](https://www.facebook.com/photo.php?fbid=10158393513440797&set=gm.1409892779205719&type=1&theater) protected cycle lanes have recently been installed on a road which is less than 10 m in width), but other sections may be best suited to different types of traffic measure, such as point closures, cycleway installation on a nearby parallel alignment, or conversion to one way streets in order to free up space for cycle lanes.
+
+The layer was generated by identifying the roads (classified by reference number) on which there is greatest cycling potential across the region and joining them together.
 Specifically we first selected all road segments with cycle potential in the 85th percentile and above, which also belong to classified trunk roads, A roads or B roads.

### 2020-06-11 code/boundaries2.R
Close #96

@@ -100,6 +100,10 @@ tm_shape(regions) + tm_polygons("Level") + tm_shape(regions_centroids) + tm_text
 regions_dft_centroids = sf::st_point_on_surface(regions_dft)
 tm_shape(regions_dft) + tm_polygons("Level") + tm_shape(regions_dft_centroids) + tm_text("Name", size = 0.7)
 
+
+
+
+
 # identify regions completely within other regions
 region_matches = sf::st_intersects(regions_dft)
 summary({nmatches = lengths(region_matches)})
@@ -109,8 +113,38 @@ regions_dft$Name[regions_within]
 regions_within_names = c("Leicester", "Nottingham", "Derby", "Stoke-on-Trent")
 regions_dft$url = paste0("<a href='test'>test</a>")
 pvars = c("url", "Name")
+
 tm_shape(regions_dft) + tm_polygons("Level", popup.vars = pvars) + tm_shape(regions_dft_centroids) + tm_text("Name", size = 0.7)
+
+
 
 saveRDS(regions_dft, "regions_dft.Rds")
 piggyback::pb_upload("regions_dft.Rds", "cyipt/cyipt-phase-1-data")
 piggyback::pb_upload("lads.Rds", "cyipt/cyipt-phase-1-data")
+
+# Add more attributes -----------------------------------------------------
+
+regions_dft = readRDS("regions_dft.Rds")
+msoa_pct = sf::read_sf("https://github.com/npct/pct-outputs-national/raw/master/commute/msoa/z_all.geojson")
+msoa_centroids = sf::st_centroid(msoa_pct)
+msoas_joined = st_join(msoa_centroids, regions["Name"])
+plot(msoas_joined["Name"])
+summary(msoas_joined$train_tube)
+summary(jsoa)
+msoas_aggregated = msoas_joined %>% 
+  sf::st_drop_geometry() %>% 
+  group_by(Name) %>% 
+  summarise(
+    n_commuters = sum(all),
+    n_pt = sum(train_tube + bus),
+    pt_cycle_shift_govtarget = round(sum(- govtarget_sipt)),
+    pt_cycle_shift_ebike = round(sum(- ebike_sipt))
+  )
+
+regions_census = inner_join(regions_dft, msoas_aggregated)
+plot(regions_census)
+saveRDS(regions_census, "regions_census.Rds")
+regions_top_12 = regions_census %>% 
[…]

### 2020-06-11 code/infographic.R

@@ -1,0 +1,171 @@
+# Aim: create infographic of top 12 zones by PT/cycling potential
+
+regions = readRDS("regions_top_12.Rds")
+regions$html_name = tolower(gsub(pattern = " |, ", replacement = "-", regions$Name))
+regions$url = paste0("<a href='", regions$html_name ,"' target='_blank'>", regions$Name, "</a>")
+i = 1
+top_routes_all = NULL
+
+# for(i in seq(nrow(regions))) {
+# for(i in 1:4) {
+for(i in 1:6) {
+  n = regions$html_name[i]
+  top_routes = sf::read_sf(file.path("popupCycleways/release", n, "top_routes.geojson"))[1]
+  top_routes$Layer = "Top ranked new cycleways"
+  cohesive_network = sf::read_sf(file.path("popupCycleways/release", n, "cohesive_network.geojson"))[1]
+  names(cohesive_network)[1] = "name"
+  cohesive_network$Layer = "Cohesive network"
+  cycleways = sf::read_sf(file.path("popupCycleways/release", n, "cycleways.geojson"))[1]
+  names(cycleways)[1] = "name"
+  cycleways$Layer = "Cycleways"
+  top_routes_region = rbind(top_routes, cohesive_network, cycleways)
+  top_routes_region$Region = regions$Name[i]
+  top_routes_all = rbind(top_routes_all, top_routes_region)
+}
+
+unique(top_routes_all$Region)
+# plot(top_routes_all)
+top_routes_all$lwd = case_when(
+  top_routes_all$Layer == "Cycleways" ~ 1,
+  top_routes_all$Layer == "Top ranked new cycleways" ~ 3,
+  top_routes_all$Layer == "Cohesive network" ~ 1
+)
+top_routes_all$Region = gsub("Greater ", "", top_routes_all$Region)
+top_routes_all$Region = gsub(" City Region", "", top_routes_all$Region)
+top_routes_all$Region = case_when(
+  top_routes_all$Region == "North East" ~ "Newcastle",
+  top_routes_all$Region == "West Yorkshire" ~ "Leeds",
+  top_routes_all$Region == "West Midlands" ~ "Birmingham",
+  TRUE ~ top_routes_all$Region
+)
+summary(as.factor(top_routes_all$Region))
+
+top_routes_all$Layer = factor(
+  top_routes_all$Layer,
[…]

### 2020-06-11 code/write-tests.R

@@ -1,0 +1,143 @@
+library(sf)
+library(pct)
+l = pct::get_pct_routes_fast(region = "london")
+write_sf(l, "l.gpkg")
+saveRDS(l, "l.Rds")
+
+bench::press(
+  n = c(10, 1000, 10000, nrow(l)),
+  {
+    bench::mark(check = FALSE,
+                # readRDS("l.Rds"),
+                # read_sf("l.gpkg"),
+                saveRDS(l[1:n, ], paste0(runif(n = 1), ".Rds")),
+                write_sf(l[1:n, ], paste0(runif(n = 1), ".gpkg"))
+    )
+  }
+)
+bench::mark(check = FALSE,
+            readRDS("l.Rds"),
+            read_sf("l.gpkg"),
+            saveRDS(l, paste0(runif(n = 1), ".Rds")),
+            write_sf(l, paste0(runif(n = 1), ".gpkg"))
+            )
+
+# library(rgdal)
+# install.packages("rgdal") # fail with:
+# /usr/lib/R/etc/Makeconf:168: recipe for target 'inverser.o' failed
+# inverser.c:5:10: fatal error: projects.h: No such file or directory
+# #include <projects.h>
+# ^~~~~~~~~~~~
+#   compilation terminated.
+# /usr/lib/R/etc/Makeconf:168: recipe for target 'inverser.o' failed
+# make: *** [inverser.o] Error 1
+# ERROR: compilation failed for package ‘rgdal’
+# * removing ‘/home/robin/R/x86_64-pc-linux-gnu-library/3.6/rgdal’
+# * restoring previous ‘/home/robin/R/x86_64-pc-linux-gnu-library/3.6/rgdal’
+
+
+# remotes::install_version(package = "rgdal", version = "1.4-8") # works
+# remotes::install_version(package = "rgdal", version = "1.4-7") # works
+# remotes::install_version(package = "rgdal", version = "1.5-8") # fails
+
+# remotes::install_cran("pct")
+# remotes::install_github("r-spatial/sf")
+# library(sf)
+# l = pct::get_pct_routes_fast(region = "london")
+#
+# # test writing both ways
+# f = function(x) file.path(tempdir(), paste0(x, ".gpkg"))
+# f("l1")
+# tm = system.time(
[…]

### 2020-06-11 code/index.Rmd
Add updated landing page text

@@
 
-The [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/) is a research project and web application supported by the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) and [Sustrans](https://www.sustrans.org.uk/) to help identify promising locations for road space reallocation in England.
+[Sustrans](https://www.sustrans.org.uk/) and the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) commissioned ITS Leeds to develop the [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/) to help identifying promising locations for new cycleways in England.
+The immediate purpose of this project is to inform bids for Tranche 2 of the [Emergency Active Travel Fund](https://www.gov.uk/government/publications/emergency-active-travel-fund-local-transport-authority-allocations), particularly for authorities which have not yet developed Local Cycling and Walking Investment Plans ([LCWIP](https://www.gov.uk/government/publications/local-cycling-and-walking-infrastructure-plans-technical-guidance-and-tools)).
+Tranche 2 will fund temporary or permanent infrastructure schemes aimed at increasing active travel and helping to shift people away from public transport given capacity constraints imposed by social distancing.
+
+For each combined and local authority, the tool identifies existing cycleways, promising locations for new cycleways on roads with spare space ('top ranked new cycleways') and what a cohesive cycling network might look like in the longer-term.
+
+# How does the tool identify 'top ranked new cycleways'?
+
+The tool identifies priority locations for new cycleways, ranking roads by their ‘cycling potential’ estimated using the Propensity to Cycle Tool ([PCT](https://www.pct.bike/)).
+This represents the estimated number of cycling trips along this road travelling to or from work or school assuming the Government’s [aim](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy) to double cycling by 2025 is met.
+
+The 'top ranked new cycleways' represents the roads with the highest cycling potential which also have spare space; that is, are either wide or have two or more road lanes in one direction.
+We have restricted the analysis to focusing on roads with spare lanes since these have the capacity to accommodate new cycleways whilst maintaining two-way traffic.
+New cycleways on these routes may therefore be faster to deliver than those requiring roads to be closed to traffic or introducing one-way systems.
+In addition, these are likely to represent the key arterial routes into town and city centres; hence new cycleways on these roads could result in large increases in cycling volumes.
+
+# How does the tool identify what a 'cohesive network' might look like?
+
+The tool also identifies what a ‘cohesive network’ for cycling might look like if we were to consider a wider range of interventions such as closing roads to motorised traffic and creating one-way systems.
+This has been estimated by joining up roads that have a high cycling potential on some or all their length.
+The ‘cohesive network’ might help authorities considering area-wide interventions such as pedestrian and cycle zones or modal filters.
+
+# How can I find the results for my area?
+
+Results for each combined and local authorities can be accessed by clicking on the relevant area from the map below or search the authority’s name from the list below.
+This opens a separate webpage presenting a map of the areas identifying ‘existing cycleways’, ‘top ranked new cycleways’ and the ‘cohesive network’.
+It also presents a table providing various data about the top ranked new cycleways including the cycling potential, length of continuous cycle network (including the existing network), current speed limit and whether the road is wide and/or has spare road lanes.
+
+# How should I use these results?
+
+Authorities may wish to use results from the tool to help prioritise their bids for Tranche 2 of the Emergency Active Travel Fund, complementing local knowledge and stakeholder engagement.
+
+For example, authorities considering new cycleways may want to review the ‘top ranked new cycleways’ to inform their decision where best to invest.
+The table below the maps allows users to rank potential new cycleways by various metrics including:
+
+- cycling potential (default metric to rank schemes)
+- total length of cycling infrastructure created
+- cycling potential x infrastructure length
+
+In addition, the table provides information about road speed limits, width and road lanes which might help users filter out locations which are less desirable for them.
+For example:
+
+- Trunk roads with speed limits of 40mph or over may be less suitable for cycleways given their strategic importance for heavy goods vehicles or potentially due to safety concerns for cyclists (particularly for temporary cycleways).
+- Authorities may wish to prioritise new cycleways on wide roads rather than reallocating road lanes from motorists to cyclists.
+
+Additionally, authorities considering area-wide measures may wish to consider the data for the ‘cohesive network’.
+Currently the tool cannot identify specific interventions to promote cycling (e.g.
+such as pedestrian and cycle zones or modal filters); however, it should give an indication of where such measures might add greatest value.
+
+# What are the key limitations of the tool?
+
+- Currently the tool has been developed primarily to help authorities prioritise the best locations for new cycleways.
+While it provides some indication of what a cohesive network might look like, it doesn’t seek to prioritise what is the best type of intervention for a given location.
+- The ranking of cycleways may not represent a proxy for which schemes are likely to deliver the best value for money.
+As a consequence this tool doesn’t remove the need to use the Department for Transport’s Active Mode Appraisal Toolkit (AMAT) to assess whether proposed schemes deliver value for money.
+- The best locations for new cycleways may not always be on roads with spare lanes.
+For example, in some circumstances it may be preferable for new cycleways to be along quiet roads running parallel to major arterial routes.
+- The results for the ‘top ranked new cycleways’ doesn’t identify routes under 500m with multiple changes of road name and doesn’t identify short routes (under 500m) linking up existing cycleways.
+To address this limitation, the tool includes results labelled ‘spare lane(s)’ and ‘estimated width >10m’ which plug these gaps.
+These layers are hidden by default but can be found by clicking the ‘layer’ button in the top left corner of the map.
+- The road network data is from 2017 and so will be slightly out-of-date.
+- The model estimates the cycling potential of roads based on the estimated number of trips to and from work and school.
+Currently it is not possible to estimate cycling potential accounting for all journey purposes including shopping and leisure trips, which would significantly increase the estimated cycling potential.
+- The PCT uses data from the 2011 Census and 2011 Cycle to School Census to estimate cycling potential.
+Hence the estimated cycling potential may represent an underestimate if there has been significant construction active in the area since 2011.
+- The tool currently only identifies priority on-road routes for new cycleways.
+There may also be desirable off-road locations for new cycleways including alongside old railway lines or across open areas.
+- Currently the tool allows for new cycleways on trunk roads with speeds of 40mph or more.
+While in some limited cases this may be desirable (e.g.
+where trunk roads are causing significant air pollution), in many cases it may not be desirable to reallocate road space from trunk roads to new cycleways.
+To address this, we have identify which top ranked routes have speed limits of 40mph or more in the tables.
+
+# Where can I find further information about the tool?
+
+The technical specification for the Rapid Cycleway Prioritisation Tool can be found here [include hyperlink], providing further detail about the assumptions underlying the tool.
+
+A webinar about the Rapid Cycleway Prioritisation Tool is being held on 17th June 2020 from 10:30-11:45.
+Sign-up details can be found [here](https://register.gotowebinar.com/register/3517000554526190862?source=dft) and recording of the webinar will be available on the website after the event.
+
+Data about the tools informing the model can be found here:
+
+- The Propensity to Cycle Tool
+- The Cycling Infrastructure Prioritisation Tool
+- OpenStreetMap
+
+For expert users, the code underlying the project can be accessed here.
+If you would like to support the project with technical input, please provide feedback via the GitHub issue tracker.
+Data generated by the project can be downloaded from GitHub (e.g. see GeoJSON files for London [here](https://github.com/cyipt/popupCycleways/tree/gh-pages/release/greater-london)).
+
+The results from this project have been delivered rapidly and should be used alongside local knowledge and stakeholder engagement.
+
 The tool provides an evidence base to support investment in cycling infrastructure, in response to new pressures placed on transport and public health systems by the COVID-19 crisis, at local and regional levels.

### 2020-06-11 code/index.Rmd
Add additional links

@@
 - The ranking of cycleways may not represent a proxy for which schemes are likely to deliver the best value for money.
-As a consequence this tool doesn’t remove the need to use the Department for Transport’s Active Mode Appraisal Toolkit (AMAT) to assess whether proposed schemes deliver value for money.
+As a consequence this tool doesn’t remove the need to use the Department for Transport’s Active Mode Appraisal Toolkit ([AMAT](https://www.gov.uk/government/publications/tag-social-and-distributional-impacts-worksheets)) to assess whether proposed schemes deliver value for money.
 - The best locations for new cycleways may not always be on roads with spare lanes.
@@
 
-The technical specification for the Rapid Cycleway Prioritisation Tool can be found here [include hyperlink], providing further detail about the assumptions underlying the tool.
+The technical specification for the Rapid Cycleway Prioritisation Tool can be found [here](https://github.com/cyipt/popupCycleways/releases/download/0.3/report.pdf), providing further detail about the assumptions underlying the tool.
 
@@
 
-- The Propensity to Cycle Tool
-- The Cycling Infrastructure Prioritisation Tool
-- OpenStreetMap
+- The [Propensity to Cycle Tool](https://www.pct.bike/)
+- The [Cycling Infrastructure Prioritisation Tool](https://www.cyipt.bike)
+- [OpenStreetMap](https://www.openstreetmap.org/#map=7/52.486/-2.428&layers=C)
 
-For expert users, the code underlying the project can be accessed here.
-If you would like to support the project with technical input, please provide feedback via the GitHub issue tracker.
+For expert users, the code underlying the project can be accessed [here](https://github.com/cyipt/popupCycleways).
+If you would like to support the project with technical input, please provide feedback via the [GitHub issue tracker](https://github.com/cyipt/popupCycleways/issues/14).
 Data generated by the project can be downloaded from GitHub (e.g. see GeoJSON files for London [here](https://github.com/cyipt/popupCycleways/tree/gh-pages/release/greater-london)).

### 2020-06-11 code/index.Rmd
Fix formatting

@@
 
-The tool identifies priority locations for new cycleways, ranking roads by their ‘cycling potential’ estimated using the Propensity to Cycle Tool ([PCT](https://www.pct.bike/)).
-This represents the estimated number of cycling trips along this road travelling to or from work or school assuming the Government’s [aim](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy) to double cycling by 2025 is met.
+The tool identifies priority locations for new cycleways, ranking roads by their 'cycling potential' estimated using the Propensity to Cycle Tool ([PCT](https://www.pct.bike/)).
+This represents the estimated number of cycling trips along this road travelling to or from work or school assuming the Government's [aim](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy) to double cycling by 2025 is met.
 
@@
 
-The tool also identifies what a ‘cohesive network’ for cycling might look like if we were to consider a wider range of interventions such as closing roads to motorised traffic and creating one-way systems.
+The tool also identifies what a 'cohesive network' for cycling might look like if we were to consider a wider range of interventions such as closing roads to motorised traffic and creating one-way systems.
 This has been estimated by joining up roads that have a high cycling potential on some or all their length.
-The ‘cohesive network’ might help authorities considering area-wide interventions such as pedestrian and cycle zones or modal filters.
+The 'cohesive network' might help authorities considering area-wide interventions such as pedestrian and cycle zones or modal filters.
 
@@
 
-Results for each combined and local authorities can be accessed by clicking on the relevant area from the map below or search the authority’s name from the list below.
-This opens a separate webpage presenting a map of the areas identifying ‘existing cycleways’, ‘top ranked new cycleways’ and the ‘cohesive network’.
+Results for each combined and local authorities can be accessed by clicking on the relevant area from the map below or search the authority's name from the list below.
+This opens a separate webpage presenting a map of the areas identifying 'existing cycleways', 'top ranked new cycleways' and the 'cohesive network'.
 It also presents a table providing various data about the top ranked new cycleways including the cycling potential, length of continuous cycle network (including the existing network), current speed limit and whether the road is wide and/or has spare road lanes.
@@
 
-For example, authorities considering new cycleways may want to review the ‘top ranked new cycleways’ to inform their decision where best to invest.
+For example, authorities considering new cycleways may want to review the 'top ranked new cycleways' to inform their decision where best to invest.
 The table below the maps allows users to rank potential new cycleways by various metrics including:
@@
 
-Additionally, authorities considering area-wide measures may wish to consider the data for the ‘cohesive network’.
+Additionally, authorities considering area-wide measures may wish to consider the data for the 'cohesive network'.
 Currently the tool cannot identify specific interventions to promote cycling (e.g.
@@
 - Currently the tool has been developed primarily to help authorities prioritise the best locations for new cycleways.
-While it provides some indication of what a cohesive network might look like, it doesn’t seek to prioritise what is the best type of intervention for a given location.
+While it provides some indication of what a cohesive network might look like, it doesn't seek to prioritise what is the best type of intervention for a given location.
 - The ranking of cycleways may not represent a proxy for which schemes are likely to deliver the best value for money.
-As a consequence this tool doesn’t remove the need to use the Department for Transport’s Active Mode Appraisal Toolkit ([AMAT](https://www.gov.uk/government/publications/tag-social-and-distributional-impacts-worksheets)) to assess whether proposed schemes deliver value for money.
+As a consequence this tool doesn't remove the need to use the Department for Transport's Active Mode Appraisal Toolkit ([AMAT](https://www.gov.uk/government/publications/tag-social-and-distributional-impacts-worksheets)) to assess whether proposed schemes deliver value for money.
 - The best locations for new cycleways may not always be on roads with spare lanes.
 For example, in some circumstances it may be preferable for new cycleways to be along quiet roads running parallel to major arterial routes.
-- The results for the ‘top ranked new cycleways’ doesn’t identify routes under 500m with multiple changes of road name and doesn’t identify short routes (under 500m) linking up existing cycleways.
-To address this limitation, the tool includes results labelled ‘spare lane(s)’ and ‘estimated width >10m’ which plug these gaps.
-These layers are hidden by default but can be found by clicking the ‘layer’ button in the top left corner of the map.
+- The results for the 'top ranked new cycleways' doesn't identify routes under 500m with multiple changes of road name and doesn't identify short routes (under 500m) linking up existing cycleways.
+To address this limitation, the tool includes results labelled 'spare lane(s)' and 'estimated width >10m' which plug these gaps.
+These layers are hidden by default but can be found by clicking the 'layer' button in the top left corner of the map.
 - The road network data is from 2017 and so will be slightly out-of-date.
@@
 The rational for the tool is that, in the context of COVID-19, big decisions around where to prioritise investment on the road network need to be made quickly.
-The tool provides a ‘first pass’ assessment of promising locations for new cycleways which have high latent demand for cycling and where road space can be reallocated whilst maintaining two-way traffic flows for motorists.
+The tool provides a 'first pass' assessment of promising locations for new cycleways which have high latent demand for cycling and where road space can be reallocated whilst maintaining two-way traffic flows for motorists.
 
@@
 The cycling potential of each road is based on the Government Target (equality) scenario from the Propensity to Cycle Tool. 
-The figures given represent the numbers of one way cycle trips to work and school that could be expected on each road, in a scenario where the [ Government’s aim of an overall doubling of cycling trips has been met](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy).
+The figures given represent the numbers of one way cycle trips to work and school that could be expected on each road, in a scenario where the [ Government's aim of an overall doubling of cycling trips has been met](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy).
 Our figures include data on travel to work from 2011 Census data and travel to school, assuming cycling grows evenly nationwide accounting for distance and hilliness [@lovelace_propensity_2017; @goodman_scenarios_2019]
@@
 -->
-The definition of ‘cycleway’ for the purpose of this layer includes shared use paths but excludes cycle lanes painted on roads.
+The definition of 'cycleway' for the purpose of this layer includes shared use paths but excludes cycle lanes painted on roads.
 The quality of the cycle paths is not reported in the data and the presence of a cycleway cannot guarantee that there is adequate provision, because some cycleways included in this dataset are of poor quality and/or are on paths shared by pedestrians that have limited capacity for high cycle traffic while observing physical distancing guidelines.

### 2020-06-11 code/index.Rmd
Update landing page

@@
 
-# Introduction
+# Introduction {-}
 
@@
 
-The results from this project have been delivered rapidly and should be used alongside local knowledge and stakeholder engagement.
+<!-- The focus of the tool is road space reallocation on strategic corridors, a key way in which transport systems can be transformed to overcome capacity constraints on public transport modes and prevent gridlock during the phased 'restart'.
+-->
 
-The tool provides an evidence base to support investment in cycling infrastructure, in response to new pressures placed on transport and public health systems by the COVID-19 crisis, at local and regional levels.
-The rational for the tool is that, in the context of COVID-19, big decisions around where to prioritise investment on the road network need to be made quickly.
-The tool provides a 'first pass' assessment of promising locations for new cycleways which have high latent demand for cycling and where road space can be reallocated whilst maintaining two-way traffic flows for motorists.
+<!-- Reallocating road space on strategic corridors is just one of the measures available for transport authorities. -->
+<!-- Evidence provided by the tool can also support a range of temporary, experimental and permanent treatments that will enable rapid and safe uptake of cycling. -->
+<!-- New [statutory guidance on network management in response to COVID-19](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities/traffic-management-act-2004-network-management-in-response-to-covid-19) outlines a range of measures, including the creation of new cycleways, reducing speed limits, widening of footways and closing roads to motor traffic via 'modal filters'.
+-->
+<!-- Road space reallocation is suited to larger roads and can have a large impact quickly, especially where moving large numbers of people away from public transport and capacity is a priority. -->
+<!-- Furthermore, low but gradually increasing levels of peak hour motor traffic mean that there is a window of opportunity to reallocate lanes without causing congestion in the short term.  -->
 
-The Rapid Cycleway Prioritisation Tool aims to help produce evidence-based answers the question: **where to reallocate road space on strategic corridors to enable uptake of cycling?**
-The focus of the tool is road space reallocation on strategic corridors, a key way in which transport systems can be transformed to overcome capacity constraints on public transport modes and prevent gridlock during the phased 'restart'.
+<!-- The tool identifies roads that have 'spare space', defined as roads with a spare lane or a width greater than 10 metres and high cycling potential, based on scenarios of mode shift to cycling for work and school in the [Propensity to Cycle Tool](https://www.pct.bike/).
+-->
+<!-- The 10 m width is an indicator of where it may be possible to introduce new light segregation, depending on levels of on street parking and other factors, without altering the navigable network for motor vehicles.
+-->
+<!-- On roads that are around 10 m wide or less, not highlighted by the tool, other interventions may be appropriate. -->
+<!-- Options include 'modal filters' which prevent through traffic for motor vehicles, making roads one way for motor traffic (as implemented successfully with an Experimental Traffic Regulation Order in [Torrington Place](https://www.camden.gov.uk/torrington-tavistock-consultation), London, for example) or removing [central markings](https://doi.org/10.1016/j.aap.2014.08.015) [@shackel_influence_2014] may be required.^[ -->
+<!-- The road width required for new cycleways depends on a range of factors including the presence of on street parking.  -->
+<!-- When on street parking is not an issue new cycleways supported by light segregation measures may be possible on narrower roads, as illustrated by the new trial cycleway on [South Road in Lancaster](https://www.facebook.com/photo/?fbid=10158393513440797&set=gm.1409892779205719&__cft__[0]=AZWl60wQjK9GqKMN_1t25YvwumbO3srWVvigT0H8NpP6n_8oaqPUKCLOW-u5ubg7Lg4agxSS7qNwVzEDCsBZYegxRVBGgHky-AkNLR-DJYyytsJpvi16VVEG8aofuaq9-S1ajmZA24d922GG8aapZ1LWv1dJ6N5ROym1TVK9rdB7mQ&__tn__=EH-R), which involved re-allocating carraigeway space on both sides of a ~9 m wide two way road using plastic wands.
+-->
+<!-- A photo of this intervention can be seen [here](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/lancashire-lane.jpeg). -->
+<!-- ] -->
+<!-- Which roads to prioritise for road space reallocation and the most effective designs will also depend on factors such as on street parking levels, safety considerations and symbolic and aesthetic considerations not present in the analysis.
+-->
 
-Reallocating road space on strategic corridors is just one of the measures available for transport authorities.
-Evidence provided by the tool can also support a range of temporary, experimental and permanent treatments that will enable rapid and safe uptake of cycling.
-New [statutory guidance on network management in response to COVID-19](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities/traffic-management-act-2004-network-management-in-response-to-covid-19) outlines a range of measures, including the creation of new cycleways, reducing speed limits, widening of footways and closing roads to motor traffic via 'modal filters'.
-Road space reallocation is suited to larger roads and can have a large impact quickly, especially where moving large numbers of people away from public transport and capacity is a priority.
-Furthermore, low but gradually increasing levels of peak hour motor traffic mean that there is a window of opportunity to reallocate lanes without causing congestion in the short term.
+<!-- The cycling potential of each road is based on the Government Target (equality) scenario from the Propensity to Cycle Tool.  -->
+<!-- The figures given represent the numbers of one way cycle trips to work and school that could be expected on each road, in a scenario where the [ Government's aim of an overall doubling of cycling trips has been met](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy).
+-->
+<!-- Our figures include data on travel to work from 2011 Census data and travel to school, assuming cycling grows evenly nationwide accounting for distance and hilliness [@lovelace_propensity_2017; @goodman_scenarios_2019] -->
+<!-- Data on road widths and number of lanes is derived from the Cycling Infrastructure Prioritisation Toolkit (see [cyipt.bike](https://www.cyipt.bike/)). -->
 
-<!-- # Input data -->
+<!-- # How to use the tool -->
 
-The tool identifies roads that have 'spare space', defined as roads with a spare lane or a width greater than 10 metres and high cycling potential, based on scenarios of mode shift to cycling for work and school in the [Propensity to Cycle Tool](https://www.pct.bike/).
-The 10 m width is an indicator of where it may be possible to introduce new light segregation, depending on levels of on street parking and other factors, without altering the navigable network for motor vehicles.
-On roads that are around 10 m wide or less, not highlighted by the tool, other interventions may be appropriate.
-Options include 'modal filters' which prevent through traffic for motor vehicles, making roads one way for motor traffic (as implemented successfully with an Experimental Traffic Regulation Order in [Torrington Place](https://www.camden.gov.uk/torrington-tavistock-consultation), London, for example) or removing [central markings](https://doi.org/10.1016/j.aap.2014.08.015) [@shackel_influence_2014] may be required.^[
-The road width required for new cycleways depends on a range of factors including the presence of on street parking. 
-When on street parking is not an issue new cycleways supported by light segregation measures may be possible on narrower roads, as illustrated by the new trial cycleway on [South Road in Lancaster](https://www.facebook.com/photo/?fbid=10158393513440797&set=gm.1409892779205719&__cft__[0]=AZWl60wQjK9GqKMN_1t25YvwumbO3srWVvigT0H8NpP6n_8oaqPUKCLOW-u5ubg7Lg4agxSS7qNwVzEDCsBZYegxRVBGgHky-AkNLR-DJYyytsJpvi16VVEG8aofuaq9-S1ajmZA24d922GG8aapZ1LWv1dJ6N5ROym1TVK9rdB7mQ&__tn__=EH-R), which involved re-allocating carraigeway space on both sides of a ~9 m wide two way road using plastic wands.
-A photo of this intervention can be seen [here](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/lancashire-lane.jpeg).
-]
-Which roads to prioritise for road space reallocation and the most effective designs will also depend on factors such as on street parking levels, safety considerations and symbolic and aesthetic considerations not present in the analysis.
+<!-- To use the evidence generated for this project: -->
 
-The cycling potential of each road is based on the Government Target (equality) scenario from the Propensity to Cycle Tool. 
-The figures given represent the numbers of one way cycle trips to work and school that could be expected on each road, in a scenario where the [ Government's aim of an overall doubling of cycling trips has been met](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy).
-Our figures include data on travel to work from 2011 Census data and travel to school, assuming cycling grows evenly nationwide accounting for distance and hilliness [@lovelace_propensity_2017; @goodman_scenarios_2019]
-Data on road widths and number of lanes is derived from the Cycling Infrastructure Prioritisation Toolkit (see [cyipt.bike](https://www.cyipt.bike/)).
+<!-- - Type in the name of the authority of interest into the search bar above the table below and click on the associated URL. -->
+<!-- - Explore the results in the map (you will find a link to a full screen map) by panning, zooming and clicking on the layers.
+See additional layers by clicking on the layers icon in the top left of the map.
+-->
+<!-- - Rank the 'top roads' in various ways (including by length and cycling potential) in the table below, to explore options highlighted by the tool to help prioritise road space reallocation schemes.
+-->
+<!-- - Interpret the results with care based on other sources of information and an understanding of the limitations of the analysis (see 'Known issues and limitations' below). -->
+<!-- - Refer to the description of the map layers below when exploring the results. -->
 
-# How to use the tool
+<!-- # Basemaps -->
 
-To use the evidence generated for this project:
+<!-- There are also five basemaps to choose from. The default basemap is grey and provides an uncluttered view.  -->
+<!-- The second basemap option is a thematic map which shows all existing cycleways, including on-road cycle lanes.  -->
+<!-- We then have two basemap options showing cycling potential for travel to work and travel to school from the Propensity to Cycle Tool.  -->
+<!-- This allows our results to be compared with cycling potentials across the wider road network.  -->
+<!-- These both follow the Government Target scenario; this is the same scenario as is used to derive the map layers described above.  -->
+<!-- Finally, a satellite image is also available as a basemap. -->
 
-- Type in the name of the authority of interest into the search bar above the table below and click on the associated URL.
-- Explore the results in the map (you will find a link to a full screen map) by panning, zooming and clicking on the layers.
-See additional layers by clicking on the layers icon in the top left of the map.
-- Rank the 'top roads' in various ways (including by length and cycling potential) in the table below, to explore options highlighted by the tool to help prioritise road space reallocation schemes.
-- Interpret the results with care based on other sources of information and an understanding of the limitations of the analysis (see 'Known issues and limitations' below).
-- Refer to the description of the map layers below when exploring the results.
+<!-- # Known issues and limitations -->
 
-# Map layers
+<!-- - The 'Top ranked new cycleways' highlighted in the map-based results were produced automatically based on a simple score combined cycling potential for school and commute trips.
+**The results should be interpreted based on local knowledge and an understanding of the limitations of the data.
+The results are designed to complement not replace other factors such as feedback from the local community and long-term strategic cycle network plans.** -->
+<!-- - The estimates of cycling potential only include school and work travel.
+Accounting for increases in cycling for other trip purposes such as for shopping and leisure would increase cycling potential substantially.
+-->
+<!-- - The road network data was taken from the Cycling Infrastructure Prioritisation Toolkit (CyIPT) project which was completed in 2017, so may be out of date. -->
+<!-- - The results are based on the existing road network. They do not flag potential routes that could be constructed, e.g. along old railway tracks or across open areas. -->
+<!-- - The results can highlight major roads that have a speed limit above 40mph that may currently be unsuitable for new cycleways created with light segregation.
+Roads not controlled by local authorities on the SRN can also be highlighted (although motorways are excluded); we do not alter the results based on who is in control of the roads.
+-->
+<!-- - Other than the Cohesive network layer, which highlights what a joined-up cycle network could look like if roads could be made one way to motor traffic and opened-up to cycling via modal filters, the results do not provide evidence on the full range of interventions.
+Identifying locations for filtered permeability interventions in residential areas, for example, requires data on residential networks not contained in this analysis.
+-->
 
-The maps contain five layers (two of which are hidden by default) that provide the results of our analysis, overlaying the basemaps. 
-Clicking on the lines on the map for any of these layers will reveal a pop-up box containing more information about the road or segment of interest.
-
-The three main layers are arranged in order of planning timescales:
-
-1.
-The **Existing cycleways** layer provides an approximation of where cycling infrastructure exists **currently** (based on OpenStreetMap data downloaded in May 2020) and is intended to help identify gaps in the network that could be filled with new interventions.
-2.
-The **Top ranked new cycleways** layer is the central result of the analysis, providing a list of roads that have high cycling potential, a minimum threshold length, and spare space according to our definition.
-These may be strong candidates for reallocation of road space **immediately or in the near future** to create new cycleways on strategic corridors under the Emergency Active Travel Fund.
-3.
-The **Cohesive network** layer is intended to show what a joined-up cycle network could look like if we were to consider using a wider range of interventions to design new cycleways, such as closing roads to motorised traffic or creating one-way systems.
-Composed of roads that have high cycling potential on some or all of their length, the layer is designed to guide **long term planning**, alongside pre-existing plans such LCWIPs [@departmentfortransport_cycling_2017], towards a continuous and coherent network.
-
-These main layers, and two other layers that informed the results but which are not shown by default, are described in detail below.
-
-## 1. Existing cycleways
-
-On the basis that plans should build on and extend existing provision, this layer shows existing cycleways.
-This may be helpful to view in connection with the other layers, because it can reveal where gaps in the existing cycle network could be filled by new infrastructure.
-<!-- This layer does not include some on-road cycle lanes or cycleways less than 500m in length, but these can be viewed in one of the basemap options (see below for more information about the basemaps).
+<!-- If you are interested in the code underlying the project and would like to support the project with technical input, please provide feedback via the [GitHub issue tracker (requires a GitHub account)](https://github.com/cyipt/popupCycleways/issues).
 -->
-The definition of 'cycleway' for the purpose of this layer includes shared use paths but excludes cycle lanes painted on roads.
-The quality of the cycle paths is not reported in the data and the presence of a cycleway cannot guarantee that there is adequate provision, because some cycleways included in this dataset are of poor quality and/or are on paths shared by pedestrians that have limited capacity for high cycle traffic while observing physical distancing guidelines.
-Also, because of the variable time lag between new cycleways being built and their appearance on OSM, not all existing cycleways are included in this layer (although most established cycleways in most areas are).
-The data was accessed from OpenStreetMap in May 2020 so is up-to-date in many areas.
-
-## 2. Top ranked new cycleways
-
-These are the 'top n' road sections in terms of cycling potential with 'spare space' as defined above.
-They may be suitable candidates for road space reallocation to create new cycleways. 
-These roads have either a spare lane or an average width greater than 10 m, and are ranked according to their mean cycling potential. 
-<!-- For routes that are shorter than 1km in length, we use 'km_cycled' instead of mean cycling potential - this is simply the mean cycling potential multiplied by the length of the route.  -->
-Top ranked new cycleways are listed in the table below the map, so you can see more information on them. 
-The number of Top ranked new cycleways varies depending on the size of the region.
-
-We used the following criteria for the Top ranked new cycleways:
-
-- A road which already has an existing cycleway alongside the majority of its length cannot be a top route.
-- Unnamed roads cannot be a top route.
-- Routes less than 500m in length cannot be a top route.
-
-Some road sections highlighted by the tool may not currently be suitable for new cycleways.
-Road sections that are on strategic and major road networks (SRN and MRN respectively) and which had a high speed limit (40 mph+) before the COVID-19 crisis can be in the top routes.
-To highlight such roads, the speed limits of road sections are listed in the table of the top ranked new cycleway locations to inform prioritisation and speed reduction interventions.
-
-## 3. Cohesive network
-
-This layer represents what a joined-up cycle network could look like if new cycleways were placed along the key corridors of high potential. 
-Unlike the 'top ranked new cycleways' layer, the cohesive network comprises all of the major high cycle potential corridors, including sections where the roads are narrower. 
-It shows what a joined-up cycle network could look like if there were no constraints on road space, along some of the most important corridors.
-
-The cohesive network layer includes sections where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
-It was created to highlight the need for a cohesive cycle network that does not 'give up' at awkward junctions.
-Cohesive networks are a key to getting people cycling [@buehler_bikeway_2016].
-
-This layer supports consideration of **long term** plans and alternatives to road space reallocation.
-Some sections of the cohesive network may still be suitable candidates for road-space reallocation (for example, [in Lancaster](https://www.facebook.com/photo.php?fbid=10158393513440797&set=gm.1409892779205719&type=1&theater) protected cycle lanes have recently been installed on a road which is less than 10 m in width), but other sections may be best suited to different types of traffic measure, such as point closures, cycleway installation on a nearby parallel alignment, or conversion to one way streets in order to free up space for cycle lanes.
-
-The layer was generated by identifying the roads (classified by reference number) on which there is greatest cycling potential across the region and joining them together.
-Specifically we first selected all road segments with cycle potential in the 85th percentile and above, which also belong to classified trunk roads, A roads or B roads.
-These segments are then joined together to form a continuous network.
-
-## 4. Spare lane(s) (not shown by default)
-
-These are routes that have more than one vehicle lane in either direction and have a cycling potential above the minimum threshold (which varies according to region as it needs to be higher in e.g.
-Greater London than in rural areas).
-This layer can be used to identify where there may be spare lanes on the network that could be reallocated to cycling (or pedestrianisation) beyond those highlighted in the Top ranked new cycleways layer.
-
-## 5. Width >= 10m (not shown by default)
-
-These routes are at least 10m in width and have a cycling potential above the minimum threshold. 
-Like the Spare lane(s) layer, this layer is intended to highlight road sections on which increasing space for cycling and walking by reallocating road space may be a viable option.
-
-# Basemaps
-
-There are also five basemaps to choose from. The default basemap is grey and provides an uncluttered view. 
-The second basemap option is a thematic map which shows all existing cycleways, including on-road cycle lanes. 
-We then have two basemap options showing cycling potential for travel to work and travel to school from the Propensity to Cycle Tool. 
-This allows our results to be compared with cycling potentials across the wider road network. 
-These both follow the Government Target scenario; this is the same scenario as is used to derive the map layers described above. 
-Finally, a satellite image is also available as a basemap.
-
-# Known issues and limitations
-
-<!-- We have developed this evidence base rapidly and -->
-
-- The 'Top ranked new cycleways' highlighted in the map-based results were produced automatically based on a simple score combined cycling potential for school and commute trips.
-**The results should be interpreted based on local knowledge and an understanding of the limitations of the data.
-The results are designed to complement not replace other factors such as feedback from the local community and long-term strategic cycle network plans.**
-- The estimates of cycling potential only include school and work travel.
-Accounting for increases in cycling for other trip purposes such as for shopping and leisure would increase cycling potential substantially.
-- The road network data was taken from the Cycling Infrastructure Prioritisation Toolkit (CyIPT) project which was completed in 2017, so may be out of date.
-- The results are based on the existing road network. They do not flag potential routes that could be constructed, e.g. along old railway tracks or across open areas.
-- The results can highlight major roads that have a speed limit above 40mph that may currently be unsuitable for new cycleways created with light segregation.
-Roads not controlled by local authorities on the SRN can also be highlighted (although motorways are excluded); we do not alter the results based on who is in control of the roads.
-- Other than the Cohesive network layer, which highlights what a joined-up cycle network could look like if roads could be made one way to motor traffic and opened-up to cycling via modal filters, the results do not provide evidence on the full range of interventions.
-Identifying locations for filtered permeability interventions in residential areas, for example, requires data on residential networks not contained in this analysis.
-
-<!-- Like many projects that have been initiated in response to COVID-19 we have developed the analy -->
-
-If you are interested in the code underlying the project and would like to support the project with technical input, please provide feedback via the [GitHub issue tracker (requires a GitHub account)](https://github.com/cyipt/popupCycleways/issues).
-To see and edit the data in OpenStreetMap on which the results are based, see [osm.org](https://www.openstreetmap.org/#map=7/52.486/-2.428&layers=C) (requires an OSM login to [edit the data](https://www.cyclestreets.net/edit/tutorial/)).
+<!-- To see and edit the data in OpenStreetMap on which the results are based, see [osm.org](https://www.openstreetmap.org/#map=7/52.486/-2.428&layers=C) (requires an OSM login to [edit the data](https://www.cyclestreets.net/edit/tutorial/)).
+-->
 
@@
 
-# Acknowledgments
+<!-- # Map layers -->
 
-Many thanks to Dr Malcolm Morgan ([University of Leeds](http://www.leeds.ac.uk/)) and Martin Lucas-Smith ([CycleStreets.net](https://www.cyclestreets.net/)) for providing substantial input into methodological and web development aspects of the project respectively.
+<!-- The maps contain five layers (two of which are hidden by default) that provide the results of our analysis, overlaying the basemaps.  -->
+<!-- Clicking on the lines on the map for any of these layers will reveal a pop-up box containing more information about the road or segment of interest. -->
 
-Thanks to Jeremy Clarke, Rabina Nawaz, John Sweetman and Richard Mace from the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) for vital input into the tool's design.
-Thanks to Kevin McCann from [Sustrans](https://www.sustrans.org.uk/) for guidance that improved the tool from a policy perspective.
-Thanks to Roger Geffen and others from [CyclingUK](https://www.cyclinguk.org/) for input into prototype versions of the tool.
+<!-- The three main layers are arranged in order of planning timescales: -->
 
-Thanks to developers of open source software, especially to developers of the [R language](https://www.r-project.org/contributors.html) and the packages [sf](https://github.com/r-spatial/sf) and [tmap](https://github.com/mtennekes/tmap) which underpinned the analysis.
-Thanks also to all the contributors to OpenStreetMap who provided open data on which much of the analysis was based, allowing publication of the results as open data under the conditions of the Open Data Commons Open Database License ([ODbL 1.0](https://www.openstreetmap.org/copyright)).
+<!-- 1.
+The **Existing cycleways** layer provides an approximation of where cycling infrastructure exists **currently** (based on OpenStreetMap data downloaded in May 2020) and is intended to help identify gaps in the network that could be filled with new interventions.
+-->
+<!-- 2.
+The **Top ranked new cycleways** layer is the central result of the analysis, providing a list of roads that have high cycling potential, a minimum threshold length, and spare space according to our definition.
+These may be strong candidates for reallocation of road space **immediately or in the near future** to create new cycleways on strategic corridors under the Emergency Active Travel Fund.
+-->
+<!-- 3.
+The **Cohesive network** layer is intended to show what a joined-up cycle network could look like if we were to consider using a wider range of interventions to design new cycleways, such as closing roads to motorised traffic or creating one-way systems.
+Composed of roads that have high cycling potential on some or all of their length, the layer is designed to guide **long term planning**, alongside pre-existing plans such LCWIPs [@departmentfortransport_cycling_2017], towards a continuous and coherent network.
+-->
+
+<!-- These main layers, and two other layers that informed the results but which are not shown by default, are described in detail below. -->
+
+<!-- ## 1. Existing cycleways -->
+
+<!-- On the basis that plans should build on and extend existing provision, this layer shows existing cycleways. -->
+<!-- This may be helpful to view in connection with the other layers, because it can reveal where gaps in the existing cycle network could be filled by new infrastructure. -->
+<!-- The definition of 'cycleway' for the purpose of this layer includes shared use paths but excludes cycle lanes painted on roads. -->
+<!-- The quality of the cycle paths is not reported in the data and the presence of a cycleway cannot guarantee that there is adequate provision, because some cycleways included in this dataset are of poor quality and/or are on paths shared by pedestrians that have limited capacity for high cycle traffic while observing physical distancing guidelines.
+-->
+<!-- Also, because of the variable time lag between new cycleways being built and their appearance on OSM, not all existing cycleways are included in this layer (although most established cycleways in most areas are).
+-->
+<!-- The data was accessed from OpenStreetMap in May 2020 so is up-to-date in many areas. -->
+
+<!-- ## 2. Top ranked new cycleways -->
+
+<!-- These are the 'top n' road sections in terms of cycling potential with 'spare space' as defined above. -->
+<!-- They may be suitable candidates for road space reallocation to create new cycleways.  -->
+<!-- These roads have either a spare lane or an average width greater than 10 m, and are ranked according to their mean cycling potential.  -->
+<!-- Top ranked new cycleways are listed in the table below the map, so you can see more information on them.  -->
+<!-- The number of Top ranked new cycleways varies depending on the size of the region. -->
+
+<!-- We used the following criteria for the Top ranked new cycleways:  -->
+
+<!-- - A road which already has an existing cycleway alongside the majority of its length cannot be a top route. -->
+<!-- - Unnamed roads cannot be a top route. -->
+<!-- - Routes less than 500m in length cannot be a top route. -->
+
+<!-- Some road sections highlighted by the tool may not currently be suitable for new cycleways. -->
+<!-- Road sections that are on strategic and major road networks (SRN and MRN respectively) and which had a high speed limit (40 mph+) before the COVID-19 crisis can be in the top routes. -->
+<!-- To highlight such roads, the speed limits of road sections are listed in the table of the top ranked new cycleway locations to inform prioritisation and speed reduction interventions. -->
+
+<!-- ## 3. Cohesive network -->
+
+<!-- This layer represents what a joined-up cycle network could look like if new cycleways were placed along the key corridors of high potential.  -->
+<!-- Unlike the 'top ranked new cycleways' layer, the cohesive network comprises all of the major high cycle potential corridors, including sections where the roads are narrower.  -->
+<!-- It shows what a joined-up cycle network could look like if there were no constraints on road space, along some of the most important corridors. -->
+
+<!-- The cohesive network layer includes sections where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic. -->
+<!-- It was created to highlight the need for a cohesive cycle network that does not 'give up' at awkward junctions. -->
+<!-- Cohesive networks are a key to getting people cycling [@buehler_bikeway_2016]. -->
+
+<!-- This layer supports consideration of **long term** plans and alternatives to road space reallocation. -->
+<!-- Some sections of the cohesive network may still be suitable candidates for road-space reallocation (for example, [in Lancaster](https://www.facebook.com/photo.php?fbid=10158393513440797&set=gm.1409892779205719&type=1&theater) protected cycle lanes have recently been installed on a road which is less than 10 m in width), but other sections may be best suited to different types of traffic measure, such as point closures, cycleway installation on a nearby parallel alignment, or conversion to one way streets in order to free up space for cycle lanes.
+-->
+
+<!-- The layer was generated by identifying the roads (classified by reference number) on which there is greatest cycling potential across the region and joining them together. -->
+<!-- Specifically we first selected all road segments with cycle potential in the 85th percentile and above, which also belong to classified trunk roads, A roads or B roads. -->
+<!-- These segments are then joined together to form a continuous network. -->
+
+<!-- ## 4. Spare lane(s) (not shown by default) -->
+
+<!-- These are routes that have more than one vehicle lane in either direction and have a cycling potential above the minimum threshold (which varies according to region as it needs to be higher in e.g.
+Greater London than in rural areas).
+-->
+<!-- This layer can be used to identify where there may be spare lanes on the network that could be reallocated to cycling (or pedestrianisation) beyond those highlighted in the Top ranked new cycleways layer.
+-->
+
+<!-- ## 5. Width >= 10m (not shown by default) -->
+
+<!-- These routes are at least 10m in width and have a cycling potential above the minimum threshold.  -->
+<!-- Like the Spare lane(s) layer, this layer is intended to highlight road sections on which increasing space for cycling and walking by reallocating road space may be a viable option. -->
+
+<!-- # Acknowledgments -->
+
+<!-- Many thanks to Dr Malcolm Morgan ([University of Leeds](http://www.leeds.ac.uk/)) and Martin Lucas-Smith ([CycleStreets.net](https://www.cyclestreets.net/)) for providing substantial input into methodological and web development aspects of the project respectively.
+-->
+
+<!-- Thanks to Jeremy Clarke, Rabina Nawaz, John Sweetman and Richard Mace from the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) for vital input into the tool's design.
+-->
+<!-- Thanks to Kevin McCann from [Sustrans](https://www.sustrans.org.uk/) for guidance that improved the tool from a policy perspective. -->
+<!-- Thanks to Roger Geffen and others from [CyclingUK](https://www.cyclinguk.org/) for input into prototype versions of the tool. -->
+
+<!-- Thanks to developers of open source software, especially to developers of the [R language](https://www.r-project.org/contributors.html) and the packages [sf](https://github.com/r-spatial/sf) and [tmap](https://github.com/mtennekes/tmap) which underpinned the analysis.
+-->
+<!-- Thanks also to all the contributors to OpenStreetMap who provided open data on which much of the analysis was based, allowing publication of the results as open data under the conditions of the Open Data Commons Open Database License ([ODbL 1.0](https://www.openstreetmap.org/copyright)).
 Data (c) [OpenStreetMap](https://www.openstreetmap.org) contributors.
+-->
 
-# References
+<!-- # References -->

### 2020-06-11 code/index.Rmd
Tweak text for index

@@
 The tool also identifies what a 'cohesive network' for cycling might look like if we were to consider a wider range of interventions such as closing roads to motorised traffic and creating one-way systems.
-This has been estimated by joining up roads that have a high cycling potential on some or all their length.
+Unlike the 'top ranked new cycleways' layer, the cohesive network comprises all of the major high cycle potential corridors, including sections where the roads are narrower.
+
+The layer was generated by joining up roads that have a high cycling potential on some or all their length.
 The 'cohesive network' might help authorities considering area-wide interventions such as pedestrian and cycle zones or modal filters.

### 2020-06-12 code/index.Rmd
Add link to travel to school data

@@
 - total length of cycling infrastructure created
-- cycling potential x infrastructure length
+- cycling potential multiplied by road section length
 
@@
 Currently it is not possible to estimate cycling potential accounting for all journey purposes including shopping and leisure trips, which would significantly increase the estimated cycling potential.
-- The PCT uses data from the 2011 Census and 2011 Cycle to School Census to estimate cycling potential.
-Hence the estimated cycling potential may represent an underestimate if there has been significant construction active in the area since 2011.
+- The PCT uses data from the 2011 Census and 2011 [National School Census](https://www.sciencedirect.com/science/article/pii/S2214140518301257) to estimate cycling potential.
+Hence the estimated cycling potential may represent an underestimate if new developments have been completed in the area since 2011.
 - The tool currently only identifies priority on-road routes for new cycleways.

### 2020-06-12 conversation.Rmd
Updated results - leeds + london
Add conversation draft

--- 2020-05-12 conversation.Rmd
@@
 ---
-title: "Pop-up cycleways provide a peak into post COVID-19 transport futures"
+title: "Cities must act to secure the future of urban cycling: our research shows how"
+# title: "Pop-up cycleways provide a peak into post COVID-19 transport futures"
+output: github_document
 ---
 
-As nations prepare to emerge from months of COVID-19 induced lockdowns, it is becoming clear life cannot simply return to normal.
-The pandemic has transformed all sectors of the economy, not least transport.
+Cities worldwide are preparing for the long transition out of [lockdown](https://www.bsg.ox.ac.uk/research/research-projects/coronavirus-government-response-tracker).
+Physical distancing measures will be in place for many months, with impacts on all walks of life, not least transport.
 
-In the coming weeks and months, walking and cycling will play a vital role in enabling the reopening of many countries, as continued social distancing measures will mean that public transport options are likely to run at a fraction of their intended capacity.
-There is a window of opportunity here that could be missed if roads become busy and dangerous again before new active transport habits become embedded.
+With public transport options running at low capacity and emerging evidence of the role of [air quality](https://www.sciencedirect.com/science/article/pii/S0048969720321215) and [exercise](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7118448/) in mitigating the risks of COVID-19, solutions are needed more than ever.
 
-Local and national governments are scrambling to act.
-In the UK, £2 billion investment in measures to promote walking and cycling has been announced.
-£250 million of this is for "emergency interventions to make cycling and walking safer" as [Transport Secretary Grant Shapps put it](https://www.gov.uk/government/speeches/transport-secretarys-statement-on-coronavirus-covid-19-9-may-2020).
-New [guidance from the Department for Transport](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities/traffic-management-act-2004-network-management-in-response-to-covid-19) advocates "measures to reallocate road space to people walking and cycling".
+New cycleways are being introduced in [many cities](https://twitter.com/robinlovelace/status/1268834853472088066), allowing healthy habits started during the lockdown to continue.
+Transport authorities must act fast, however, to take advantage of the current [cycling boom](https://www.bbc.co.uk/news/business-52564351) while [traffic levels are still below normal](https://www.forbes.com/sites/carltonreid/2020/05/07/car-traffic-rise-of-11-shows-urgent-need-for-pop-ups-to-protect-cyclists-and-pedestrians/amp/), and to avoid [gridlock](https://static1.squarespace.com/static/5d30896202a18c0001b49180/t/5ee090e5635f3e0513867b40/1591775464939/Report_1+-+Congestion+ahead_+a+faster+route+is+now+available.pdf).
+
+<image id="341146" align="centre" caption="New cycleway on space reclaimed from motor traffic as part of the COVID-19 response in Park Lane, London." source="Transport for London." />
+
+Our research at the University of Leeds has identified roads where there is both space and demand for cycling infrastructure.
+
+Our [methods](https://osf.io/7wjb6) have been used in a nationwide project funded by the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) and transport charity [Sustrans](https://www.sustrans.org.uk/) to help relieve immediate pressures on the transport system and create long term change.
+The result is the [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/).
+This is a free and open tool to help ensure that the government's [Emergency Active Travel Fund](https://www.gov.uk/government/publications/emergency-active-travel-fund-local-transport-authority-allocations) is spent where it is most needed, for maximum long term benefit.
+
+## New routes
+
+The [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/) provides an interactive map for every transport authority in England.
+The main result is a list of road sections that have both high demand and sufficient space for cycling.
+
+These are represented as blue lines in the map of Leeds below, many of which are included in the council's [new cycleway plans](https://news.leeds.gov.uk/news/leeds-city-council-announces-emergency-walking-and-cycling-plans-in-response-to-covid-19).
+Additional layers in the map include existing off road cycleways (shown in green), which in many cities are disjointed or of variable quality, and, in purple, a vision of what a joined-up cycle network could look like.
+
+<image id="340974" align="centre" source="University of Leeds" caption="The map output of the Rapid Cycleway Prioritisation Tool for West Yorkshire, zoomed in on Leeds" />
+
+This new tool builds on [our previous work](https://www.jtlu.org/index.php/jtlu/article/view/862), carried out with other universities, to help authorities identify and develop [strategic cycle networks](https://npct.github.io/pct-shiny/regions_www/www/static/03d_other_reports/2019-use-of-pct-report.pdf).
+We created a national dataset of roads, including estimates of the number of lanes in either direction and road width.
+
+This can help identify where roads have space that could be reallocated to widen pavements or to rapidly introduce new cycleways.
+One of the first of has been installed outside a hospital in [Leicester](https://www.leicestermercury.co.uk/news/leicester-news/light-lockdown-traffic-leicester-allows-4083625).
+Community engagement through projects such as [Widen My Path](https://www.widenmypath.com/), which provides a forum for comments on where more space for walking and cycling is most needed, is another vital part of this process.
+
+<tc-read-more id="135248"></tc-read-more>
+
+The tool's focus on road space reallocation is due to the importance of speed and capacity.
+Light segregation measures such as plastic bollards or [wands](https://www.bbc.co.uk/news/uk-england-leeds-52577554) can be implemented much faster than constructing new cycleways from scratch.
+Wide, direct and continuous cycleways, of the type created by road space reallocation on wide roads, are needed for capacity and to ensure physical distancing guidelines can be followed.
+Finally, road space reallocation has environmental benefits, representing '[zero carbon infrastructure](https://www.ice.org.uk/news-and-insight/the-infrastructure-blog/october-2019/net-zero-needs-zero-carbon-infrastructure)'.
+
+An important [finding](https://osf.io/7wjb6) is that many cities have wide roads with spare space, as shown the maps of six major cities below.
+Notably, none of these yet have a joined-up cycle network.
+An example of the type of road highlighted by our tool is [Kirkstall Road](https://www.openstreetmap.org/way/7991370#map=15/53.8049/-1.5718) in Leeds, which has high cycling potential and sufficient width for new cycleways.
+Kirkstall road is already part of plans by [Leeds City Council](https://news.leeds.gov.uk/news/new-measures-to-support-social-distancing-for-people-walking-and-cycling-in-leeds) to become a [trial cycleway](https://www.transportxtra.com/publications/local-transport-today/news/65382/leeds-to-pilot-enhanced-cycling-safety-u).
+
+Our methods, based on open data and code, could be used in cities worldwide.
+Given the global nature of the challenges we face, open and collaborative research is vital.
+The potential for international application can be seen in [research we carried out](https://atfutures.github.io/upthat/articles/adaptation.html) for the World Health Organisation on possible cycling uptake in cities in low income countries.
+
+<image id="341363" align="centre" source="Robin Lovelace" caption="Maps showing existing, disjointed cycleway networks (green), potential cycleway routes on wide roads according to the Rapid Cycleway Prioritisation Tool (blue) and cohesive networks (purple) in 6 major cities." zoomable="true" />
+
+## Planning for the future
+
+The UK government has announced £2 billion of investment in measures to promote walking and cycling in England over the next five years.
+£250 million has been allocated for emergency interventions [to make cycling and walking safer](https://www.gov.uk/government/speeches/transport-secretarys-statement-on-coronavirus-covid-19-9-may-2020).
+
+Similar commitments are being made in [Scotland](https://www.transport.gov.scot/news/10-million-to-support-pop-up-active-travel-infrastructure/) and [Wales](https://gov.wales/councils-challenged-change-wales-transport-environment-during-coronavirus-lockdown).
+Local authorities urgently need to decide how this funding should be spent.
+
+<tweet url="https://twitter.com/laura_laker/status/1257660564098162689" />
+
+If action is prioritised in places where there is a long-term need for cycle improvements, there is a greater chance that these developments can become permanent.
+In Paris, COVID-19 related measures aim to contribute towards Mayor Anne Hidalgo's long-term [Plan Velo](https://www.curbed.com/2020/1/15/21065343/bike-paris-cycling-anne-hidalgo).
 
@@
 
-Local authorities urgently need answers on how this funding should be spent, and where road space reallocation should be prioritised.
-New cyclinginfrastructure which is developed based on rigorous analysis of the best available data, in combination with local knowledge and community engagement, are more likely to be effective than plans based on intuition alone.
-Recognising the importance of a strong and actionable evidence-base, and the window of opportunity, we used our experience developing tools that are already being used by local authorities, notably the multi-institution PCT and the CyIPT, to develop methods and outputs taylored for the new transport planning landscape.
-
-<image id="334100" align="centre" caption="Screenshot from the website salfordliveablestreets.commonplace.is to support local responses to the COVID-19 pandemic, including the prioritisation of pop-up active transport infrastructure." />
-
-In a pre-print [paper](https://osf.io/7wjb6) on methods to prioritise pop-up road space reallocation schemes, we have classified pop-up interventions into three categories.
-Firstly, in streets that are wide or have multiple lanes, roadspace can be reallocated to widen pavements or introduce pop-up cycleways, such as outside a hospital in [Leicester](https://www.leicestermercury.co.uk/news/leicester-news/light-lockdown-traffic-leicester-allows-4083625).
-Secondly, physical barriers can be used to make routes for walking and cycling [faster and more direct](https://cyclehighways.eu/design-and-build/infrastructure/filtered-permeability.html) than the routes for motor vehicles.
-This can be seen in Tower Hamlets, London, where simple barriers like planters are used to good effect.
-Thirdly, roads can be closed altogether to motor vehicles, as is happening in [Deansgate, Manchester](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175).
-
-The first of these three options is the most simple to implement,  with relatively few implications for traffic flow.
-We have mapped out **link** the top locations for this reallocation of road space in ten major cities across England, with the intention of rolling out these methods across the country in the near future.
-Having identified roads where a spare lane could be re-purposed for cycling, we have assessed the cycle potential of these streets based on census data for travel to work and cycling uptake scenarios under the UK Government’s national target of doubling cycling levels by 2025.
-The length of road segments is also used to filter the top schemes.
-
-The use of open data and publicly accessible models is a key aspect of this research.
-Given the global nature of the challenges we face, open and collaborative research is vital so that methods can be applied by as wide a user base as possible.
-The potential for international application can be seen for example in [research for the World Health Organisation](https://atfutures.github.io/upthat/articles/adaptation.html) modelling cycling uptake in cities in low income countries.
-Within the UK, [50+ Local Authorities](https://npct.github.io/pct-shiny/regions_www/www/static/03d_other_reports/2019-use-of-pct-report.pdf) are now using the PCT to inform their strategic cycle network plans.
-
-Results from our [methods](https://osf.io/7wjb6) show that many cities have 'spare lanes' under the current levels of traffic in city centres and near key worker destinations such as hospitals.
-These results could support the prioritisation of vital resources dedicated to pop-up infrastructure.
-With the situation rapidly developing, some of the roads shown here, such as [Kirkstall Road](https://www.transportxtra.com/publications/local-transport-today/news/65382/leeds-to-pilot-enhanced-cycling-safety-u) (in the top 10 results our methods generated for Leeds), have already been identified by Local Authorities as targets for intervention.
-As the recovery continues, we hope our methods will support cities to take the opportunities that the crisis presents for active travel.
-
-<image id="334113" align="centre" caption="Roads in Leeds with at least one spare lane that could be reallocate to provide space for active travel, and locations of hospitals" />
-
-<tweet url="https://twitter.com/laura_laker/status/1257660564098162689" />
-
-With the rush to act, there is a great need for evidence of where new interventions should be prioritised.
-As with the medical science, research is needed now.
-Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in is effective, safe, and placed where it is most needed.
-Agile research can respond to rapidly changing conditions, as government priorities shift and different sectors of the economy reopen.
-Measures can be undertaken on a temporary basis, but if action is prioritised in places where there is a long-term need for cycle improvements, there is a greater chance that they become permanent.
-In Paris, Covid-19 related measures aim to contribute towards Mayor Anne Hidalgo's long-term [Plan Velo](https://www.curbed.com/2020/1/15/21065343/bike-paris-cycling-anne-hidalgo).
+New cycling infrastructure is more likely to be effective when it is developed based on analysis of the best available data, in combination with vital local knowledge.
+City planners, politicians and citizens need to act to ensure that transport interventions made during the crisis are of maximum benefit now and in the post-pandemic world.
 
@@
 
-<!-- ![](https://raw.githubusercontent.com/cyipt/tempCycleways/master/README_files/figure-gfm/manchester-1.png) -->
+<!-- ![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/README_files/figure-gfm/manchester-1.png) -->
 

### 2020-06-12 report.Rmd
Typo fix II
grouping methods

@@
 
-The rationale was to provide interactive evidence that supports scheme prioritisation based on a range of variables.
-Length, for example, is important for creating a joined-up and continuous, a vital component of cycling strategies [@parkin_designing_2018].
+The rational was to provide interactive evidence that supports scheme prioritisation based on a range of variables.
+Length, for example, is important for creating a joined-up and continuous network, a vital component of cycling strategies [@parkin_designing_2018].
 
@@
 Only roads with a continuous length above a threshold distance (500 m) identified as candidates for inclusion in the Top ranked new cycleways layer. 
-Roads with an existing cycleway along 75% or more of their length were excluded (we selected such a high value because the presence of a cycleway in OSM does not guarantee that this is a high quality, continuous route).
+Roads with an existing cycleway along 80% or more of their length were excluded (we selected such a high value because the presence of a cycleway in OSM does not guarantee that this is a high quality, continuous route).
 Roads were ranked in terms of their cycling potential and highest ranking roads were included in the roads.

### 2020-06-13 code/index.Rmd
Final changes for release

@@
 
-The technical specification for the Rapid Cycleway Prioritisation Tool can be found [here](https://github.com/cyipt/popupCycleways/releases/download/0.3/report.pdf), providing further detail about the assumptions underlying the tool.
+The technical specification for the Rapid Cycleway Prioritisation Tool can be found [here](https://github.com/cyipt/popupCycleways/releases/download/v1/report.pdf), providing further detail about the assumptions underlying the tool.
 
@@
 If you would like to support the project with technical input, please provide feedback via the [GitHub issue tracker](https://github.com/cyipt/popupCycleways/issues/14).
-Data generated by the project can be downloaded from GitHub (e.g. see GeoJSON files for London [here](https://github.com/cyipt/popupCycleways/tree/gh-pages/release/greater-london)).
+Data generated by the project can be downloaded from GitHub (e.g. see GeoJSON files for London [here](https://github.com/cyipt/popupCycleways/tree/gh-pages/v1/greater-london)).
 
@@
 To use the tool, click on the region of interest below and click on the link ([www.cyipt.bike/rapid/west-yorkshire](https://www.cyipt.bike/rapid/west-yorkshire/) for West Yorkshire, for example).
-Explore the results in the new page that appears with reference to the guidance on this page and the [Rapid Cycleway Prioritisation Tool report](https://github.com/cyipt/popupCycleways/releases/download/0.3/report.pdf).
+Explore the results in the new page that appears with reference to the guidance on this page and the [Rapid Cycleway Prioritisation Tool report](https://github.com/cyipt/popupCycleways/releases/download/v1/report.pdf).
 
@@
 ```{r}
-dn = "popupCycleways/release" # dirctory name
+dn = "popupCycleways/v1" # dirctory name
 built_regions = list.files(dn)
@@
 
-<!-- # Acknowledgments -->
+# Acknowledgments
 
-<!-- Many thanks to Dr Malcolm Morgan ([University of Leeds](http://www.leeds.ac.uk/)) and Martin Lucas-Smith ([CycleStreets.net](https://www.cyclestreets.net/)) for providing substantial input into methodological and web development aspects of the project respectively.
--->
+Many thanks to Dr Malcolm Morgan ([University of Leeds](http://www.leeds.ac.uk/)) and Martin Lucas-Smith ([CycleStreets.net](https://www.cyclestreets.net/)) for providing substantial input into methodological and web development aspects of the project respectively.
 
-<!-- Thanks to Jeremy Clarke, Rabina Nawaz, John Sweetman and Richard Mace from the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) for vital input into the tool's design.
--->
-<!-- Thanks to Kevin McCann from [Sustrans](https://www.sustrans.org.uk/) for guidance that improved the tool from a policy perspective. -->
-<!-- Thanks to Roger Geffen and others from [CyclingUK](https://www.cyclinguk.org/) for input into prototype versions of the tool. -->
+Thanks to Jeremy Clarke, Rabina Nawaz, John Sweetman and Richard Mace from the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) for vital input into the tool's design.
+Thanks to Kevin McCann from [Sustrans](https://www.sustrans.org.uk/) for guidance that improved the tool from a policy perspective.
+Thanks to Roger Geffen and others from [CyclingUK](https://www.cyclinguk.org/) for input into prototype versions of the tool.
 
-<!-- Thanks to developers of open source software, especially to developers of the [R language](https://www.r-project.org/contributors.html) and the packages [sf](https://github.com/r-spatial/sf) and [tmap](https://github.com/mtennekes/tmap) which underpinned the analysis.
--->
-<!-- Thanks also to all the contributors to OpenStreetMap who provided open data on which much of the analysis was based, allowing publication of the results as open data under the conditions of the Open Data Commons Open Database License ([ODbL 1.0](https://www.openstreetmap.org/copyright)).
+Thanks to developers of open source software, especially to developers of the [R language](https://www.r-project.org/contributors.html) and the packages [sf](https://github.com/r-spatial/sf) and [tmap](https://github.com/mtennekes/tmap) which underpinned the analysis.
+Thanks also to all the contributors to OpenStreetMap who provided open data on which much of the analysis was based, allowing publication of the results as open data under the conditions of the Open Data Commons Open Database License ([ODbL 1.0](https://www.openstreetmap.org/copyright)).
 Data (c) [OpenStreetMap](https://www.openstreetmap.org) contributors.
--->
 

### 2020-06-13 report.Rmd
Final changes for release

@@
 ---
-title: "Rapid Cycleway Prioritisation Tool"
-subtitle: "Technical specification"
-author: Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds
+title: "Rapid Cycleway Prioritisation Tool: Technical specification"
+subtitle: "Version 1.0, June 2020, Institute for Transport Studies, University of Leeds"
+author: Dr Robin Lovelace & Dr Joey Talbot
 # output: bookdown::word_document2
@@
 browseURL("report.pdf")
+piggyback::pb_upload("report.pdf")
+piggyback::pb_download_url("report.pdf")
 browseURL("report.docx")
@@
 # Introduction
+
+[Sustrans](https://www.sustrans.org.uk/) and the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) commissioned ITS Leeds to develop the [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/) to help identifying promising locations for new cycleways in England.
+The immediate purpose of this project is to inform bids for Tranche 2 of the [Emergency Active Travel Fund](https://www.gov.uk/government/publications/emergency-active-travel-fund-local-transport-authority-allocations), particularly for authorities which have not yet developed Local Cycling and Walking Investment Plans ([LCWIP](https://www.gov.uk/government/publications/local-cycling-and-walking-infrastructure-plans-technical-guidance-and-tools)).
 
@@
 
-The approach was designed to identify roads and road segments where investment in new cycleways should be prioritised using a transparent methodology and nationally consistent datasets, alongside planning authority expertise and local knowledge.
+The approach was designed to identify roads and road segments where investment in new cycleways should be prioritised using a transparent methodology and nationally consistent datasets, alongside planning authority expertise and local knowledge [@lovelace_methods_2020].
 Strong and actionable datasets already exist to support strategic cycle network planning, notably the Propensity to Cycle Tool (PCT), a national planning support system.

### 2020-06-15 rapid-slides.Rmd
Update rapid slides

---
title: "The Rapid Cycleway Prioritisation Tool"
subtitle: "`r emojifont::emoji('rocket')`<br/>methods and results"
author: "Robin Lovelace and Joey Talbot"
institute: "University of Leeds, [Institute for Transport Studies](https://environment.leeds.ac.uk/transport)"
date: "Green Transport Recovery Series, 2020-06-17"
output:
  xaringan::moon_reader:
    css: [default, robot, robot-fonts]
    lib_dir: libs
    nature: 
      beforeInit: "https://platform.twitter.com/widgets.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, echo=FALSE}
knitr::include_graphics("https://pbs.twimg.com/media/EToi6UVXQAEmY2h?format=jpg&name=large")
```

<!-- 14h30-15h30 lecture part (this could already drift towards the practical part but doesn't have to ...) -->

<!-- 15:30-16h00 coffee break -->

<!-- 16h00-17h00 hands-on workshop (could be somewhat open-ended, we'll have the room all afternoon - but if we would go into extra time, we should make a short 'cut' around 17h such that those who have/want to go home can do so). -->

<!-- Sally: see it as an opportunity - vital for change -->

<!-- Lucy: Propensity to cycle tool -->

<!-- It's brilliant, magic, helps imagine what the future could be like -->

---

# Motivations

---

# Data

---

# Methods

---

# Results

---

# Demo

---

---

# Links I

- Sheffield [infrastructure](https://www.cyclesheffield.org.uk/2018/01/01/cyclesheffield-response-to-portobello-trippet-lane-cycle-route-designs/)

- Article on Yorkshire Voice on the Leeds-Bradford [Cycle Superhighway](http://www.yorkshire-voice.com/mixed-views-as-29m-cycle-superhighway-linking-leeds-and-bradford-nears-completion/)

- A detailed article on the same cycleway by [Mark Strong](https://aseasyasridingabike.wordpress.com/2016/08/11/a-visit-to-the-leeds-bradford-cycle-superhighway/)

- Academic article on thinking at level of cycleway networks (Buehler and Dill [2015](https://www.tandfonline.com/doi/full/10.1080/01441647.2015.1069908?journalCode=ttrv20))

- The Propensity to Cycle Tool ([pct.bike](https://www.pct.bike/))

- The Cycling Infrastructure Prioritisation Toolkit ([CyIPT.bike](https://www.cyipt.bike/existing/#13/53.4690/-2.2506/mapnik))

- Consultation on the Torrington Place / Tavistock Place 'pop-up' cycleway from [Cambden.gov.uk](https://www.camden.gov.uk/documents/20142/3452947/Consultation+leaflet+FINAL.pdf/f628d6e8-c47b-82f1-cb40-8e24db78bea6)

---

# Links II

- Map of pop-up interventions in response to COVID-19 from [CyclingUK](https://www.cyclinguk.org/covid-19-safe-space-social-distancing)

- Maps of roads with 'spare lanes' in 10 cities from our our 'cyipt' [website](https://www.cyclinguk.org/article/10-cities-where-pop-bike-lanes-could-benefit-millions-explore-maps)

- New statutory [guidance from DfT on road space reallocation](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities)

- Analysis to support [road space reallocation in Manchester](https://github.com/cyipt/popupCycleways/blob/master/code/mcr-safe-streets-analysis.md#analysis-of-cycling-potential-around-key-corridors-in-greater-manchester)

- Image of new [pop-up going in in Greater Manchester](https://github.com/cyipt/popupCycleways/issues/27)

---

# Limitations and next steps

- Source: Adam Reynolds on a [GitHub issue](https://github.com/luukvdmeer/sfnetworks/issues/13)

![](https://user-images.githubusercontent.com/1948692/74239070-0429a200-4ccf-11ea-8ce0-5530e84db2ff.jpeg)

---

# Thanks!

Contact: me at

- r. lovelace at leeds ac dot uk (email)

- [`@robinlovelace`](https://twitter.com/robinlovelace) (Twitter)

--

- Check out www.pct.bike

--

- www.cyipt.bike

--

- Pop-up cycleway info: watch this space

```{r, echo=FALSE, eval=FALSE}
# export code from slides
knitr::purl("tds-regional-street.Rmd")
```

### 2020-06-15 rapid-slides.Rmd
Update slides

@@
 ---
+
+```{r, include=FALSE}
+library(dplyr)
+```
 
@@
 
+background-image: url(https://pbs.twimg.com/media/EZvQNJlWoAE4D_c?format=jpg&name=4096x4096)
+
 # Motivations
+
+--
+
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+
+1) reduced capacity on public transport 2) need for space so physical distancing guidelines can be followed
+3) healthy restart 4) need to act fast
+
+---
+
+### How to cope with reduced public transport capacity post COVID-19?
+
+Plans from Leeds City Council responding to national [guidance](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities) and [funding](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) for 'pop-up' cycleways:
+
+![](https://pbs.twimg.com/media/EZ_-A0dXgAAlBzt?format=png&name=900x900)
+
+---
+
+## Spatial networks can support planning
+
+### Transport networks are spatial networks
+
+Spatial network representation in R:
+
+![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/mapout.png)
+
+---
+
+background-image: url("https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/facet-output.png")
+background-position: right
+background-size: contain
+
+--
+
+### sf + igraph used in project to support investment in 'pop-ups' nationwide
+
+Every city in England
+
+--
+
+Results in Leeds:
+
+```{r, echo=FALSE, results='markup'}
+res_leeds = sf::read_sf("https://www.cyipt.bike/rapid/west-yorkshire/top_routes.geojson")
+res_leeds %>% 
+  arrange(desc(mean_cycling_potential)) %>% 
+  slice(1:5) %>%
+  sf::st_drop_geometry() %>% 
+  dplyr::select(name) %>% 
+  as.data.frame()
+```
+
+Cycleway planned for
+
+Kirkstall Road (source: [BBC](https://www.bbc.co.uk/news/uk-england-leeds-52577554))
+
+Free and open results:
+
+[www.cyipt.bike/rapid](https://www.cyipt.bike/rapid)
+
+Source code: [github.com/cyipt/popupCycleways](https://github.com/cyipt/popupCycleways)
+
+<!-- ![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/facet-output.png) -->
+
+```{r, out.width="70%", echo=FALSE}
+# knitr::include_graphics("https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/facet-output.png")
+```
+
+---
+
+## Ingredients to prioritise transport infrastructure
+
+.left-column[
+
+- Current travel patterns
+- Road attribute data 
+- Potential usage levels
+- Impact on overall network
+
+**Spatial networks can help, especially with the last two**
+
+Source: https://www.pct.bike/
+
+]
+
+.right-column[
+
+![](figs/pct.png)
+
+]
 

### 2020-06-15 rapid-slides.Rmd
Update slides

@@
 title: "The Rapid Cycleway Prioritisation Tool"
-subtitle: "`r emojifont::emoji('rocket')`<br/>methods and results"
+subtitle: "🛣🚲🚀<br/>methods and results<br/>[cyipt.bike/rapid](https://www.cyipt.bike/rapid/)"
 author: "Robin Lovelace and Joey Talbot"
-institute: "University of Leeds, [Institute for Transport Studies](https://environment.leeds.ac.uk/transport)"
+institute: "University of Leeds, [Institute for Transport Studies](https://environment.leeds.ac.uk/transport). [Source code](https://github.com/cyipt/popupCycleways/blob/master/rapid-slides.Rmd)."
 date: "Green Transport Recovery Series, 2020-06-17"
@@
 
-```{r, echo=FALSE}
-knitr::include_graphics("https://pbs.twimg.com/media/EToi6UVXQAEmY2h?format=jpg&name=large")
-```
-
-<!-- 14h30-15h30 lecture part (this could already drift towards the practical part but doesn't have to ...) -->
-
-<!-- 15:30-16h00 coffee break -->
-
-<!-- 16h00-17h00 hands-on workshop (could be somewhat open-ended, we'll have the room all afternoon - but if we would go into extra time, we should make a short 'cut' around 17h such that those who have/want to go home can do so).
--->
-
-<!-- Sally: see it as an opportunity - vital for change -->
-
-<!-- Lucy: Propensity to cycle tool -->
-
-<!-- It's brilliant, magic, helps imagine what the future could be like -->
-
----
-
 background-image: url(https://pbs.twimg.com/media/EZvQNJlWoAE4D_c?format=jpg&name=4096x4096)
+background-position: bottom
+background-size: contain
 
@@
 
-### How to cope with reduced public transport capacity post COVID-19?
+# How does the tool identify ‘top ranked new cycleways’?
 
-Plans from Leeds City Council responding to national [guidance](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities) and [funding](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) for 'pop-up' cycleways:
-
-![](https://pbs.twimg.com/media/EZ_-A0dXgAAlBzt?format=png&name=900x900)
+![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/mapout.png)
 
@@
 
-## Spatial networks can support planning
+# How does the tool identify what a ‘cohesive network’ might look like?
 
-### Transport networks are spatial networks
+---
 
-Spatial network representation in R:
+# How can I find the results for my area?
 
-![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/mapout.png)
+---
+
+# How should I use these results?
 
@@
 
-### sf + igraph used in project to support investment in 'pop-ups' nationwide
+Every city in England
 
-Every city in England
+## Results
+
+---
+
+# Case study: Leeds
+
+.right-column[
 
@@
 
-Free and open results:
+See the results at:
 
-[www.cyipt.bike/rapid](https://www.cyipt.bike/rapid)
+[www.cyipt.bike/rapid/west-yorkshire](https://www.cyipt.bike/rapid/west-yorkshire)
 
-Source code: [github.com/cyipt/popupCycleways](https://github.com/cyipt/popupCycleways)
+]
 
-<!-- ![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/facet-output.png) -->
+.left-column[
 
-```{r, out.width="70%", echo=FALSE}
-# knitr::include_graphics("https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/facet-output.png")
-```
+![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/mapout.png)
+
+]
 
@@
 
-## Ingredients to prioritise transport infrastructure
+---
 
-.left-column[
+## What are the key limitations of the tool?
 
-- Current travel patterns
-- Road attribute data 
-- Potential usage levels
-- Impact on overall network
+---
 
-**Spatial networks can help, especially with the last two**
+## Where can I find further information about the tool?
 
-Source: https://www.pct.bike/
+---
 
-]
+## How to cope with reduced public transport capacity post COVID-19?
 
-.right-column[
+Plans from Leeds City Council responding to national [guidance](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities) and [funding](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) for 'pop-up' cycleways:
 
-![](figs/pct.png)
-
-]
+![](https://pbs.twimg.com/media/EZ_-A0dXgAAlBzt?format=png&name=900x900)
 

### 2020-06-15 rapid-slides.Rmd
Update slides

@@
 
-# How does the tool identify ‘top ranked new cycleways’?
+### How to identify ‘top ranked new cycleways’?
 
 ![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/mapout.png)
+
+---
+
+Tablular results for Leeds
+
+```{r, echo=FALSE, out.height="100%"}
+res_leeds = sf::read_sf("https://www.cyipt.bike/rapid/west-yorkshire/top_routes.geojson")
+res_table = res_leeds %>%
+  sf::st_drop_geometry() %>% 
+  arrange(desc(mean_cycling_potential))
+# %>% 
+#   select(
+#     Name = name,
+#     Reference = ref,
+#     `Length (m)` = group_length,
+#     `Cycling potential` = mean_cycling_potential,
+#     `Length * potential (km)` = km_cycled,
+#     # `SRN` = srn,
+#     `Speed limit` = speed_limit,
+#     `Mean estimated width (m)` = mean_width,
+#     `Majority spare lane(s)?` = majority_spare_lane
+#     ) 
+DT::datatable(res_table, height = 600)
+```
 
@@
 
-.right-column[
-
---
+.left-column[
 
@@
 
-See the results at:
-
-[www.cyipt.bike/rapid/west-yorkshire](https://www.cyipt.bike/rapid/west-yorkshire)
-
 ]
 
-.left-column[
+.right-column[
 
 ![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/mapout.png)
+
+See the results at [www.cyipt.bike/rapid/west-yorkshire](https://www.cyipt.bike/rapid/west-yorkshire)
 

### 2020-06-15 rapid-slides.Rmd
Add top ranked cycleways slide

@@
 
-![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/mapout.png)
+<!-- ![](figures/results-top-leeds.png) -->
+
+- All road sections with spare lanes or estimated width > 10 m (CyIPT)
+- Of those, which have a minum continuous length of at least 500 m?
+- Rank the roads by cycling potential (commute + school, Government Target)
+- Show the top n, where n depends on the size of the region (70 in London)
+
+![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-top-leeds.png)
 

### 2020-06-15 rapid-slides.Rmd
Add cohesive network slide

@@
 
-# How does the tool identify what a ‘cohesive network’ might look like?
+### How to identify a ‘cohesive network’?
+
+- 'Long term layer'
+- Intended to show what a joined-up cycle network could look like if we were to consider new cycleways by either closing roads to motorised traffic or creating one-way systems
+- Find  roads that have high cycling potential on some or all of their length
+- Join them up!
+
+![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-cohesive-leeds.png)
 

### 2020-06-15 rapid-slides.Rmd
Add existing cycleways slide

--- 2020-06-15 rapid-slides.Rmd
@@
 
+### Existing cycleways
+
+- Aims to identify **immediate** issues
+- From OpenStreetMap, March 2020
+- Contains sections of variable quality
+- Aim to highligh where gaps need to be filled
+
+![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-cycleways-leeds.png)
+
+---
+
 ### How to identify ‘top ranked new cycleways’?
@@
 
+- Aims to identify **short term priorities**
 - All road sections with spare lanes or estimated width > 10 m (CyIPT)
@@
 
-# How does the tool identify what a ‘cohesive network’ might look like?
+### How to identify a ‘cohesive network’?
+
+- Aims to support **long term planning**
+- Intended to show what a joined-up cycle network could look like if we were to consider new cycleways by either closing roads to motorised traffic or creating one-way systems
+- Find  roads that have high cycling potential on some or all of their length
+- Join them up
+
+![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-cohesive-leeds.png)
 

### 2020-06-15 rapid-slides.Rmd
Add results for regions
Complete slides

@@
 3) healthy restart 4) need to act fast
+
+---
+
+Plans from Leeds City Council responding to national [guidance](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities) and [funding](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) for 'pop-up' cycleways:
+
+![](https://pbs.twimg.com/media/EZ_-A0dXgAAlBzt?format=png&name=900x900)
 
@@
 
-# How can I find the results for my area?
+### How can I find the results for my area?
+
+.left-column[
+
+Head to [cyipt.bike/rapid/](https://www.cyipt.bike/rapid/)
+
+]
+
+.right-column[
+
+![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-regions.png)
+]
 
@@
 # How should I use these results?
+
+- Aim: to support evidence-based prioritisation of bids for Tranche 2 of the Emergency Active Travel Fund
+- Complementing local knowledge and stakeholder engagement
+- Scoping of options
+- Order of implementations
+
+Investment should depend on many factors including:
+
+- cycling potential (default metric to rank schemes)
+- total length of cycling infrastructure created
+- cycling potential multiplied by road section length
 
@@
 
-# Case study: Leeds
+# Case study demo: Leeds
 
@@
 
-## What are the key limitations of the tool?
+## Limitations
+
+- Narrow focus on road space reallocation options
+- No economic appraisal
+  - Use the Active Mode Appraisal Toolkit ([AMAT](https://www.gov.uk/government/publications/tag-social-and-distributional-impacts-worksheets)) 
+- The best locations for new cycleways may not always have space
+  - Search for parallel routes, which may be available
+- Many options not in top routes
+  - see cohesive network and look for other option
+- The road network data is from 2017 and widths are estimates
+  - Compare with real world knowledge
+- Only work and school trips included
+  - Use local knowledge of new developments and other 'trip attractors
+- Prioritises routes in large cities in authorities with multiple settlements
+  - Use alternative sources of information, including the PCT
 
@@
 
----
+- See the technical specification [here](https://github.com/cyipt/popupCycleways/releases/download/v1/report.pdf)
 
-## How to cope with reduced public transport capacity post COVID-19?
+Data about the tools informing the model can be found here:
 
-Plans from Leeds City Council responding to national [guidance](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities) and [funding](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) for 'pop-up' cycleways:
+- The [Propensity to Cycle Tool](https://www.pct.bike/)
+- The [Cycling Infrastructure Prioritisation Tool](https://www.cyipt.bike)
+- [OpenStreetMap](https://www.openstreetmap.org/#map=7/52.486/-2.428&layers=C)
 
-![](https://pbs.twimg.com/media/EZ_-A0dXgAAlBzt?format=png&name=900x900)
+- Code underlying the project can be accessed [here](https://github.com/cyipt/popupCycleways).
+- Support the project with technical input via the [GitHub issue tracker](https://github.com/cyipt/popupCycleways/issues/14)
+- Data generated by the project can be downloaded from GitHub (e.g. see GeoJSON files for London [here](https://github.com/cyipt/popupCycleways/tree/gh-pages/v1/greater-london)).
 
@@
 
-# Data
-
----
-
-# Methods
-
----
-
-# Results
-
----
-
-# Demo
-
----
-
----
-
-# Links I
-
-- Sheffield [infrastructure](https://www.cyclesheffield.org.uk/2018/01/01/cyclesheffield-response-to-portobello-trippet-lane-cycle-route-designs/)
-
-- Article on Yorkshire Voice on the Leeds-Bradford [Cycle Superhighway](http://www.yorkshire-voice.com/mixed-views-as-29m-cycle-superhighway-linking-leeds-and-bradford-nears-completion/)
-
-- A detailed article on the same cycleway by [Mark Strong](https://aseasyasridingabike.wordpress.com/2016/08/11/a-visit-to-the-leeds-bradford-cycle-superhighway/)
+# Links
 
@@
 
----
-
-# Links II
-
 - Map of pop-up interventions in response to COVID-19 from [CyclingUK](https://www.cyclinguk.org/covid-19-safe-space-social-distancing)
-
-- Maps of roads with 'spare lanes' in 10 cities from our our 'cyipt' [website](https://www.cyclinguk.org/article/10-cities-where-pop-bike-lanes-could-benefit-millions-explore-maps)
 
@@
 
-# Limitations and next steps
+# Next steps
 
-- Source: Adam Reynolds on a [GitHub issue](https://github.com/luukvdmeer/sfnetworks/issues/13)
+- Many future directions of travel including evidence to support [Low Traffic Neighbourhoods](https://github.com/luukvdmeer/sfnetworks/issues/13)
 
@@
 
-- www.cyipt.bike
-
---
-
-- Pop-up cycleway info: watch this space
-
-```{r, echo=FALSE, eval=FALSE}
-# export code from slides
-knitr::purl("tds-regional-street.Rmd")
-```
+- www.cyipt.bike/rapid

### 2020-06-15 rapid-slides.Rmd
Tweak slides

@@
 ---
+
+```{r, engine='bash', eval=FALSE, echo=FALSE}
+cp rapid-slides.html popupCycleways/v1/slides/index.html
+cp -Rv libs popupCycleways/v1/slides/
+cd popupCycleways
+git status
+git add -A
+git commit -am 'Add slides'
+git push
+cd -
+```
+
+```{r, eval=FALSE, echo=FALSE}
+# Aim: make pdf slides
+pagedown::chrome_print("rapid-slides.html")
+browseURL("rapid-slides.pdf")
+piggyback::pb_upload("rapid-slides.pdf")
+piggyback::pb_download_url("rapid-slides.pdf")
+```
 
@@
 
----
-
 ## Limitations
@@
 - www.cyipt.bike/rapid
+
+<!-- Tweet: @ launch -->
+
+<!-- The Rapid Cycleway Prioritisation Tool is finally launched 🚀 -->
+
+<!-- It provides evidence to support road space reallocation for active travel in response to #COVID19. -->
+
+<!-- Article: -->
+<!-- Preprint of methods: -->
+<!-- Web tool (free and #opensource):  -->

### 2020-06-15 rapid-slides.Rmd
Update slides, close cyipt/icicle#18

@@
 title: "The Rapid Cycleway Prioritisation Tool"
-subtitle: "🛣🚲🚀<br/>methods and results<br/>[cyipt.bike/rapid](https://www.cyipt.bike/rapid/)"
+subtitle: "methods and results<br/>[cyipt.bike/rapid](https://www.cyipt.bike/rapid/)"
 author: "Robin Lovelace and Joey Talbot"
-institute: "University of Leeds, [Institute for Transport Studies](https://environment.leeds.ac.uk/transport). [Source code](https://github.com/cyipt/popupCycleways/blob/master/rapid-slides.Rmd)."
+institute: "University of Leeds, [Institute for Transport Studies](https://environment.leeds.ac.uk/transport).<br/>Publicly available [slides](https://www.cyipt.bike/rapid/slides/).
+Reproducible [code](https://github.com/cyipt/popupCycleways/blob/master/rapid-slides.Rmd).
+Paper [preprint](https://osf.io/7wjb6/)."
 date: "Green Transport Recovery Series, 2020-06-17"
@@
 ```{r, engine='bash', eval=FALSE, echo=FALSE}
+# 🛣🚲🚀
 cp rapid-slides.html popupCycleways/v1/slides/index.html
@@
 
---
-
 <br>
@@
 1) reduced capacity on public transport 2) need for space so physical distancing guidelines can be followed
-3) healthy restart 4) need to act fast
+3) healthy restart 4) need to act fast. Image credit: [TfL](https://twitter.com/W_Bradley/status/1268836932219478016).
 
@@
 
-Plans from Leeds City Council responding to national [guidance](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities) and [funding](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) for 'pop-up' cycleways:
+Plans from Leeds City Council responding to national [guidance](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities) and [funding](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) for 'pop-up' cycleways (image credit: [Leeds City Council](https://news.leeds.gov.uk/news/leeds-city-council-announces-emergency-walking-and-cycling-plans-in-response-to-covid-19)):
 
@@
 ![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-cycleways-leeds.png)
+Image credit: [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/)
 
@@
 ![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-top-leeds.png)
+
+Image credit: [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/)
 
@@
 ![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-cohesive-leeds.png)
+Image credit: [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/)
 
@@
   - Use alternative sources of information, including the PCT
+  - Do additional analysis, e.g. based on data downloaded from the Rapid tool or the PCT
 
@@
 
-# Next steps
+## Next steps
 
-- Many future directions of travel including evidence to support [Low Traffic Neighbourhoods](https://github.com/luukvdmeer/sfnetworks/issues/13)
+Many future directions of travel including evidence to support [Low Traffic Neighbourhoods](https://github.com/luukvdmeer/sfnetworks/issues/13)
 
@@
 <!-- Web tool (free and #opensource):  -->
+
+<!-- CWIS technical reports -->
+<!-- Cost per intervention type -->

### 2020-06-17 rapid-slides.Rmd
Final update to slides, heads-up @joeytalbot

@@
 
---
-
-Every city in England
-
-## Results
-
 ---
@@
 
-- Consultation on the Torrington Place / Tavistock Place 'pop-up' cycleway from [Cambden.gov.uk](https://www.camden.gov.uk/documents/20142/3452947/Consultation+leaflet+FINAL.pdf/f628d6e8-c47b-82f1-cb40-8e24db78bea6)
+- Consultation on the Torrington Place / Tavistock Place 'pop-up' cycleway from [Camden.gov.uk](https://www.camden.gov.uk/documents/20142/3452947/Consultation+leaflet+FINAL.pdf/f628d6e8-c47b-82f1-cb40-8e24db78bea6)
 

### 2020-06-18 rapid-slides.Rmd
Tweaks

@@
 git status
+git diff
 git add -A

### 2020-06-22 article.tex
Add latex version

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Methods to prioritise pop-up active transport infrastructure and their application in a national cycleway prioritisation tool},
  pdfauthor={Robin Lovelace, Malcolm Morgan, Joey Talbot, Martin-Lucas-Smith},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\title{Methods to prioritise pop-up active transport infrastructure and their application in a national cycleway prioritisation tool}
\author{Robin Lovelace, Malcolm Morgan, Joey Talbot, Martin-Lucas-Smith}
\date{}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

In the context of reduced public transport capacity in the wake of the COVID-19 pandemic, governments are scrambling to enable walking and cycling while adhering to physical distancing guidelines.
A range of pop-up options exist, include road space reallocation, which represents a `quick win' for cities with `spare space' along continuous road sections that have high latent cycling potential.
We developed methods to condense the complexity of city networks down to the most promising roads for road space reallocation schemes.
The resulting Rapid Cycleway Prioritisation Tool has been deployed for all cities England to help prioritise emergency funds for new cycleways nationwide.
The methods and concepts could be used to support investment in pop-up infrastructure in cities worldwide.

\hypertarget{research-questions-and-hypothesis}{%
\section{RESEARCH QUESTIONS AND HYPOTHESIS}\label{research-questions-and-hypothesis}}

Much attention has focused on the impacts of COVID-19 on long-distance travel patterns (e.g.~Iacus et al. 2020; Jittrapirom and Tanaksaranond 2020) but short distance travel patterns have also changed.
There has been a notable increase in cycling in some areas (Harrabin 2020) due to the need increased need for exercise close to home for mental and physical health (Jim\a'enez-Pav\a'on, Carbonell-Baeza, and Lavie 2020) and a reduction in public transport options (e.g.~Tian et al. 2020).
The second reason is particularly important given that many `key workers' are low paid, with limited access to private automobiles.

Local and national governments are working out how best to respond.
Many options are available to ensure that citizens can benefit from outdoor activity while minimising health risks, ranging from the hand sanitiser provision to the creation of extra active transport space (Freeman and Eykelbosh 2020).
Installation of `pop-up' active transport infrastructure has been endorsed and implemented in many places (Laker 2020).
The Scottish government, for example, has provided £10 million ``to keep key workers moving'' by ``reallocating road space to better enable this shift and make it safer for people who choose to walk, cycle or wheel for essential trips or for exercise'' (Transport Scotland 2020).
On 9\textsuperscript{th} May 2020, the UK government announced a £250 million package for pop-up active transport infrastructure (Reid 2020).
Significantly, alongside this funding comes updated \href{https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities/traffic-management-act-2004-network-management-in-response-to-covid-19}{statutory guidance} on pop-up infrastructure and safety.
Evidence is needed to ensure that such investment is spent effectively and where it is most needed.

Most pop-up active transport infrastructure can be classified into three broad categories:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  `filtered permeability', where measures such as point closures or contraflow cycle lanes are used to design street networks in which routes for cyclists are faster and more direct than routes for drivers, e.g.~as \href{https://twitter.com/TowerHamletsNow/status/1257564043856019458}{shown} in {[}Tower Hamlets{]}
\item
  banning cars and to pedestrianise streets, as in New York's `Open Streets' scheme (Litman 2020)
\item
  the reallocation of one or more lanes on wide roads to create pop-up cycleways and pavements (Orsman 2020).
\end{enumerate}

The focus of this article is on the third category.
The research question is:

\begin{quote}
How can automated data analysis and interactive visualisation methods help prioritise the reallocation of road space for pop-up active transport infrastructure?
\end{quote}

Because of the recent, localised and often ad-hoc nature of pop-up infrastructure, it is difficult to make, let alone test, hypotheses related to the research question.
Our broad hypothesis is that digital tools based on open data, and crowdsourcing such as the interactive map used to support community-level responses to COVID-19 in Salford (Salford City Council 2020), illustrated in Figure 1, can lead to more effective use of resources allocated to pop-up interventions.

\begin{figure}
\includegraphics[width=1\linewidth]{figures/widenmypath-bristol} \includegraphics[width=1\linewidth]{figures/saferstreets} \caption{Screenshots from the websites widenmypath.com (top) and salfordliveablestreets.commonplace.is (bottom) to support local responses to the COVID-19 pandemic, including the prioritisation of pop-up active transport infrastructure.}\label{fig:commonplace}
\end{figure}

\hypertarget{methods-and-data}{%
\section{METHODS AND DATA}\label{methods-and-data}}

Two key datasets were used for the project:

\begin{itemize}
\tightlist
\item
  Estimates of cycling potential at the street segment level from the UK Department for Transport funded Propensity to Cycle Tool (PCT) project (Goodman et al. 2019; Lovelace et al. 2017)
\item
  Data derived from OpenStreetMap, with several new variables added to support cycling infrastructure planning (see www.cyipt.bike for an overview)
\end{itemize}

Datasets from the PCT and CyIPT project were merged, resulting in crucial variables summarised in Figure 1.
Cycling potential is defined as the number of one-way journeys to work and to school, under a scenario in which the government aim of doubling cycling levels is met.
This does not include other types of journey such as leisure and shopping.

Roads are classified by speed limit because this has been shown to be a key factor associated with the incidence of severe injuries and fatalities of cyclists (Chen and Shen 2016), with odds of cyclist injury on 20 mph roads in London found to be 21\% lower than on 30 mph roads (Aldred et al. 2018).
Therefore the suitability of roads for cycle infrastructure, the preferred degree of physical segregation, or the necessity to reduce traffic speeds could all be influenced by current speed limits.

\begin{table}[H]
\centering
\begin{tabular}{lllll}
\toprule
 & 20 mph or less & 30 mph & 40+ mph & Overall\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{N. observations}}\\
\hspace{1em}N. observations & (N=4777) & (N=19241) & (N=3105) & (N=27123)\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Highway type}}\\
\hspace{1em}\hspace{1em}bridleway & 53 (1.1\%) & 0 (0\%) & 0 (0\%) & 53 (0.2\%)\\
\hspace{1em}cycleway & 1214 (25.4\%) & 0 (0\%) & 0 (0\%) & 1214 (4.5\%)\\
\hspace{1em}footway & 1332 (27.9\%) & 0 (0\%) & 0 (0\%) & 1332 (4.9\%)\\
\hspace{1em}other & 187 (3.9\%) & 715 (3.7\%) & 1608 (51.8\%) & 2510 (9.3\%)\\
\hspace{1em}pedestrian/living\_street & 22 (0.5\%) & 0 (0\%) & 0 (0\%) & 22 (0.1\%)\\
\hspace{1em}primary & 14 (0.3\%) & 1432 (7.4\%) & 955 (30.8\%) & 2401 (8.9\%)\\
\hspace{1em}residential & 605 (12.7\%) & 8604 (44.7\%) & 0 (0\%) & 9209 (34.0\%)\\
\hspace{1em}secondary & 28 (0.6\%) & 1408 (7.3\%) & 176 (5.7\%) & 1612 (5.9\%)\\
\hspace{1em}service & 1018 (21.3\%) & 7 (0.0\%) & 0 (0\%) & 1025 (3.8\%)\\
\hspace{1em}tertiary & 170 (3.6\%) & 5206 (27.1\%) & 290 (9.3\%) & 5666 (20.9\%)\\
\hspace{1em}unclassified & 134 (2.8\%) & 1869 (9.7\%) & 75 (2.4\%) & 2078 (7.7\%)\\
\hspace{1em}bus\_guideway & 0 (0\%) & 0 (0\%) & 1 (0.0\%) & 1 (0.0\%)\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Cycling potential}}\\
\hspace{1em}Mean (SD) & 39.0 (67.8) & 36.9 (66.7) & 40.7 (55.1) & 37.7 (65.7)\\
\hspace{1em}Median [Min, Max] & 14.0 [1.00, 810] & 15.0 [1.00, 810] & 24.0 [1.00, 519] & 16.0 [1.00, 810]\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Width (m)}}\\
\hspace{1em}Mean (SD) & 6.62 (2.71) & 7.34 (2.28) & 8.84 (2.36) & 7.41 (2.42)\\
\hspace{1em}Median [Min, Max] & 7.00 [1.00, 21.0] & 7.00 [1.00, 24.0] & 9.00 [2.00, 21.0] & 7.00 [1.00, 24.0]\\
\hspace{1em}Missing & 1688 (35.3\%) & 1225 (6.4\%) & 447 (14.4\%) & 3360 (12.4\%)\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{N. lanes}}\\
\hspace{1em}1 & 3840 (80.4\%) & 1555 (8.1\%) & 406 (13.1\%) & 5801 (21.4\%)\\
\hspace{1em}2 & 937 (19.6\%) & 16979 (88.2\%) & 2266 (73.0\%) & 20182 (74.4\%)\\
\hspace{1em}3 & 0 (0\%) & 490 (2.5\%) & 289 (9.3\%) & 779 (2.9\%)\\
\hspace{1em}4+ & 0 (0\%) & 217 (1.1\%) & 144 (4.6\%) & 361 (1.3\%)\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{geographic-subsetting}{%
\subsection{Geographic subsetting}\label{geographic-subsetting}}

The region of analysis may seem like a basic consideration: most cities have well-defined administrative zones.
In Leeds and many other cities, it makes sense to focus on the region directly surrounding the city centre, in a kind of `geographical triage' to omit from the analysis pop-up options in the outskirts, focus valuable attention on the routes that are most likely to serve the highest number of people, and ensure that road sections outside administrative areas but close to key destinations are included.

Figure \ref{fig:gsub} shows three broad strategies for geographic subsetting: based on administrative boundaries, distance from the centre, and distance from the centre and key destinations.
Major hospitals are used to illustrate the third strategy, as many key workers need to get to hospitals.
Schools could also be used here as an example of a key destination that may not fit well within administrative zones.
The latter case (Figure \ref{fig:gsub}, right) shows that administrative boundaries can exclude important roads.
The definition of `city centres' and `key destinations' is straightforward in clearly defined and well-understood city planning contexts.
In contexts where the method must be deployed nationwide, however the use of such subsetting approaches was found to be problematic, so the Rapid Cycleway Prioritisation Tool for England (v1) uses the first subsetting option, but subsets by larger regional boundaries to encourage regional collaboration on cycleway network design.

\begin{figure}
\includegraphics[width=1\linewidth]{article_files/figure-latex/gsub-1} \caption{Illustration of geographic subsetting based on administrative boundaries (left), distance to a central point (middle) and distance to city centre or key destinations (right). Radii of 5 km, 8 km and 10 km are shown for reference (note that some roads within 10 km of the center are outside the regional boundary).}\label{fig:gsub}
\end{figure}

\hypertarget{road-attributes}{%
\subsection{Road attributes}\label{road-attributes}}

Pop-up cycleways can be placed either on the side of wide roads (as is the case on \href{https://www.lancasterguardian.co.uk/news/uk-news/mixed-reactions-new-lancaster-pop-cycle-lanes-busy-city-centre-road-2875909}{South Road, Lancaster}) or in an entire lane that has been closed to motor traffic (as is the case on \href{https://metro.co.uk/2020/05/14/road-turns-giant-cycle-lane-make-social-distancing-easier-12703847/}{Park Lane}, London).
Accordingly, we defined `spare space' as either roads on which there is more than one lane in either direction or lane width above a threshold (set at 10 m based on feedback from engineers and the observation that South Road has a width of \textasciitilde9 m yet can just fit cycleways protected by plastic `wands').
This definition assumes no alteration of the navigable network for motor vehicles.

To identify road sections with a spare lane we developed a simple algorithm that takes the OSM variable \href{https://wiki.openstreetmap.org/wiki/Key:lanes}{\texttt{lanes}} if it is present and, if not, derives the number from the highway type and presence/absence of bus lanes.
Width estimates were taken from the CyIPT tool (see \href{https://www.cyipt.bike/}{www.cyipt.bike} for details).
All segments defined as having a spare space using this method are shown in Figure \ref{fig:levels} (left).

\hypertarget{attribute-filtering-and-grouping}{%
\subsection{Attribute filtering and grouping}\label{attribute-filtering-and-grouping}}

To ensure our route recommendations could achieve sufficient coherency, we undertook several stages of road segment filtering and grouping.
Segments were grouped by road reference number (i.e.~`A' or `B' road number) and by proximing, within a 100 m buffer.
Filtering then removed groups without distance weighted mean width \textgreater= 10 m or spare lanes along the majority of their length, and groups with distance-weighted mean cycling potential below a minimum threshold.

Segments without a reference number were subjected to stricter filtering criteria, to prevent the inclusion of unwanted short segments on side streets.
For all segments, a final round of grouping (ignoring previous groups) with a 100 m buffer was then used to remove groups with length below 500 m.
This step removed short sections distant from any others, thus improving the coherency of the results.
Finally, road names were used to identify continuous road sections with the same name of length \textgreater= 500m.
Groups containing five or more different named roads were labeled ``Unnamed road.''
An example of the impact of grouping strategy is shown in Figure \ref{fig:levels}.
The resulting network shows that grouping the segments first then filtering based on mean group-level attributes results in a more cohesive network than filtering individual segments then grouping the results.

\begin{figure}
\includegraphics[width=1\linewidth]{article_files/figure-latex/levels-1} \caption{Illustration of the 'group then filter' method to identify roads with spare space that meet threshold values for length and cycling potential. The right hand panel contains roads on which the majority of segments have spare space (including segments that may not on their own be estimated to have spare space), coloured by group membership.}\label{fig:levels}
\end{figure}

\hypertarget{selection-of-top-routes}{%
\subsection{Selection of top routes}\label{selection-of-top-routes}}

Top routes were selected from the results of the previous steps.
These must not be labeled ``Unnamed road'' or have existing cycleways along more than 80\% of their length.
A high threshold was chosen here because the presence of an existing cycleway on OSM does not mean that this is necessarily a high quality cycleway.
Continuity of cycle provision is important for creating high quality networks (Parkin 2018).

\hypertarget{findings}{%
\section{FINDINGS}\label{findings}}

The results of the method applied to the city of Leeds are shown in Figure \ref{fig:res} (see \href{https://www.cyipt.bike/rapid/west-yorkshire/}{cyipt.bike/rapid} for interactive version) and Table 1.
We found that analysis of open transport network data, alongside careful selection of parameters, can generate plausible results for the prioritisation of pop-up cycle infrastructure.
Reducing the 85,000 road segments for Leeds down to a handful of candidate segments with more than 1 lane near key destinations has great potential to support policy-makers, especially when decisions need to be made fast.

\begin{figure}
\includegraphics[width=1\linewidth]{figures/results-top-leeds} \caption{Results, showing road segments with a spare lane (light blue) and road groups with a minium threshold length, 1km in this case (dark blue). The top 10 road groups are labelled.}\label{fig:res}
\end{figure}

\begin{table}

\caption{\label{tab:unnamed-chunk-2}The top 10 candidate roads for space reallocation for pop-up lane reallocation interventions. Roads with 'spare lanes' identified using methods presented in the paper are ranked by cycling potential under the Government Target scenario, representing a doubling in commuter and school cycling levels compared with 2011 levels.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Name & Length (m) & Potential (Government Target) & Km/day (length * potential)\\
\midrule
Headingley Lane & 971 & 546 & 530\\
A660 & 718 & 414 & 297\\
Woodhouse Lane & 2438 & 372 & 907\\
A65 & 787 & 238 & 187\\
Kirkstall Road & 4407 & 237 & 1044\\
\addlinespace
Clay Pit Lane & 2235 & 231 & 516\\
Low Road & 516 & 194 & 100\\
Chapeltown Road & 1744 & 163 & 284\\
Roundhay Road & 909 & 161 & 146\\
Dewsbury Road & 538 & 157 & 84\\
\bottomrule
\end{tabular}
\end{table}

After initially developing the method for a single city, we applied the methods nationwide.
An illustration of the scale of the results is shown in Figure \ref{fig:facet}, which shows the results for six major cities in England, including existing cycleway and `cohesive network' layers, described on the tool's website \href{https://www.cyipt.bike/rapid/}{cyipt.bike/rapid}.
Local authorities are planning new pop-up cycleways informed by a range of sources of evidence, including the Rapid Cycleway Prioritisation Tool, and in many cases the plans match the routes highlighted by our tool.\footnote{
  \href{https://www.bbc.co.uk/news/uk-england-leeds-52577554}{Kirkstall Road} in Leeds and \href{https://www.se16.com/6208-work-starts-on-54m-cycleway-along-jamaica-road}{Jamaica Road} in London are a couple of examples.
  Many more examples can be found on posts mentioning the tool on \href{https://twitter.com/search?q=cyipt.bike\%2Frapid}{social media}.}

\begin{figure}
\includegraphics[width=1\linewidth]{figures/facet-output} \caption{Maps showing existing, disjointed cycleway networks (green), potential cycleway routes on wide roads according to the Rapid Cycleway Prioritisation Tool (blue) and cohesive networks (purple) in 6 major cities}\label{fig:facet}
\end{figure}

The approach is not without limitations.
Its reliance on data rather than community engagement represents a rather top-down approach to transport planning.
To overcome this issue, future work could seek to incorporate the results such as those presented above into a participatory map of the type shown in Figure \ref{fig:commonplace}.
Further work could also extend the method in various ways, for example by refining estimates of cycling potential based on new parameters such as proximity to key destinations and estimates of road width.
We welcome feedback on the results and methods at \href{https://github.com/cyipt/popupCycleways}{github.com/cyipt/popupCycleways}.

A major advantage of the approach is that it is scalable.
It would be feasible to internationalise the approach, given sufficient computer and developer resource to overcome the key data barriers: lack to cycling potential data and lack of road width data that are specific to the UK.
In summary, the methods presented can help lock-in the benefits of COVID-19 related cycling booms long term.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-aldred_cycling_2018}{}%
Aldred, Rachel, Anna Goodman, John Gulliver, and James Woodcock. 2018. ``Cycling Injury Risk in London: A Case-Control Study Exploring the Impact of Cycle Volumes, Motor Vehicle Volumes, and Road Characteristics Including Speed Limits.'' \emph{Accident Analysis \& Prevention} 117 (August): 75--84. \url{https://doi.org/10.1016/j.aap.2018.03.003}.

\leavevmode\hypertarget{ref-chen_built_2016}{}%
Chen, Peng, and Qing Shen. 2016. ``Built Environment Effects on Cyclist Injury Severity in Automobile-Involved Bicycle Crashes.'' \emph{Accident Analysis \& Prevention} 86: 239--46.

\leavevmode\hypertarget{ref-freeman_covid19_2020}{}%
Freeman, Shirra, and Angela Eykelbosh. 2020. ``COVID-19 and Outdoor Safety: Considerations for Use of Outdoor Recreational Spaces.'' BC Centre for Disease Control.

\leavevmode\hypertarget{ref-goodman_scenarios_2019}{}%
Goodman, Anna, Ilan Fridman Rojas, James Woodcock, Rachel Aldred, Nikolai Berkoff, Malcolm Morgan, Ali Abbas, and Robin Lovelace. 2019. ``Scenarios of Cycling to School in England, and Associated Health and Carbon Impacts: Application of the `Propensity to Cycle Tool'.'' \emph{Journal of Transport \& Health} 12 (March): 263--78. \url{https://doi.org/10.1016/j.jth.2019.01.008}.

\leavevmode\hypertarget{ref-harrabin_boom_2020}{}%
Harrabin, Roger. 2020. ``Boom Time for Bikes as Virus Changes Lifestyles.'' \emph{BBC News}, May.

\leavevmode\hypertarget{ref-iacus_estimating_2020}{}%
Iacus, Stefano Maria, Fabrizio Natale, Carlos Satamaria, Spyridon Spyratos, and Michele Vespe. 2020. ``Estimating and Projecting Air Passenger Traffic During the COVID-19 Coronavirus Outbreak and Its Socio-Economic Impact.'' \emph{arXiv:2004.08460 {[}Physics, Stat{]}}, April. \url{http://arxiv.org/abs/2004.08460}.

\leavevmode\hypertarget{ref-jimenez-pavon_physical_2020}{}%
Jim\a'enez-Pav\a'on, David, Ana Carbonell-Baeza, and Carl J. Lavie. 2020. ``Physical Exercise as Therapy to Fight Against the Mental and Physical Consequences of COVID-19 Quarantine: Special Focus in Older People.'' \emph{Progress in Cardiovascular Diseases}, March. \url{https://doi.org/10.1016/j.pcad.2020.03.009}.

\leavevmode\hypertarget{ref-jittrapirom_exploratory_2020}{}%
Jittrapirom, Peraphan, and Garavig Tanaksaranond. 2020. ``An Exploratory Survey on the Perceived Risk of COVID-19 and Travelling.'' Preprint. SocArXiv. \url{https://doi.org/10.31235/osf.io/v3g5d}.

\leavevmode\hypertarget{ref-laker_world_2020}{}%
Laker, Laura. 2020. ``World Cities Turn Their Streets over to Walkers and Cyclists.'' \emph{The Guardian}, April.

\leavevmode\hypertarget{ref-litman_pandemicresilient_2020}{}%
Litman, Todd. 2020. ``Pandemic-Resilient Community Planning.'' Victoria Transport Policy Institute.

\leavevmode\hypertarget{ref-lovelace_propensity_2017}{}%
Lovelace, Robin, Anna Goodman, Rachel Aldred, Nikolai Berkoff, Ali Abbas, and James Woodcock. 2017. ``The Propensity to Cycle Tool: An Open Source Online System for Sustainable Transport Planning.'' \emph{Journal of Transport and Land Use} 10 (1). \url{https://doi.org/10.5198/jtlu.2016.862}.

\leavevmode\hypertarget{ref-orsman_covid_2020}{}%
Orsman, B. 2020. ``Covid 19 Coronavirus: Social Distancing Cones Rolled Out Across Auckland.'' \emph{NZ Herald}, April.

\leavevmode\hypertarget{ref-parkin_designing_2018}{}%
Parkin, John. 2018. \emph{Designing for Cycle Traffic: International Principles and Practice}. ICE Publishing.

\leavevmode\hypertarget{ref-reid_government_2020}{}%
Reid, Carlton. 2020. ``U.K. Government Boosts Bicycling and Walking with Ambitious Billion Post-Pandemic Plan.'' \emph{Forbes}.

\leavevmode\hypertarget{ref-salfordcitycouncil_salford_2020}{}%
Salford City Council. 2020. ``Salford Liveable Streets.'' https://salfordliveablestreets.commonplace.is.

\leavevmode\hypertarget{ref-tian_investigation_2020}{}%
Tian, Huaiyu, Yonghong Liu, Yidan Li, Chieh-Hsi Wu, Bin Chen, Moritz U. G. Kraemer, Bingying Li, et al. 2020. ``An Investigation of Transmission Control Measures During the First 50 Days of the COVID-19 Epidemic in China.'' \emph{Science} 368 (6491): 638--42. \url{https://doi.org/10.1126/science.abb6105}.

\leavevmode\hypertarget{ref-transportscotland_10_2020}{}%
Transport Scotland. 2020. `` Million to Support Pop-up Active Travel Infrastructure.'' https://www.transport.gov.scot/news/10-million-to-support-pop-up-active-travel-infrastructure/.
\end{cslreferences}

\end{document}

### 2020-06-22 article.tex
Update paper for #103

@@
 
-Datasets from the PCT and CyIPT project were merged, resulting in crucial variables summarised in Figure 1.
+Datasets from the PCT and CyIPT project were merged, resulting in crucial variables summarised in Table 1.
 Cycling potential is defined as the number of one-way journeys to work and to school, under a scenario in which the government aim of doubling cycling levels is met.
@@
 
-\begin{table}[H]
+\begin{table}
+
+\caption{\label{tab:t1}Summary of the road segment dataset for Leeds.}
 \centering
-\begin{tabular}{lllll}
+\begin{tabular}[t]{lllll}
 \toprule
@@
 \multicolumn{5}{l}{\textbf{N. observations}}\\
-\hspace{1em}N. observations & (N=4777) & (N=19241) & (N=3105) & (N=27123)\\
+\hspace{1em} & 4777 & 19241 & 3105 & 27123\\
 \addlinespace[0.3em]
@@
 
-The results of the method applied to the city of Leeds are shown in Figure \ref{fig:res} (see \href{https://www.cyipt.bike/rapid/west-yorkshire/}{cyipt.bike/rapid} for interactive version) and Table 1.
+The results of the method applied to the city of Leeds are shown in Figure \ref{fig:res} (see \href{https://www.cyipt.bike/rapid/west-yorkshire/}{cyipt.bike/rapid} for interactive version) and Table 2.
 We found that analysis of open transport network data, alongside careful selection of parameters, can generate plausible results for the prioritisation of pop-up cycle infrastructure.

### 2020-06-22 article.tex
Update latex version

@@
   pdftitle={Methods to prioritise pop-up active transport infrastructure and their application in a national cycleway prioritisation tool},
-  pdfauthor={Robin Lovelace, Malcolm Morgan, Joey Talbot, Martin-Lucas-Smith},
+  pdfauthor={Robin Lovelace, Joey Talbot, Malcolm Morgan, Martin-Lucas-Smith},
   hidelinks,
@@
 \title{Methods to prioritise pop-up active transport infrastructure and their application in a national cycleway prioritisation tool}
-\author{Robin Lovelace, Malcolm Morgan, Joey Talbot, Martin-Lucas-Smith}
+\author{Robin Lovelace, Joey Talbot, Malcolm Morgan, Martin-Lucas-Smith}
 \date{}
@@
 \item
-  `filtered permeability', where measures such as point closures or contraflow cycle lanes are used to design street networks in which routes for cyclists are faster and more direct than routes for drivers, e.g.~as \href{https://twitter.com/TowerHamletsNow/status/1257564043856019458}{shown} in {[}Tower Hamlets{]}
+  Measures such as point closures or contraflow cycle lanes, which can be used to promote `filtered permeability', a strategy in which street networks are redesigned so that routes for cyclists are faster and more direct than routes for drivers.
+An example of this is \href{https://twitter.com/TowerHamletsNow/status/1257564043856019458}{shown} in {[}Tower Hamlets{]}.
 \item
-  banning cars and to pedestrianise streets, as in New York's `Open Streets' scheme (Litman 2020)
+  Measures to close roads entirely to motor traffic or to particular classes of vehicle, either permanently or at certain times of day, as in New York's `Open Streets' scheme (Litman 2020).
 \item
-  the reallocation of one or more lanes on wide roads to create pop-up cycleways and pavements (Orsman 2020).
+  Measures to reallocate space on wide roads to create new cycleways or to widen pavements (Orsman 2020).
 \end{enumerate}
@@
 Because of the recent, localised and often ad-hoc nature of pop-up infrastructure, it is difficult to make, let alone test, hypotheses related to the research question.
-Our broad hypothesis is that digital tools based on open data, and crowdsourcing such as the interactive map used to support community-level responses to COVID-19 in Salford (Salford City Council 2020), illustrated in Figure 1, can lead to more effective use of resources allocated to pop-up interventions.
+Our broad hypothesis is that digital tools based on open data, combined with crowdsourcing such as the interactive map used to support community-level responses to COVID-19 in Salford (Salford City Council 2020), illustrated in Figure 1, can lead to more effective use of resources allocated to pop-up interventions.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{figures/widenmypath-bristol} \includegraphics[width=1\linewidth]{figures/saferstreets} \caption{Screenshots from the websites widenmypath.com (top) and salfordliveablestreets.commonplace.is (bottom) to support local responses to the COVID-19 pandemic, including the prioritisation of pop-up active transport infrastructure.}\label{fig:commonplace}
+\includegraphics[width=1\linewidth]{figures/widenmypath} \includegraphics[width=1\linewidth]{figures/saferstreets} \caption{Screenshots from the websites widenmypath.com (top) which includes top cycle route recommendations generated using the methods outlined in this paper in an open web forum, and salfordliveablestreets.commonplace.is (bottom) to support local responses to the COVID-19 pandemic, including the prioritisation of pop-up active transport infrastructure.}\label{fig:commonplace}
 \end{figure}
@@
 
-Roads are classified by speed limit because this has been shown to be a key factor associated with the incidence of severe injuries and fatalities of cyclists (Chen and Shen 2016), with odds of cyclist injury on 20 mph roads in London found to be 21\% lower than on 30 mph roads (Aldred et al.
+Roads are classified by speed limit because this has been shown to be a key factor associated with the incidence of severe injuries and fatalities of cyclists (Chen and Shen 2016), with odds of cyclist injury on 20 mph (32.2 kmph) roads in London found to be 21\% lower than on 30 mph (48.3 kmph) roads (Aldred et al.
 2018).
@@
 \toprule
- & 20 mph or less & 30 mph & 40+ mph & Overall\\
+ & 20 mph (32.2 kmph) or less & 30 mph (48.3 kmph) & 40+ mph (64.4 kmph) & Overall\\
 \midrule

### 2020-06-22 article.tex
Close #107

@@
 
-\caption{\label{tab:t1}Summary of the road segment dataset for Leeds.}
+\caption{\label{tab:t1}Summary of the road segment dataset for Leeds. Units of speed are in miles per hour (mph), with 20, 30 and 40 mph representing 32, 48 and 64 kilometers per hour respectively.}
 \centering
@@
 \toprule
- & 20 mph (32.2 kmph) or less & 30 mph (48.3 kmph) & 40+ mph (64.4 kmph) & Overall\\
+ & 20 mph or less & 30 mph & 40+ mph & Overall\\
 \midrule

### 2020-06-22 article.tex
Final tweaks

@@
 We developed methods to condense the complexity of city networks down to the most promising roads for road space reallocation schemes.
-The resulting Rapid Cycleway Prioritisation Tool has been deployed for all cities England to help prioritise emergency funds for new cycleways nationwide.
+The resulting Rapid Cycleway Prioritisation Tool has been deployed for all local and regional transport authorities in England to help prioritise emergency funds for new cycleways nationwide.
 The methods and concepts could be used to support investment in pop-up infrastructure in cities worldwide.
@@
 2020; Jittrapirom and Tanaksaranond 2020) but short distance travel patterns have also changed.
-There has been a notable increase in cycling in some areas (Harrabin 2020) due to the need increased need for exercise close to home for mental and physical health (Jim\a'enez-Pav\a'on, Carbonell-Baeza, and Lavie 2020) and a reduction in public transport options (e.g.~Tian et al.
+There has been a notable increase in cycling in some areas (Harrabin 2020) due to the increased need for exercise close to home for mental and physical health (Jim\a'enez-Pav\a'on, Carbonell-Baeza, and Lavie 2020) and a reduction in public transport options (e.g.~Tian et al.
 2020).
@@
 Local and national governments are working out how best to respond.
-Many options are available to ensure that citizens can benefit from outdoor activity while minimising health risks, ranging from the hand sanitiser provision to the creation of extra active transport space (Freeman and Eykelbosh 2020).
+Many options are available to ensure that citizens can benefit from outdoor activity while minimising health risks, ranging from hand sanitiser provision to the creation of extra active transport space (Freeman and Eykelbosh 2020).
 Installation of `pop-up' active transport infrastructure has been endorsed and implemented in many places (Laker 2020).
@@
 \item
-  Measures to close roads entirely to motor traffic or to particular classes of vehicle, either permanently or at certain times of day, as in New York's `Open Streets' scheme (Litman 2020).
+  Measures to close roads entirely to cars, either permanently or at certain times of day, as in New York's `Open Streets' scheme (Litman 2020).
 \item
@@
 Its reliance on data rather than community engagement represents a rather top-down approach to transport planning.
-To overcome this issue, future work could seek to incorporate the results such as those presented above into a participatory map of the type shown in Figure \ref{fig:commonplace}.
-Further work could also extend the method in various ways, for example by refining estimates of cycling potential based on new parameters such as proximity to key destinations and estimates of road width.
+The incorporation of our results into participatory maps such as the one presented in Figure \ref{fig:commonplace} will help to mitigate this limitation.
+Further work could also extend the method in various ways, for example by refining estimates of cycling potential based on additional parameters such as proximity to key destinations.
 We welcome feedback on the results and methods at \href{https://github.com/cyipt/popupCycleways}{github.com/cyipt/popupCycleways}.

### 2020-06-22 article.tex
Mention units in caption

@@
 \includegraphics[width=1\linewidth]{article_files/figure-latex/gsub-1} \caption{Illustration of geographic subsetting based on administrative boundaries (left), distance to a central point (middle) and distance to city centre or key destinations (right).
-Radii of 5 km, 8 km and 10 km are shown for reference (note that some roads within 10 km of the center are outside the regional boundary).}\label{fig:gsub}
+Radii of 5 km, 8 km and 10 km are shown for reference (note that some roads within 10 km of the center are outside the regional boundary).
+Units of speed are in miles per hour (mph), with 20, 30 and 40 mph representing 32, 48 and 64 kilometers per hour respectively.}\label{fig:gsub}
 \end{figure}

### 2020-06-23 article.tex
Update paper
Fix typos in .tex and .md versions

@@
 The resulting Rapid Cycleway Prioritisation Tool has been deployed for all local and regional transport authorities in England to help prioritise emergency funds for new cycleways nationwide.
-The methods and concepts could be used to support investment in pop-up infrastructure in cities worldwide.
+The approach could be used to support investment in pop-up infrastructure in cities worldwide.
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{figures/widenmypath} \includegraphics[width=1\linewidth]{figures/saferstreets} \caption{Screenshots from the websites widenmypath.com (top) which includes top cycle route recommendations generated using the methods outlined in this paper in an open web forum, and salfordliveablestreets.commonplace.is (bottom) to support local responses to the COVID-19 pandemic, including the prioritisation of pop-up active transport infrastructure.}\label{fig:commonplace}
+\includegraphics[width=1\linewidth]{figures/jpg/widenmypath} \includegraphics[width=1\linewidth]{figures/jpg/saferstreets} \caption{Screenshots from the websites widenmypath.com (top) which includes top cycle route recommendations generated using the methods outlined in this paper in an open web forum, and salfordliveablestreets.commonplace.is (bottom) to support local responses to the COVID-19 pandemic, including the prioritisation of pop-up active transport infrastructure.}\label{fig:commonplace}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{figures/results-top-leeds} \caption{Results, showing road segments with a spare lane (light blue) and road groups with a minium threshold length, 1km in this case (dark blue).
+\includegraphics[width=1\linewidth]{figures/results-top-leeds} \caption{Results of the Rapid Cycleway Prioritisation Tool, showing road segments with a spare lane (light blue) and road groups with a minium threshold length, 1km in this case (dark blue).
 The top 10 road groups are labelled.}\label{fig:res}
@@
 A major advantage of the approach is that it is scalable.
-It would be feasible to internationalise the approach, given sufficient computer and developer resource to overcome the key data barriers: lack to cycling potential data and lack of road width data that are specific to the UK.
+It would be feasible to internationalise the approach, given sufficient computer and developer resource to overcome the key data barriers: lack of cycling potential and road width data.
 In summary, the methods presented can help lock-in the benefits of COVID-19 related cycling booms long term.

### 2020-06-23 article.tex
Minor fix in description of findings

@@
 We found that analysis of open transport network data, alongside careful selection of parameters, can generate plausible results for the prioritisation of pop-up cycle infrastructure.
-Reducing the 85,000 road segments for Leeds down to a handful of candidate segments with more than 1 lane near key destinations has great potential to support policy-makers, especially when decisions need to be made fast.
+Reducing tens of thousands of road segments Leeds down to a handful of promising named road sections has great potential to support policy-makers, especially when decisions need to be made fast.
 
